---
title: 'Sovereign ETF Intelligence'
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: 
      version: 4
      bg: '#F8F9FA'
      fg: '#1A1A1A'
      primary: '#003049'
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(PerformanceAnalytics)
library(scales)
library(here)

if(!exists('project_tree')) source(here('project_tree.R'))
if(file.exists(here(project_tree$scripts$advanced_reporting))) {
  source(here(project_tree$scripts$advanced_reporting))
}
```

Main Dashboard
=====================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
### Global Controls

```{r}
checkboxInput('outlier_adj', 'Enable Outlier Management', value = TRUE)
helpText('Sovereign Intelligence Suite v3.2')
helpText('Status: Technical Signal Engine Active')
```

Row {data-height=600}
-----------------------------------------------------------------------
### Analytical Sigma: Statistical Shock Analysis (Daily Z-Score)
```{r}
renderPlot({ p_sigma })
```

Row {data-height=400}
-----------------------------------------------------------------------
### Nominal Performance View: Actual YTD % (Price-Based)
```{r}
renderPlot({ p_quad_nominal })
```
### Annualized Efficiency Frontier
```{r}
renderPlot({ p_eff })
```

Technicals Page
=====================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
### Technical Logic

**Exhaustion Signal:** Price stretch (>10MA) combined with extreme Sigma (>2Z).

**Strength Buy:** Price < 10MA in a long-term uptrend (200MA > 0).

```{r}
cats <- sort(unique(standard_metadata$category))
checkboxGroupInput('tech_cats', 'Select Strategies:', choices = cats, selected = cats)
```

Row
-----------------------------------------------------------------------
### Mean Reversion & Exhaustion Quadrant (Sigma vs 10MA Stretch)

```{r}
filtered_tech <- reactive({
  unified_mom_df %>% filter(category %in% input$tech_cats)
})

renderPlot({
  plot_technical_exhaustion_quadrant(filtered_tech())
})
```

Row {data-height=350}
-----------------------------------------------------------------------
### Technical Pulse: Momentum Clusters & Action Signals

```{r}
renderTable({
  filtered_tech() %>% 
    select(ticker, momentum_cluster, action_signal, dist_10, dist_50, dist_200, trend_regime) %>% 
    mutate(
      `10d dist` = scales::percent(dist_10, accuracy = 0.1),
      `50d dist` = scales::percent(dist_50, accuracy = 0.1),
      `200d dist` = scales::percent(dist_200, accuracy = 0.1)
    ) %>%
    select(ticker, momentum_cluster, action_signal, `10d dist`, `50d dist`, `200d dist`, trend_regime) %>%
    rename(Ticker = ticker, Cluster = momentum_cluster, Action = action_signal, Regime = trend_regime)
})
```

Momentum Quadrant (Z)
=====================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
### Statistical Filters

```{r}
checkboxGroupInput('quad_cats', 'Select Strategies:', choices = cats, selected = cats)
```

Row
-----------------------------------------------------------------------
### Interactive Momentum Developing Quadrant (Z-Scores)

```{r}
filtered_mom <- reactive({
  unified_mom_df %>% filter(category %in% input$quad_cats)
})

renderPlot({
  plot_nominal_quadrant(filtered_mom())
})
```

Row {data-height=250}
-----------------------------------------------------------------------
### Data Pulse (Z-Scores)

```{r}
renderTable({
  filtered_mom() %>% 
    select(ticker, category, z_ytd, z_score) %>% 
    arrange(desc(z_ytd)) %>%
    rename(Ticker = ticker, Strategy = category, 'Z (YTD)' = z_ytd, 'Z (Daily)' = z_score)
})
```
