---
title: "Calendar Year Analog Projection"
subtitle: "Trend vs Reversal Probability from Historical Analogs"
author: "Sovereign Intelligence"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    fig_width: 11
    fig_height: 6
  pdf_document:
    latex_engine: xelatex
    toc: true
    fig_width: 11
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center"
)

library(tidyverse)
library(scales)
library(here)

source(here("scripts/00_init_universe.R"))
source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
source(here("utility/plot_calendar_perf.R"))
source(here("utility/plot_calendar_analog.R"))

xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))
xts_rel <- readRDS(here("02_data_processed/xts_rel.rds"))
rt      <- build_regime_table(xts_ret[, "SPY"], t_fall = 0.10)

TICKERS <- c("SPY", "QQQ", "GLD", "TLT", "XLK", "SMH", "XLF", "XLI")
```

---

## 1 · The Question

> *"Given what has happened so far this year, how likely is it that the current
> trend continues to year-end — and how likely is a reversal?"*

This report addresses that question using two complementary tools:

1. **Transition matrices** — what fraction of historical Up/Down years were
   followed by another Up/Down year, conditional on the market regime
2. **Historical analog matching** — find the calendar years whose YTD return
   (measured at today's day-of-year) most closely resembles the current year,
   then project forward using the distribution of where those years ended up

Both tools are honest about sample size: with ~15 years of data per ticker,
confidence intervals are wide. The value is in **directional signal and
regime context**, not precise probability estimates.

---

## 2 · Academic Background

### 2.1 What the literature says about annual return predictability

| Finding | Source | Implication |
|---|---|---|
| 12-month momentum persists at 3–12m horizons | Jegadeesh & Titman (1993) | Trend signal valid within a calendar year |
| 3–5 year reversals after sustained runs | De Bondt & Thaler (1985) | Fade long winning streaks across years |
| 12-month seasonal autocorrelation | Heston & Sadka (2008) | Same-month-last-year has predictive power |
| Momentum crashes in Recovery regimes | Daniel & Moskowitz (2016) | Regime conditions the trend/reversal call |
| Window dressing inflates year-end winner prices | Lakonishok et al. (1991) | Q4 momentum may overstate alpha for growth ETFs |
| Minimum 3–5 years to distinguish skill from luck | Grinold & Kahn | Annual t-stats are wide; treat probabilities as soft |

### 2.2 Why the annual horizon is special

Calendar year boundaries create **artificial discontinuities**:

- Institutional flows reset at year-end (window dressing, tax-loss harvesting)
- Performance measurement resets to 0% on January 1
- Momentum measured at 12-month lags has documented seasonality (Heston & Sadka)

This means the calendar-year trend/reversal question is **not the same** as a
rolling 12-month question. The year-boundary itself affects behaviour.

### 2.3 Regime conditioning matters

Daniel & Moskowitz (2016) show momentum strategies crash specifically in
**Recovery regimes** — when the market rebounds sharply after a Fall.
In those environments, what looked like a strong trend (e.g. defensive
outperformance in the Fall) reverses violently.

This motivates computing transition probabilities **separately** for each
of our three regimes: Fall, Recovery, and Consolidation.

---

## 3 · Method A — Transition Matrices

### 3.1 How it works

For each ticker, we:

1. Compute the full-year calendar return for every year in the history
2. Classify each year as **Up** (≥ 0%) or **Down** (< 0%)
3. Build a 2×2 transition matrix: P(next year Up | this year Up), etc.
4. Repeat separately for years that occurred in each regime

The lag-1 autocorrelation $\rho$ of annual returns gives a single summary
statistic:

- $\rho > 0$ → **momentum** (trend more likely than reversal)
- $\rho < 0$ → **mean reversion** (reversal more likely than trend)
- $|\rho| \leq 0.1$ → **no signal** (coin-flip at annual horizon)

### 3.2 Results

```{r transitions, results='asis'}
for (tk in TICKERS) {
  cat(sprintf("\n\n#### %s\n\n", tk))
  cat("```\n")
  compute_transitions(xts_ret[, tk], rt)
  cat("```\n\n")
}
```

### 3.3 Interpretation guide

The transition matrix is useful for **framing the prior** before looking at
the current year's YTD. Key patterns to look for:

- **P(Up|Up) > 60%** — meaningful momentum tendency; current up-year is more
  likely to finish up than down
- **P(Up|Down) > 60%** — strong recovery pattern; down years tend to bounce
- **Regime divergence** — if P(Up|Up) in Consolidation >> P(Up|Up) in Fall,
  the regime matters more than the prior year's direction

> **Caution:** With ~15 years of data, a "70%" estimate has a 95% confidence
> interval of roughly ±25%. These are soft signals, not actuarial tables.

---

## 4 · Method B — Historical Analog Matching

### 4.1 How it works

**Step 1 — Establish position in the year**
Find today's trading day number within the calendar year (e.g. day 55 of ~252).

**Step 2 — Compute each historical year's YTD at that same day**
For every past year, compute the cumulative return from Jan 1 to trading day 55.

**Step 3 — Find analogs**
Select years where the YTD return at that day was within ±6 percentage points
of the current year's YTD. If fewer than 3 analogs exist, widen the window
in 2pp steps until at least 3 are found.

**Step 4 — Project forward**
For each analog year, observe what happened from day 55 to year-end.
Compute:

$$P(\text{Trend}) = \frac{\text{analog years that ended in same direction as current YTD}}{\text{total analog years}}$$

**Step 5 — Build the cone**
Plot the full-year paths of analog years. From today's day forward, shade the
10–90th percentile band (outer cone) and 25–75th percentile band (inner cone).
The dashed line is the median analog path.

### 4.2 What the chart shows

```
Grey lines    = all historical years (context)
Blue lines    = analog years (similar YTD at today's point in the year)
Inner cone    = 25–75th percentile range of where analogs went from here
Outer cone    = 10–90th percentile range
Dashed blue   = median analog path forward
Red line      = current year (stops at today)
Dotted vert.  = "today" marker
Annotation    = P(Trend), P(Reversal), analog year list, current regime
```

### 4.3 Why analog matching is better than autocorrelation alone

Autocorrelation uses all years equally. Analog matching **conditions on
magnitude**: a year that is +15% YTD in March is not the same as a year
that is +1% YTD in March, even though both are "Up". The analog method
naturally respects this.

It is closest in spirit to the **ensemble forecast** approach used in
meteorology — where today's atmospheric state is matched to historical
states, and the distribution of subsequent outcomes is used as the forecast.

---

## 5 · Analog Projection Charts

```{r analog-plots, fig.height=7, results='asis'}
for (tk in TICKERS) {
  cat(sprintf("\n\n### %s\n\n", tk))
  print(plot_calendar_analog(xts_ret[, tk], rt))
  cat("\n\n")
}
```

---

## 6 · Calendar Year History (Reference)

Full calendar year paths for all tickers — useful for visually identifying
which years were the analogs and how they evolved.

```{r cal-history, fig.height=9}
plot_multi_calendar_perf(xts_ret[, TICKERS], ncol = 2)
```

---

## 7 · Limitations & Honest Caveats

| Limitation | Why it matters |
|---|---|
| **Small sample** | ~15 annual observations per ticker. A "70% P(trend)" has ±25% confidence interval. |
| **Non-stationarity** | Regime structure, correlations, and volatility change over time. Historical analogs may not reflect current market structure. |
| **Calendar boundary artefact** | Transition probabilities are measured at Dec 31 — a date that has institutional significance (rebalancing, flows) not shared by other dates. |
| **Partial year extrapolation** | The analog method assumes the remainder of the year is drawn from the same distribution as historical analog years. Structural breaks (policy changes, exogenous shocks) can invalidate this. |
| **No factor conditioning** | Analogs are matched on raw return only, not on macro regime, valuation, or rate environment. Adding regime conditioning (from `rt`) would improve signal. |

---

## 8 · How to Use in Practice

**Step 1 — Check the regime first**
The current regime (shown in the annotation box) tells you which part of
the transition matrix is relevant. A Consolidation regime with P(trend)=70%
is more reliable than the same number in a Fall regime.

**Step 2 — Look at the cone width**
A wide cone = high uncertainty = do not trade on the signal alone.
A narrow cone with most analogs clustered = higher confidence.

**Step 3 — Check n analogs**
If n < 5, the probabilities are almost meaningless statistically.
Use the chart as qualitative context only.

**Step 4 — Combine with regime analysis**
Cross-reference with `plot_rel_calendar_perf()` to see whether the ticker's
current year performance is being driven by absolute return or by alpha vs SPY.
A ticker that is up primarily because SPY is up (small gap between panels)
is more vulnerable to reversal in a regime shift.

---

## 9 · Scripts & Reproducibility

| File | Purpose |
|---|---|
| `utility/plot_calendar_analog.R` | `compute_transitions()` and `plot_calendar_analog()` |
| `utility/plot_calendar_perf.R` | `plot_calendar_perf()`, `plot_multi_calendar_perf()`, `plot_rel_calendar_perf()` |
| `scripts_spy_dd_regime/spy_dd_regime.R` | `build_regime_table()` |

To refresh this report with today's data:

```r
rmarkdown::render("Rmd/calendar_analog_report.Rmd")
```

To run interactively for a single ticker:

```r
source("utility/plot_calendar_analog.R")
rt <- build_regime_table(xts_ret[, "SPY"], t_fall = 0.10)

compute_transitions(xts_ret[, "GLD"], rt)
plot_calendar_analog(xts_ret[, "GLD"], rt)
plot_calendar_analog(xts_ret[, "GLD"], rt, window_pct = 0.10)  # wider window
```

---

*Report generated: `r format(Sys.Date(), "%d %B %Y")` |
Data: daily log returns via `xts_ret_returns.rds` |
Regime engine: `build_regime_table()` with T_FALL = 10%*
