---
title: "Phase 2.2 — Overlay Enhancer Screen"
subtitle: "Baseline: 6040Classic (60% SPY + 40% IEF)"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
params:
  w_eq:           0.60
  w_fi:           0.40
  ir_floor:       0.20
  fall_tolerance: -0.02
  roll_win:       63
  n_select:       8
  rho_max:        1.00
---

```{css, echo=FALSE}
.finding-box {
  background: #f0f9ff;
  border-left: 4px solid #2563eb;
  padding: 10px 16px;
  margin: 12px 0;
  font-size: 0.92em;
}
.type-a { background:#dcfce7; border-left:4px solid #16a34a; padding:6px 14px; margin:6px 0; }
.type-b { background:#fef9c3; border-left:4px solid #ca8a04; padding:6px 14px; margin:6px 0; }
.type-c { background:#ede9fe; border-left:4px solid #7c3aed; padding:6px 14px; margin:6px 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 6
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("project_tree"))       source(here("project_tree.R"))
if (!exists("etf_metadata"))       source(here(project_tree$scripts$init))
if (!exists("xts_ret"))            source(here(project_tree$scripts$wrangle))
if (!exists("build_regime_table")) source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
if (!exists("rt") || !is.data.frame(rt))
  rt <- build_regime_table(xts_ret[, "SPY"])
if (!exists("plot_xts_pair"))      source(here("utility/plot_xts_pair.R"))

# Override script-level params from Rmd params before sourcing
W_EQ           <- params$w_eq
W_FI           <- params$w_fi
IR_FLOOR       <- params$ir_floor
FALL_TOLERANCE <- params$fall_tolerance
ROLL_WIN       <- as.integer(params$roll_win)
N_SELECT       <- as.integer(params$n_select)
RHO_MAX        <- params$rho_max
BASELINE_LABEL <- sprintf("%d/%d Classic",
                           round(params$w_eq * 100), round(params$w_fi * 100))

# Source the screen — auto-run block is guarded by knitr.in.progress
source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))
# After sourcing, these objects are available:
#   enhancer_candidates  — Stage 1 watchlist + trigger_type (A/B)
#   enhancer_scores      — Stage 2 context scores + trigger_type_final (A/B/C)
#   selected_enhancers   — Stage 3 final selection
#
# To replicate interactively in RStudio:
#   enhancer_candidates <- build_enhancer_candidates(xts_ret, rt)
#   enhancer_scores     <- score_current_context(enhancer_candidates, xts_ret, rt)
#   selected_enhancers  <- select_enhancers(enhancer_scores, xts_ret)
```

---

## 1. Framework

<div class="finding-box">
**Phase 1.0 baseline fixed:** `r BASELINE_LABEL` = `r round(params$w_eq*100)`% SPY + `r round(params$w_fi*100)`% IEF, buy-and-hold.

An enhancer at weight *w* means: **(1 − w) × `r BASELINE_LABEL` + w × TICKER**

All spreads, IR, and excess returns on this page are computed vs `r BASELINE_LABEL`, not raw SPY.
Regime detection uses SPY via `build_regime_table()` — unchanged.
</div>

### Trigger types

<div class="type-a">**Type A — Unconditional.** Fall excess vs baseline ≥ `r params$fall_tolerance*100`pp. Hold at fixed weight at all times.</div>
<div class="type-b">**Type B — Regime-conditional.** Fall excess < `r params$fall_tolerance*100`pp. Hold only during Recovery + Consolidation; revert to `r BASELINE_LABEL` in Fall.</div>
<div class="type-c">**Type C — 200DMA signal.** Type B ticker with live `trend_signals` data. Use 200DMA trend filter instead of hard regime switch — finer entry/exit, whipsaw cost applies.</div>

---

## 2. Stage 1 — Candidate Watchlist

```{r stage1-table}
enhancer_candidates %>%
  mutate(
    full_ir         = round(full_ir, 2),
    full_excess_ann = scales::percent(full_excess_ann, accuracy = 0.1),
    fall_excess_ann = scales::percent(fall_excess_ann, accuracy = 0.1),
    recov_excess_ann= scales::percent(recov_excess_ann, accuracy = 0.1),
    cons_excess_ann = scales::percent(cons_excess_ann, accuracy = 0.1)
  ) %>%
  select(
    Ticker       = ticker,
    Trigger      = trigger_type,
    Name         = name,
    `Asset Class`= asset_class,
    `Full IR`    = full_ir,
    `Full Excess`= full_excess_ann,
    `Fall α`     = fall_excess_ann,
    `Recov α`    = recov_excess_ann,
    `Cons α`     = cons_excess_ann,
    Days         = n_days
  ) %>%
  kbl(caption = sprintf("Stage 1 candidates vs %s (IR floor = %.2f)", BASELINE_LABEL, IR_FLOOR)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), font_size = 12) %>%
  row_spec(which(enhancer_candidates$trigger_type == "A"),
           background = "#f0fdf4") %>%
  row_spec(which(enhancer_candidates$trigger_type == "B"),
           background = "#fefce8")
```

**`r nrow(enhancer_candidates)` candidates admitted** —
`r sum(enhancer_candidates$trigger_type == "A")` Type A (unconditional),
`r sum(enhancer_candidates$trigger_type == "B")` Type B (regime-conditional).

---

## 3. Stage 2 — Context Scores

Current SPY regime: **`r unique(enhancer_scores$cur_regime)`**

```{r stage2-table}
enhancer_scores %>%
  mutate(
    context_score  = round(context_score, 1),
    rolling_ir     = round(rolling_ir, 2),
    regime_excess  = scales::percent(regime_excess, accuracy = 0.1),
    spread_momentum= round(spread_momentum, 5),
    trend_f        = if_else(is.na(trend_on), "n/a",
                             if_else(trend_on == 1, "✓ ON", "✗ OFF"))
  ) %>%
  select(
    Ticker         = ticker,
    `Trigger Final`= trigger_label,
    `Ctx Score`    = context_score,
    `Roll IR 63d`  = rolling_ir,
    `Regime α`     = regime_excess,
    `200DMA`       = trend_f,
    Momentum       = spread_momentum,
    `Asset Class`  = asset_class
  ) %>%
  kbl(caption = sprintf("Stage 2 context scores — regime: %s | baseline: %s",
                        unique(enhancer_scores$cur_regime), BASELINE_LABEL)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), font_size = 12) %>%
  row_spec(which(enhancer_scores$trigger_type_final == "A"), background = "#f0fdf4") %>%
  row_spec(which(enhancer_scores$trigger_type_final == "B"), background = "#fefce8") %>%
  row_spec(which(enhancer_scores$trigger_type_final == "C"), background = "#f5f3ff")
```

---

## 4. Stage 3 — Selected Enhancers

```{r stage3-table}
selected_enhancers %>%
  mutate(
    context_score   = round(context_score, 1),
    full_ir         = round(full_ir, 2),
    fall_excess_ann = scales::percent(fall_excess_ann, accuracy = 0.1),
    recov_excess_ann= scales::percent(recov_excess_ann, accuracy = 0.1),
    rolling_ir      = round(rolling_ir, 2),
    regime_excess   = scales::percent(regime_excess, accuracy = 0.1),
    trend_f         = if_else(is.na(trend_on), "n/a",
                              if_else(trend_on == 1, "✓ ON", "✗ OFF"))
  ) %>%
  select(
    `#`            = rank,
    Ticker         = ticker,
    Trigger        = trigger_label,
    `Ctx Score`    = context_score,
    `Full IR`      = full_ir,
    `Fall α`       = fall_excess_ann,
    `Recov α`      = recov_excess_ann,
    `Roll IR 63d`  = rolling_ir,
    `Regime α`     = regime_excess,
    `200DMA`       = trend_f,
    `Asset Class`  = asset_class
  ) %>%
  kbl(caption = sprintf("Stage 3 selection — top %d enhancers vs %s",
                        nrow(selected_enhancers), BASELINE_LABEL)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), font_size = 13) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(which(selected_enhancers$trigger_type_final == "A"), background = "#f0fdf4") %>%
  row_spec(which(selected_enhancers$trigger_type_final == "B"), background = "#fefce8") %>%
  row_spec(which(selected_enhancers$trigger_type_final == "C"), background = "#f5f3ff")
```

---

## 5. Sleeve Wealth Charts

One chart per asset class. Grey line = `r BASELINE_LABEL`. Coloured lines = selected enhancers. Background thin grey = other Stage 1 candidates in the same sleeve.

```{r sleeve-charts, echo=TRUE, fig.height=7, fig.width=13}
plot_enhancer_sleeves(
  selected_enhancers,
  xts_ret,
  rt,
  w_eq                = W_EQ,
  w_fi                = W_FI,
  baseline_label      = BASELINE_LABEL,
  show_candidates     = TRUE,
  enhancer_candidates = enhancer_candidates
)
```

---

## 6. Pair Analysis — Selected Enhancers vs `r BASELINE_LABEL`

Each panel = 3-plot patchwork (wealth | spread | drawdown). Regime shading: pink = Fall, green = Recovery.

```{r pair-plots, echo=TRUE, fig.height=21, fig.width=13, results='asis'}
for (i in seq_len(nrow(selected_enhancers))) {
  row <- selected_enhancers[i, ]
  tk  <- row$ticker
  trig <- row$trigger_label

  cat(sprintf("\n### %d. %s — %s\n\n", i, tk, trig))

  baseline_xts <- .build_baseline(xts_ret, W_EQ, W_FI, BASELINE_LABEL)
  pfx <- sprintf("[#%d | Score %.1f | %s | %s regime]",
                 row$rank, row$context_score, trig, row$cur_regime)

  plot_xts_pair(
    xts1         = xts_ret[, tk],
    xts2         = baseline_xts,
    label1       = tk,
    label2       = BASELINE_LABEL,
    regime_tbl   = rt,
    title_prefix = pfx
  )
}
```

---

## 7. Parameters used

```{r params-table, echo=FALSE}
tibble(
  Parameter      = c("Baseline", "w_eq", "w_fi", "ir_floor", "fall_tolerance",
                     "roll_win", "n_select", "rho_max"),
  Value          = c(BASELINE_LABEL,
                     params$w_eq, params$w_fi, params$ir_floor, params$fall_tolerance,
                     params$roll_win, params$n_select, params$rho_max),
  Description    = c(
    "Phase 1.0 SAA baseline label",
    "SPY weight in baseline",
    "IEF weight in baseline",
    "Min full-period IR vs baseline for Stage 1 admission",
    "Max Fall excess loss vs baseline for Type A classification",
    "Rolling window for IR and spread momentum (days)",
    "Max tickers selected in Stage 3",
    "Max pairwise ρ between selected tickers (1.0 = off)"
  )
) %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped","condensed"), font_size = 12)
```
