# ── Render from console ───────────────────────────────────────────────────────
# Full equity universe (all g_eq_tickers in xts_ret):
#   rmarkdown::render(
#     here::here("Rmd/eq_factsheets_batch.Rmd"),
#     output_file = here::here("03_reports/eq_factsheets_batch.html")
#   )
#
# Quick test — 3 tickers, no 8-panel (fast, ~2 min):
#   rmarkdown::render(..., params = list(tickers = c("XLK","QQQ","GLD")))
#
# Full run with 8-panel factsheets (~45 min):
#   rmarkdown::render(..., params = list(show_full_factsheet = TRUE))
#
# Custom benchmark:
#   rmarkdown::render(..., params = list(bmk = "URTH", t_fall = 0.10))
# ─────────────────────────────────────────────────────────────────────────────

---
title: "Equity ETF Factsheets — Full Universe"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 2
    theme: flatly
    highlight: tango
    fig_caption: true
params:
  bmk:               "SPY"
  t_fall:            0.10
  tickers:           !r NULL   # NULL = full g_eq_tickers; or c("XLK","QQQ","GLD")
  show_full_factsheet: false    # true = include 8-panel combined (slow); false = skip
---

```{css, echo=FALSE}
.caption      { color: #1B3A6B; font-size: 0.82em; font-style: italic;
                border-left: 3px solid #1B3A6B; padding-left: 6px; margin-top: 4px; }
.section-div  { border-top: 2px solid #dee2e6; margin: 48px 0 24px 0; }
.ticker-badge { background: #F0F4FA; border-left: 4px solid #1B2A4A;
                padding: 7px 14px; margin-bottom: 14px; font-size: 0.88em; color: #333; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

select <- dplyr::select

# ── params fallback (interactive / render-without-params) ─────────────────────
if (!exists("params")) params <- list(bmk = "SPY", t_fall = 0.10,
                                      tickers = NULL, show_full_factsheet = FALSE)

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
if (!exists("etf_metadata"))
  source(here("scripts/00_init_universe.R"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

source(here("key_plots/chart_regime_rel_overlay.R"))   # always reload — contains pp fix

source(here("utility/factsheet_functions.R"))

# ── Regime table (benchmark) ──────────────────────────────────────────────────
rt <- build_regime_table(xts_ret[, params$bmk], params$t_fall)

# ── Equity tickers: in universe AND in xts_ret, excluding the benchmark ───────
eq_tickers <- intersect(g_eq_tickers, colnames(xts_ret))
eq_tickers <- eq_tickers[eq_tickers != params$bmk]
if (!is.null(params$tickers)) eq_tickers <- intersect(params$tickers, eq_tickers)
```

## Universe overview

```{r overview-table}
overview <- etf_metadata %>%
  filter(ticker %in% eq_tickers) %>%
  select(ticker, name, pf_function) %>%
  arrange(pf_function, ticker)

kable(overview,
      col.names = c("Ticker", "Name", "Portfolio Function"),
      caption   = paste0("Equity ETFs included in this report (n = ", nrow(overview),
                         ")  |  Benchmark: ", params$bmk)) %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, monospace = TRUE, color = "#1B2A4A") %>%
  column_spec(3, color = "#555555")
```

---

```{r batch-factsheets, results='asis', fig.width=13}
# ── Direct loop — no knit_child nesting ───────────────────────────────────────
# Per ticker: markdown header + stats table + triple overlay + regime stats.
# 8-panel factsheet only when show_full_factsheet = TRUE.
# All plots printed directly; knitr handles figure output automatically.

for (tk in eq_tickers) {

  m        <- etf_metadata[etf_metadata$ticker == tk, ]
  tk_name  <- if (nrow(m) > 0) m$name[1]        else tk
  tk_class <- if (nrow(m) > 0) m$asset_class[1] else "Equity"
  tk_pfn   <- if (nrow(m) > 0) m$pf_function[1] else "\u2014"
  col_data <- xts_ret[, tk]
  valid    <- !is.na(col_data)
  tk_start <- if (any(valid)) format(min(index(col_data[valid]))) else "\u2014"
  tk_end   <- if (any(valid)) format(max(index(col_data[valid]))) else "\u2014"

  # ── Section header ─────────────────────────────────────────────────────────
  cat(sprintf(
    "\n\n## %s \u2014 %s\n\n<div class='ticker-badge'><strong>%s</strong> &nbsp;|&nbsp; Class: %s &nbsp;|&nbsp; Function: %s &nbsp;|&nbsp; Period: %s \u2192 %s &nbsp;|&nbsp; Benchmark: %s</div>\n\n",
    tk, tk_name, tk, tk_class, tk_pfn, tk_start, tk_end, params$bmk
  ))

  # ── Stats table ────────────────────────────────────────────────────────────
  tryCatch({
    r1  <- as.numeric(coredata(xts_ret[, tk]))
    r2  <- as.numeric(coredata(xts_ret[, params$bmk]))
    idx <- which(!is.na(r1) & !is.na(r2))
    r1  <- r1[idx]; r2 <- r2[idx]; rs <- r1 - r2
    ann <- function(r) (prod(1+r))^(252/length(r)) - 1
    vol <- function(r) sd(r, na.rm=TRUE) * sqrt(252)
    mdd <- function(r) { w <- cumprod(1+r); min((w - cummax(w))/cummax(w)) }
    stats_tbl <- tibble::tibble(
      Series       = c(tk, params$bmk, paste0(tk, " \u2212 ", params$bmk)),
      `Ann Return` = scales::percent(c(ann(r1), ann(r2), ann(rs)), accuracy=0.1),
      `Ann Vol`    = scales::percent(c(vol(r1), vol(r2), vol(rs)), accuracy=0.1),
      `Sharpe`     = round(c(ann(r1)/vol(r1), ann(r2)/vol(r2), ann(rs)/vol(rs)), 2),
      `Max DD`     = scales::percent(c(mdd(r1), mdd(r2), mdd(rs)), accuracy=0.1)
    )
    print(kable(stats_tbl, align=c("l","r","r","r","r"),
                caption=paste0("Summary statistics \u2014 ", tk, " vs ", params$bmk)) %>%
      kable_styling(full_width=FALSE,
                    bootstrap_options=c("hover","condensed","bordered"),
                    position="center", font_size=12) %>%
      row_spec(0, bold=TRUE, background="#1B2A4A", color="white") %>%
      row_spec(1, background="#F0F4FA") %>%
      row_spec(3, bold=TRUE) %>%
      column_spec(1, bold=TRUE, color="#1B2A4A"))
  }, error = function(e) cat("> Stats error for", tk, ":", conditionMessage(e), "\n"))

  # ── Plot 1: triple overlay + rolling 12m relret ────────────────────────────
  knitr::opts_chunk$set(fig.height = 9)
  tryCatch(
    print(plot_regime_triple(tk, bmk = params$bmk, xts_ret = xts_ret)),
    error = function(e) {
      print(ggplot() + annotate("text", x=.5, y=.5, colour="grey50", size=3.5,
        label=paste0("Triple error (", tk, "): ", conditionMessage(e))) + theme_void())
    }
  )

  # ── Plot 2: regime overlay ext ─────────────────────────────────────────────
  knitr::opts_chunk$set(fig.height = 8)
  tryCatch(
    print(plot_regime_overlay_ext(tk, bmk = params$bmk, xts_ret = xts_ret)),
    error = function(e) {
      print(ggplot() + annotate("text", x=.5, y=.5, colour="grey50", size=3.5,
        label=paste0("Overlay error (", tk, "): ", conditionMessage(e))) + theme_void())
    }
  )

  # ── Plot 3: own regime stats ───────────────────────────────────────────────
  knitr::opts_chunk$set(fig.height = 5)
  tryCatch(
    print(plot_regime_stats(xts_ret[, tk], asset_name = tk)),
    error = function(e) {
      print(ggplot() + annotate("text", x=.5, y=.5, colour="grey50", size=3.5,
        label=paste0("Regime stats error (", tk, "): ", conditionMessage(e))) + theme_void())
    }
  )

  # ── Plot 4: 8-panel factsheet (optional — slow) ────────────────────────────
  if (isTRUE(params$show_full_factsheet)) {
    knitr::opts_chunk$set(fig.height = 22)
    tryCatch({
      fs_out <- suppressMessages(
        fs_render(tk, bmk = params$bmk, ret_xts = xts_ret, rt = rt)
      )
      print(fs_out$combined)
    }, error = function(e) {
      print(ggplot() + annotate("text", x=.5, y=.5, colour="grey50", size=3.5,
        label=paste0("Factsheet error (", tk, "): ", conditionMessage(e))) + theme_void())
    })
  }

  cat("\n\n<div class='section-div'></div>\n\n")
}
```

---

```{r footer}
cat(sprintf(
  "_Generated %s · %d equity ETFs · Benchmark: %s · T\\_fall = %.0f%%_",
  format(Sys.Date(), "%Y-%m-%d"),
  length(eq_tickers),
  params$bmk,
  params$t_fall * 100
))
```
