# ── Render from console ───────────────────────────────────────────────────────
# Full equity universe (all g_eq_tickers in xts_ret):
#   rmarkdown::render(
#     here::here("Rmd/eq_factsheets_batch.Rmd"),
#     output_file = here::here("03_reports/eq_factsheets_batch.html")
#   )
#
# Custom benchmark or subset via params:
#   rmarkdown::render(..., params = list(bmk = "URTH", t_fall = 0.10))
# ─────────────────────────────────────────────────────────────────────────────

---
title: "Equity ETF Factsheets — Full Universe"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 2
    theme: flatly
    highlight: tango
    fig_caption: true
params:
  bmk:    "SPY"
  t_fall: 0.10
---

```{css, echo=FALSE}
.caption      { color: #1B3A6B; font-size: 0.82em; font-style: italic;
                border-left: 3px solid #1B3A6B; padding-left: 6px; margin-top: 4px; }
.section-div  { border-top: 2px solid #dee2e6; margin: 48px 0 24px 0; }
.ticker-badge { background: #F0F4FA; border-left: 4px solid #1B2A4A;
                padding: 7px 14px; margin-bottom: 14px; font-size: 0.88em; color: #333; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

select <- dplyr::select

# ── params fallback (interactive / render-without-params) ─────────────────────
if (!exists("params")) params <- list(bmk = "SPY", t_fall = 0.10)

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
if (!exists("etf_metadata"))
  source(here("scripts/00_init_universe.R"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

if (!exists("plot_regime_overlay_ext") || !exists("plot_regime_triple"))
  source(here("key_plots/chart_regime_rel_overlay.R"))

source(here("utility/factsheet_functions.R"))

# ── Regime table (benchmark) ──────────────────────────────────────────────────
rt <- build_regime_table(xts_ret[, params$bmk], params$t_fall)

# ── Equity tickers: in universe AND in xts_ret, excluding the benchmark ───────
eq_tickers <- intersect(g_eq_tickers, colnames(xts_ret))
eq_tickers <- eq_tickers[eq_tickers != params$bmk]
```

## Universe overview

```{r overview-table}
overview <- etf_metadata %>%
  filter(ticker %in% eq_tickers) %>%
  select(ticker, name, pf_function) %>%
  arrange(pf_function, ticker)

kable(overview,
      col.names = c("Ticker", "Name", "Portfolio Function"),
      caption   = paste0("Equity ETFs included in this report (n = ", nrow(overview),
                         ")  |  Benchmark: ", params$bmk)) %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, monospace = TRUE, color = "#1B2A4A") %>%
  column_spec(3, color = "#555555")
```

---

```{r batch-factsheets, results='asis'}
# ── Per-ticker section template ───────────────────────────────────────────────
# Each ticker gets:
#   1. Metadata badge (class / function / date range)
#   2. Summary stats table
#   3. Triple overlay  (plot_regime_triple) — cumret + rolling 12m relret
#   4. Regime overlay  (plot_regime_overlay_ext) — BMK bars + ticker line + rel area
#   5. 8-panel factsheet (fs_render combined)
#   6. Own regime stats (plot_regime_stats)
#
# knit_child runs each block in the parent environment so variables defined in
# earlier sub-chunks (fs_out) are available in later ones within the same ticker.

template <- paste0("

## {{tk}} \u2014 {{tk_name}}

<div class='ticker-badge'>
<strong>{{tk}}</strong> &nbsp;|&nbsp;
Class: {{tk_class}} &nbsp;|&nbsp;
Function: {{tk_pfn}} &nbsp;|&nbsp;
Period: {{tk_start}} \u2192 {{tk_end}} &nbsp;|&nbsp;
Benchmark: {{bmk}}
</div>

", "\x60\x60\x60", "{r s1-{{tk}}, fig.show='asis'}
.fs_out_{{tk}} <- tryCatch(
  suppressMessages(
    fs_render('{{tk}}', bmk = params$bmk, ret_xts = xts_ret, rt = rt)
  ),
  error = function(e) { message('fs_render failed for {{tk}}: ', conditionMessage(e)); NULL }
)
if (!is.null(.fs_out_{{tk}})) {
  kable(.fs_out_{{tk}}$stats,
        align   = c('l','r','r','r','r','r'),
        caption = paste0('Summary statistics \u2014 {{tk}} vs {{bmk}}')) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c('hover','condensed','bordered'),
                  position = 'center', font_size = 12) %>%
    row_spec(0, bold = TRUE, background = '#1B2A4A', color = 'white') %>%
    row_spec(1, background = '#F0F4FA') %>%
    row_spec(3, bold = TRUE) %>%
    column_spec(1, bold = TRUE, color = '#1B2A4A') %>%
    print()
} else { cat('> Stats table unavailable for {{tk}}.\n') }
", "\x60\x60\x60", "

", "\x60\x60\x60", "{r s1b-{{tk}}, fig.height=9, fig.cap='Triple overlay + rolling 12m relative return \u2014 {{tk}} vs {{bmk}}'}
tryCatch(
  print(plot_regime_triple('{{tk}}', bmk = '{{bmk}}', xts_ret = xts_ret)),
  error = function(e) ggplot() +
    annotate('text', x=.5, y=.5, colour='grey50', size=3.5,
             label = paste0('Triple error ({{tk}}): ', conditionMessage(e))) + theme_void()
)
", "\x60\x60\x60", "

", "\x60\x60\x60", "{r s2-{{tk}}, fig.height=8, fig.cap='Regime overlay \u2014 {{tk}} vs {{bmk}}'}
tryCatch(
  print(plot_regime_overlay_ext('{{tk}}', bmk = '{{bmk}}', xts_ret = xts_ret)),
  error = function(e) ggplot() +
    annotate('text', x=.5, y=.5, colour='grey50', size=3.5,
             label = paste0('Overlay error ({{tk}}): ', conditionMessage(e))) + theme_void()
)
", "\x60\x60\x60", "

", "\x60\x60\x60", "{r s3-{{tk}}, fig.height=22, fig.cap='8-panel factsheet \u2014 {{tk}} vs {{bmk}}'}
if (!is.null(.fs_out_{{tk}})) {
  print(.fs_out_{{tk}}$combined)
} else {
  tryCatch(
    print(suppressMessages(
      fs_render('{{tk}}', bmk = params$bmk, ret_xts = xts_ret, rt = rt)
    )$combined),
    error = function(e) ggplot() +
      annotate('text', x=.5, y=.5, colour='grey50', size=3.5,
               label = paste0('Factsheet error ({{tk}}): ', conditionMessage(e))) + theme_void()
  )
}
", "\x60\x60\x60", "

", "\x60\x60\x60", "{r s4-{{tk}}, fig.height=5, fig.cap='Own regime statistics \u2014 {{tk}}'}
tryCatch(
  print(plot_regime_stats(xts_ret[, '{{tk}}'], asset_name = '{{tk}}')),
  error = function(e) ggplot() +
    annotate('text', x=.5, y=.5, colour='grey50', size=3.5,
             label = paste0('Regime stats error ({{tk}}): ', conditionMessage(e))) + theme_void()
)
", "\x60\x60\x60", "

<div class='section-div'></div>

")

# ── Render ────────────────────────────────────────────────────────────────────
sections <- lapply(eq_tickers, function(tk) {

  m        <- etf_metadata[etf_metadata$ticker == tk, ]
  tk_name  <- if (nrow(m) > 0) m$name[1]        else tk
  tk_class <- if (nrow(m) > 0) m$asset_class[1] else "Equity"
  tk_pfn   <- if (nrow(m) > 0) m$pf_function[1] else "—"

  col_data <- xts_ret[, tk]
  valid    <- !is.na(col_data)
  tk_start <- if (any(valid)) format(min(index(col_data[valid]))) else "—"
  tk_end   <- if (any(valid)) format(max(index(col_data[valid]))) else "—"

  expanded <- knitr::knit_expand(
    text     = template,
    tk       = tk,
    tk_name  = tk_name,
    tk_class = tk_class,
    tk_pfn   = tk_pfn,
    tk_start = tk_start,
    tk_end   = tk_end,
    bmk      = params$bmk
  )

  knitr::knit_child(text = expanded, envir = environment(), quiet = TRUE)
})

cat(paste(unlist(sections), collapse = "\n"))
```

---

```{r footer}
cat(sprintf(
  "_Generated %s · %d equity ETFs · Benchmark: %s · T\\_fall = %.0f%%_",
  format(Sys.Date(), "%Y-%m-%d"),
  length(eq_tickers),
  params$bmk,
  params$t_fall * 100
))
```
