---
title: "`r paste('Exante Hypothesis —', params$bmk, 'Benchmark')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
date: "`r Sys.Date()`"
params:
  bmk:      "SPY"   # Benchmark ticker — any column present in xts_ret
  w_blend:  0.5     # Benchmark weight in the blend (1 - w_blend goes to ticker)
  ret_eps:  0.005   # Return improvement threshold for role boundary (annualised)
  risk_eps: 0.005   # DD improvement threshold for role boundary
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 8
)

library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(scales)
library(ggrepel)
library(here)

# PerformanceAnalytics masks dplyr::select — restore dplyr version
select <- dplyr::select

# ── Data ──────────────────────────────────────────────────────────────────────
xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
source(here("scripts/00_init_universe.R"))
source(here("scripts_exante_hypothesis/exante_engine.R"))
source(here("scripts_exante_hypothesis/exante_visuals.R"))

# ── Validate benchmark ────────────────────────────────────────────────────────
if (!(params$bmk %in% colnames(xts_ret)))
  stop(sprintf("Benchmark '%s' not found in xts_ret. Available: %s",
               params$bmk,
               paste(head(colnames(xts_ret), 10), collapse = ", ")))

# ── Build metrics ─────────────────────────────────────────────────────────────
exante_tbl <- build_exante_metrics(
  xts_ret  = xts_ret,
  bmk      = params$bmk,
  horizons = c("full", "1Q", "1Y"),
  w_blend  = params$w_blend,
  ret_eps  = params$ret_eps,
  risk_eps = params$risk_eps
)

exante_tbl <- enrich_with_metadata(exante_tbl, etf_metadata)

# ── Convenience subsets ───────────────────────────────────────────────────────
tbl_full <- exante_tbl %>% filter(horizon == "full")
tbl_1Q   <- exante_tbl %>% filter(horizon == "1Q")
tbl_1Y   <- exante_tbl %>% filter(horizon == "1Y")

bmk_label   <- params$bmk
blend_pct   <- round(params$w_blend * 100)
ticker_pct  <- 100 - blend_pct
```

---

## 1. Hypothesis & Framework

The **Exante Hypothesis** formalises a practical question every portfolio allocator faces:

> *Given a benchmark (`r bmk_label`), does adding a ticker to a blended portfolio
> make things better or worse — and on which dimension?*

We analyse every ticker in the universe as a candidate to blend
**`r blend_pct`% `r bmk_label` / `r ticker_pct`% Ticker**.
For each ticker we compute two improvement metrics relative to holding `r bmk_label` alone:

| Metric | Formula | Sign convention |
|---|---|---|
| **delta_ret** | Ann. Return(blend) − Ann. Return(`r bmk_label`) | + = adds return |
| **delta_dd** | MaxDD(blend) − MaxDD(`r bmk_label`) | + = *less* severe drawdown |
| **delta_vol** | Ann. Vol(blend) − Ann. Vol(`r bmk_label`) | − = less volatile |

These two primary metrics place every ticker in one of four quadrants:

| Role | delta_ret | delta_dd | Interpretation |
|---|---|---|---|
| **Dominant** | > 0 | > 0 | Improves return AND reduces max drawdown |
| **Enhancer** | > 0 | ≤ 0 | Lifts return but does not protect on the downside |
| **Stabilizer** | ≤ 0 | > 0 | Pays return cost in exchange for lower drawdown |
| **Detractor** | ≤ 0 | ≤ 0 | Worsens both dimensions — avoid |

Analysis is conducted across three investment horizons:

- **Full** — entire available history (~`r nrow(xts_ret)` trading days)
- **1Y** — non-overlapping 252-day windows (`r nrow(tbl_1Y %>% distinct(n_windows))` windows)
- **1Q** — non-overlapping 63-day windows (`r tbl_1Q$n_windows[1]` windows)

```{r data-summary, echo=TRUE}
cat(sprintf("Universe  : %d tickers\n",    n_distinct(exante_tbl$ticker)))
cat(sprintf("Benchmark : %s\n",            params$bmk))
cat(sprintf("Blend     : %d%% %s + %d%% Ticker\n",
            blend_pct, params$bmk, ticker_pct))
cat(sprintf("Date range: %s → %s  (%d trading days)\n",
            as.character(start(xts_ret)),
            as.character(end(xts_ret)),
            nrow(xts_ret)))
cat(sprintf("Thresholds: ret_eps = %.3f | risk_eps = %.3f\n",
            params$ret_eps, params$risk_eps))
```

---

## 2. The Anti-Diversification Insight

Before the results, a key hypothesis inherited from the Long/Short analysis:

> **Conventional diversification dogma says:** lower correlation to `r bmk_label`
> → better drawdown protection.
>
> **Our finding:** *this does not hold for MaxDD in a long-only blend.*
> Correlation ≈ 0 can *worsen* portfolio drawdown relative to high-correlation tickers.

**The mechanism:**

- **Corr ≈ 1** — the blended portfolio moves nearly like `r bmk_label`;
  DD ≈ `r bmk_label`'s DD (no benefit, but no extra harm).
- **Corr ≈ 0** — two *independent* drawdown clocks run simultaneously.
  If the ticker also carries a large standalone DD, the portfolio suffers
  `r bmk_label`'s DD *and* the ticker's DD at different times —
  a blended MaxDD that is *worse* than the benchmark alone.
- **Corr ≈ −1** — genuine hedge; MaxDD shrinks. Rare for most ETFs vs `r bmk_label`.

The key driver of `delta_dd` is therefore **the ticker's own standalone MaxDD**,
not just its correlation. Plot P2 tests this directly.

---

## 3. Full-Sample Classification

### P1 — 2×2 Quadrant: Every Ticker Positioned

```{r p1-quadrant, fig.height=9}
print(plot_exante_quadrant(exante_tbl,
                            horizon_sel = "full",
                            label_n     = 8,
                            ret_eps     = params$ret_eps,
                            risk_eps    = params$risk_eps))
```

### P4 — Role Distribution Across Horizons

```{r p4-counts, fig.height=5}
print(plot_role_counts(exante_tbl))
```

```{r role-summary-table}
role_tbl <- exante_tbl %>%
  count(horizon, role) %>%
  group_by(horizon) %>%
  mutate(pct = scales::percent(n / sum(n), accuracy = 1)) %>%
  ungroup() %>%
  mutate(label = paste0(n, " (", pct, ")")) %>%
  select(horizon, role, label) %>%
  pivot_wider(names_from = role, values_from = label, values_fill = "0 (0%)") %>%
  mutate(horizon = factor(horizon, levels = c("1Q", "1Y", "full"))) %>%
  arrange(horizon)

knitr::kable(role_tbl,
             caption = sprintf("Role count (%% of universe) by horizon — %s benchmark",
                               bmk_label),
             align = "lcccc")
```

**Key observations from the full-sample classification:**

```{r obs-text, results='asis'}
n_dom  <- sum(tbl_full$role == "Dominant")
n_enh  <- sum(tbl_full$role == "Enhancer")
n_stab <- sum(tbl_full$role == "Stabilizer")
n_det  <- sum(tbl_full$role == "Detractor")
n_tot  <- nrow(tbl_full)

cat(sprintf(
  "- **%d tickers (%.0f%%)** are Dominant — improve both return and drawdown simultaneously.\n",
  n_dom, n_dom / n_tot * 100))
cat(sprintf(
  "- **%d tickers (%.0f%%)** are Stabilizers — accept a return cost to reduce drawdown.\n",
  n_stab, n_stab / n_tot * 100))
cat(sprintf(
  "- **%d tickers (%.0f%%)** are Enhancers — add return but worsen drawdown exposure.\n",
  n_enh, n_enh / n_tot * 100))
cat(sprintf(
  "- **%d tickers (%.0f%%)** are Detractors — worse on both dimensions.\n",
  n_det, n_det / n_tot * 100))
```

---

## 4. Anti-Diversification Test

### P2 — Correlation vs Drawdown Cost

```{r p2-corr-dd, fig.height=8}
print(plot_corr_dd(exante_tbl, horizon_sel = "full", show_loess = TRUE))
```

**Reading this plot:**

- The **dotted arrow** shows what diversification dogma predicts: as correlation
  to `r bmk_label` falls, blended drawdown should improve (positive delta_dd).
- The **loess curve** shows the empirical relationship.
- **Point size** = ticker's standalone MaxDD — larger = more painful own history.

If the loess curve is flat or U-shaped for correlation ∈ [0, 1], that confirms
*near-zero correlation does not guarantee DD protection in practice.*
The mechanism is visible in the colour coding — Detractors clustered at low
correlation tend to be large points (high standalone DD).

---

## 5. Horizon Sensitivity

### P3 — Role Stability Across 1Q / 1Y / Full

```{r p3-heatmap, fig.height=11}
print(plot_exante_heatmap(exante_tbl, sort_by = "role_full"))
```

Tickers whose role **does not change** across all three horizons are the most
reliable candidates for strategic allocation decisions against `r bmk_label`.
Tickers that flip between Enhancer and Stabilizer are horizon-sensitive and
require tactical judgment on the investment period.

### P6 — Consistency Within the 1Q Horizon

```{r p6-stability, fig.height=8}
print(plot_horizon_stability(exante_tbl, horizon_sel = "1Q", top_n = 32))
```

A bar reaching 100% means the ticker played its full-sample role in *every*
single quarter. A bar near 50% signals a coin-flip — the role alternates with
market regimes.

---

## 6. The Vol-DD Dissociation

### P5 — Does Vol Reduction Translate to DD Reduction?

```{r p5-vol-dd, fig.height=8}
print(plot_vol_dd_tradeoff(exante_tbl, horizon_sel = "full"))
```

The dashed diagonal represents the theoretical case where MaxDD shrinks in
exact proportion to volatility reduction.
Points **below** the diagonal reveal the dissociation: the ticker reduces
annualised volatility (a conventional "diversification benefit") but *fails* to
reduce — or actively worsens — the portfolio's maximum drawdown vs `r bmk_label`.

These are the most dangerous tickers from a client-experience perspective:
they appear safe on vol-based metrics (Sharpe, tracking error) but deliver the
full force of drawdown episodes.

---

## 7. Asset Class Breakdown

```{r asset-class-table}
ac_tbl <- tbl_full %>%
  filter(!is.na(asset_class)) %>%
  count(asset_class, role) %>%
  group_by(asset_class) %>%
  mutate(pct = scales::percent(n / sum(n), accuracy = 1)) %>%
  ungroup() %>%
  mutate(label = paste0(n, " (", pct, ")")) %>%
  select(asset_class, role, label) %>%
  pivot_wider(names_from = role, values_from = label, values_fill = "—") %>%
  arrange(asset_class)

knitr::kable(ac_tbl,
             caption = sprintf("Role distribution by asset class — %s benchmark (full sample)",
                               bmk_label),
             align = "lcccc")
```

```{r pf-function-table}
pf_tbl <- tbl_full %>%
  filter(!is.na(pf_function)) %>%
  count(pf_function, role) %>%
  group_by(pf_function) %>%
  mutate(pct = scales::percent(n / sum(n), accuracy = 1)) %>%
  ungroup() %>%
  mutate(label = paste0(n, " (", pct, ")")) %>%
  select(pf_function, role, label) %>%
  pivot_wider(names_from = role, values_from = label, values_fill = "—") %>%
  arrange(pf_function)

knitr::kable(pf_tbl,
             caption = sprintf("Role distribution by portfolio function — %s benchmark (full sample)",
                               bmk_label),
             align = "lcccc")
```

---

## 8. Full Results Table

```{r full-table}
tbl_full %>%
  arrange(role, desc(delta_dd)) %>%
  mutate(
    corr_full   = round(corr_full, 2),
    delta_ret   = scales::percent(delta_ret,   accuracy = 0.1),
    delta_dd    = scales::percent(delta_dd,    accuracy = 0.1),
    delta_vol   = scales::percent(delta_vol,   accuracy = 0.1),
    dd_tick     = scales::percent(dd_tick,     accuracy = 0.1),
    ret_tick    = scales::percent(ret_tick,    accuracy = 0.1),
    consistency = scales::percent(consistency, accuracy = 1)
  ) %>%
  select(ticker, role, corr_full, delta_ret, delta_dd, delta_vol,
         ret_tick, dd_tick, consistency, asset_class, pf_function) %>%
  knitr::kable(
    caption = sprintf("Full-sample Exante classification — %s benchmark, all tickers",
                      bmk_label),
    align = "llrrrrrrrlll"
  )
```

---

## 9. Methodology Notes

**Blend construction:**
Daily returns: `R_blend = {r blend_pct}% × R_{r bmk_label} + {r ticker_pct}% × R_ticker`.
The blend weight `w_blend` is a parameter; setting it to 0.3 / 0.7 / etc. traces
an efficient frontier between pure `r bmk_label` and pure ticker.

**Annualised return (sub-period horizons):**
For 1Q windows (63 days), the period return is scaled by `252 / 63 = 4` to annualise.
This extrapolation is aggressive for short windows; the *median across all windows*
is used to smooth noise.

**MaxDD definition:**
`MaxDD = min(Wealth_t / cummax(Wealth) − 1)` over the window.
Values ≤ 0; `delta_dd > 0` means the blend had a *less negative* (shallower) MaxDD.

**Role classification thresholds:**
Neutral zone = ±`r params$ret_eps * 100` pp/year for `delta_ret`,
±`r params$risk_eps * 100` pp for `delta_dd`.

**Consistency score:**
Fraction of sub-period windows where the ticker's role matched its full-sample role.
1.0 = perfectly stable; 0.5 = regime-alternating.

**To reproduce or change benchmark:**
```r
render_exante_report(bmk = "AGG")          # Fixed income benchmark
render_exante_report(bmk = "QQQ")          # Nasdaq benchmark
render_exante_report(bmk = "GLD", w_blend = 0.3)  # Gold at 30% weight
```
