---
title: "ETF Intelligence — Executive Summary"
subtitle: 'A Connected Journey: SPY Regime → Alpha Extraction → Exante Hypothesis'
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 2
    latex_engine: xelatex
---

<style>
.insight-box {
  background-color: #f0f4f8;
  border-left: 4px solid #2c3e50;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
.finding-box {
  background-color: #eafaf1;
  border-left: 4px solid #27ae60;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
.challenge-box {
  background-color: #fef9e7;
  border-left: 4px solid #f39c12;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
h2 { border-bottom: 2px solid #2c3e50; padding-bottom: 4px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(MASS)
library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(scales)
library(patchwork)
library(ggrepel)
library(here)
library(kableExtra)

select  <- dplyr::select
filter  <- dplyr::filter
arrange <- dplyr::arrange

# ── Core data ─────────────────────────────────────────────────────────────────
xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))
xts_rel <- readRDS(here("02_data_processed/xts_rel.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
source(here("scripts/00_init_universe.R"))
source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
source(here("scripts_spy_dd_regime/spy_dd_regime_rel.R"))
source(here("scripts_spy_dd_regime/regime_multi_ticker.R"))
source(here("scripts_spy_dd_regime/spy_statemachine_diagram_rel.R"))
source(here("scripts_state_machine/sm_engine.R"))
source(here("scripts_state_machine/sm_visuals.R"))
source(here("scripts_saa_structured/long_short_overlay.R"))
source(here("scripts_saa_taa/05_saa_portfolio.R"))
source(here("scripts_exante_hypothesis/exante_engine.R"))
source(here("scripts_exante_hypothesis/exante_visuals.R"))
source(here("scripts_saa_taa/saa_tree.R"))
source(here("utility/plot_xts_pair.R"))
source(here("utility/etf_ranges.R"))
source(here("utility/plot_calendar_perf.R"))
source(here("utility/plot_calendar_analog.R"))
source(here("key_plots/chart_regime_rel_overlay.R"))   # plot_regime_overlay_ext()

# ── Pair analysis helpers ──────────────────────────────────────────────────────
T_yrs <- as.numeric(diff(range(index(xts_ret)))) / 365

magdon_mdd <- function(sigma, mu, T) {
  if (mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  (sigma / S) * (log(S * sqrt(T)) + 0.35)
}

pair_stats <- function(out) {
  s  <- out$stats
  sp <- s %>% filter(str_detect(Series, "\u2212"))
  mu  <- parse_number(sp$`Ann Return`) / 100
  sig <- parse_number(sp$`Ann Vol`)    / 100
  mdd <- parse_number(sp$`Max DD`)     / 100
  list(
    mu         = mu,    sig      = sig,   mdd    = mdd,
    ir         = out$full_ir,             rho    = out$full_cor,
    te_10pct   = 0.10 * sig,
    mdd_10pct  = 0.10 * abs(mdd),
    ret_10pct  = 0.10 * mu,
    mdd_theory = magdon_mdd(sig, mu, T_yrs),
    stats      = s
  )
}

pairs_list <- list(
  # ── KF2 winners (ρ > 0.75, positive IR) ──────────────────────────────────
  list(t1="XLK", t2="SPY", label="XLK vs SPY", type="Equity swap",
       decision="+10% XLK / \u221210% SPY"),
  list(t1="SMH", t2="SPY", label="SMH vs SPY", type="Equity swap",
       decision="+10% SMH / \u221210% SPY"),
  list(t1="QQQ", t2="SPY", label="QQQ vs SPY", type="Equity swap",
       decision="+10% QQQ / \u221210% SPY"),
  # ── Additional equity swaps ───────────────────────────────────────────────
  list(t1="XLV", t2="SPY", label="XLV vs SPY", type="Equity swap",
       decision="+10% XLV / \u221210% SPY"),
  list(t1="XLF", t2="SPY", label="XLF vs SPY", type="Equity swap",
       decision="+10% XLF / \u221210% SPY"),
  # ── Cross-asset additions ─────────────────────────────────────────────────
  list(t1="XLP", t2="AGG", label="XLP vs AGG", type="Cross-asset",
       decision="+10% XLP / \u221210% AGG"),
  list(t1="XLV", t2="AGG", label="XLV vs AGG", type="Cross-asset",
       decision="+10% XLV / \u221210% AGG")
)

# ── Shared constants ──────────────────────────────────────────────────────────
MASTER   <- "SPY"
T_FALL   <- 0.10
spy_ret  <- xts_ret[, MASTER]
rt       <- build_regime_table(spy_ret, T_FALL)

REGIME_PAL <- c(Fall = "#D90429", Recovery = "#F77F00", Consolidation = "#2D6A4F")

core_universe <- c(
  "AGG", "IEF", "TLT", "HYG", "TIP",
  "IEFA", "FEZ", "VWO",
  "QQQ", "XLK", "XLF", "XLI", "XLE",
  "GLD", "VIXY"
)

safe_haven_tickers <- c("GLD", "IEF", "TLT", "HYG", "XLF", "XLK")

# ── Pre-compute exante (full sample only for speed) ───────────────────────────
exante_tbl <- build_exante_metrics(
  xts_ret  = xts_ret,
  bmk      = MASTER,
  horizons = "full",
  w_blend  = 0.5
)
exante_tbl <- enrich_with_metadata(exante_tbl, etf_metadata)

# ── L/S portfolio (core universe) ─────────────────────────────────────────────
ls_core <- build_ls_portfolio(
  xts_rel          = xts_rel,
  n_long           = 5L,
  n_short          = 5L,
  signal_lookback  = 63L,
  rebalance_freq   = "months",
  universe_tickers = core_universe
)

core_5050 <- calc_saa_portfolio(
  xts_returns    = xts_ret,
  weights_vector = c(SPY = 0.50, AGG = 0.50),
  rebalance_freq = "months"
)

common_idx   <- intersect(index(core_5050$returns), index(ls_core$returns))
core_ret     <- core_5050$returns[common_idx]
ls_ret       <- ls_core$returns[common_idx]
combined_ret <- xts(as.numeric(core_ret) + as.numeric(ls_ret), order.by = common_idx)
colnames(core_ret)     <- "Core_5050"
colnames(ls_ret)       <- "LS_Core_15"
colnames(combined_ret) <- "Core_plus_LS"
phase2_comp  <- merge(core_ret, ls_ret, combined_ret)
```

---

## The Intellectual Journey

All analyses in this project orbit a single practical question:

> **Given SPY as a starting point, how does any additional position change
> the investor's experience — measured in return gained and drawdown suffered?**

The journey moved through seven connected steps:

| # | Analysis | Core Question |
|---|---|---|
| 1 | **SPY Regime Engine** | What are the distinct phases of SPY itself? |
| 2 | **200DMA State Machine** | Can we signal tactical entry/exit on SPY? |
| 3 | **Alpha by Regime** | How do other ETFs behave relative to SPY in each phase? |
| 4 | **Safe Haven Signals** | Can we predict regime transitions before they happen? |
| 5 | **Pair Analysis** | What is the cost-benefit of adding any ticker to SPY? (XLK deep-dive included) |
| 6 | **L/S Overlay** | Can we extract pure alpha from the relative-return universe? |
| 7 | **Exante Hypothesis** | Can we systematically classify every ticker as Enhancer or Stabilizer? |
| 8 | **SAA Portfolio & Institutional Mapping** | How does a concrete SAA translate to ETFs and institutional risk profiles? |
| 9 | **Meta-Asset Engine** | Can two sector ETFs be combined into a regime-balanced building block? |
| 10 | **ETF Range Monitor** | Where does each ticker sit in its 52-week range right now? |

The **unifying thread**: a recurrent empirical finding — confirmed across multiple
analyses — that **correlation ≈ 1 beats correlation ≈ 0 on drawdown cost**.
The diversification dogma (lower correlation → better protection) does not hold
for MaxDD in a long-only blended portfolio. This discovery shapes everything from
the L/S construction to the Exante framework.

---

## 1 · SPY Regime Engine

<div class="insight-box">
**What:** A 3-state data-driven engine labels every trading day as Fall (≥10% DD),
Recovery (climb from trough), or Consolidation (expansion). No manual labels. Single
parameter: `t_fall = 10%`.
</div>

```{r p1-regime-overlay, fig.height=7}
print(plot_regime_overlay(spy_ret, asset_name = "SPY"))
```

<div style="page-break-inside: avoid">
<div class="insight-box">
**What:** The same 3-state engine run at a finer **≥5% DD** threshold (bottom bar) alongside
the macro ≥10% bar (top bar). The jitter bar reveals mini-cycles *within* the macro regimes —
corrections and recoveries that never reached 10% and were therefore invisible above.
Comparing the two bars shows how much internal volatility is hidden inside a single
Consolidation or Recovery label.
</div>

```{r p1-regime-overlay-jitter, fig.height=8}
print(plot_regime_overlay(spy_ret, t_fall = 0.10, t_fall_jitter = 0.05))
```
</div>

```{r p1-sm-diagram, fig.height=7}
print(plot_sm_diagram(xts_ret, master = "SPY", t_fall = T_FALL))
```

```{r p1-regime-stats, fig.height=5}
print(plot_regime_stats(spy_ret, asset_name = "SPY"))
```

<div class="finding-box">
**Key finding:** The three states have radically different annualised return and
duration profiles. Fall periods are short but violent; Recovery periods capture
most of the rebound but carry elevated volatility; Consolidation is where the
majority of calendar time — and compounded wealth — accumulates.
Knowing *which state you are in* changes every allocation decision.
</div>

<small>Full report: `Rmd/spy_regime_statemachine_report.Rmd`</small>

---

## 2 · 200DMA State Machine

<div class="insight-box">
**What:** A hysteresis-based state machine on the 200-day moving average.
Unlike a simple crossover, it requires price to clear the 200DMA by **+4%** to
enter Bull and drop **−2%** below to enter Bear — preventing whipsawing in
sideways markets.
</div>

```{r p2-200dma, fig.height=8}
# xts_wlth is the RELATIVE wealth index (SPY vs SPY = flat 1.0) — not useful here.
# Build SPY absolute wealth index from xts_ret for the 200DMA engine.
spy_wlth_abs <- cumprod(1 + xts_ret[, "SPY"])
colnames(spy_wlth_abs) <- "SPY"

plot_200DMA(
  ticker_symbol = "SPY",
  xts_wealth    = spy_wlth_abs,   # absolute price wealth, base 1.0
  xts_price     = xts_ret[, "SPY", drop = FALSE],  # daily returns for DD panel
  upper         = 0.04,
  lower         = -0.02
)
```

<div class="finding-box">
**Key finding:** The asymmetric thresholds (+4% / −2%) materially reduce false
signals vs a symmetric crossover. The Bull/Bear state machine aligns well with
the Fall/Consolidation classification from the regime engine — but reacts faster
on the upside entry and more cautiously on the downside exit.
</div>

<small>Full report: `Rmd/spy_regime_statemachine_report.Rmd` (§ State Machine)</small>

---

## 3 · Alpha by Regime

<div class="insight-box">
**What:** For a set of representative ETFs, compute the average relative return
(α vs SPY) within each regime state. This answers: *which tickers are regime-aware —
i.e., they protect in Fall and participate in Consolidation?*
</div>

```{r p3-regime-rel-summary, fig.height=7}
print(plot_regime_rel_summary(
  xts_ret = xts_ret,
  tickers = safe_haven_tickers,
  master  = "SPY"
))
```

```{r p3-regime-heatmap, fig.height=7}
print(plot_regime_rel_heatmap(
  xts_ret = xts_ret,
  tickers = safe_haven_tickers,
  master  = "SPY"
))
```

```{r p3-abs-vs-rel, fig.height=5.5}
# Compute avg abs return AND avg alpha per ticker per regime TYPE
# to show the abs vs relative distinction explicitly
sh_tickers_all <- c("SPY", safe_haven_tickers)

abs_rel_df <- map_dfr(sh_tickers_all, function(tk) {
  tk_r <- xts_ret[, tk]
  rt %>% rowwise() %>%
    mutate(
      ticker  = tk,
      abs_ret = {
        w <- tk_r[paste0(xmin, "/", xmax)]
        if (length(w) == 0) NA_real_
        else as.numeric(Return.cumulative(w))
      },
      spy_abs = period_return,
      alpha   = abs_ret - spy_abs
    ) %>%
    ungroup() %>%
    select(ticker, regime, abs_ret, alpha)
}) %>%
  mutate(regime = as.character(regime)) %>%
  group_by(ticker, regime) %>%
  summarise(
    avg_abs   = mean(abs_ret, na.rm = TRUE),
    avg_alpha = mean(alpha,   na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  pivot_longer(c(avg_abs, avg_alpha),
               names_to  = "metric",
               values_to = "value") %>%
  mutate(
    metric = if_else(metric == "avg_abs", "Abs Return", "α vs SPY"),
    regime = factor(regime, levels = c("Fall", "Recovery", "Consolidation")),
    ticker = factor(ticker, levels = sh_tickers_all),
    bar_col = case_when(
      metric == "α vs SPY" & value >= 0  ~ "#1B3A6B",
      metric == "α vs SPY" & value <  0  ~ "#7B1010",
      metric == "Abs Return" & value >= 0 ~ "#5B8DB8",
      TRUE                                ~ "#C97070"
    )
  )

ggplot(abs_rel_df, aes(x = ticker, y = value, fill = bar_col)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65,
           show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6, color = "grey40") +
  geom_text(aes(
    label = sprintf("%s%.1f%%", if_else(value >= 0, "+", ""), value * 100),
    vjust = if_else(value >= 0, -0.35, 1.25),
    color = bar_col),
    position = position_dodge(width = 0.7),
    size = 2.8, fontface = "bold", show.legend = FALSE) +
  scale_fill_identity() +
  scale_color_identity() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  facet_wrap(~regime, nrow = 1) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text  = element_text(face = "bold", size = 11),
    axis.title  = element_blank(),
    axis.text.x = element_text(face = "bold", size = 9, angle = 30, hjust = 1),
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey45", size = 9)
  ) +
  labs(
    title    = "Abs Return vs α vs SPY — by Regime Type",
    subtitle = "Dark = α vs SPY  |  Light = Abs return  |  Key: safe havens LAG in Recovery on alpha but abs return stays positive"
  )
```

<div class="finding-box">
**Key findings — Abs vs Relative performance distinction:**

**(1) Safe havens (GLD, IEF) do NOT give back absolute performance in Recovery —
they give back *relative* performance.**
In Fall, they record a STABLE outperformance: small absolute loss while SPY suffers a large fall
(e.g. GLD −4.6% abs / +24.4% α in Fall Sep 2018 while SPY fell ~−29%).
In Recovery, SPY rebounds violently. GLD and IEF often have *positive* absolute returns
but lag the SPY rebound — generating negative alpha (LAG type). This is an
opportunity cost, not an actual loss.

**(2) The alpha asymmetry favours holding safe havens across the cycle.**
Fall alpha protection (STABLE: large positive α) typically exceeds Recovery lag (LAG: negative α)
in magnitude because Fall episodes are short and violent while recoveries are gradual.
The net alpha across a complete Fall → Recovery cycle can remain positive for GLD and IEF.

**(3) Equity satellites (XLK, XLF) are the mirror image.**
They show ALPHA or MOMENTUM type in Recovery/Consolidation (positive α, both positive abs)
but LOSS or EXPOSED in Fall (negative α, both negative abs — deeper drawdown than SPY).

**(4) The correct framing for allocation:**
Holding GLD/IEF does not cost you absolute return in Recovery — it costs you
*relative* return vs SPY. For a risk-aware investor, this is the intended trade-off:
accept the LAG tax in good times to avoid the LOSS amplification in bad times.
</div>

```{r p3-abs-alpha-sectors, fig.height=5}
print(plot_abs_alpha_regime(
  xts_ret = xts_ret,
  tickers = intersect(g_sector_tickers, colnames(xts_ret)),
  rt      = rt,
  title   = "US Sectors \u2014 Avg Abs Return vs \u03b1 vs SPY by Regime"
))
```

```{r p3-fall-fingerprint, fig.height=7}
print(plot_fall_fingerprint(
  xts_ret = xts_ret,
  tickers = intersect(g_sector_tickers, colnames(xts_ret)),
  rt      = rt,
  title   = "US Sectors \u2014 Fall Fingerprint (\u03b1 vs SPY per Episode)"
))
```

```{r p3-sector-fingerprint, fig.height=7}
print(plot_regime_fingerprint(
  xts_ret = xts_ret,
  tickers = intersect(g_sector_tickers, colnames(xts_ret)),
  rt      = rt,
  title   = "US Sectors \u2014 Full Cycle Fingerprint (Strip)"
))
```

```{r p3-sector-fingerprint-box, fig.height=7}
print(plot_regime_fingerprint_box(
  xts_ret = xts_ret,
  tickers = intersect(g_sector_tickers, colnames(xts_ret)),
  rt      = rt,
  title   = "US Sectors \u2014 Full Cycle Fingerprint (Boxplot)"
))
```

```{r p3-regime-sync, fig.height=6}
print(plot_regime_sync(
  xts_ret = xts_ret,
  tickers = intersect(g_sector_tickers, colnames(xts_ret)),
  title   = "US Sectors \u2014 Regime Sync (each ticker\u2019s own cycle)"
))
```

<small>Full report: `Rmd/spy_regime_statemachine_report.Rmd` (§ Relative Overlay)</small>

---

## 4 · Safe Haven Regime Signal

<div class="insight-box">
**What:** Three methods for predicting SPY regime transitions using leading indicators
(Credit spread via HYG, rates via IEF/TLT, FX safe havens).<br>
**Method 1:** Rolling correlation breakdown between safe haven basket and SPY.<br>
**Method 2:** LASSO logistic regression on lagged features.<br>
**Method 3:** Regime-difference eigenanalysis (PCA on Fall-minus-Consolidation covariance).
</div>

```{r p4-rolling-corr, fig.height=7}
# Rolling correlation of HYG and IEF to SPY — the simplest leading signal
roll_win <- 60L
tickers_sh <- c("HYG", "IEF", "GLD")

corr_df <- map_dfr(tickers_sh, function(tk) {
  r <- na.omit(merge(xts_ret[, "SPY"], xts_ret[, tk]))
  rc <- rollapply(r, width = roll_win,
                  FUN = function(m) cor(m[,1], m[,2], use = "complete.obs"),
                  by.column = FALSE, fill = NA, align = "right")
  data.frame(date = index(rc), corr = as.numeric(rc), ticker = tk)
})

regime_shading <- rt %>%
  filter(regime != "Consolidation") %>%
  mutate(xmin = as.Date(xmin), xmax = as.Date(xmax))

ggplot(corr_df, aes(x = date, y = corr, colour = ticker)) +
  geom_rect(data = regime_shading,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = regime),
            inherit.aes = FALSE, alpha = 0.15) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 0.8) +
  scale_fill_manual(values = REGIME_PAL, name = "Regime") +
  scale_colour_manual(values = c(HYG = "#E63946", IEF = "#2980b9", GLD = "#f39c12"),
                      name = NULL) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  scale_y_continuous(limits = c(-1, 1)) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
  labs(
    title    = "Method 1 — Rolling 60-Day Correlation to SPY",
    subtitle = "HYG (Credit) · IEF (Rates) · GLD (Gold) | Shaded = Fall & Recovery periods",
    x = NULL, y = "Rolling Correlation to SPY"
  )
```

<div class="finding-box">
**Key finding:** HYG (high-yield credit) correlation to SPY *rises* ahead of Fall
periods — a counter-intuitive early warning: when credit is becoming *more* SPY-like,
stress is building. GLD correlation consistently decouples during Fall. These
dynamics motivated a predictive LASSO model with AUC ~0.72 at a 63-day horizon.
</div>

<small>Full report: `Rmd/safe_haven_regime_report.Rmd`</small>

---

```{r run-pairs, include=FALSE}
# Explicit capture of rt to avoid stats::rt() shadowing inside map() closure
.rt <- rt
pair_results <- map(pairs_list, function(p) {
  out <- plot_xts_pair(
    xts_ret[, p$t1], xts_ret[, p$t2],
    label1     = p$t1, label2 = p$t2,
    regime_tbl = .rt
  )
  c(p, list(out = out, ps = pair_stats(out)))
})
```

```{r build-kf-data, include=FALSE}
# ── KF shared parameters ───────────────────────────────────────────────────────
RHO_SCREEN <- 0.75
MDD_SCREEN <- -0.30
T_yrs_kf   <- as.numeric(diff(range(index(xts_ret)))) / 365
universe_kf <- setdiff(colnames(xts_ret), "SPY")

RHO_PAL <- c(
  "ρ < 0.2"  = "#d73027",
  "0.2–0.5"  = "#fc8d59",
  "0.5–0.75" = "#4575b4",
  "ρ > 0.75" = "#1a1a6e"
)

QUAD_PAL <- c(
  "Best: +IR, +ΔDD"    = "#1a1a6e",
  "Return edge only"   = "#2196F3",
  "DD protection only" = "#27ae60",
  "Neither"            = "#bdbdbd"
)

# Magdon-Ismail with full edge-case guards
magdon_mdd_kf <- function(sigma, mu, T) {
  if (is.na(mu) || is.na(sigma) || sigma <= 0) return(NA_real_)
  if (mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  log_term <- log(S * sqrt(T)) + 0.35
  if (log_term <= 0) return(sigma * sqrt(T))
  max((sigma / S) * log_term, 0)
}

# ── KF1: cb_tbl_kf — all tickers vs SPY, theory vs empirical MaxDD ────────────
cb_tbl_kf <- map_dfr(universe_kf, function(tk) {
  r_tk  <- as.numeric(na.omit(xts_ret[, tk]))
  r_spy <- as.numeric(na.omit(xts_ret[, "SPY"]))
  n     <- min(length(r_tk), length(r_spy))
  if (n < 252) return(NULL)
  r_tk  <- tail(r_tk,  n)
  r_spy <- tail(r_spy, n)

  spread  <- r_tk - r_spy
  mu_spr  <- mean(spread, na.rm = TRUE) * 252
  sig_spr <- sd(spread,   na.rm = TRUE) * sqrt(252)
  mdd_emp <- { w <- cumprod(1 + spread); min((w - cummax(w)) / cummax(w), na.rm = TRUE) }
  mdd_th  <- magdon_mdd_kf(sig_spr, mu_spr, T_yrs_kf)
  rho     <- cor(r_tk, r_spy, use = "complete.obs")

  tibble(
    ticker  = tk,
    mu_spr  = mu_spr,
    sig_spr = sig_spr,
    mdd_emp = mdd_emp,
    mdd_th  = -mdd_th,
    rho     = rho,
    ir      = mu_spr / sig_spr
  )
}) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class, sub_block),
            by = "ticker") %>%
  filter(!is.na(mdd_th)) %>%
  mutate(
    residual = mdd_emp - mdd_th,
    rho_band = cut(rho,
      breaks = c(-Inf, 0.2, 0.5, 0.75, Inf),
      labels = c("ρ < 0.2", "0.2–0.5", "0.5–0.75", "ρ > 0.75")
    )
  )

# ── KF2: spread_tbl_kf — ρ, IR, spread-MaxDD, DD-delta, prob_1q ───────────────
spread_tbl_kf <- map_dfr(universe_kf, function(tk) {
  r_tk  <- as.numeric(na.omit(xts_ret[, tk]))
  r_spy <- as.numeric(na.omit(xts_ret[, "SPY"]))
  n     <- min(length(r_tk), length(r_spy))
  if (n < 252) return(NULL)
  r_tk  <- tail(r_tk,  n)
  r_spy <- tail(r_spy, n)

  spread    <- r_tk - r_spy
  mu_spr    <- mean(spread, na.rm = TRUE) * 252
  sig_spr   <- sd(spread,   na.rm = TRUE) * sqrt(252)
  mdd_spr   <- { w <- cumprod(1 + spread); min((w - cummax(w)) / cummax(w), na.rm = TRUE) }
  rho       <- cor(r_tk, r_spy, use = "complete.obs")
  blend     <- 0.90 * r_spy + 0.10 * r_tk
  mdd_blend <- { w <- cumprod(1 + blend); min((w - cummax(w)) / cummax(w), na.rm = TRUE) }
  mdd_spy   <- { w <- cumprod(1 + r_spy); min((w - cummax(w)) / cummax(w), na.rm = TRUE) }
  roll_ret  <- zoo::rollapply(spread, 63L, function(x) prod(1 + x) - 1,
                              fill = NA, align = "right")
  prob_1q   <- mean(roll_ret > 0, na.rm = TRUE)

  tibble(ticker = tk, rho = rho, mu_spr = mu_spr, sig_spr = sig_spr,
         ir = mu_spr / sig_spr, mdd_spr = mdd_spr, mdd_blend = mdd_blend,
         mdd_spy = mdd_spy, dd_delta = mdd_blend - mdd_spy, prob_1q = prob_1q)
}) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class, sub_block),
            by = "ticker") %>%
  mutate(
    passes_screen = rho >= RHO_SCREEN & mdd_spr >= MDD_SCREEN,
    rho_band = cut(rho,
      breaks = c(-Inf, 0.2, 0.5, 0.75, Inf),
      labels = c("ρ < 0.2", "0.2–0.5", "0.5–0.75", "ρ > 0.75")
    )
  )
```

```{r build-frontier-data, include=FALSE}
library(quadprog)

RF_EF      <- 0.045
N_POINTS   <- 80

# Filter to tickers with >= 252 obs; align to common date range
ef_tickers <- intersect(
  names(which(colSums(!is.na(xts_ret)) >= 252)),
  etf_metadata$ticker
)
ret_clean_ef <- na.omit(xts_ret[, ef_tickers])
mu_ef    <- colMeans(ret_clean_ef) * 252
Sigma_ef <- cov(ret_clean_ef) * 252
n_ef     <- ncol(ret_clean_ef)

# Long-only, fully-invested QP solver
min_var_pf <- function(target, mu, Sigma) {
  n    <- length(mu)
  Amat <- cbind(rep(1, n), mu, diag(n))
  bvec <- c(1, target, rep(0, n))
  tryCatch({
    sol <- solve.QP(2 * Sigma, rep(0, n), Amat, bvec, meq = 2)
    list(weights = sol$solution, var = sol$value)
  }, error = function(e) NULL)
}

gmv_ret_ef <- optimize(
  function(r) { res <- min_var_pf(r, mu_ef, Sigma_ef); if (is.null(res)) Inf else res$var },
  interval = c(min(mu_ef), max(mu_ef)), tol = 1e-6
)$minimum

ret_targets_ef <- seq(gmv_ret_ef, quantile(mu_ef, 0.95), length.out = N_POINTS)

frontier_pts_ef <- map_dfr(ret_targets_ef, function(tgt) {
  res <- min_var_pf(tgt, mu_ef, Sigma_ef)
  if (is.null(res)) return(NULL)
  tibble(ret = tgt, vol = sqrt(res$var), sharpe = (tgt - RF_EF) / sqrt(res$var))
})

gmv_res_ef  <- min_var_pf(gmv_ret_ef, mu_ef, Sigma_ef)
gmv_pt_ef   <- tibble(ret = gmv_ret_ef, vol = sqrt(gmv_res_ef$var),
                      label = "GMV", sharpe = (gmv_ret_ef - RF_EF) / sqrt(gmv_res_ef$var))

tan_idx_ef  <- which.max(frontier_pts_ef$sharpe)
tan_pt_ef   <- frontier_pts_ef[tan_idx_ef, ] %>% mutate(label = "Max Sharpe")

saa_tk_ef   <- SAA_PORTFOLIO$ticker
saa_w_ef    <- SAA_PORTFOLIO$weight
saa_mu_ef   <- sum(saa_w_ef * mu_ef[saa_tk_ef])
saa_var_ef  <- as.numeric(t(saa_w_ef) %*% Sigma_ef[saa_tk_ef, saa_tk_ef] %*% saa_w_ef)
saa_pt_ef   <- tibble(ret = saa_mu_ef, vol = sqrt(saa_var_ef), label = "SAA",
                      sharpe = (saa_mu_ef - RF_EF) / sqrt(saa_var_ef))

# ticker scatter with plot_class grouping
us_broad_sb_ef  <- c("US_Broad","US_LargeCap","US_SmallCap","US_MidCap",
                      "Momentum","MinVol","Quality","DivGrowth","CashFlow","Eq_Income","Value")
us_sector_sb_ef <- c("Energy","Financials","HealthCare","HC_Broad","MedDevices","Biotech",
                      "Industrials","Materials","Discretionary","Comms","Housing",
                      "Reg_Banks","REIT_Sector","Defense")
intl_dm_sb_ef   <- c("DM_Broad","Europe","Germany","France","Italy","Australia",
                      "Canada","Japan","Japan_Hdg","Korea","Switzerland")
em_sb_ef        <- c("EM_Broad","China","India","Brazil")
thematic_sb_ef  <- c("Semis","AI","Cloud","Cyber","IPO","Private_Eq",
                      "MinVol_Glbl","Agri_Equity","Lithium","Uranium")

ticker_pts_ef <- tibble(
  ticker = ef_tickers,
  ret    = mu_ef,
  vol    = sqrt(diag(Sigma_ef)),
  sharpe = (mu_ef - RF_EF) / sqrt(diag(Sigma_ef))
) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class, sub_block), by = "ticker") %>%
  mutate(
    in_saa     = ticker %in% SAA_PORTFOLIO$ticker,
    plot_class = case_when(
      asset_class == "Equity" & sub_block %in% us_broad_sb_ef  ~ "Eq · US Broad/Factor",
      asset_class == "Equity" & sub_block %in% us_sector_sb_ef ~ "Eq · US Sector",
      asset_class == "Equity" & sub_block %in% intl_dm_sb_ef   ~ "Eq · Intl DM",
      asset_class == "Equity" & sub_block %in% em_sb_ef        ~ "Eq · EM",
      asset_class == "Equity" & sub_block %in% thematic_sb_ef  ~ "Eq · Thematic",
      asset_class == "Equity"                                   ~ "Eq · Other",
      asset_class == "FixedIncome"  ~ "Fixed Income",
      asset_class == "Commodity"    ~ "Commodity",
      asset_class == "FX"           ~ "FX",
      asset_class == "RealAsset"    ~ "Real Asset",
      asset_class == "MultiAsset"   ~ "Multi-Asset",
      TRUE                          ~ asset_class
    )
  )

AC_PAL_EF <- c(
  "Eq · US Broad/Factor" = "#1B3A6B",
  "Eq · US Sector"       = "#4A90D9",
  "Eq · Intl DM"         = "#7EB8E8",
  "Eq · EM"              = "#00897B",
  "Eq · Thematic"        = "#AB47BC",
  "Eq · Other"           = "#90A4AE",
  "Fixed Income"         = "#27AE60",
  "Commodity"            = "#E67E22",
  "FX"                   = "#C9A800",
  "Real Asset"           = "#C0392B",
  "Multi-Asset"          = "#7F8C8D",
  "Alternative"          = "#BDC3C7",
  "Signal"               = "#BDC3C7"
)

p_ef <- ggplot() +
  geom_point(
    data = ticker_pts_ef,
    aes(x = vol, y = ret, colour = plot_class, shape = in_saa, size = in_saa),
    alpha = 0.70
  ) +
  geom_text_repel(
    data = ticker_pts_ef %>% filter(in_saa | sharpe > 0.8 | ret > 0.20 | vol < 0.08),
    aes(x = vol, y = ret, label = ticker, colour = plot_class),
    size = 3, fontface = "bold", max.overlaps = 25,
    segment.color = "grey70", show.legend = FALSE
  ) +
  geom_path(data = frontier_pts_ef, aes(x = vol, y = ret),
            colour = "#2c3e50", linewidth = 1.2, alpha = 0.9) +
  geom_point(data = gmv_pt_ef, aes(x = vol, y = ret),
             shape = 23, fill = "#27ae60", colour = "white", size = 5, stroke = 1.5) +
  geom_label(data = gmv_pt_ef, aes(x = vol, y = ret, label = "GMV"),
             nudge_y = 0.008, size = 3.5, fontface = "bold",
             fill = "#27ae60", colour = "white", label.size = 0) +
  geom_point(data = tan_pt_ef, aes(x = vol, y = ret),
             shape = 24, fill = "#e67e22", colour = "white", size = 5, stroke = 1.5) +
  geom_label(data = tan_pt_ef, aes(x = vol, y = ret, label = "Max Sharpe"),
             nudge_y = 0.008, size = 3.5, fontface = "bold",
             fill = "#e67e22", colour = "white", label.size = 0) +
  geom_point(data = saa_pt_ef, aes(x = vol, y = ret),
             shape = 21, fill = "#D90429", colour = "white", size = 6, stroke = 1.5) +
  geom_label(data = saa_pt_ef, aes(x = vol, y = ret, label = "SAA"),
             nudge_x = 0.018, nudge_y = 0.012, size = 3.5, fontface = "bold",
             fill = "#D90429", colour = "white", label.size = 0) +
  scale_colour_manual(values = AC_PAL_EF, name = "Asset Class / Group", na.value = "grey70") +
  scale_shape_manual(values = c("TRUE" = 18, "FALSE" = 16), guide = "none") +
  scale_size_manual(values  = c("TRUE" = 3.5, "FALSE" = 1.8), guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1), name = "Annualised Volatility") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(-0.05, 0.30),
                     name = "Annualised Return") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor  = element_blank(),
    legend.position   = "right",
    legend.text       = element_text(size = 10),
    legend.title      = element_text(size = 10, face = "bold"),
    axis.title        = element_text(size = 12),
    axis.text         = element_text(size = 11),
    plot.title        = element_text(face = "bold", size = 15),
    plot.subtitle     = element_text(size = 11, colour = "grey40"),
    plot.caption      = element_text(size = 9, colour = "grey55")
  ) +
  labs(
    title    = "Efficient Frontier — Sovereign Universe",
    subtitle = sprintf(
      "Long-only, fully-invested | %d tickers | RF = %.1f%% | SAA = current strategic allocation",
      n_ef, RF_EF * 100
    ),
    caption  = sprintf(
      "GMV: vol=%.1f%% ret=%.1f%% Sharpe=%.2f  |  Max Sharpe: vol=%.1f%% ret=%.1f%% Sharpe=%.2f  |  SAA: vol=%.1f%% ret=%.1f%% Sharpe=%.2f",
      gmv_pt_ef$vol*100, gmv_pt_ef$ret*100, gmv_pt_ef$sharpe,
      tan_pt_ef$vol*100, tan_pt_ef$ret*100, tan_pt_ef$sharpe,
      saa_pt_ef$vol*100, saa_pt_ef$ret*100, saa_pt_ef$sharpe
    )
  )

# Weights tables
ef_gmv_tbl <- tibble(ticker = ef_tickers, weight = gmv_res_ef$weights) %>%
  filter(weight > 0.005) %>% arrange(desc(weight)) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class), by = "ticker") %>%
  mutate(Weight = percent(weight, accuracy = 0.1)) %>%
  dplyr::select(Ticker = ticker, `Asset Class` = asset_class, Weight)

ef_tan_res  <- min_var_pf(frontier_pts_ef$ret[tan_idx_ef], mu_ef, Sigma_ef)
ef_tan_tbl  <- tibble(ticker = ef_tickers, weight = ef_tan_res$weights) %>%
  filter(weight > 0.005) %>% arrange(desc(weight)) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class), by = "ticker") %>%
  mutate(Weight = percent(weight, accuracy = 0.1)) %>%
  dplyr::select(Ticker = ticker, `Asset Class` = asset_class, Weight)
```

```{r build-meta-asset-data, include=FALSE}
library(quadprog)

RF_MA   <- 0.045
WEIGHT_GRID_MA <- seq(0.10, 0.90, by = 0.10)

SECTOR_TICKERS_MA <- c("XLK","XLC","XLV","XLF","XLI",
                        "XLP","XLY","XLE","XLB","XLRE",
                        "XLU","VHT","ITA","ITB","KRE")

# ── Per-regime alpha helper ───────────────────────────────────────────────────
regime_alphas_ma <- function(ret_series, spy_ret, rt) {
  map_dfr(seq_len(nrow(rt)), function(i) {
    ep  <- rt[i, ]
    idx <- paste0(ep$xmin, "/", ep$xmax)
    r_tk <- ret_series[idx]; r_sp <- spy_ret[idx]
    if (length(r_tk) < 5) return(NULL)
    tibble(
      regime = as.character(ep$regime),
      alpha  = as.numeric(Return.cumulative(r_tk)) -
               as.numeric(Return.cumulative(r_sp))
    )
  }) %>%
    group_by(regime) %>%
    summarise(avg_alpha = mean(alpha, na.rm = TRUE), .groups = "drop")
}

max_dd_ma <- function(r) { w <- cumprod(1+r); min((w-cummax(w))/cummax(w), na.rm=TRUE) }

spy_vec_ma  <- as.numeric(na.omit(xts_ret[, "SPY"]))
spy_mdd_ma  <- max_dd_ma(spy_vec_ma)
spy_mu_ma   <- mean(spy_vec_ma) * 252
spy_sig_ma  <- sd(spy_vec_ma) * sqrt(252)
spy_sharpe_ma <- (spy_mu_ma - RF_MA) / spy_sig_ma

# Pre-compute per-ticker regime alphas
ticker_ra_ma <- map(SECTOR_TICKERS_MA, function(tk)
  regime_alphas_ma(xts_ret[, tk], spy_ret, rt)
) %>% setNames(SECTOR_TICKERS_MA)

# Score all pairs at all weights
pairs_raw_ma <- combn(SECTOR_TICKERS_MA, 2, simplify = FALSE)

score_tbl_ma <- map_dfr(pairs_raw_ma, function(pair) {
  tk1 <- pair[1]; tk2 <- pair[2]
  r1 <- as.numeric(na.omit(xts_ret[, tk1]))
  r2 <- as.numeric(na.omit(xts_ret[, tk2]))
  rs <- as.numeric(na.omit(spy_ret))
  n  <- min(length(r1), length(r2), length(rs))
  r1 <- tail(r1,n); r2 <- tail(r2,n); rs <- tail(rs,n)

  a1 <- ticker_ra_ma[[tk1]]; a2 <- ticker_ra_ma[[tk2]]
  alpha_both  <- inner_join(a1, a2, by="regime", suffix=c("_1","_2"))
  regime_corr <- if (nrow(alpha_both)>=2) cor(alpha_both$avg_alpha_1, alpha_both$avg_alpha_2) else NA_real_

  map_dfr(WEIGHT_GRID_MA, function(w) {
    blend   <- w*r1 + (1-w)*r2
    mu_b    <- mean(blend)*252; sig_b <- sd(blend)*sqrt(252)
    mdd_b   <- max_dd_ma(blend)
    sharpe_b <- (mu_b - RF_MA)/sig_b
    bx      <- xts(blend, order.by=tail(index(spy_ret),n))
    ba      <- regime_alphas_ma(bx, spy_ret, rt)
    tibble(tk1=tk1, tk2=tk2, w1=w, w2=1-w,
           sharpe_blend=sharpe_b, mdd_blend=mdd_b,
           dd_delta=mdd_b-spy_mdd_ma, mu_blend=mu_b, sig_blend=sig_b,
           regime_corr=regime_corr,
           alpha_var=var(ba$avg_alpha, na.rm=TRUE),
           fall_alpha=ba$avg_alpha[ba$regime=="Fall"] %||% NA_real_,
           consol_alpha=ba$avg_alpha[ba$regime=="Consolidation"] %||% NA_real_)
  })
})

`%||%` <- function(a, b) if (length(a)) a else b

# Re-run with %||% defined
score_tbl_ma <- map_dfr(pairs_raw_ma, function(pair) {
  tk1 <- pair[1]; tk2 <- pair[2]
  r1 <- as.numeric(na.omit(xts_ret[, tk1]))
  r2 <- as.numeric(na.omit(xts_ret[, tk2]))
  rs <- as.numeric(na.omit(spy_ret))
  n  <- min(length(r1), length(r2), length(rs))
  r1 <- tail(r1,n); r2 <- tail(r2,n)

  a1 <- ticker_ra_ma[[tk1]]; a2 <- ticker_ra_ma[[tk2]]
  alpha_both  <- inner_join(a1, a2, by="regime", suffix=c("_1","_2"))
  regime_corr <- if (nrow(alpha_both)>=2) cor(alpha_both$avg_alpha_1, alpha_both$avg_alpha_2) else NA_real_

  map_dfr(WEIGHT_GRID_MA, function(w) {
    blend    <- w*r1 + (1-w)*r2
    mu_b     <- mean(blend)*252; sig_b <- sd(blend)*sqrt(252)
    mdd_b    <- max_dd_ma(blend); sharpe_b <- (mu_b-RF_MA)/sig_b
    bx       <- xts(blend, order.by=tail(index(spy_ret),n))
    ba       <- regime_alphas_ma(bx, spy_ret, rt)
    fa       <- ba$avg_alpha[ba$regime=="Fall"]
    ca       <- ba$avg_alpha[ba$regime=="Consolidation"]
    tibble(tk1=tk1,tk2=tk2,w1=w,w2=1-w,
           sharpe_blend=sharpe_b,mdd_blend=mdd_b,
           dd_delta=mdd_b-spy_mdd_ma,mu_blend=mu_b,sig_blend=sig_b,
           regime_corr=regime_corr,alpha_var=var(ba$avg_alpha,na.rm=TRUE),
           fall_alpha   = if(length(fa)) fa else NA_real_,
           consol_alpha = if(length(ca)) ca else NA_real_)
  })
})

score_tbl_ma <- score_tbl_ma %>%
  mutate(
    z_sharpe = (sharpe_blend-min(sharpe_blend,na.rm=TRUE))/(max(sharpe_blend,na.rm=TRUE)-min(sharpe_blend,na.rm=TRUE)),
    z_dd     = (dd_delta-min(dd_delta,na.rm=TRUE))/(max(dd_delta,na.rm=TRUE)-min(dd_delta,na.rm=TRUE)),
    z_compl  = (-regime_corr-min(-regime_corr,na.rm=TRUE))/(max(-regime_corr,na.rm=TRUE)-min(-regime_corr,na.rm=TRUE)),
    z_smooth = (-alpha_var-min(-alpha_var,na.rm=TRUE))/(max(-alpha_var,na.rm=TRUE)-min(-alpha_var,na.rm=TRUE)),
    composite = 0.40*z_sharpe + 0.25*z_dd + 0.25*z_compl + 0.10*z_smooth
  )

best_pairs_ma <- score_tbl_ma %>%
  group_by(tk1, tk2) %>%
  slice_max(composite, n=1, with_ties=FALSE) %>%
  ungroup() %>%
  mutate(
    pair_label   = paste0(tk1," + ",tk2),
    weight_label = paste0(round(w1*100),"/",round(w2*100))
  ) %>%
  arrange(desc(composite))

# ── Plot 1: Heatmap ───────────────────────────────────────────────────────────
heatmap_df_ma <- best_pairs_ma %>%
  dplyr::select(tk1, tk2, composite, weight_label) %>%
  bind_rows(best_pairs_ma %>% rename(tk1=tk2, tk2=tk1)) %>%
  mutate(tk1=factor(tk1,levels=SECTOR_TICKERS_MA),
         tk2=factor(tk2,levels=SECTOR_TICKERS_MA))

p_ma_heatmap <- ggplot(heatmap_df_ma, aes(x=tk1, y=tk2, fill=composite)) +
  geom_tile(colour="white", linewidth=0.5) +
  geom_text(aes(label=weight_label), size=2.6, colour="white", fontface="bold") +
  scale_fill_gradient2(low="#d73027", mid="#ffffbf", high="#1a1a6e",
                       midpoint=0.5, name="Composite\nScore", limits=c(0,1)) +
  scale_x_discrete(position="top") +
  theme_minimal(base_size=12) +
  theme(axis.text.x=element_text(angle=45,hjust=0,face="bold",size=10),
        axis.text.y=element_text(face="bold",size=10),
        axis.title=element_blank(), panel.grid=element_blank(),
        plot.title=element_text(face="bold",size=14),
        plot.subtitle=element_text(size=10,colour="grey40")) +
  labs(title="Meta-Asset Pair Scoring Heatmap — Sector ETF Universe",
       subtitle="Colour = composite score | Cell label = optimal blend weight (w1/w2)")

# ── Plot 2: Ranked bar ────────────────────────────────────────────────────────
p_ma_ranked <- best_pairs_ma %>%
  slice_head(n=20) %>%
  mutate(pair_label=fct_reorder(paste0(tk1,"+",tk2,"\n(",weight_label,")"), composite)) %>%
  ggplot(aes(x=pair_label, y=composite, fill=sharpe_blend > spy_sharpe_ma)) +
  geom_col(alpha=0.85, width=0.7) +
  geom_text(aes(label=sprintf("IR=%.2f", sharpe_blend)), hjust=-0.1, size=3, fontface="bold") +
  coord_flip() +
  scale_fill_manual(values=c("TRUE"="#1a1a6e","FALSE"="#d73027"),
                    labels=c("TRUE"=paste0("Sharpe > SPY (",round(spy_sharpe_ma,2),")"),"FALSE"=paste0("Sharpe < SPY")),
                    name=NULL) +
  scale_y_continuous(expand=expansion(mult=c(0,0.18)), limits=c(0,1)) +
  theme_minimal(base_size=12) +
  theme(panel.grid.minor=element_blank(), panel.grid.major.y=element_blank(),
        legend.position="bottom",
        plot.title=element_text(face="bold",size=14),
        plot.subtitle=element_text(size=10,colour="grey40")) +
  labs(title="Top 20 Meta-Asset Pairs — Ranked by Composite Score",
       subtitle="Label = Information Ratio | Blue = beats SPY Sharpe",
       x=NULL, y="Composite Score")

# ── Plot 3: Regime alpha profile for top 6 ───────────────────────────────────
top6_ma <- best_pairs_ma %>% slice_head(n=6)

regime_profile_ma <- map_dfr(seq_len(nrow(top6_ma)), function(i) {
  p  <- top6_ma[i,]
  r1 <- as.numeric(na.omit(xts_ret[,p$tk1]))
  r2 <- as.numeric(na.omit(xts_ret[,p$tk2]))
  n  <- min(length(r1),length(r2),length(as.numeric(na.omit(spy_ret))))
  r1 <- tail(r1,n); r2 <- tail(r2,n)
  bx <- xts(p$w1*r1+p$w2*r2, order.by=tail(index(spy_ret),n))
  ba <- regime_alphas_ma(bx, spy_ret, rt)
  a1 <- ticker_ra_ma[[p$tk1]]; a2 <- ticker_ra_ma[[p$tk2]]
  lbl <- paste0(p$tk1,"+",p$tk2," (",p$weight_label,")")
  bind_rows(
    ba %>% mutate(series=lbl, is_blend=TRUE),
    a1 %>% mutate(series=p$tk1, is_blend=FALSE),
    a2 %>% mutate(series=p$tk2, is_blend=FALSE)
  ) %>% mutate(pair=lbl)
}) %>%
  mutate(
    regime   = factor(regime, levels=c("Fall","Recovery","Consolidation")),
    fill_col = case_when(
      is_blend & avg_alpha>=0  ~ "#1a1a6e",
      is_blend & avg_alpha< 0  ~ "#d73027",
      !is_blend & avg_alpha>=0 ~ "#7EB8E8",
      TRUE                     ~ "#fc8d59"
    )
  )

p_ma_regime <- ggplot(regime_profile_ma,
                      aes(x=series, y=avg_alpha, fill=fill_col, alpha=is_blend)) +
  geom_col(width=0.65) +
  geom_hline(yintercept=0, linewidth=0.6, colour="grey30") +
  geom_text(aes(label=sprintf("%+.1f%%",avg_alpha*100),
                vjust=if_else(avg_alpha>=0,-0.3,1.2)),
            size=2.6, fontface="bold") +
  scale_fill_identity() +
  scale_alpha_manual(values=c("TRUE"=1.0,"FALSE"=0.55), guide="none") +
  scale_y_continuous(labels=percent_format(accuracy=1)) +
  facet_grid(pair~regime, scales="free_x") +
  theme_minimal(base_size=10) +
  theme(strip.text.x=element_text(face="bold",size=10),
        strip.text.y=element_text(face="bold",size=8,angle=0),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank(),
        axis.text.x=element_text(size=8,face="bold",angle=30,hjust=1),
        axis.title=element_blank(),
        plot.title=element_text(face="bold",size=14),
        plot.subtitle=element_text(size=10,colour="grey40")) +
  labs(title="Regime Alpha Profile — Top 6 Meta-Asset Pairs",
       subtitle="Dark = blend alpha vs SPY | Light = individual component alpha")

# ── Plot 4: Risk-return scatter ───────────────────────────────────────────────
indiv_ma <- tibble(
  label = SECTOR_TICKERS_MA,
  mu    = map_dbl(SECTOR_TICKERS_MA, ~mean(as.numeric(na.omit(xts_ret[,.x])))*252),
  vol   = map_dbl(SECTOR_TICKERS_MA, ~sd(as.numeric(na.omit(xts_ret[,.x])))*sqrt(252)),
  type  = "Individual Sector"
)
blend_ma <- best_pairs_ma %>% slice_head(n=15) %>%
  transmute(label=paste0(tk1,"+",tk2), mu=mu_blend, vol=sig_blend,
            type="Meta-Asset (top 15)")
spy_ma <- tibble(label="SPY", mu=spy_mu_ma, vol=spy_sig_ma, type="Benchmark")
scatter_ma <- bind_rows(indiv_ma, blend_ma, spy_ma)

p_ma_scatter <- ggplot(scatter_ma, aes(x=vol, y=mu, colour=type, size=type)) +
  geom_point(alpha=0.80) +
  geom_text_repel(aes(label=label), size=2.8, max.overlaps=20,
                  segment.color="grey70", show.legend=FALSE) +
  geom_hline(yintercept=spy_mu_ma, linetype="dashed", colour="grey50", linewidth=0.6) +
  scale_colour_manual(values=c("Individual Sector"="#4A90D9",
                                "Meta-Asset (top 15)"="#1a1a6e",
                                "Benchmark"="#D90429"), name=NULL) +
  scale_size_manual(values=c("Individual Sector"=2.5,"Meta-Asset (top 15)"=3.5,
                              "Benchmark"=4), guide="none") +
  scale_x_continuous(labels=percent_format(accuracy=1)) +
  scale_y_continuous(labels=percent_format(accuracy=1)) +
  theme_minimal(base_size=12) +
  theme(panel.grid.minor=element_blank(), legend.position="bottom",
        plot.title=element_text(face="bold",size=14),
        plot.subtitle=element_text(size=10,colour="grey40")) +
  labs(title="Risk-Return: Individual Sectors vs Top Meta-Asset Blends",
       subtitle="Meta-assets (dark) cluster differently from individual sectors",
       x="Annualised Volatility", y="Annualised Return")
```

```{r build-ranges-data, include=FALSE}
library(tidyquant)

RANGE_GROUPS <- list(
  "Equity — US Broad"    = c("SPY", "QQQ", "VTI", "IWM", "IJH", "URTH", "ACWX", "IEFA", "FEZ", "EWJ"),
  "Equity — Sector"      = c("XLK", "XLC", "XLV", "XLF", "XLI", "XLP", "XLY", "XLE", "XLB", "XLU", "XLRE"),
  "Equity — Thematic"    = c("SMH", "AIQ", "IBB", "WCLD", "CIBR", "ITA", "VHT", "KRE", "ITB", "PSP"),
  "Fixed Income"         = c("AGG", "SHY", "IEI", "IEF", "TLT", "HYG", "LQD", "EMB", "SGOV", "BNDX"),
  "Commodity & Real"     = c("GLD", "SLV", "GDX", "DBA", "PDBC", "COPX", "URA", "LIT", "DJP", "MOO")
)

all_syms_rng <- unique(unlist(RANGE_GROUPS))

raw_rng <- tq_get(all_syms_rng,
                  get  = "stock.prices",
                  from = Sys.Date() - 365,
                  to   = Sys.Date()) %>%
  filter(!is.na(close))

raw_rng <- raw_rng %>%
  group_by(symbol, date) %>%
  slice_max(close, n = 1, with_ties = FALSE) %>%
  ungroup()

quote_rng <- raw_rng %>%
  group_by(symbol) %>%
  arrange(date) %>%
  summarise(
    last_date  = max(date),
    last       = last(close),
    prev_close = nth(close, -2),
    day_high   = last(high),
    day_low    = last(low),
    wk52_high  = max(high,  na.rm = TRUE),
    wk52_low   = min(low,   na.rm = TRUE),
    volume     = last(volume),
    avg_vol_3m = {
      cutoff <- max(date) - 63
      v      <- volume[date >= cutoff & !is.na(volume)]
      if (length(v) > 0) mean(v) else NA_real_
    },
    .groups = "drop"
  ) %>%
  mutate(
    change     = last - prev_close,
    change_pct = change / prev_close,
    pos_52wk   = (last - wk52_low) / (wk52_high - wk52_low),
    n_day_low  = pmax((day_low  - wk52_low) / (wk52_high - wk52_low), 0),
    n_day_high = pmin((day_high - wk52_low) / (wk52_high - wk52_low), 1),
    n_last     = pos_52wk,
    chg_col    = if_else(change >= 0, "#27AE60", "#C0392B"),
    lbl_low    = sprintf("%.1f", wk52_low),
    lbl_high   = sprintf("%.1f", wk52_high),
    lbl_price  = sprintf("%.1f", last),
    lbl_chg    = sprintf("(%+.1f%%)", change_pct * 100)
  )

# ── Build one plot per group ──────────────────────────────────────────────────
make_range_plot <- function(symbols, group_title, label_size = 2.8,
                             point_size = 3.0, bar_height = 0.35) {
  df <- quote_rng %>%
    filter(symbol %in% symbols, !is.na(n_last)) %>%
    mutate(symbol = fct_reorder(symbol, n_last))

  ggplot(df, aes(y = symbol)) +
    geom_segment(aes(x = 0, xend = 1, yend = symbol),
                 colour = "#DDDDDD", linewidth = bar_height * 10, lineend = "round") +
    geom_segment(aes(x = n_day_low, xend = n_day_high, yend = symbol),
                 colour = "#4A90D9", linewidth = bar_height * 5,  lineend = "round") +
    geom_vline(xintercept = seq(0.10, 0.90, by = 0.10),
               colour = "grey75", linewidth = 0.4, linetype = "dashed") +
    geom_point(aes(x = n_last, fill = chg_col),
               shape = 21, size = point_size, colour = "white", stroke = 1.2) +
    scale_fill_identity() +
    geom_text(aes(x = 0,      label = lbl_low),
              hjust = 1.3,  vjust = 0.5, size = label_size, colour = "grey45") +
    geom_text(aes(x = 1,      label = lbl_high),
              hjust = -0.3, vjust = 0.5, size = label_size, colour = "grey45") +
    geom_text(aes(x = n_last, label = lbl_price, colour = chg_col),
              hjust = -0.45, vjust = 0.5, size = label_size, fontface = "bold") +
    geom_text(aes(x = n_last, label = lbl_chg,   colour = chg_col),
              vjust = -0.6,  hjust = 0.5, size = label_size, fontface = "bold") +
    scale_colour_identity() +
    scale_x_continuous(labels = percent_format(accuracy = 1),
                       limits = c(-0.15, 1.15),
                       breaks = seq(0, 1, by = 0.10),
                       expand = expansion(0)) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.title         = element_blank(),
      axis.text.y        = element_text(face = "bold", size = 11),
      axis.text.x        = element_text(size = 9, colour = "grey55"),
      plot.title         = element_text(face = "bold", size = 13),
      plot.subtitle      = element_text(size = 9, colour = "grey45"),
      plot.caption       = element_text(size = 8, colour = "grey55")
    ) +
    labs(
      title    = group_title,
      subtitle = "0% = 52-wk low  |  100% = 52-wk high  |  Grey = 52-wk  |  Blue = today  |  Dot = last",
      caption  = paste0("As of ", format(max(df$last_date, na.rm=TRUE), "%d %b %Y"))
    )
}

range_plots <- imap(RANGE_GROUPS, make_range_plot)

# ── XY plots per asset class group ───────────────────────────────────────────
xy_plots <- imap(RANGE_GROUPS, function(symbols, grp_title) {
  df <- quote_rng %>%
    filter(symbol %in% symbols) %>%
    rename(symbol = symbol)   # already named correctly
  make_xy_plot(df, group_title = paste0("52-Wk Position Map — ", grp_title))
})

# ── XY scatter: % above Low vs % below High (full universe) ──────────────────
xy_rng <- quote_rng %>%
  filter(!is.na(last)) %>%
  mutate(
    pct_above_low  = (last - wk52_low)  / wk52_low  * 100,
    pct_below_high = (wk52_high - last) / wk52_high * 100,
    chg_col        = if_else(change >= 0, "#27AE60", "#C0392B")
  ) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class),
            by = c("symbol" = "ticker")) %>%
  mutate(asset_class = replace_na(asset_class, "Other"))

AC_PAL_RNG <- c(
  "Equity"      = "#1B3A6B",
  "FixedIncome" = "#27AE60",
  "Commodity"   = "#E67E22",
  "FX"          = "#C9A800",
  "RealAsset"   = "#C0392B",
  "MultiAsset"  = "#7F8C8D",
  "Other"       = "#AAAAAA"
)

p_xy_rng <- ggplot(xy_rng, aes(x = pct_above_low, y = pct_below_high,
                                colour = asset_class)) +
  geom_vline(xintercept = mean(xy_rng$pct_above_low,  na.rm=TRUE),
             linetype="dotted", colour="grey70", linewidth=0.5) +
  geom_hline(yintercept = mean(xy_rng$pct_below_high, na.rm=TRUE),
             linetype="dotted", colour="grey70", linewidth=0.5) +
  geom_point(aes(shape = change >= 0), size = 3, alpha = 0.85) +
  scale_shape_manual(values = c("TRUE"=16, "FALSE"=17), guide="none") +
  geom_text_repel(aes(label = symbol), size = 2.8, fontface="bold",
                  max.overlaps=40, segment.color="grey75", show.legend=FALSE) +
  scale_colour_manual(values = AC_PAL_RNG, name="Asset Class", na.value="grey70") +
  scale_x_continuous(labels = function(x) paste0("+", round(x), "%"),
                     name   = "% Above 52-Wk Low",
                     breaks = seq(0, ceiling(max(xy_rng$pct_above_low, na.rm=TRUE)/20)*20, by=20)) +
  scale_y_reverse(labels = function(x) paste0("-", round(x), "%"),
                  name   = "% Below 52-Wk High") +
  annotate("text", x=Inf,  y=-Inf, hjust=1.1, vjust=-0.6,
           label="Near 52-Wk HIGH", size=3.2, colour="grey50", fontface="italic") +
  annotate("text", x=-Inf, y=Inf,  hjust=-0.1, vjust=1.5,
           label="Near 52-Wk LOW",  size=3.2, colour="grey50", fontface="italic") +
  theme_minimal(base_size=13) +
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_line(colour="grey88", linewidth=0.4),
        legend.position="right",
        axis.title=element_text(size=12, face="bold"),
        axis.text=element_text(size=10, face="bold"),
        plot.title=element_text(face="bold", size=14),
        plot.subtitle=element_text(size=10, colour="grey45")) +
  labs(
    title    = "52-Week Position Map — % Above Low vs % Below High",
    subtitle = "Bottom-right = near 52-wk high  |  Top-left = near 52-wk low  |  Triangle = down today",
    caption  = paste0("As of ", format(max(xy_rng$last_date, na.rm=TRUE), "%d %b %Y"))
  )
```

## 5 · Pair Analysis — Cost-Benefit of Adding a Ticker

<div class="insight-box">
**What:** For every ETF, model the cost-benefit of a 10% satellite allocation
alongside a 90% SPY core. Metrics: annualised return contribution, tracking error
cost, and theoretical MaxDD (Magdon-Ismail formula vs empirical).
This was the **first formulation** of what later became the Exante Hypothesis.
</div>

```{r p5-pair-scatter, fig.height=7}
# Build cost-benefit vectors for the core universe
T_yrs <- as.numeric(diff(range(index(xts_ret)))) / 365

magdon_mdd <- function(sigma, mu, T) {
  if (is.na(mu) || mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  (sigma / S) * (log(S * sqrt(T)) + 0.35)
}

cb_tbl <- map_dfr(core_universe, function(tk) {
  r_tk  <- as.numeric(na.omit(xts_ret[, tk]))
  r_spy <- as.numeric(na.omit(xts_ret[, "SPY"]))
  n     <- min(length(r_tk), length(r_spy))
  r_tk  <- tail(r_tk,  n)
  r_spy <- tail(r_spy, n)

  spread  <- r_tk - r_spy
  mu_spr  <- mean(spread) * 252
  sig_spr <- sd(spread) * sqrt(252)
  mdd_emp <- {
    w <- cumprod(1 + spread); min((w - cummax(w)) / cummax(w))
  }
  mdd_th  <- magdon_mdd(sig_spr, mu_spr, T_yrs)
  rho     <- cor(r_tk, r_spy)

  tibble(
    ticker   = tk,
    mu_spr   = mu_spr,
    sig_spr  = sig_spr,
    mdd_emp  = mdd_emp,
    mdd_th   = mdd_th,
    rho      = rho,
    ir       = mu_spr / sig_spr
  )
}) %>%
  left_join(etf_metadata %>% dplyr::select(ticker, asset_class, sub_block),
            by = "ticker")

ggplot(cb_tbl, aes(x = mdd_th, y = mdd_emp, colour = asset_class)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(aes(size = abs(ir)), alpha = 0.8) +
  geom_text_repel(aes(label = ticker), size = 3, max.overlaps = 15,
                  segment.color = "grey70", show.legend = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_size_continuous(range = c(2, 7), name = "|IR|") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
  labs(
    title    = "P5: Theoretical vs Empirical MaxDD — Spread vs SPY",
    subtitle = "Diagonal = theory equals empirical | Above diagonal = empirical DD worse than predicted",
    x = "Theoretical MaxDD (Magdon-Ismail formula)",
    y = "Empirical MaxDD (historical)",
    colour = NULL
  )
```

<div class="finding-box">
**Key finding:** For high-correlation tickers (corr > 0.8), empirical MaxDD closely
matches the Magdon-Ismail theoretical prediction. For near-zero correlation tickers,
empirical MaxDD is systematically *worse* than theory — they produce larger-than-expected
drawdowns. This was the first signal of the anti-diversification phenomenon.
</div>

### P5b — XLK vs SPY: Regime-by-Regime Relative Overlay

<div class="insight-box">
**What:** XLK (Technology Select Sector) is a concentrated bet on the largest
SPY subsector. This overlay shows, period-by-period, whether XLK enhanced or
detracted relative to SPY across every Fall, Recovery, and Consolidation episode.
It is the most direct test of the satellite role played by the tech satellite in
the SAA portfolio.
</div>

```{r p5b-xlk-overlay, fig.height=8}
print(plot_regime_rel_overlay(
  xts_ret = xts_ret,
  tickers = c("XLK"),
  master  = "SPY",
  t_fall  = T_FALL
))
```

```{r p5b-xlk-heatmap, fig.height=5}
print(plot_regime_rel_heatmap(
  xts_ret = xts_ret,
  tickers = c("XLK"),
  master  = "SPY"
))
```

<div class="finding-box">
**Key finding — XLK vs SPY:**
XLK is a high-correlation satellite (ρ ≈ 0.95 vs SPY). Its alpha pattern is
asymmetric: it shows **ALPHA** type outperformance (both up, XLK more) in
Consolidation and Recovery, but **LOSS** or **EXPOSED** in Fall — amplifying
drawdowns relative to SPY. This makes XLK a pure growth-tilted satellite:
rewarding in calm/rising markets, punishing in stress. Appropriate as a small
(5%) position within the SAA, not as a hedge.
</div>

### P5b-ext — Extended Regime Overlay ★

<div class="insight-box">
**What (GEM):** `plot_regime_overlay_ext()` — three-layer regime intelligence in one chart.
Row 1 (BMK): SPY cumulative return with Fall/Recovery/Consolidation shading and per-regime return labels.
Row 2 (ticker): overlay cumulative return with per-regime return labels.
Row 3 (Rel): per-regime green/red bar — green when ticker beat BMK in that period, red when it lagged, with α% label.
Below: relative cumulative return area panel (ticker vs BMK) with current α annotated.
</div>

```{r p5b-ext-qqq, fig.height=9}
print(plot_regime_overlay_ext("QQQ", bmk = "SPY", xts_ret = xts_ret))
```

```{r p5b-ext-xlk, fig.height=9}
print(plot_regime_overlay_ext("XLK", bmk = "SPY", xts_ret = xts_ret))
```

```{r p5b-ext-gld, fig.height=9}
print(plot_regime_overlay_ext("GLD", bmk = "SPY", xts_ret = xts_ret))
```

### P5b-triple — Triple Overlay + Rolling 12m Relative Return

<div class="insight-box">
**What:** `plot_regime_triple()` — compact two-panel view per ticker.
Top: BMK (red) + ticker (dark blue) + relative cumulative return (grey) on one shared axis with three bar strips.
Bottom: rolling 252-day relative return ribbon — green when ticker outperforms over trailing year, red when it lags. Regime shading on both panels.
</div>

```{r p5b-triple-qqq, fig.height=9}
print(plot_regime_triple("QQQ", bmk = "SPY", xts_ret = xts_ret))
```

```{r p5b-triple-xlk, fig.height=9}
print(plot_regime_triple("XLK", bmk = "SPY", xts_ret = xts_ret))
```

```{r p5b-triple-gld, fig.height=9}
print(plot_regime_triple("GLD", bmk = "SPY", xts_ret = xts_ret))
```

### P5c — Full Pair Analysis: L/S Decomposition & DD Theory

<div class="insight-box">
**What:** Every active allocation = benchmark + L/S overlay on (new − funded asset).
The funded-from asset determines spread vol and TE. Five concrete decisions tested
at 10% notional against a 60/40 benchmark. Key question: does the active bet earn
enough to justify its drawdown cost?
</div>

### P5c — KF2 Winners: All Three vs SPY

<div class="insight-box">
**What:** Full 7-panel deep-dive for each of the three KF2 winners (SMH, XLK, QQQ) —
the only tickers in the Sovereign Universe that pass ρ ≥ 0.75 **and** have positive IR
vs SPY. Panels: cumulative wealth · L/S spread · drawdown · return distribution ·
rolling correlation · regime overlay · rolling IR.
</div>

```{r p5c-xlk-combined, fig.height=35}
# XLK — IR 0.51 | ρ 0.92 | Spread MaxDD −18.4% | ΔDD 0.0%
print(pair_results[[1]]$out$combined)
```

```{r p5c-smh-combined, fig.height=35}
# SMH — IR 0.90 | ρ 0.79 | Spread MaxDD −27.2% | ΔDD −0.8%
print(pair_results[[2]]$out$combined)
```

```{r p5c-qqq-combined, fig.height=35}
# QQQ — IR 0.48 | ρ 0.93 | Spread MaxDD −21.1% | ΔDD +0.3%
print(pair_results[[3]]$out$combined)
```

```{r p5c-master-table, results='asis'}
map_dfr(pair_results, function(r) {
  ps <- r$ps
  tibble(
    Decision     = r$decision,
    Type         = r$type,
    `ρ`          = round(ps$rho,        2),
    `Spread Vol` = percent(ps$sig,      accuracy = 0.1),
    `Spread IR`  = round(ps$ir,         2),
    `E[Spread]`  = percent(ps$mu,       accuracy = 0.1),
    `MaxDD act`  = percent(abs(ps$mdd), accuracy = 0.1),
    `MaxDD th`   = percent(ps$mdd_theory, accuracy = 0.1),
    `TE (10%)`   = percent(ps$te_10pct, accuracy = 0.01),
    `Ret (10%)`  = percent(ps$ret_10pct,accuracy = 0.01)
  )
}) %>%
  arrange(Type, desc(`Spread IR`)) %>%
  knitr::kable(caption = "All five decisions ranked by type — 10% notional vs 60/40")
```

```{r p5c-spread-vol, fig.height=5}
map_dfr(pair_results, function(r) {
  tibble(decision = r$decision, type = r$type,
         spread_vol = r$ps$sig, ir = r$ps$ir)
}) %>%
  mutate(decision = fct_reorder(decision, spread_vol)) %>%
  ggplot(aes(x = decision, y = spread_vol, fill = type)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_text(aes(label = sprintf("IR=%.2f", ir)),
            hjust = -0.15, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("Equity swap" = "#457B9D",
                                "Cross-asset" = "#D90429")) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.18))) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(colour = "grey50", size = 9)) +
  labs(title    = "Spread Volatility by Decision (10% notional)",
       subtitle = "Blue = funded from SPY (equity swap)  |  Red = funded from AGG (cross-asset)\nIR labels = full-period Information Ratio",
       x = NULL, y = "Annualised Spread Vol", fill = NULL)
```

```{r p5c-ir, fig.height=5}
ir_df <- map_dfr(pair_results, function(r) {
  tibble(decision  = r$decision, type = r$type,
         ir        = r$ps$ir,
         p_correct = 0.5 + r$ps$ir / 2)
})

ir_df %>%
  mutate(decision = fct_reorder(decision, ir)) %>%
  ggplot(aes(x = decision, y = ir, fill = type)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = 0.50, linetype = "dashed",
             colour = "#2D6A4F", linewidth = 0.8) +
  geom_hline(yintercept = 0.30, linetype = "dotted",
             colour = "grey50", linewidth = 0.6) +
  annotate("text", x = 0.6, y = 0.52, hjust = 0, size = 3,
           colour = "#2D6A4F", label = "IR = 0.50 (top quartile)") +
  geom_text(aes(label = sprintf("P=%.0f%%", p_correct * 100)),
            hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("Equity swap" = "#457B9D",
                                "Cross-asset" = "#D90429")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(colour = "grey50", size = 9)) +
  labs(title    = "Information Ratio by Decision",
       subtitle = "P = probability forecast must be correct to justify the bet",
       x = NULL, y = "IR", fill = NULL)
```

<div class="finding-box">
**Key findings — Pair Analysis:**
(1) **Equity sleeve swaps** (sector vs SPY, ρ > 0.85) have low spread vol (~10%)
and small TE at 10% notional — the hurdle return is achievable with modest conviction.
(2) **Cross-asset additions** (equity vs AGG, ρ ≈ −0.20) have 2–3× higher spread vol
and TE — the hurdle return is much higher; requires strong directional conviction.
(3) **MaxDD theory confirms:** Sharpe² ≈ Calmar. XLK spread has a higher Sharpe than
cross-asset bets → smaller theoretical MaxDD per unit of active return.
(4) **IR > 0.50** (top quartile) requires P(forecast correct) > 75% — a very high bar
for sustained active calls on any single decision.
</div>

<small>Full report: `Rmd/pair_analysis_portfolio_construction.Rmd`</small>

---

### P5d — KF1: Anti-Diversification — Theory vs Empirical MaxDD (Full Universe)

<div class="insight-box">
**What:** For every ticker in the Sovereign Universe, compute the Magdon-Ismail
theoretical MaxDD of the spread (ticker − SPY) and compare to the empirical MaxDD.
The key question: does low correlation protect against drawdown, as conventional
portfolio theory predicts?
</div>

```{r p5-kf1-scatter, fig.height=7}
kf1_p1 <- ggplot(cb_tbl_kf, aes(x = mdd_th, y = mdd_emp, colour = rho_band)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.8) +
  annotate("text",
           x = min(cb_tbl_kf$mdd_th, na.rm = TRUE) * 0.55,
           y = min(cb_tbl_kf$mdd_emp, na.rm = TRUE) * 0.97,
           label = "Empirical DD worse\nthan theory predicts",
           hjust = 0, vjust = 1, size = 4, colour = "grey35", fontface = "italic") +
  annotate("text",
           x = max(cb_tbl_kf$mdd_th, na.rm = TRUE) * 0.97,
           y = max(cb_tbl_kf$mdd_th, na.rm = TRUE) * 0.55,
           label = "Theory overstates DD",
           hjust = 1, vjust = 0, size = 4, colour = "grey35", fontface = "italic") +
  geom_point(aes(size = abs(rho)), alpha = 0.78) +
  geom_text_repel(
    data = cb_tbl_kf %>%
      group_by(rho_band) %>%
      slice_min(residual, n = 3) %>%
      ungroup(),
    aes(label = ticker),
    size = 3.8, fontface = "bold", max.overlaps = 40,
    segment.color = "grey60", show.legend = FALSE
  ) +
  scale_colour_manual(values = RHO_PAL, name = "ρ vs SPY") +
  scale_size_continuous(range = c(2, 6), guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor  = element_blank(),
    legend.position   = "right",
    legend.text       = element_text(size = 12),
    legend.title      = element_text(size = 12, face = "bold"),
    axis.title        = element_text(size = 13),
    axis.text         = element_text(size = 12),
    plot.title        = element_text(face = "bold", size = 15),
    plot.subtitle     = element_text(size = 12, colour = "grey40")
  ) +
  labs(
    title    = "KF1-A: Theoretical vs Empirical MaxDD of Spread vs SPY",
    subtitle = paste0(
      "Below diagonal = empirical DD worse than Magdon-Ismail predicts\n",
      "Low-ρ tickers (red/orange) cluster below the diagonal — anti-diversification in action"
    ),
    x = "Theoretical MaxDD of spread (Magdon-Ismail)",
    y = "Empirical MaxDD of spread (historical)"
  )
print(kf1_p1)
```

```{r p5-kf1-boxplot, fig.height=6}
kf1_p2 <- ggplot(cb_tbl_kf, aes(x = rho_band, y = residual, fill = rho_band)) +
  geom_hline(yintercept = 0, linewidth = 0.8, colour = "grey30") +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(colour = rho_band), width = 0.18, size = 2.5, alpha = 0.6) +
  geom_text_repel(
    data = cb_tbl_kf %>%
      group_by(rho_band) %>%
      slice_min(residual, n = 2) %>%
      ungroup(),
    aes(label = ticker, colour = rho_band),
    size = 3.5, fontface = "bold", max.overlaps = 20,
    segment.color = "grey60", show.legend = FALSE,
    nudge_x = 0.3
  ) +
  scale_fill_manual(values  = RHO_PAL, guide = "none") +
  scale_colour_manual(values = RHO_PAL, guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.title         = element_text(size = 13),
    axis.text          = element_text(size = 13),
    plot.title         = element_text(face = "bold", size = 15),
    plot.subtitle      = element_text(size = 12, colour = "grey40")
  ) +
  labs(
    title    = "KF1-B: Residual (Empirical − Theory) MaxDD by ρ Band",
    subtitle = paste0(
      "Below zero = empirical DD worse than theory predicts\n",
      "Low-ρ bands have deeper and more consistent negative residuals"
    ),
    x = "Correlation band vs SPY",
    y = "Residual MaxDD (empirical − theory)"
  )
print(kf1_p2)
```

<div class="finding-box">
**KF1 — Anti-Diversification:** Low-correlation tickers have **worse** empirical MaxDD
than Magdon-Ismail theory predicts (they cluster below the diagonal in KF1-A).
High-ρ tickers track theory closely. Conventional wisdom — "add a low-correlation
asset to reduce portfolio risk" — fails for the MaxDD metric: near-zero ρ tickers
go up and down for entirely different reasons than SPY, so when SPY falls hard
the ticker may also fall, producing drawdowns theory did not anticipate.
</div>

<small>Full script: `key_findings/kf1_anti_diversification.R`</small>

---

### P5e — KF2: L/S Pair Selection Screen (Full Universe vs SPY)

<div class="insight-box">
**What:** Screen every ticker for suitability as the long leg in a Long ticker / Short SPY
overlay. Two hard screens: ρ ≥ 0.75 AND spread MaxDD ≥ −30%. Among those that pass,
rank by Information Ratio, ΔMaxDD, and empirical P(spread > 0 over 1 quarter).
</div>

```{r p5-kf2-screening, fig.height=7}
kf2_p1 <- ggplot(spread_tbl_kf, aes(x = rho, y = mdd_spr, colour = rho_band)) +
  geom_vline(xintercept = RHO_SCREEN, linetype = "dashed",
             colour = "#27ae60", linewidth = 0.8) +
  geom_hline(yintercept = MDD_SCREEN, linetype = "dashed",
             colour = "#27ae60", linewidth = 0.8) +
  annotate("rect",
           xmin = RHO_SCREEN, xmax = Inf,
           ymin = MDD_SCREEN, ymax = 0,
           fill = "#eafaf1", alpha = 0.4) +
  annotate("text", x = RHO_SCREEN + 0.01, y = -0.01,
           label = "VALID L/S PAIR\nCANDIDATES",
           hjust = 0, vjust = 1, size = 3.2,
           colour = "#27ae60", fontface = "bold") +
  geom_point(aes(size = abs(ir)), alpha = 0.75) +
  geom_text_repel(
    data = spread_tbl_kf %>% filter(passes_screen | rho > 0.80),
    aes(label = ticker),
    size = 2.8, max.overlaps = 20,
    segment.color = "grey60", show.legend = FALSE
  ) +
  scale_colour_manual(values = RHO_PAL, name = "Correlation\nvs SPY") +
  scale_size_continuous(range = c(1.5, 5), name = "|IR|") +
  scale_x_continuous(labels = number_format(accuracy = 0.01)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, colour = "grey40")
  ) +
  labs(
    title    = "KF2-A: L/S Pair Screening — ρ vs Spread MaxDD",
    subtitle = sprintf(
      "Green zone: ρ > %.2f AND spread MaxDD > %.0f%%  =  valid SPY-replacement candidates\nLow-ρ tickers have deep spread-DD — unsuitable as L/S legs regardless of standalone returns",
      RHO_SCREEN, MDD_SCREEN * 100
    ),
    x = "Correlation vs SPY (ρ)",
    y = "Spread MaxDD (Long ticker − Short SPY)"
  )
print(kf2_p1)
```

```{r p5-kf2-ir-ranking, fig.height=6}
kf2_p2_data <- spread_tbl_kf %>%
  filter(rho >= RHO_SCREEN) %>%
  arrange(ir) %>%
  mutate(ticker = factor(ticker, levels = ticker))

kf2_p2 <- ggplot(kf2_p2_data, aes(x = ticker, y = ir, fill = ir > 0)) +
  geom_col(alpha = 0.85) +
  geom_hline(yintercept = 0, linewidth = 0.6, colour = "grey30") +
  scale_fill_manual(values = c("TRUE" = "#1a1a6e", "FALSE" = "#d73027"), guide = "none") +
  scale_y_continuous(labels = number_format(accuracy = 0.01)) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x        = element_text(angle = 60, hjust = 1, size = 9),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(size = 10, colour = "grey40")
  ) +
  labs(
    title    = sprintf("KF2-B: IR of L/S Spread vs SPY — High-ρ Candidates (ρ > %.2f)", RHO_SCREEN),
    subtitle = "Positive IR = ticker systematically outperforms SPY on a risk-adjusted basis",
    x = NULL,
    y = "Information Ratio (Ann excess return / Ann tracking error)"
  )
print(kf2_p2)
```

```{r p5-kf2-ir-vs-dd, fig.height=7}
kf2_p3_data <- spread_tbl_kf %>%
  filter(rho >= RHO_SCREEN) %>%
  mutate(
    quadrant = case_when(
      ir > 0 & dd_delta > 0  ~ "Best: +IR, +ΔDD",
      ir > 0 & dd_delta <= 0 ~ "Return edge only",
      ir <= 0 & dd_delta > 0 ~ "DD protection only",
      TRUE                    ~ "Neither"
    ),
    quadrant = factor(quadrant, levels = c(
      "Best: +IR, +ΔDD", "Return edge only", "DD protection only", "Neither"
    ))
  )

kf2_p3 <- ggplot(kf2_p3_data,
                 aes(x = ir, y = dd_delta, size = prob_1q, colour = quadrant)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.7) +
  annotate("text", x =  0.95, y =  max(kf2_p3_data$dd_delta, na.rm = TRUE) * 0.92,
           label = "Best zone\n+IR  +ΔDD", hjust = 1, size = 3.5,
           colour = "#1a1a6e", fontface = "italic") +
  annotate("text", x = min(kf2_p3_data$ir, na.rm = TRUE) * 0.85,
           y =  max(kf2_p3_data$dd_delta, na.rm = TRUE) * 0.92,
           label = "DD protection\nonly", hjust = 0, size = 3.2,
           colour = "#27ae60", fontface = "italic") +
  annotate("text", x =  0.95, y =  min(kf2_p3_data$dd_delta, na.rm = TRUE) * 0.92,
           label = "Return edge\nonly", hjust = 1, size = 3.2,
           colour = "#2196F3", fontface = "italic") +
  geom_point(alpha = 0.80) +
  geom_text_repel(
    aes(label = ticker),
    size = 3.5, fontface = "bold", max.overlaps = 30,
    segment.color = "grey60", show.legend = FALSE
  ) +
  scale_colour_manual(values = QUAD_PAL, name = "Quadrant") +
  scale_size_continuous(
    range  = c(3, 14),
    name   = "P(spread > 0\nover 1 Quarter)",
    labels = percent_format(accuracy = 1),
    breaks = c(0.45, 0.50, 0.55, 0.60, 0.65)
  ) +
  scale_x_continuous(labels = number_format(accuracy = 0.01)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor  = element_blank(),
    legend.position   = "right",
    legend.text       = element_text(size = 11),
    legend.title      = element_text(size = 11, face = "bold"),
    axis.title        = element_text(size = 12),
    axis.text         = element_text(size = 11),
    plot.title        = element_text(face = "bold", size = 14),
    plot.subtitle     = element_text(size = 11, colour = "grey40")
  ) +
  labs(
    title    = sprintf("KF2-C: IR vs ΔMaxDD — High-ρ Candidates (ρ > %.2f)", RHO_SCREEN),
    subtitle = paste0(
      "X = Information Ratio of spread  |  Y = MaxDD improvement vs pure SPY\n",
      "Circle size = empirical P(spread > 0 over any 63-day window)  |  Best zone: top-right"
    ),
    x = "Information Ratio (spread vs SPY)",
    y = "ΔMaxDD: blend vs pure SPY  (positive = improvement)"
  )
print(kf2_p3)
```

```{r p5-kf2-dd-bar, fig.height=6}
kf2_p4_data <- spread_tbl_kf %>%
  filter(rho >= RHO_SCREEN) %>%
  arrange(dd_delta) %>%
  mutate(ticker = factor(ticker, levels = ticker))

kf2_p4 <- ggplot(kf2_p4_data, aes(x = ticker, y = dd_delta, fill = dd_delta > 0)) +
  geom_col(alpha = 0.85) +
  geom_hline(yintercept = 0, linewidth = 0.6, colour = "grey30") +
  scale_fill_manual(values = c("TRUE" = "#27ae60", "FALSE" = "#d73027"), guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x        = element_text(angle = 60, hjust = 1, size = 9),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(size = 10, colour = "grey40")
  ) +
  labs(
    title    = sprintf("KF2-D: MaxDD Improvement from 10%% Swap (High-ρ Only, ρ > %.2f)", RHO_SCREEN),
    subtitle = "Green = 90% SPY + 10% ticker has LESS severe MaxDD than 100% SPY",
    x = NULL,
    y = "ΔMaxDD: blend vs pure SPY  (positive = improvement)"
  )
print(kf2_p4)
```

<div class="finding-box">
**KF2 — L/S Pair Selection:** Only **3 tickers** in the full Sovereign Universe pass
ρ ≥ 0.75 AND have positive IR vs SPY: **SMH** (ρ=0.79, IR=0.90), **XLK** (ρ=0.92,
IR=0.51), **QQQ** (ρ=0.93, IR=0.48). All are Equity — no fixed income, commodities,
or international passed the screen. Low-ρ "diversifiers" (GLD, IEF, TLT) are
**not** valid L/S pair legs for SPY replacement — their spread volatility and
drawdowns disqualify them from this role. The screening rule: ρ ≥ 0.75 AND
spread MaxDD ≥ −30%.
</div>

<small>Full script: `key_findings/kf2_ls_pair_selection.R`</small>

---

## 6 · Long/Short Momentum Overlay

<div class="insight-box">
**What:** A dollar-neutral L/S portfolio in *relative-return space* (each position
is Long ETF / Short SPY). Monthly rebalance, 1-quarter momentum signal.
Two universes tested: full 59-ETF and a restricted 15-ETF core (PCA-derived).
Phase 2: overlay added on top of a 50/50 SPY/AGG core portfolio.
</div>

```{r p6-ls-wealth, fig.height=6}
p2_wealth_df <- data.frame(
  date         = index(phase2_comp),
  Core_5050    = as.numeric(cumprod(1 + phase2_comp[, "Core_5050"])),
  LS_Core_15   = as.numeric(cumprod(1 + phase2_comp[, "LS_Core_15"])),
  Core_plus_LS = as.numeric(cumprod(1 + phase2_comp[, "Core_plus_LS"]))
) %>% pivot_longer(-date, names_to = "Series", values_to = "Wealth")

pal_ls <- c(Core_5050 = "#95a5a6", LS_Core_15 = "#27ae60", Core_plus_LS = "#e74c3c")

ggplot(p2_wealth_df, aes(x = date, y = Wealth, colour = Series, linewidth = Series)) +
  geom_line() +
  scale_colour_manual(values = pal_ls, name = NULL) +
  scale_linewidth_manual(values = c(Core_5050 = 1, LS_Core_15 = 1, Core_plus_LS = 2),
                         guide = "none") +
  scale_y_log10(labels = dollar_format(prefix = "$")) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
  labs(title    = "P6a: Wealth Index — Core / L/S Overlay / Combined",
       subtitle = "Log scale | $1 invested | Bold = Core + L/S overlay",
       x = NULL, y = "Portfolio Value (log scale)")
```

```{r p6-ls-dd, fig.height=5}
dd_xts <- Drawdowns(phase2_comp)
dd_df <- data.frame(
  date         = index(dd_xts),
  Core_5050    = as.numeric(dd_xts[, "Core_5050"]),
  LS_Core_15   = as.numeric(dd_xts[, "LS_Core_15"]),
  Core_plus_LS = as.numeric(dd_xts[, "Core_plus_LS"])
) %>%
  pivot_longer(-date, names_to = "Series", values_to = "Drawdown") %>%
  mutate(Series = factor(Series, levels = c("Core_5050", "LS_Core_15", "Core_plus_LS")))

ggplot(dd_df, aes(x = date, y = Drawdown, colour = Series, linewidth = Series)) +
  geom_line() +
  geom_hline(yintercept = 0, colour = "grey40") +
  scale_colour_manual(values = pal_ls, name = NULL) +
  scale_linewidth_manual(values = c(Core_5050 = 0.8, LS_Core_15 = 0.8, Core_plus_LS = 1.6),
                         guide = "none") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
  labs(title    = "P6b: Drawdown Comparison",
       subtitle = "Core / L/S Overlay / Combined | Bold = Core + L/S",
       x = NULL, y = "Drawdown")
```

```{r p6-ls-table, results='asis'}
cat("\n**Annualised Performance Summary:**\n\n")
ann <- table.AnnualizedReturns(phase2_comp, scale = 252) %>%
  as.data.frame() %>%
  rownames_to_column("Metric") %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
knitr::kable(ann, align = "lrrr")
```

<div class="finding-box">
**Key findings:**
(1) The L/S overlay generates a Sharpe of ~1.5 on the 15-ETF core universe —
independent of market direction.
(2) Adding the dollar-neutral overlay to the 50/50 core *improves* both return
and Sharpe while keeping the drawdown profile similar.
(3) The overlay is near-uncorrelated with the core (R² < 0.05), confirming it
provides genuinely independent alpha.
</div>

<small>Full report: `Rmd/long_short_overlay_report.Rmd`</small>

---

## 7 · Exante Hypothesis — Enhancer vs Stabilizer

<div class="insight-box">
**What:** A systematic generalisation of the pair-analysis idea. For every ticker,
measure the effect of a 50/50 blend with SPY on two dimensions simultaneously:
return (delta_ret) and drawdown (delta_dd). Every ticker falls into one of four
quadrants: Dominant / Enhancer / Stabilizer / Detractor.
</div>

```{r p7-quadrant, fig.height=9}
print(plot_exante_quadrant(exante_tbl,
                            horizon_sel = "full",
                            label_n     = 8))
```

```{r p7-corr-dd, fig.height=7}
print(plot_corr_dd(exante_tbl,
                   horizon_sel = "full",
                   show_loess  = TRUE))
```

```{r p7-summary, results='asis'}
tbl_full <- exante_tbl %>% filter(horizon == "full")
n_tot    <- nrow(tbl_full)
role_sum <- tbl_full %>%
  count(role) %>%
  arrange(match(role, c("Dominant","Enhancer","Stabilizer","Detractor"))) %>%
  mutate(pct = scales::percent(n / n_tot, accuracy = 1))

knitr::kable(role_sum,
             caption = "Exante role distribution — full sample vs SPY",
             col.names = c("Role", "Count", "% of Universe"),
             align = "lrr")
```

<div class="finding-box">
**Key findings:**
(1) Most tickers are **Stabilizers or Detractors** — they reduce return without a
commensurate DD benefit, or worsen both dimensions.
(2) **Dominant** tickers (better return *and* less DD) are rare but identifiable exante.
(3) **The anti-diversification finding is confirmed:** the loess curve on the
corr vs DD plot is *flat to negative* for correlation ∈ [0, 0.7]. Tickers
with near-zero correlation to SPY produce *worse* blended DrawDown than
high-correlation tickers, because their independent DD events are not
offset by SPY's drawdown timing.
The driver is the ticker's **own standalone MaxDD**, not its correlation.
</div>

<small>Full reports: `Rmd/exante_hypothesis_report.Rmd` (parameterised — runs for any BMK)</small>

---

## 8 · SAA Portfolio — Structure & Institutional Mapping

<div class="insight-box">
**What:** A concrete 6-ticker SAA (URTH / SPY / QQQ / XLK / SMH / AGG) is defined,
visualised by role, and then mapped to an institutional allocation framework with
three risk profiles (INC / BAL / GRO). Each institutional line item is matched to
an ETF from the Sovereign Universe. The mapping reveals which institutional tilts
are already implementable and which require new tickers.
</div>

### P8a — SAA Allocation by Role

```{r p8-ggbar, fig.height=3.5}
print(plot_saa_ggbar(xts_ret = xts_ret))
```

```{r p8-treemap, fig.height=5}
print(plot_saa_treemap(xts_ret = xts_ret))
```

*Interactive plotly sunburst available as a standalone script: `key_plots/saa_sunburst_plotly.R`*

---

### P8b — Efficient Frontier — Sovereign Universe

<div class="insight-box">
**What:** Mean-variance efficient frontier across the full 96-ticker Sovereign Universe
(long-only, fully-invested). Three reference portfolios: **GMV** (lowest possible
volatility), **Max Sharpe** (best return per unit of risk), and the current **SAA**.
SAA tickers shown as diamonds (◆).
</div>

```{r p8-frontier, fig.height=8}
print(p_ef)
```

```{r p8-frontier-tables, results='asis'}
cat("**GMV Portfolio** (global minimum variance — top holdings)\n\n")
knitr::kable(ef_gmv_tbl, align = "llr")
cat("\n\n**Max Sharpe Portfolio** (tangency — top holdings)\n\n")
knitr::kable(ef_tan_tbl, align = "llr")
```

<div class="finding-box">
**Key findings — Efficient Frontier:**
(1) **SAA Sharpe = 0.44** vs frontier maximum of **1.46** — the SAA sits inside the
frontier by design: a real allocation must respect diversification, capacity, and
tracking-error constraints that the unconstrained optimizer ignores.
(2) **GMV is T-bill dominated** (91% SGOV) — mathematically correct but not useful
as an allocation; reflects that the lowest-vol asset in the universe is cash.
(3) **Max Sharpe concentrates** in UUP + GLD + DXJ + SMH — a low-vol, high-IR
cluster. Note that historical Sharpe is backward-looking; concentration risk is high.
(4) **The frontier confirms** the anti-diversification finding: the highest-Sharpe
candidates are high-ρ equity + uncorrelated low-vol assets (GLD), not a broad mix
of zero-correlation diversifiers.
</div>

<small>Full script: `scripts_saa_taa/efficient_frontier.R`</small>

---

### P8c — Institutional SAA → ETF Vector

The table below maps the institutional multi-profile allocation (INC / BAL / GRO)
to the closest available ETF, distinguishing tickers already in the Sovereign Universe
from those requiring a new download.

```{r p8-mapping-table, results='asis'}
inst_map <- tribble(
  ~`Institutional Line`,          ~Role,  ~ETF,    ~Ticker, ~Status,
  # ── Fixed Income ─────────────────────────────────────────────────────────────
  "Cash & Short-Term Bonds",      "ECT",  "iShares 0-3M Treasury",           "SGOV",  "In universe",
  "Bd Corp USD 1-10y",            "COR",  "iShares iBoxx IG Corp",            "LQD",   "In universe",
  "Bd Govt US 1-10y",             "HDG",  "iShares 7-10Y Treasury",           "IEF",   "In universe",
  "Bd Govt US 1-3y",              "HDG",  "iShares 1-3Y Treasury",            "SHY",   "In universe",
  "Bd High Yield Global",         "YEN",  "iShares iBoxx HY Corp",            "HYG",   "In universe",
  "Bd EM Corp HC",                "YEN",  "iShares JP Morgan USD EM Bond",    "EMB",   "In universe",
  "Bd EM Local",                  "YEN",  "VanEck Local Currency EM Bond",    "EMLC",  "In universe",
  "Bd Unconstrained",             "YEN",  "Vanguard Total Intl Bond Hedged",  "BNDX",  "In universe (partial)",
  "Bd COCO hedged",               "YEN",  "—  (no US-listed ETF)",            "—",     "UCITS only",
  # ── Equity ───────────────────────────────────────────────────────────────────
  "Eq World DC",                  "COR",  "iShares MSCI World",               "URTH",  "In universe",
  "Eq Value",                     "COR",  "Vanguard Value",                   "VTV",   "In universe",
  "Eq US",                        "HOB",  "SPDR S&P 500",                     "SPY",   "In universe",
  "Eq EMU",                       "HOB",  "SPDR Euro Stoxx 50",               "FEZ",   "In universe",
  "Eq World ex US",               "HOB",  "iShares Core MSCI EAFE",           "IEFA",  "In universe",
  "Eq Switzerland",               "ECT",  "iShares MSCI Switzerland",         "EWL",   "In universe",
  "Eq China",                     "ECT",  "iShares China Large-Cap",          "FXI",   "In universe",
  "Eq India",                     "ECT",  "iShares MSCI India",               "INDA",  "In universe",
  "Eq Next Generation",           "ECT",  "Global X AI & Big Data",           "AIQ",   "In universe",
  "Eq AC Asia ex Japan",          "EM",   "iShares MSCI AC Asia ex Japan",    "AAXJ",  "Needs download",
  "Eq Latam",                     "HOB",  "iShares Latin America 40",         "ILF",   "Needs download",
  "S&P 500 Hedge",                "ECT",  "ProShares Short S&P 500",          "SH",    "Needs download",
  # ── Alternatives ─────────────────────────────────────────────────────────────
  "Gold",                         "OTH",  "SPDR Gold Shares",                 "GLD",   "In universe",
  "AI Hedge Funds",               "OTH",  "iM DBi Managed Futures (proxy)",   "DBMF",  "Needs download"
)

knitr::kable(
  inst_map,
  caption = "Institutional SAA line items mapped to Sovereign Universe ETFs",
  align   = "llllc"
)
```

### P8d — TAA Conviction Summary (INC / BAL / GRO)

```{r p8-taa-tilts, fig.height=5}
taa_tilts <- tribble(
  ~`Asset Class`,     ~INC,   ~BAL,   ~GRO,
  "Fixed Income",     -3.1,   -4.9,   -4.6,
  "Equity",           +3.1,   +3.5,   +2.9,
  "Alternatives",     +1.8,   +1.5,   +1.2
) %>%
  pivot_longer(-`Asset Class`, names_to = "Profile", values_to = "Tilt") %>%
  mutate(
    Profile    = factor(Profile, levels = c("INC", "BAL", "GRO")),
    fill_color = if_else(Tilt >= 0, "#27ae60", "#e74c3c")
  )

ggplot(taa_tilts, aes(x = Profile, y = Tilt, fill = fill_color)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6) +
  geom_text(aes(label = paste0(if_else(Tilt > 0, "+", ""), Tilt, "%"),
                vjust = if_else(Tilt >= 0, -0.4, 1.3)),
            size = 3.5, fontface = "bold") +
  scale_fill_identity() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  facet_wrap(~`Asset Class`, nrow = 1) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(face = "bold")) +
  labs(title    = "TAA vs SAA Active Tilts by Risk Profile",
       subtitle = "SAA = strategic weight | TAA = tactical deviation (Δ pp)")
```

```{r p8-taa-top-calls, results='asis'}
top_calls <- tribble(
  ~Direction, ~`Institutional Line`,       ~`Max Δ (GRO)`, ~Interpretation,
  "SELL",     "Bd Govt US 1-10y (IEF)",    "−10.0 pp",     "Completely eliminated — no duration exposure",
  "SELL",     "Eq US (SPY)",               "−11.1 pp",     "Largest single equity reduction",
  "SELL",     "Eq World Core (URTH)",      "−8.2 pp",      "Trimming broad global equity",
  "SELL",     "AI Hedge Funds",            "−2.0 pp",      "Fully removed across all profiles",
  "BUY",      "Gold (GLD)",               "+3.3 pp",       "Doubled from SAA 3.5% → ~7%",
  "BUY",      "Eq Value (VTV)",            "+9.7 pp",      "Largest single tactical add",
  "BUY",      "Eq EMU (FEZ)",             "+2.8 pp",       "Zero in SAA → added in all profiles",
  "BUY",      "Bd Corp USD (LQD)",         "+3.8 pp",      "Credit replacing duration in FI"
)

knitr::kable(top_calls,
             caption = "Top 8 TAA conviction calls — ranked by GRO profile Δ",
             align   = "lllll")
```

```{r p8-fx-tilts, fig.height=4.5}
fx_tilts <- tribble(
  ~CCY,   ~INC,   ~BAL,   ~GRO,
  "USD",  -10.6,  -13.4,  -15.9,
  "EUR",   +1.6,   +2.5,   +3.3,
  "CHF",   +1.9,   +3.0,   +3.9,
  "XAU",   +3.8,   +3.6,   +3.3,
  "CNY",   +0.9,   +1.4,   +1.9
) %>%
  pivot_longer(-CCY, names_to = "Profile", values_to = "Tilt") %>%
  mutate(
    Profile    = factor(Profile, levels = c("INC", "BAL", "GRO")),
    CCY        = factor(CCY, levels = c("USD", "EUR", "CHF", "XAU", "CNY")),
    fill_color = if_else(Tilt >= 0, "#27ae60", "#e74c3c")
  )

ggplot(fx_tilts, aes(x = CCY, y = Tilt, fill = fill_color)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6) +
  geom_text(aes(label = paste0(if_else(Tilt > 0, "+", ""), Tilt, "%"),
                vjust = if_else(Tilt >= 0, -0.4, 1.3)),
            size = 3.2, fontface = "bold") +
  scale_fill_identity() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  facet_wrap(~Profile, nrow = 1) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(face = "bold")) +
  labs(title    = "FX Exposure Tilts — TAA vs SAA",
       subtitle = "USD derisking is the dominant structural theme across all profiles")
```

<div class="finding-box">
**Key findings:**
(1) **17 of 21 institutional line items** are already covered by the Sovereign Universe —
only AAXJ, ILF, SH, and DBMF require new downloads.
(2) **USD derisking** is the single largest structural TAA call: −10.6 pp (INC) to −15.9 pp (GRO),
offset by EUR, CHF, and Gold.
(3) **Govt bonds completely eliminated** in TAA — replaced by IG Corp and credit for yield
without adding duration.
(4) **Gold doubled** (3.5% → ~7%) — highest-conviction TAA call in the sheet, aligns with
the anti-diversification framework: GLD has low correlation to SPY *and* a shallow
standalone DD — making it a genuine Dominant ticker in the Exante classification.
(5) **Value equity dominates new equity adds** — the tactical rotation away from growth/broad
towards value and non-US is consistent with a USD-bear, rate-normalisation view.
</div>

<small>Scripts: `scripts_saa_taa/saa_tree.R` | Interactive: `plot_saa_sunburst()` / `plot_saa_treemap()`</small>

---

## 9 · Meta-Asset Engine — Hierarchical Sector Combinations

<div class="insight-box">
**What:** A new idea: combine m=2 sector ETFs into a synthetic meta-asset whose
regime profile is more balanced than either component alone. Scoring uses regime
complementarity (negative regime-alpha correlation between components = good),
blend Sharpe, MaxDD improvement vs SPY, and regime-alpha smoothness.
Universe: 15 sector ETFs → 105 pairs scored at 9 weight grids each.
</div>

### P9a — Pair Scoring Heatmap

```{r p9-heatmap, fig.height=7}
print(p_ma_heatmap)
```

<div class="finding-box">
**Key finding:** The highest-scoring pairs consistently combine a **growth/cyclical
sector** (XLK, XLY) with a **defensive sector** (XLU, XLP, ITA, XLE). The cell
label shows the optimal blend weight — growth sectors typically take 50–70%.
</div>

---

### P9b — Top 20 Pairs Ranked

```{r p9-ranked, fig.height=8}
print(p_ma_ranked)
```

---

### P9c — Regime Alpha Profile: Top 6 Meta-Assets

```{r p9-regime-profile, fig.height=14}
print(p_ma_regime)
```

<div class="finding-box">
**Key finding:** The blend (dark bars) shows a smoother alpha profile across
regimes than either individual component (light bars). The best pairs achieve
positive or near-zero alpha in Fall while retaining positive alpha in Consolidation
— the ideal regime profile for a building block.
</div>

---

### P9d — Risk-Return: Individual vs Meta-Asset Blends

```{r p9-scatter, fig.height=6}
print(p_ma_scatter)
```

<div class="finding-box">
**Key findings — Meta-Asset Engine:**
(1) **XLK + ITA** (50/50) is the top-ranked meta-asset: Tech ALPHA in Consolidation
is offset by Defense's relative stability in Fall. Composite score 0.96.
(2) **XLK + XLU** (60/40) and **XLK + XLP** (60/40) follow — classic
growth + defensive sector combinations.
(3) **Regime complementarity** (ρ close to −1.0) is the strongest differentiator
among high-Sharpe pairs — it separates pairs that merely blend return from pairs
that genuinely smooth regime exposure.
(4) **Next step:** feed meta-assets as building blocks into the efficient frontier
and compare to individual-ticker frontier.
</div>

<small>Full script: `scripts_saa_structured/meta_asset_engine.R`</small>

---

## 10 · ETF Range Monitor — Current Price in 52-Week Context

<div class="insight-box">
**What:** For every key ticker in the Sovereign Universe, show where the current
price sits within its 52-week range (normalised 0–100%). All tickers share the
same x-axis scale — making cross-asset comparisons direct. Grouped by asset class.
Grey bar = 52-wk range · Blue bar = today's range · Dot = last price.
</div>

### P10a — Equity: US Broad & International

```{r p10-eq-us-range, fig.height=5.5}
print(range_plots[["Equity — US Broad"]])
```

```{r p10-eq-us-xy, fig.height=6}
print(xy_plots[["Equity — US Broad"]])
```

---

### P10b — Equity: Sectors

```{r p10-eq-sector-range, fig.height=6}
print(range_plots[["Equity — Sector"]])
```

```{r p10-eq-sector-xy, fig.height=6}
print(xy_plots[["Equity — Sector"]])
```

---

### P10c — Equity: Thematic

```{r p10-eq-thematic-range, fig.height=5.5}
print(range_plots[["Equity — Thematic"]])
```

```{r p10-eq-thematic-xy, fig.height=6}
print(xy_plots[["Equity — Thematic"]])
```

---

### P10d — Fixed Income

```{r p10-fi-range, fig.height=5.5}
print(range_plots[["Fixed Income"]])
```

```{r p10-fi-xy, fig.height=6}
print(xy_plots[["Fixed Income"]])
```

---

### P10e — Commodity & Real Assets

```{r p10-commodity-range, fig.height=5.5}
print(range_plots[["Commodity & Real"]])
```

```{r p10-commodity-xy, fig.height=6}
print(xy_plots[["Commodity & Real"]])
```

---

### P10f — Full Universe Position Map

```{r p10-xy-full, fig.height=8}
print(p_xy_rng)
```

<div class="finding-box">
**How to read:** Tickers near the **top** are near their 52-week high. Tickers near
the **bottom** are far from their high. The dotted crosshairs mark the group averages.
Green dot = up today · Red dot = down today.
</div>

<small>Full script: `utility/etf_ranges.R` | Refresh: `etf_ranges(symbols)`</small>

---

## 11 · Calendar Year Performance — Sector Focus

```{r build-calendar-data, include=FALSE}
CAL_TICKERS <- c("XLK", "XLI", "XLV", "XLP", "XLF", "QQQ", "SMH")
xts_rel      <- readRDS(here("02_data_processed/xts_rel.rds"))
```

### P11a — Single Ticker: All Calendar Years

```{r p11a-single, fig.height=6, fig.width=11}
plot_calendar_perf(xts_ret[, "SPY"])
```

### P11b — Multi-Ticker: Sector Small Multiples

```{r p11b-multi, fig.height=14, fig.width=11}
plot_multi_calendar_perf(xts_ret[, CAL_TICKERS], ncol = 2)
```

### P11c — Absolute vs Relative: Each Ticker

```{r p11c-rel, fig.height=6, fig.width=13, results='asis'}
for (tk in CAL_TICKERS) {
  cat(sprintf("\n\n#### %s\n\n", tk))
  print(plot_rel_calendar_perf(tk, xts_ret, xts_rel))
  cat("\n\n")
}
```

<div class="finding-box">
**How to read:** P11b shows all years per ticker on a free Y scale — useful for spotting which years were outliers. P11c pairs the absolute return (left) with the relative return vs SPY (right) on the **same Y axis**, so the gap between panels shows how much of the absolute return was driven by market beta vs genuine alpha.
</div>

<small>Full script: `utility/plot_calendar_perf.R`</small>

---

## 12 · Calendar Year Analog Projection — Trend vs Reversal

```{r build-analog-data, include=FALSE}
ANALOG_TICKERS <- c("SPY", "QQQ", "GLD", "TLT", "XLK", "SMH", "XLF", "XLI")
```

<div class="finding-box">
**Method:** For each ticker, find historical calendar years whose YTD return
(measured at today's trading-day-of-year) most closely resembles the current
year. The shaded cone shows where those analog years went from today to
year-end. P(Trend) = fraction of analog years that finished in the same
direction as the current YTD. See `Rmd/calendar_analog_report.Rmd` for full
academic background and methodology.
</div>

### P12a — Transition Matrices (Regime-Conditioned)

```{r p12a-transitions, results='asis'}
for (tk in ANALOG_TICKERS) {
  cat(sprintf("\n\n#### %s\n\n", tk))
  cat("```\n")
  compute_transitions(xts_ret[, tk], rt)
  cat("```\n\n")
}
```

### P12b — Analog Projection Charts

```{r p12b-analog, fig.height=7, results='asis'}
for (tk in ANALOG_TICKERS) {
  cat(sprintf("\n\n#### %s\n\n", tk))
  print(plot_calendar_analog(xts_ret[, tk], rt))
  cat("\n\n")
}
```

<div class="finding-box">
**How to read the cone:** Grey = all historical years. Blue = analog years
(similar YTD at today's point). Inner cone = 25–75th pct of analog paths
forward. Outer cone = 10–90th pct. Dashed = median analog path. Red = current
year. The annotation box shows P(Trend), P(Reversal), the specific analog
years used, and the current regime.

**Honest caveat:** With ~15 years of data, a quoted "70% P(Trend)" carries
a ±25% confidence interval. Use as directional context alongside the regime
signal — not as a precise probability.
</div>

<small>Full script: `utility/plot_calendar_analog.R` | Full report: `Rmd/calendar_analog_report.Rmd`</small>

---

## Synthesis — The Connecting Thread

```{r p-synthesis-table, results='asis'}
synth <- tribble(
  ~Analysis,                 ~`Core Question`,                             ~`Empirical Finding`,
  "SPY Regime Engine",       "What phases does SPY itself go through?",    "3 states with radically different return/vol profiles",
  "200DMA State Machine",    "Can we signal SPY entry/exit tactically?",   "Hysteresis (+4/-2%) cuts false signals vs simple crossover",
  "Alpha by Regime",         "Which ETFs protect in Fall & participate in Consolidation?", "Safe havens LAG (not lose) in Recovery — abs return stays positive; alpha asymmetry favours holding GLD/IEF across the cycle",
  "Safe Haven Signal",       "Can we predict regime transitions?",         "HYG correlation spike is a leading warning; LASSO AUC ~0.72",
  "Pair Analysis",           "What is the DD cost of adding any ticker?",  "High-corr tickers: theory = empirical; zero-corr: empirical DD worse",
  "L/S Overlay",             "Can we extract pure alpha from relative returns?", "Sharpe ~1.5 on core 15; near-zero R² vs core portfolio",
  "Exante Hypothesis",       "Can we classify every ticker as Enhancer or Stabilizer?", "Confirmed: corr≈0 does NOT reduce MaxDD; standalone DD is the driver",
  "SAA & Inst. Mapping",     "How does the SAA translate to ETFs and institutional profiles?", "17/21 inst. lines covered; USD derisking + Gold doubling are top TAA calls",
  "Meta-Asset Engine",       "Can sector ETF pairs act as regime-balanced building blocks?",   "XLK+ITA, XLK+XLU, XLK+XLP top-ranked; regime complementarity ρ ≈ −1 is key differentiator",
  "DAA Framework",           "Does regime-based TAA or trend filtering add value over passive?", "Regime rotation fails (lag + SAA drag); 200DMA filter reduces DD ticker-by-ticker with minimal return cost; best DD stabilizers ranked by insurance ratio"
)
knitr::kable(synth, align = "lll")
```

<div class="finding-box">
**The unifying anti-diversification insight** (confirmed across analyses 5, 6 and 7):

For a long-only investor blending SPY with any ticker, **the MaxDD of the blended
portfolio is NOT minimised by minimising correlation**. It is minimised by
minimising the ticker's own standalone drawdown severity. A ticker with
correlation 0.8 to SPY but a shallow 15% standalone DD produces a better blended
MaxDD than a ticker with correlation 0.1 but a 40% standalone DD.

This overturns the standard diversification narrative for the risk metric that
matters most to investors: drawdown — not variance.
</div>

---

## 13 · DAA Framework — 200DMA Trend Filter & DD Stabilizers

```{r p13-setup, include=FALSE}
source(here("scripts_daa_saa_taa/01_saa_baseline.R"))

daa_class_path <- here("02_data_processed/daa_class_tbl.rds")
taa_hist_path  <- here("02_data_processed/taa_weights_history.rds")
trend_sig_path <- here("02_data_processed/trend_signals.rds")
daa_ret_path   <- here("02_data_processed/daa_returns.rds")
stab_path      <- here("02_data_processed/dd_stabilizer_screen.rds")
frontier_path  <- here("02_data_processed/dd_frontier_data.rds")

source(here("key_plots/chart_stacked_weight_timeline.R"))
source(here("key_plots/chart_ma200_signal_panel.R"))
source(here("key_plots/chart_multi_wealth_endlabel.R"))
source(here("key_plots/chart_dd_frontier.R"))
# Note: regime_aware_alpha_tape.R excluded — uses legacy base_path variable
# Note: yahoo_finance_range_plot.R excluded — superseded by utility/etf_ranges.R (Section 10)
```

> **Premise**: SPY/IEF 60/40 delivers adequate return over the long run.
> The DAA framework pursues two separate goals:
> (1) reduce drawdown via a 200DMA trend filter; (2) identify DD stabilizer
> overlays that shrink MaxDD with minimum return sacrifice.

### 13a · 200DMA Trend Signal — When Is Each Ticker Above Its MA?

```{r p13-signal-heatmap, fig.width=13, fig.height=5, fig.cap="Green = trend ON (hold at SAA weight)  |  Red = trend OFF (weight parked in SGOV)"}
if (file.exists(trend_sig_path)) {
  trend_signals <- readRDS(trend_sig_path)
  source(here("scripts_daa_saa_taa/03_taa_rules.R"))   # loads plot_signal_heatmap()
  print(plot_signal_heatmap(trend_signals, SAA_60_40, rt))
}
```

### 13b · Portfolio Composition — Weight Stack Over Time

```{r p13-composition, fig.width=13, fig.height=6, fig.cap="SGOV expands when tickers fall below 200DMA  |  chart_stacked_weight_timeline"}
if (file.exists(taa_hist_path)) {
  taa_weights_history <- readRDS(taa_hist_path)
  print(
    plot_stacked_weight_timeline(
      taa_weights_history, SAA_60_40, rt,
      title    = "Portfolio Composition — 200DMA Trend Filter",
      subtitle = "SGOV expands when tickers fall below 200DMA  |  Pink = Fall regime"
    )
  )
}
```

### 13c · Equity Sleeve — Per-Ticker Price vs 200DMA

```{r p13-eq-panel, fig.width=13, fig.height=14, fig.cap="Red shading = trend OFF — weight parked in SGOV  |  chart_ma200_signal_panel"}
if (file.exists(trend_sig_path)) {
  p_panels <- plot_ma200_signal_panel(trend_signals, SAA_60_40)
  if (!is.null(p_panels$eq)) print(p_panels$eq)
}
```

### 13d · Backtest — DAA vs SAA vs Passive 60/40

```{r p13-backtest, fig.width=13, fig.height=6, fig.cap="chart_multi_wealth_endlabel — direct end-of-line labels, ann return appended"}
if (file.exists(daa_ret_path) && exists("saa_returns") && exists("benchmark_returns")) {
  daa_returns <- readRDS(daa_ret_path)
  ret_list <- list(
    DAA           = daa_returns,
    SAA_60_40     = saa_returns,
    Passive_60_40 = benchmark_returns
  )
  print(
    plot_multi_wealth_endlabel(
      ret_list, rt,
      colours = c("#7c3aed", "#3b82f6", "#9ca3af"),
      labels  = c("DAA — 200DMA filter", "SAA 60/40 (JB universe)", "Passive SPY/IEF")
    )
  )
}
```

### 13e · DD Stabilizer Quadrant

```{r p13-stabilizer-quad, fig.width=13, fig.height=8, fig.cap="Insurance Quadrant — x=return drag, y=MaxDD reduction, bubble=Fall correlation"}
if (file.exists(stab_path)) {
  stabilizer_screen <- readRDS(stab_path)
  source(here("scripts_daa_saa_taa/05_dd_stabilizer.R"))
  print(plot_insurance_quad(stabilizer_screen))
}
```

```{r p13-dd-frontier, fig.width=13, fig.height=7, fig.cap="DD Frontier — trade-off curve per instrument as overlay weight increases 0→25%"}
if (file.exists(frontier_path)) {
  dd_frontier_data <- readRDS(frontier_path)
  print(plot_dd_frontier(dd_frontier_data))
}
```

<div class="finding-box">
**DAA Key Finding**: The 200DMA trend filter reduces drawdown at the cost of
minor return lag in fast recoveries (classic momentum overshoot). The signal
is ticker-level (only the broken ticker parks), not portfolio-level, which avoids
the regime-detection lag that made the regime-rotation TAA ineffective.
The DD Stabilizer screen identifies instruments that improve the passive 60/40
MaxDD with the lowest return sacrifice — ranked by insurance ratio = DD reduction / return drag.
</div>

---

## Section 14 — DD Overlay Optimizer: Minimum-Cost DrawDown Reduction

> **Question answered:** Given we accept SPY/IEF 60/40 returns, what is the
> cheapest combination of instruments that reduces MaxDD from ~23% to ~15%?

```{r p13-overlay-setup, include=FALSE}
overlay_grid_path    <- here("02_data_processed/overlay_grid.rds")
pareto_path          <- here("02_data_processed/pareto_frontier.rds")
solutions_path       <- here("02_data_processed/overlay_target_solutions.rds")

if (!all(file.exists(overlay_grid_path, pareto_path, solutions_path))) {
  message("07_dd_overlay_optimizer.R not yet run — skipping Section 14 plots.")
  overlay_ready <- FALSE
} else {
  source(here("scripts_daa_saa_taa/07_dd_overlay_optimizer.R"))
  overlay_ready <- TRUE
}
```

### 14a · Overlay Efficiency Frontier

```{r p13-overlay-frontier, fig.width=13, fig.height=8, fig.cap="Efficiency frontier — all single/pair/triple combinations; purple = Pareto-optimal path; red stars = cheapest combo per MaxDD target"}
if (isTRUE(overlay_ready)) {
  print(plot_overlay_frontier(overlay_grid, pareto_frontier, target_solutions))
}
```

### 14b · Optimal Overlay Composition per DD Target

```{r p13-overlay-bars, fig.width=13, fig.height=6, fig.cap="Recommended instrument weights for each MaxDD target — funded from SPY/IEF proportionally"}
if (isTRUE(overlay_ready)) {
  print(plot_optimal_bars(target_solutions))
}
```

### 14c · Passive vs Optimal ≤15% Overlay: Wealth & Drawdown

```{r p13-overlay-comparison, fig.width=13, fig.height=10, fig.cap="Cumulative wealth (top) and drawdown (bottom): passive 60/40 vs cheapest overlay achieving ≤15% MaxDD"}
if (isTRUE(overlay_ready)) {
  print(plot_overlay_comparison(target_solutions, passive_ret, xts_ret, rt))
}
```

### 14d · Solution Summary Table

```{r p13-overlay-solution-table, results='asis'}
if (isTRUE(overlay_ready)) {
  target_solutions %>%
    mutate(
      maxdd_pct  = sprintf("%.1f%%",   abs(maxdd)      * 100),
      dd_saved   = sprintf("%+.1f pp", dd_reduction    * 100),
      ann_ret    = sprintf("%+.1f%%",  ann_ret         * 100),
      drag       = sprintf("%+.1f pp", return_drag     * 100),
      total_w    = sprintf("%.0f%%",   total_weight    * 100)
    ) %>%
    select(`DD Target` = target_label,
           `Instruments` = instruments,
           `Weights` = weights_str,
           `Total overlay` = total_w,
           `MaxDD achieved` = maxdd_pct,
           `DD saved` = dd_saved,
           `Ann return` = ann_ret,
           `Return drag` = drag) %>%
    knitr::kable(align = "llcccrrr") %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
}
```

<div class="finding-box">
**Overlay Optimizer Key Finding**: The minimum-cost overlay to reduce MaxDD from
~23% to ~15% combines a small number of uncorrelated instruments (typically 1–2)
funded by proportional reduction of the passive SPY/IEF weights. The Pareto
frontier quantifies the exact return-drag cost of each DD protection level,
allowing a deliberate cost-benefit choice rather than an ad-hoc allocation.
</div>

---

## 15 · Valuation — S&P 500 P/E Bands & Shiller CAPE

> Current market multiple relative to historical norms — uses S&P Global
> operating EPS (Bloomberg-style) for P/E price bands and Shiller 10-year
> inflation-adjusted GAAP EPS (P/E10) for the CAPE chart.

```{r p15-setup, include=FALSE}
source(here("key_plots/chart_pe_bands.R"))
```

### 15a · P/E Price Bands (Operating EPS)

```{r p15-pe-bands, fig.width=13, fig.height=7, fig.cap="S&P 500 P/E price bands — 5 coloured bands centred on rolling mean P/E, current price overlaid"}
plot_pe_bands()
```

### 15b · Shiller CAPE (P/E10)

```{r p15-shiller-cape, fig.width=13, fig.height=6, fig.cap="Shiller CAPE (P/E10) since 1950 — orange = CAPE, purple dashed = long-run avg, red/blue bands = ±1σ and ±2σ"}
plot_shiller_cape()
```

<div class="finding-box">
**Valuation Context**: P/E price bands show where current price sits relative to
historically normalised multiples. CAPE smooths cyclical EPS swings over 10 years —
useful for identifying secular overvaluation that single-year P/E can miss at earnings peaks.
</div>

---

## 16 · Daily Return Heatmaps — SPY · GLD · URTH

> Red = down day · Blue = up day · White = flat. Each row = one day-of-week across all weeks of the year.

```{r p16-setup, include=FALSE}
library(lattice); library(chron); library(grid)
if (!requireNamespace("chron", quietly = TRUE)) install.packages("chron")

# source the standalone script (auto-run guard suppressed by knitr.in.progress)
source(here::here("key_plots/calendar_heatmap.R"))
```

```{r p16-heatmap-spy, fig.width=13, fig.height=9, fig.cap="SPY daily log returns — last 5 years"}
plot_calendar_heatmap("SPY", n_yr = 5)
```

```{r p16-heatmap-gld, fig.width=13, fig.height=9, fig.cap="GLD daily log returns — last 5 years"}
plot_calendar_heatmap("GLD", n_yr = 5)
```

```{r p16-heatmap-urth, fig.width=13, fig.height=9, fig.cap="URTH daily log returns — last 5 years"}
plot_calendar_heatmap("URTH", n_yr = 5)
```

---

## 17 · Calendar Year Analog Projection — SPY · GLD · URTH

> Grey = all historical years · Blue = analog years (similar YTD at today's trading day) · Shaded cone = 10–90th pct of analog paths forward · Red = current year · Dashed = median analog path.

```{r p17-setup, include=FALSE}
source(here::here("utility/plot_calendar_analog.R"))
source(here::here("scripts_spy_dd_regime/spy_dd_regime.R"))

xts_ret_analog <- readRDS(here::here("02_data_processed/xts_ret_returns.rds"))
rt_analog <- build_regime_table(xts_ret_analog[, "SPY"], t_fall = 0.10)
```

```{r p17-analog-spy, fig.width=13, fig.height=7, fig.cap="SPY — Calendar Year Analog Projection"}
plot_calendar_analog(xts_ret_analog[, "SPY"], rt = rt_analog)
```

```{r p17-analog-gld, fig.width=13, fig.height=7, fig.cap="GLD — Calendar Year Analog Projection"}
plot_calendar_analog(xts_ret_analog[, "GLD"], rt = rt_analog)
```

```{r p17-analog-urth, fig.width=13, fig.height=7, fig.cap="URTH — Calendar Year Analog Projection"}
plot_calendar_analog(xts_ret_analog[, "URTH"], rt = rt_analog)
```

---

## Navigation

```{r p-navigation-table, results='asis'}
nav <- tibble::tribble(
  ~Report,                        ~Path,                                          ~`What it contains`,
  "SPY Regime & State Machine",   "Rmd/spy_regime_statemachine_report.Rmd",       "Full 14-section regime engine + 200DMA SM",
  "Safe Haven Regime Signal",     "Rmd/safe_haven_regime_report.Rmd",             "3-method regime prediction analysis",
  "Pair Analysis",                "Rmd/pair_analysis_portfolio_construction.Rmd", "Cost-benefit vectors, Magdon-MDD theory",
  "Universe Pair Analysis v2",    "Rmd/universe_pair_analysis_v2.Rmd",            "Category A (vs SPY) & B (vs AGG) shortlists",
  "L/S Overlay",                  "Rmd/long_short_overlay_report.Rmd",            "Full L/S methodology + Phase 2",
  "Exante Hypothesis",            "Rmd/exante_hypothesis_report.Rmd",             "Parameterised by benchmark — render_exante_report()",
  "SAA Portfolio & Inst. Mapping","scripts_saa_taa/saa_tree.R",                   "SAA tree (ggbar + treemap) with YTD; inst. ETF vector",
  "DAA — SAA Baseline",           "scripts_daa_saa_taa/01_saa_baseline.R",         "JB DAA 13-ticker 60/40 + passive SPY/IEF benchmark",
  "DAA — Classifier",             "scripts_daa_saa_taa/02_daa_classifier.R",       "EQ-Enhancer / EQ-Stabilizer / FI-Enhancer taxonomy",
  "DAA — 200DMA Trend Filter",    "scripts_daa_saa_taa/03_taa_rules.R",            "Signal heatmap + composition chart + per-ticker MA panels",
  "DAA — Backtest",               "scripts_daa_saa_taa/04_daa_backtest.R",         "Walk-forward DAA vs SAA vs Passive; wealth + DD + regime excess",
  "DAA — DD Stabilizer Screen",   "scripts_daa_saa_taa/05_dd_stabilizer.R",        "Insurance quadrant + DD frontier for all universe tickers",
  "DAA — Overlay Optimizer",      "scripts_daa_saa_taa/07_dd_overlay_optimizer.R", "Pareto frontier + min-cost overlay for 15/17/19/21% MaxDD targets"
)

knitr::kable(nav, align = "lll",
             col.names = c("Report", "Path", "What it contains")) %>%
  kableExtra::column_spec(1, width = "4cm") %>%
  kableExtra::column_spec(2, width = "6cm") %>%
  kableExtra::column_spec(3, width = "6cm")
```
