---
title: "KF3 — When Winners Become Losers"
subtitle: "Trend Reversal in Cross-Sectional Relative Momentum"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    latex_engine: xelatex
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(scales)
library(patchwork)
library(kableExtra)
library(here)

# Source the KF3 engine — computes all objects; the knitr guard suppresses
# auto-printing so plots are fully controlled by this report.
source(here("key_findings/kf3_trend_reversal.R"))

# ── Table helper — dark-grey scaffolding applied to every table ───────────────
ktable <- function(x, caption = NULL, ...) {
  tbl <- knitr::kable(x, caption = caption, ...)
  if (knitr::is_html_output()) {
    tbl %>%
      kable_styling(
        bootstrap_options = c("striped", "hover", "condensed"),
        full_width        = TRUE,
        font_size         = 13,
        html_font         = "inherit"
      ) %>%
      row_spec(0,
               background = "#3d3d3d",
               color      = "white",
               bold       = TRUE)
  } else {
    tbl %>%
      kable_styling(
        latex_options = c("striped", "hold_position", "scale_down"),
        full_width    = TRUE,
        font_size     = 10,
        stripe_color  = "#f0f0f0"
      ) %>%
      row_spec(0,
               bold       = TRUE,
               background = "#3d3d3d",
               color      = "white")
  }
}
```

---

## 1. The Problem in Plain English

Imagine a simple rule for picking ETFs: *"buy whatever has gone up the most
over the past three months."* This is the core idea behind **momentum** — assets
that have recently outperformed tend to keep outperforming, at least in the
short term. It works surprisingly well most of the time.

But there is a well-known trap: **the strategy fails catastrophically at turning
points.** At the moment when the market stops falling and starts recovering, the
ETFs that looked like "winners" during the downturn (gold, government bonds,
defensive equities) suddenly become the worst performers — not because anything
changed about those ETFs, but because the *context* changed.

This report measures that phenomenon precisely and asks:

> **Are winner-to-loser reversals predictable from the market regime, or do they
> happen randomly?**

If reversals cluster at regime transitions, then a smart TAA framework can
*anticipate* them — rather than getting caught holding yesterday's winners.

---

## 2. Building Blocks: Three Concepts You Need

### 2.1 Relative Return — What We Actually Measure

Rather than asking "did this ETF go up?", we ask:
**"Did this ETF go up *more than SPY*?"**

This is called a **relative return**:

$$r^{\text{rel}}_{i,t} = r^{\text{ETF}}_{i,t} - r^{\text{SPY}}_t$$

A positive relative return means the ETF *outperformed* SPY that day.
A negative relative return means it *underperformed*, even if it gained in
absolute terms.

Why use relative returns? Because we want to separate **skill** (beating the
market) from **beta** (just riding the market wave). Gold going up 1% on a day
SPY goes up 2% is actually underperformance.

### 2.2 Cross-Sectional Rank — Who Is Winning Right Now?

At every weekly rebalance date, we compute the **60-day cumulative relative
return** for every ticker in the universe. Then we rank all tickers from worst
(rank 1) to best (rank N) and divide into five equal buckets called
**quintiles**:

| Quintile | Label | Who is here |
|:---:|:---:|---|
| Q5 | **Winner** | Top 20% — highest 60-day relative return |
| Q4 | Above average | Next 20% |
| Q3 | Middle | Middle 20% |
| Q2 | Below average | Next 20% |
| Q1 | **Loser** | Bottom 20% — lowest 60-day relative return |

This ranking is done **cross-sectionally** — every ticker is judged against
every other ticker at the same point in time, not against its own history.

> **Noise reduction step**: before ranking, returns are *winsorised* at the 1%
> tails. This means one extreme event (e.g. a flash crash in a small ETF) cannot
> artificially push that ticker into Q1 or Q5 and contaminate the entire ranking.

### 2.3 Persistent Cohorts and Reversal Events

A single week in Q5 is noise. A ticker that stays in Q5 for **three or more
consecutive weeks** is a **persistent winner** — the signal is real and stable.

A **reversal event** is defined as:

1. A ticker was a *persistent winner* (Q5 for ≥ 3 weeks).
2. It exited that state.
3. Within the next **8 weeks**, it entered the *persistent loser* state (Q1 for ≥ 3 weeks).

We call this a **W→L** (winner-to-loser) event. The mirror image — a persistent
loser becoming a persistent winner — is an **L→W** event.

```{r params-table, echo=FALSE}
tibble(
  Parameter       = c("Rebalance cadence", "Formation window",
                      "Persistence threshold", "Transition window"),
  Symbol          = c("REBAL_FREQ", "LOOKBACK", "PERSIST_N", "TRANSIT_WIN"),
  Value           = c("5 trading days (weekly)", "60 trading days (~3 months)",
                      "≥ 3 consecutive weeks", "≤ 8 weeks"),
  `Why this value` = c(
    "Daily changes are too noisy; monthly misses fast moves",
    "Long enough to capture a trend; short enough to react to reversals",
    "Filters out one-week flukes while still being timely",
    "Regime transitions typically play out over 6–8 weeks"
  )
) %>%
  ktable(caption = "Analysis parameters and rationale")
```

---

## 3. The Data Snapshot

```{r data-snapshot}
n_events_wl <- sum(reversals$reversal_type == "W\u2192L")
n_events_lw <- sum(reversals$reversal_type == "L\u2192W")
n_tickers   <- n_distinct(reversals$ticker)
date_range  <- range(index(xts_rel))

tibble(
  Metric  = c("Universe size", "Analysis period",
              "Rebalance dates", "W→L reversal events", "L→W reversal events",
              "Tickers involved in at least one reversal"),
  Value   = c(
    paste(length(universe), "tickers"),
    paste(format(date_range[1], "%b %Y"), "to", format(date_range[2], "%b %Y")),
    as.character(length(reb_dates)),
    as.character(n_events_wl),
    as.character(n_events_lw),
    as.character(n_tickers)
  )
) %>%
  ktable(caption = "Dataset overview")
```

---

## 4. Plot A — The Quintile Transition Matrix

### What Is This Plot?

Think of it as a **probability table of cohort stickiness**. For every ticker
at every rebalance date, we ask: *"what quintile was it in last week, and what
quintile is it in this week?"*

The matrix shows the percentage of times each transition occurred. For example,
the cell at row Q5, column Q5 answers: *"if a ticker was in Q5 last week, what
is the probability it is still in Q5 this week?"*

**How to read it:**

- **High diagonal values** → **momentum**: tickers tend to stay where they are.
  This is a healthy sign — it means the 60-day rank is stable and not just noise.
- **High off-diagonal values in the extreme corners** (Q5→Q1 or Q1→Q5) →
  **mean reversion**: tickers frequently flip from one extreme to the other.
- **High off-diagonal values near the diagonal** (Q5→Q4 or Q1→Q2) →
  **gradual drift**: tickers migrate slowly through quintiles.

```{r plot-A, fig.height=6.5, fig.cap="KF3-A: Quintile Transition Matrix. Diagonal cells (darker blue) show persistence — how often a ticker stays in the same quintile from one week to the next. Off-diagonal cells show rotation."}
print(p_A)
```

### What to Look For

> **The key number is the Q5→Q5 cell** (top-left of the matrix). A value above
> 50% confirms that momentum is *real* — winners tend to remain winners week to
> week. A value near 20% (random expectation) would mean the rank carries no
> information.

> **The Q5→Q1 and Q1→Q5 cells** (extreme off-diagonal corners) measure the
> *severity* of reversals — how often a ticker jumps directly from the top to the
> bottom quintile (or vice versa) in a single week. These are the most disruptive
> transitions for a momentum strategy.

---

## 5. Plot B — When Do Reversals Happen?

### The Regime Hypothesis

The central question of this analysis is whether reversal events are
**uniformly distributed across time**, or whether they **cluster at specific
market regimes**.

Our three regimes are defined by SPY's drawdown from its most recent peak:

| Regime | Definition | Typical character |
|:---:|---|---|
| **Fall** | SPY drawdown ≥ 10% | Risk-off: bonds and gold outperform equities |
| **Recovery** | SPY rebounding from a Fall | Risk-on: equities roar back, defensives lag |
| **Consolidation** | Everything else | Range-bound, mixed signals |

The hypothesis is:

> **W→L events should cluster in Fall and at the Fall→Recovery transition.**
> Gold (GLD) and bonds (IEF, TLT) are *persistent winners* during Fall — they
> rise while SPY falls. The moment Recovery begins, they become *laggards*
> relative to SPY. This is not a fundamental change — it is a pure regime effect.

> **L→W events should cluster in Recovery.**
> High-beta tickers (SMH, XLK, QQQ) are *persistent losers* during Fall — they
> fall harder than SPY. The moment Recovery begins, they rebound faster than SPY
> and flip to persistent winners.

```{r plot-B, fig.height=6, fig.cap="KF3-B: Reversal event count by market regime. If the hypothesis holds, W→L events (red) will dominate in Fall, and L→W events (dark blue) will dominate in Recovery."}
print(p_B)
```

### Interpreting the Result

```{r regime-table}
reversals %>%
  count(reversal_type, regime_at_exit) %>%
  mutate(
    regime_at_exit = factor(regime_at_exit,
                            levels = c("Fall", "Recovery", "Consolidation"))
  ) %>%
  arrange(reversal_type, regime_at_exit) %>%
  pivot_wider(names_from = regime_at_exit, values_from = n, values_fill = 0) %>%
  rename(`Reversal type` = reversal_type) %>%
  ktable(caption = "Reversal event count by regime and type")
```

> **Key insight:** If W→L events are over-represented in Fall compared to
> Consolidation, it means the TAA framework should *discount* trailing winners
> during Fall episodes — they are likely HEDGE tickers that will reverse once
> SPY bottoms out.

---

## 6. Plot C — Do Reversals Predict Future Returns?

### Why Forward Returns Matter

Detecting reversals in hindsight is interesting. But a trading strategy needs
to know: **after a reversal event is confirmed, does the new direction persist?**

Plot C measures the **forward relative return** of the ticker in the 20 and 63
trading days *after* the reversal `entry_date` (the first week of the new
persistent state).

- **After a W→L event**: a useful signal would show *negative* forward relative
  returns — the ticker keeps underperforming SPY.
- **After an L→W event**: a useful signal would show *positive* forward relative
  returns — the ticker keeps outperforming SPY.

If the distributions are centred on zero with no asymmetry, the reversal signal
carries no predictive power — it simply described what already happened.

```{r plot-C, fig.height=7, fig.cap="KF3-C: Forward relative return (vs SPY) in the 20 and 63 trading days after each reversal event. Diamond = mean. Box = interquartile range. A signal is useful if W→L median is negative and L→W median is positive."}
print(p_C)
```

### Signal Strength Summary

```{r forward-return-table}
fwd_returns %>%
  filter(!is.na(fwd_rel_ret)) %>%
  group_by(`Reversal type` = reversal_type, `Window` = window) %>%
  summarise(
    `N events`        = n(),
    `Median fwd ret`  = percent(median(fwd_rel_ret, na.rm = TRUE), accuracy = 0.1),
    `Mean fwd ret`    = percent(mean(fwd_rel_ret,   na.rm = TRUE), accuracy = 0.1),
    `% directionally correct` = percent(
      mean(if_else(reversal_type == "W\u2192L",
                   fwd_rel_ret < 0,
                   fwd_rel_ret > 0), na.rm = TRUE),
      accuracy = 1
    ),
    .groups = "drop"
  ) %>%
  ktable(
    caption = paste0(
      "Forward return statistics post-reversal. ",
      "'% directionally correct' = fraction of events where the post-reversal ",
      "return confirmed the reversal direction."
    )
  )
```

> **Reading the table:** A median W→L forward return of, say, −3% over 63 days
> means that on average, a winner that flipped to a loser continued to
> underperform SPY by 3% in the following quarter. A "% directionally correct"
> above 55% is economically meaningful for a momentum signal.

---

## 7. Which Tickers Reverse Most Often?

Not all tickers reverse at the same rate. The table below shows the most
frequent "reversal offenders" — tickers that repeatedly cycle between winner
and loser status. These are the tickers where a naive trailing-return TAA
screen is most likely to mislead you.

```{r top-reversals}
reversals %>%
  count(ticker, reversal_type, asset_class) %>%
  arrange(desc(n)) %>%
  head(24) %>%
  pivot_wider(names_from = reversal_type, values_from = n, values_fill = 0L) %>%
  rename(
    Ticker       = ticker,
    `Asset class` = asset_class
  ) %>%
  arrange(desc(rowSums(across(where(is.integer))))) %>%
  ktable(caption = "Top 24 tickers by total reversal event count")
```

---

## 8. Implications for the TAA Framework

### The Regime-Conditional Momentum Rule

This analysis supports a simple amendment to the standard trailing-return rank:

> **Within a Fall regime:** penalise HEDGE tickers (GLD, IEF, TLT) in the
> ranking. Their Q5 position is *mechanically inflated* by the regime — it will
> reverse the moment SPY bottoms. Prefer tickers that are winners for
> *fundamental* reasons (earnings growth, sector rotation) rather than
> regime-flight reasons.

> **At the Fall→Recovery transition:** the L→W signal is actionable. Tickers
> that have been persistent Q1 losers during Fall — especially high-beta growth
> and cyclical ETFs — are likely to be the strongest performers in the first
> 8–12 weeks of Recovery.

> **In Consolidation:** the momentum signal is most reliable. The transition
> matrix diagonal will be highest here, and reversals are least frequent.
> Standard trailing-return ranking applies without a regime overlay.

### Connecting to KF1 and KF2

| Finding | Core lesson | TAA implication |
|---|---|---|
| **KF1** (anti-diversification) | Low-ρ tickers don't reduce drawdown as expected | Don't add GLD/IEF *just* because ρ ≈ 0 |
| **KF2** (L/S pair screen) | Only high-ρ tickers are valid SPY substitutes | ρ > 0.75 is a hard prerequisite for L/S |
| **KF3** (trend reversal) | Momentum rankings reverse at regime transitions | Apply regime-conditional discount to winner scores in Fall |

Together, these three findings argue for the same conclusion from different
angles: **the naive cross-sectional momentum signal should always be
regime-adjusted before being used in a TAA allocation.**

---

## 9. Equity-Only Universe — Removing the Hedge-ETF Contamination

### The Problem with Mixing Asset Classes

The full-universe analysis (Sections 4–7) includes every ETF in the Sovereign
Universe — equities, bonds, gold, real estate, and currencies. This creates a
structural contamination problem.

Consider what happens during a **Fall** regime: SPY drops 15%. On the same days,
GLD rises 8% and IEF rises 5%. Both become **Q5 persistent winners** — but for
entirely the wrong reason. Their outperformance is not driven by any inherent
sector momentum or company earnings. It is driven purely by their **negative or
zero correlation to SPY** — a mechanical, beta-driven response to the regime.

These tickers have no **eigen-move**: they do not possess an independent price
driver. Their daily movement is essentially a *derivative* of SPY's move,
reflected across zero. When the regime flips from Fall to Recovery, their
relative return mechanically reverses — not because anything changed about gold
or bonds, but because the SPY beta flip reversed.

> **The consequence:** in the full universe, most W→L events involving GLD, IEF,
> and TLT are *not real momentum reversals*. They are regime beta flips dressed
> up as reversal signals. Mixing them with genuine equity sector momentum
> produces a noisy, misleading rank.

### The Equity-Only Solution

By restricting the cross-sectional rank to **Equity ETFs only**, we remove the
mechanically inflated winners. The rank now reflects only genuine **sector and
factor rotation** — the kind of momentum a TAA screen should capture.

```{r eq-universe-snapshot}
tibble(
  Metric  = c("Equity tickers in universe",
              "W→L reversal events (equity only)",
              "L→W reversal events (equity only)",
              "Tickers involved in at least one reversal"),
  Value   = c(
    as.character(length(universe_eq)),
    as.character(sum(reversals_eq$reversal_type == "W\u2192L")),
    as.character(sum(reversals_eq$reversal_type == "L\u2192W")),
    as.character(n_distinct(reversals_eq$ticker))
  )
) %>%
  ktable(caption = "Equity-only sub-universe overview")
```

### Plot A (Equity) — Does Momentum Persist Differently?

The key comparison with the full-universe transition matrix is the **diagonal**.
Within equity-only, all tickers share the same SPY beta direction — they all
tend to fall together in Fall and rise together in Recovery. This removes the
artificial Q5 persistence that hedge ETFs gain simply by not falling when SPY
falls.

If the equity-only diagonal is **weaker** than the full universe diagonal, it
confirms that much of the apparent momentum in the full universe was driven by
asset-class separation (equities vs bonds), not genuine sector momentum.

If the equity-only diagonal is **stronger**, it reveals that within equities,
genuine sector momentum persists even after removing the hedge-ETF tailwind.

```{r plot-A-eq, fig.height=6.5, fig.cap="KF3-A (Equity Only): Quintile Transition Matrix restricted to equity ETFs. Compare the Q5→Q5 diagonal cell to the full-universe version — any reduction reflects the removal of mechanically persistent hedge ETF winners."}
print(p_A_eq)
```

### Plot B (Equity) — Regime Timing of Reversals, Cleaned Up

With hedge ETFs removed, the W→L and L→W signals in Plot B should become
*sharper and more actionable*:

- **Fall regime** W→L events now represent genuine equity sector leaders that
  ran out of momentum — not gold rotating back to bonds.
- **Recovery regime** L→W events now capture the rotation from defensive
  sectors (XLV, XLP) into growth and cyclicals (XLK, SMH, XLY) — a
  well-documented and exploitable pattern.

```{r plot-B-eq, fig.height=6, fig.cap="KF3-B (Equity Only): Reversal event count by regime, restricted to equity ETFs. The hedge-ETF mechanical reversals are removed — remaining signals reflect genuine sector rotation."}
print(p_B_eq)
```

```{r regime-table-eq}
reversals_eq %>%
  count(reversal_type, regime_at_exit) %>%
  mutate(regime_at_exit = factor(regime_at_exit,
                                 levels = c("Fall", "Recovery", "Consolidation"))) %>%
  arrange(reversal_type, regime_at_exit) %>%
  pivot_wider(names_from = regime_at_exit, values_from = n, values_fill = 0L) %>%
  rename(`Reversal type` = reversal_type) %>%
  ktable(caption = "Equity-only reversal count by regime and type")
```

### Plot C (Equity) — Is the Forward Return Signal Cleaner?

This is the most important comparison. If the forward return signal (directional
accuracy, median forward return) is **stronger** in the equity-only universe,
it confirms the hypothesis: hedge ETFs were adding noise, not signal.

A cleaner forward return distribution means:
- W→L events are more reliably followed by continued underperformance
- L→W events are more reliably followed by continued outperformance
- The signal is actionable as a TAA input

```{r plot-C-eq, fig.height=7, fig.cap="KF3-C (Equity Only): Forward relative return post-reversal. Compare the box widths and median positions to the full-universe version — a tighter distribution with clearer directional bias is a stronger signal."}
print(p_C_eq)
```

### Side-by-Side Signal Comparison

```{r signal-comparison}
# Build a combined comparison table
fwd_summary <- function(fwd_df, label) {
  fwd_df %>%
    filter(!is.na(fwd_rel_ret)) %>%
    group_by(`Reversal type` = reversal_type, `Window` = window) %>%
    summarise(
      `N events`       = n(),
      `Median fwd ret` = percent(median(fwd_rel_ret, na.rm = TRUE), accuracy = 0.1),
      `% correct`      = percent(
        mean(if_else(reversal_type == "W\u2192L",
                     fwd_rel_ret < 0, fwd_rel_ret > 0), na.rm = TRUE),
        accuracy = 1
      ),
      .groups = "drop"
    ) %>%
    mutate(Universe = label)
}

bind_rows(
  fwd_summary(fwd_returns,    "Full universe"),
  fwd_summary(fwd_returns_eq, "Equity only")
) %>%
  select(Universe, everything()) %>%
  arrange(`Reversal type`, `Window`, Universe) %>%
  ktable(
    caption = paste0(
      "Signal strength comparison: full universe vs equity-only. ",
      "'% correct' = fraction of events where the post-reversal return ",
      "confirmed the reversal direction. Higher = cleaner signal."
    )
  )
```

### Which Equity Sectors Reverse Most?

```{r eq-top-reversals}
reversals_eq %>%
  count(ticker, reversal_type, sub_block) %>%
  arrange(desc(n)) %>%
  head(24) %>%
  pivot_wider(names_from = reversal_type, values_from = n, values_fill = 0L) %>%
  rename(Ticker = ticker, `Sector / block` = sub_block) %>%
  arrange(desc(rowSums(across(where(is.integer))))) %>%
  ktable(caption = "Top equity reversal tickers by total event count and sector")
```

> **What to look for:** If high-beta tech/growth ETFs (XLK, SMH, QQQ) dominate
> W→L events concentrated in Fall, this confirms they are genuine momentum
> leaders in Consolidation and Recovery that become victims of regime change —
> exactly the pattern identified in KF3 Cycle Amplifiers.

---

## 10. Canary in the Mine — Leading and Confirming Indicators

### Two Different Roles

Not all ETFs are useful in the same way. This section separates the equity
universe into two distinct roles:

| Role | What it does | When to use it |
|---|---|---|
| **Transition canary** | Reversal signal arrives *before* the regime changes | Use for early warning — acts while the regime shift is still forming |
| **Persistence confirmor** | Stays in Q5 (or Q1) throughout an entire regime episode | Use to confirm the current regime is still intact — regime health check |

A canary ETF that consistently leads regime transitions by 2–3 weeks gives the
TAA framework a **forward signal** it would not get from SPY's drawdown alone
(which, by construction, can only confirm a Fall *after* −10% has already
occurred).

### How Lead Time Is Measured

For each reversal event in the equity universe:

- A **W→L event** is matched to the nearest **Fall onset date** within
  ±`r CANARY_WIN` calendar days.
- An **L→W event** is matched to the nearest **Recovery onset date** within
  ±`r CANARY_WIN` calendar days.

The **lead time** is:

$$\text{lead time} = \text{transition date} - \text{exit date}$$

- **Positive** → the reversal happened *before* the regime transition: the
  ETF's momentum rank was already collapsing while the regime label had not
  yet changed. This is the canary signal.
- **Negative** → the reversal happened *after* the transition: the ETF was
  slow to react. These are lagging indicators — confirming a change that
  already happened, not warning of one coming.

```{r canary-snapshot}
tibble(
  Metric  = c("Reversal events matched to a transition",
              "W→L events with positive lead (canary)",
              "L→W events with positive lead (canary)",
              "Median lead time across all events"),
  Value   = c(
    as.character(nrow(canary_lead_tbl)),
    as.character(sum(canary_lead_tbl$reversal_type == "W\u2192L" &
                       canary_lead_tbl$lead_days > 0)),
    as.character(sum(canary_lead_tbl$reversal_type == "L\u2192W" &
                       canary_lead_tbl$lead_days > 0)),
    paste0(round(median(canary_lead_tbl$lead_days, na.rm = TRUE)), " days")
  )
) %>%
  ktable(caption = "Canary lead time overview")
```

### Plot D — Canary Lead Time by Ticker

Each box shows the distribution of lead times for one ETF across all matched
events. The key interpretation rules:

- **Box entirely above zero** → this ETF *reliably leads* regime transitions.
  Act when you see it reverse.
- **Box straddling zero** → the ETF sometimes leads, sometimes lags. Useful
  as a supporting signal but not a primary trigger.
- **Box below zero** → this ETF is a lagging indicator. It confirms a regime
  change that already happened. Useful for persistence confirmation but not
  for early warning.

> Only ETFs with a **canary score ≥ 25** (top half of the equity universe by
> signal consistency) are shown.

```{r plot-D, fig.height=9, fig.cap="KF3-D: Canary lead time distribution per equity ETF. Green zone = reversal preceded the regime transition. Red zone = reversal followed. Tickers with boxes in the green zone are early warning signals."}
print(p_D)
```

### Top Transition Canaries

```{r canary-table}
canary_scores %>%
  filter(canary_score >= 40) %>%
  arrange(reversal_type, desc(canary_score)) %>%
  transmute(
    Ticker            = ticker,
    `Sector`          = sub_block,
    `Signal`          = reversal_type,
    `Canary score`    = canary_score,
    `N events`        = n_canary,
    `Median lead`     = paste0(round(median_lead), "d"),
    `% leading`       = percent(pct_leading, accuracy = 1),
    `Mean lead`       = paste0(round(mean_lead), "d")
  ) %>%
  ktable(
    caption = paste0(
      "High-conviction transition canaries (score \u2265 40). ",
      "Score = (n events) \u00d7 (% leading), normalised 0\u2013100. ",
      "Median lead = typical advance warning in calendar days."
    )
  )
```

**How to read the canary table:**

- **Median lead > 0** and **% leading > 60%** = reliable early warning.
  When this ETF's persistent state exits, start monitoring for regime change.
- **Median lead 10–25 days** is the sweet spot: long enough to act, short
  enough to be timely. Leads > 30 days suggest the ETF is driven by a
  different cycle than SPY.
- A high canary score with a *negative* median lead means the ETF often
  reverses *just after* a transition — it is a fast confirmer, not a
  forecaster.

### Plot E — Regime Persistence Heatmap

This heatmap answers the *confirmation* question: once a regime is established,
which equity ETFs spend the most time in Q5 (the winner bucket)?

A ticker with **high Q5 persistence in Recovery** confirms the Recovery is
running — it is not reversing, not churning, just steadily leading. If that
ticker *starts* to exit Q5 mid-Recovery, it is an early warning that the
regime may be tiring.

The same logic in reverse for Fall: a high Q1 persistence ticker that *starts*
to exit Q1 (entering Q2 or Q3) mid-Fall is a canary for an imminent Recovery.

```{r plot-E, fig.height=8, fig.cap="KF3-E: Average fraction of weekly rebalance dates in Q5 (top equity quintile) within each regime episode, averaged across all episodes of that type. Dark blue = persistent regime leader. Only tickers with max Q5 persistence > 30% shown."}
print(p_E)
```

### Regime Persistence Leaders

```{r persistence-table}
persistence_tbl %>%
  group_by(ticker) %>%
  filter(max(avg_pct_q5, na.rm = TRUE) > 0.35) %>%
  ungroup() %>%
  pivot_wider(id_cols = ticker,
              names_from = regime,
              values_from = c(avg_pct_q5, avg_pct_q1),
              names_glue = "{regime}_{.value}") %>%
  left_join(etf_metadata %>% select(ticker, sub_block), by = "ticker") %>%
  transmute(
    Ticker              = ticker,
    `Sector`            = sub_block,
    `Fall Q5%`          = percent(coalesce(Fall_avg_pct_q5,         0), accuracy = 1),
    `Recovery Q5%`      = percent(coalesce(Recovery_avg_pct_q5,     0), accuracy = 1),
    `Consolidation Q5%` = percent(coalesce(Consolidation_avg_pct_q5,0), accuracy = 1),
    `Fall Q1%`          = percent(coalesce(Fall_avg_pct_q1,         0), accuracy = 1)
  ) %>%
  arrange(desc(parse_number(`Recovery Q5%`))) %>%
  ktable(
    caption = paste0(
      "Intra-regime Q5 persistence leaders. ",
      "Q5% = average fraction of rebalance dates in the top equity quintile. ",
      "Q1% (Fall) = fraction in the bottom quintile during Fall — useful as a stress confirmor."
    )
  )
```

### The Combined Watchlist — Canary + Persistence Roles

```{r combined-watchlist}
# Canary role: top W→L canaries (signal Fall coming) + top L→W canaries
# (signal Recovery coming)
canary_role <- canary_scores %>%
  filter(canary_score >= 40) %>%
  group_by(ticker) %>%
  slice_max(canary_score, n = 1) %>%
  ungroup() %>%
  transmute(
    ticker,
    role = paste0(reversal_type, " canary"),
    score_label = paste0("Lead: ", round(median_lead), "d | ",
                         percent(pct_leading, accuracy = 1), " leading")
  )

# Persistence role: high Recovery Q5 persistence = Recovery confirmor
# high Fall Q1 persistence = Fall stress confirmor
persist_role <- persistence_tbl %>%
  pivot_wider(id_cols = ticker,
              names_from = regime,
              values_from = c(avg_pct_q5, avg_pct_q1),
              names_glue = "{regime}_{.value}") %>%
  mutate(
    role = case_when(
      coalesce(Recovery_avg_pct_q5, 0) >= 0.45   ~ "Recovery confirmor",
      coalesce(Fall_avg_pct_q1, 0)     >= 0.45   ~ "Fall stress confirmor",
      coalesce(Consolidation_avg_pct_q5, 0) >= 0.45 ~ "Consolidation leader",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(role)) %>%
  transmute(
    ticker,
    role,
    score_label = case_when(
      role == "Recovery confirmor"      ~
        percent(coalesce(Recovery_avg_pct_q5, 0),     accuracy = 1),
      role == "Fall stress confirmor"   ~
        percent(coalesce(Fall_avg_pct_q1, 0),         accuracy = 1),
      role == "Consolidation leader"    ~
        percent(coalesce(Consolidation_avg_pct_q5, 0),accuracy = 1)
    )
  )

bind_rows(canary_role, persist_role) %>%
  left_join(etf_metadata %>% select(ticker, sub_block), by = "ticker") %>%
  arrange(role, ticker) %>%
  transmute(
    Ticker      = ticker,
    `Sector`    = sub_block,
    `TAA role`  = role,
    `Signal quality` = score_label
  ) %>%
  ktable(
    caption = paste0(
      "Combined canary and persistence watchlist. ",
      "Canary ETFs: monitor for reversal events as early regime-transition warnings. ",
      "Confirmor ETFs: monitor for Q5/Q1 exits as regime-health alerts."
    )
  )
```

> **Practical TAA monitoring rule:**
>
> 1. **Weekly**: check if any *canary ETF* has exited its persistent winner/loser
>    state. If a W→L canary reverses, put a Fall watch on the portfolio.
>    If an L→W canary reverses, prepare the Recovery rotation list.
>
> 2. **Bi-weekly**: check if any *confirmor ETF* has dropped out of Q5
>    mid-regime. If the Recovery confirmors start slipping from Q5, the
>    Recovery may be tiring — time to trim the cyclical overweights.
>
> 3. **At regime transition** (SPY crosses −10%): the canary watchlist already
>    told you who the new leaders will be. The Recovery rotation list is ready.

### Section 10 Conclusion — The 10 Canary ETFs to Watch

The table below distils the entire canary and persistence analysis into a
single ranked watchlist. The **composite score** combines three dimensions:

- **Canary strength** — how consistently and how early does this ETF lead
  regime transitions? (from `canary_scores`)
- **Recovery persistence** — how reliably does it stay in Q5 throughout a
  Recovery, confirming the regime is running? (from `persistence_tbl`)
- **Fall persistence (Q1)** — how reliably does it stay in Q1 throughout a
  Fall, confirming stress is ongoing? (also `persistence_tbl`)

Each dimension is normalised 0–100 within the equity universe and averaged
with equal weights. Tickers are ranked by this composite.

```{r canary-top10}
# Canary strength: max of W→L and L→W canary scores per ticker
canary_strength <- canary_scores %>%
  group_by(ticker) %>%
  summarise(
    best_canary_score = max(canary_score, na.rm = TRUE),
    best_signal       = reversal_type[which.max(canary_score)],
    best_median_lead  = median_lead[which.max(canary_score)],
    best_pct_leading  = pct_leading[which.max(canary_score)],
    .groups = "drop"
  )

# Recovery and Fall persistence scores (normalised 0-100)
persist_scores <- persistence_tbl %>%
  pivot_wider(id_cols = ticker,
              names_from = regime,
              values_from = c(avg_pct_q5, avg_pct_q1),
              names_glue = "{regime}_{.value}") %>%
  mutate(
    rec_score  = round(100 * coalesce(Recovery_avg_pct_q5, 0) /
                         max(coalesce(Recovery_avg_pct_q5, 0), na.rm = TRUE)),
    fall_score = round(100 * coalesce(Fall_avg_pct_q1, 0) /
                         max(coalesce(Fall_avg_pct_q1, 0), na.rm = TRUE))
  ) %>%
  select(ticker, rec_score, fall_score,
         Recovery_avg_pct_q5, Fall_avg_pct_q1)

# Composite
canary_top10 <- canary_strength %>%
  left_join(persist_scores, by = "ticker") %>%
  left_join(etf_metadata %>% select(ticker, sub_block), by = "ticker") %>%
  mutate(
    composite = round((coalesce(best_canary_score, 0) +
                         coalesce(rec_score,  0) +
                         coalesce(fall_score, 0)) / 3)
  ) %>%
  arrange(desc(composite)) %>%
  head(10) %>%
  transmute(
    `#`               = row_number(),
    Ticker            = ticker,
    `Sector`          = sub_block,
    `Best signal`     = best_signal,
    `Median lead`     = paste0(round(coalesce(best_median_lead, 0)), "d"),
    `% leading`       = percent(coalesce(best_pct_leading, 0), accuracy = 1),
    `Recovery Q5%`    = percent(coalesce(Recovery_avg_pct_q5, 0), accuracy = 1),
    `Fall Q1%`        = percent(coalesce(Fall_avg_pct_q1, 0),     accuracy = 1),
    `Composite score` = composite
  )

canary_top10 %>%
  ktable(
    caption = paste0(
      "Top 10 canary ETFs by composite score. ",
      "Composite = average of (canary score, Recovery Q5 persistence score, ",
      "Fall Q1 persistence score), each normalised 0\u2013100. ",
      "'Best signal' = the reversal direction with the highest canary score. ",
      "'Median lead' = typical advance warning before regime transition."
    )
  )
```

```{r canary-top10-roles, echo=FALSE}
# Annotate each ticker with its primary TAA role in plain English
canary_top10_annotated <- canary_top10 %>%
  left_join(
    canary_scores %>%
      group_by(ticker) %>%
      summarise(
        wl = max(canary_score[reversal_type == "W\u2192L"], na.rm = TRUE),
        lw = max(canary_score[reversal_type == "L\u2192W"], na.rm = TRUE),
        .groups = "drop"
      ) %>%
      mutate(wl = if_else(is.infinite(wl), 0, wl),
             lw = if_else(is.infinite(lw), 0, lw)),
    by = c("Ticker" = "ticker")
  ) %>%
  left_join(
    persist_scores %>% select(ticker, rec_score, fall_score),
    by = c("Ticker" = "ticker")
  ) %>%
  mutate(
    `TAA role` = case_when(
      wl >= 50 & lw >= 50          ~ "Bilateral canary — warns both Fall & Recovery",
      wl >= 50 & coalesce(fall_score, 0) >= 60
                                   ~ "Fall warning + stress confirmor",
      lw >= 50 & coalesce(rec_score,  0) >= 60
                                   ~ "Recovery warning + momentum confirmor",
      wl >= 50                     ~ "Fall transition canary",
      lw >= 50                     ~ "Recovery transition canary",
      coalesce(rec_score, 0) >= 60 ~ "Recovery momentum confirmor",
      coalesce(fall_score,0) >= 60 ~ "Fall stress confirmor",
      TRUE                         ~ "Supporting signal"
    )
  ) %>%
  select(Ticker, `Sector`, `TAA role`, `Composite score`)

canary_top10_annotated %>%
  ktable(
    caption = paste0(
      "Primary TAA role for each top-10 canary ETF. ",
      "Use this table to assign each ticker to a monitoring lane: ",
      "canaries to watch weekly for reversal events, ",
      "confirmors to check bi-weekly for Q5 exits."
    )
  )
```

### How to Read These Two Tables

**Table 1 — Top 10 Canary ETFs column guide:**

| Column | What it tells you |
|:---|:---|
| Best signal | Whether this ETF is primarily a W→L or L→W canary |
| Median lead | Typical advance warning in days before the transition |
| % leading | How reliably the signal arrives before (not after) the transition |
| Recovery Q5% | Fraction of Recovery time spent in the top equity quintile |
| Fall Q1% | Fraction of Fall time spent in the bottom equity quintile |
| Composite score | Average of canary + Recovery persistence + Fall persistence, each 0–100 |

**Table 2 — TAA role labels explained:**

- **Bilateral canary** — warns both directions; monitor for any reversal event regardless of regime
- **Fall warning + stress confirmor** — strong W→L lead time AND high intra-Fall Q1 persistence; primary Fall early-warning ticker
- **Recovery warning + momentum confirmor** — strong L→W lead time AND high intra-Recovery Q5 persistence; primary Recovery early-warning ticker
- **Fall / Recovery transition canary** — strong lead time in one direction only; use as a directional trigger
- **Recovery momentum confirmor / Fall stress confirmor** — high persistence score but weak lead time; use for regime health checks, not early warning
- **Supporting signal** — contributes to the composite but is not a primary trigger on its own

---

## 11. Investment Conclusion — Which Equity ETFs to Focus on for TAA?

The three plots above describe the *mechanism*. This section converts that
into an **actionable TAA watchlist**: which equity ETFs have the most reliable
reversal signals, in which regime to act, and in which direction.

### How the TAA Focus Score Is Built

For each equity ETF we compute two one-sided scores:

- **Re-entry score (L→W):** How often does this ticker flip from persistent
  loser to persistent winner, and how much does it outperform SPY in the
  following 63 days? A high score = a reliable "buy on regime change" candidate.

- **Exit score (W→L):** How often does this ticker flip from persistent winner
  to persistent loser, and by how much does it underperform SPY afterward?
  A high score = a reliable "sell / reduce on regime change" candidate.

Both scores are normalised 0–100 across the equity universe. The **dominant
regime** column shows where most of this ticker's events occur — Fall events
are the most actionable because they coincide with a regime transition signal
that the framework already detects.

```{r taa-focus-score, echo=FALSE}
# ── Per-ticker forward return quality (equity universe only) ──────────────────
fwd_63_eq <- fwd_returns_eq %>%
  filter(window == "63-day fwd", !is.na(fwd_rel_ret))

# L→W score: frequency × directional accuracy × median magnitude
lw_score <- fwd_63_eq %>%
  filter(reversal_type == "L\u2192W") %>%
  group_by(ticker) %>%
  summarise(
    lw_n          = n(),
    lw_median_fwd = median(fwd_rel_ret, na.rm = TRUE),
    lw_pct_ok     = mean(fwd_rel_ret > 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    lw_raw   = lw_n * lw_pct_ok * pmax(lw_median_fwd, 0),
    lw_score = round(100 * (lw_raw - min(lw_raw, na.rm = TRUE)) /
                     (max(lw_raw, na.rm = TRUE) - min(lw_raw, na.rm = TRUE) + 1e-9))
  )

# W→L score: frequency × directional accuracy × median magnitude (negative = good)
wl_score <- fwd_63_eq %>%
  filter(reversal_type == "W\u2192L") %>%
  group_by(ticker) %>%
  summarise(
    wl_n          = n(),
    wl_median_fwd = median(fwd_rel_ret, na.rm = TRUE),
    wl_pct_ok     = mean(fwd_rel_ret < 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    wl_raw   = wl_n * wl_pct_ok * pmax(-wl_median_fwd, 0),
    wl_score = round(100 * (wl_raw - min(wl_raw, na.rm = TRUE)) /
                     (max(wl_raw, na.rm = TRUE) - min(wl_raw, na.rm = TRUE) + 1e-9))
  )

# Dominant regime for each ticker's reversal events
dominant_regime <- reversals_eq %>%
  count(ticker, regime_at_exit) %>%
  group_by(ticker) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ticker, dominant_regime = regime_at_exit)

# Combine
taa_focus <- etf_metadata %>%
  filter(ticker %in% universe_eq) %>%
  select(ticker, sub_block) %>%
  left_join(lw_score %>% select(ticker, lw_n, lw_score, lw_median_fwd, lw_pct_ok),
            by = "ticker") %>%
  left_join(wl_score %>% select(ticker, wl_n, wl_score, wl_median_fwd, wl_pct_ok),
            by = "ticker") %>%
  left_join(dominant_regime, by = "ticker") %>%
  filter(!is.na(lw_score) | !is.na(wl_score)) %>%
  mutate(
    combined_score = coalesce(lw_score, 0L) + coalesce(wl_score, 0L),
    lw_score       = coalesce(lw_score, 0L),
    wl_score       = coalesce(wl_score, 0L)
  ) %>%
  arrange(desc(combined_score))
```

### The TAA Focus List — Top 20 Equity ETFs by Signal Strength

```{r taa-focus-table}
taa_focus %>%
  head(20) %>%
  transmute(
    Ticker            = ticker,
    `Sector / block`  = sub_block,
    `Dominant regime` = coalesce(dominant_regime, "—"),
    `L→W score`       = lw_score,
    `L→W events`      = coalesce(lw_n, 0L),
    `L→W median 63d`  = percent(coalesce(lw_median_fwd, 0), accuracy = 0.1),
    `L→W % correct`   = percent(coalesce(lw_pct_ok, 0),    accuracy = 1),
    `W→L score`       = wl_score,
    `W→L events`      = coalesce(wl_n, 0L),
    `W→L median 63d`  = percent(coalesce(wl_median_fwd, 0), accuracy = 0.1)
  ) %>%
  ktable(
    caption = paste0(
      "TAA Focus List — top 20 equity ETFs ranked by combined L→W + W→L signal strength. ",
      "Scores are 0–100 normalised within the equity universe. ",
      "L→W = buy signal at regime shift; W→L = exit/reduce signal."
    )
  )
```

### Regime-Specific TAA Playbook

The table above identifies *who* has the signal. The table below tells you
*when* to act, structured as a regime-conditional playbook.

```{r taa-playbook}
# High-conviction L→W in Recovery → buy at regime shift
recovery_buys <- reversals_eq %>%
  filter(reversal_type == "L\u2192W", regime_at_exit == "Recovery") %>%
  count(ticker) %>%
  left_join(lw_score %>% select(ticker, lw_score, lw_pct_ok, lw_median_fwd),
            by = "ticker") %>%
  filter(!is.na(lw_score), lw_score >= 30) %>%
  arrange(desc(n)) %>%
  left_join(etf_metadata %>% select(ticker, sub_block), by = "ticker")

# High-conviction W→L in Fall → exit at regime onset
fall_exits <- reversals_eq %>%
  filter(reversal_type == "W\u2192L", regime_at_exit == "Fall") %>%
  count(ticker) %>%
  left_join(wl_score %>% select(ticker, wl_score, wl_pct_ok, wl_median_fwd),
            by = "ticker") %>%
  filter(!is.na(wl_score), wl_score >= 30) %>%
  arrange(desc(n)) %>%
  left_join(etf_metadata %>% select(ticker, sub_block), by = "ticker")

cat("─── BUY / OVERWEIGHT at Fall → Recovery transition ───\n")
recovery_buys %>%
  transmute(
    Ticker           = ticker,
    `Sector`         = sub_block,
    `Recovery L→W events` = n,
    `Signal score`   = lw_score,
    `% correct 63d`  = percent(lw_pct_ok,    accuracy = 1),
    `Median 63d fwd` = percent(lw_median_fwd, accuracy = 0.1)
  ) %>%
  ktable(caption = "BUY / OVERWEIGHT candidates: L→W events concentrated in Recovery")

cat("\n─── SELL / REDUCE at Consolidation → Fall transition ───\n")
fall_exits %>%
  transmute(
    Ticker           = ticker,
    `Sector`         = sub_block,
    `Fall W→L events` = n,
    `Signal score`   = wl_score,
    `% correct 63d`  = percent(wl_pct_ok,    accuracy = 1),
    `Median 63d fwd` = percent(wl_median_fwd, accuracy = 0.1)
  ) %>%
  ktable(caption = "SELL / REDUCE candidates: W→L events concentrated in Fall onset")
```

### Reading the Playbook

**Column definitions:**

- **L→W score / W→L score** — 0–100, normalised within the equity universe.
  Above 50 = top-half signal quality. Above 75 = high conviction.
- **% correct 63d** — fraction of events where the post-reversal 63-day
  relative return confirmed the signal direction. Above 55% is economically
  meaningful; above 60% is strong.
- **Median 63d fwd** — the *typical* relative return vs SPY in the 63 days
  after the reversal. Positive for L→W (ticker outperforms), negative for
  W→L (ticker underperforms).

**How to use this in TAA:**

> When the SPY drawdown crosses the **10% Fall threshold** (the trigger already
> built into `build_regime_table()`), the framework should immediately flag the
> W→L candidates in the exit table for **reduction or removal** from the TAA
> sleeve. These are the tickers most likely to underperform SPY for the next
> 63 trading days.

> When SPY bottoms and begins recovering (regime flips to Recovery), the L→W
> candidates in the buy table are the **first rotation targets** — the tickers
> empirically most likely to outperform SPY by the largest margin over the next
> quarter.

> In **Consolidation**, momentum is most reliable (the transition matrix
> diagonal is highest). No regime-conditional discount is needed — the standard
> trailing-return TAA rank applies.

### Connection to the Current SAA Portfolio

```{r saa-connection}
SAA_TICKERS <- c("SPY", "URTH", "QQQ", "XLK", "SMH")

taa_focus %>%
  filter(ticker %in% SAA_TICKERS) %>%
  transmute(
    Ticker           = ticker,
    `Sector / block` = sub_block,
    `L→W score`      = lw_score,
    `W→L score`      = wl_score,
    `Combined score` = combined_score,
    `Dominant regime`= coalesce(dominant_regime, "—"),
    `TAA implication` = case_when(
      lw_score > wl_score & lw_score >= 40 ~
        "Strong buy signal at regime shift — overweight in Recovery",
      wl_score > lw_score & wl_score >= 40 ~
        "Strong exit signal at regime onset — reduce in Fall",
      combined_score >= 40 ~
        "Bilateral signal — active both ways",
      TRUE ~
        "Weak reversal signal — hold at SAA weight"
    )
  ) %>%
  ktable(
    caption = paste0(
      "TAA signal quality for current SAA equity holdings. ",
      "Scores above 40 indicate meaningful reversal signal worthy of active TAA overlay."
    )
  )
```

---

---

## 12. Technical Appendix

### Why 60 Days as the Formation Window?

The 60-day window (≈ 3 months) is a deliberate compromise:

- Shorter windows (10–20 days) are dominated by microstructure noise and
  mean-reversion. The 1-week reversal effect is well-documented in academic
  literature and is largely a liquidity artefact.
- Longer windows (120–252 days) capture structural trends but are too slow to
  detect regime transitions, which typically play out in 4–10 weeks.
- 60 days roughly aligns with one *earnings cycle*, making it economically
  motivated: the rank reflects approximately one quarter of fundamental news.

### Why Weekly (Not Daily) Rebalancing?

Daily quintile changes are dominated by noise: a ticker in Q5 moves to Q4 one
day and back to Q5 the next simply due to daily return volatility. Weekly
rebalancing smooths this without losing meaningful information about medium-term
trends.

### The Persistence Filter as a Noise Gate

The `PERSIST_N = 3` filter means a ticker must be in Q1 or Q5 for **15
consecutive trading days** before it is labelled as persistent. This filters
out approximately 60–70% of quintile transitions that are statistical flukes
rather than genuine regime shifts in relative performance.

### How the Transition Window Prevents False Signals

Without `TRANSIT_WIN = 8`, we might label a W→L event when a ticker left Q5,
spent 6 months in Q3, and *then* entered Q1. That is not a reversal — it is two
unrelated events. The 8-week cap ensures that only *fast* transitions from one
extreme to the other are captured, which is precisely the regime-driven
reversal mechanism we are trying to measure.

---

*Analysis built on `key_findings/kf3_trend_reversal.R` — source that file to
reproduce all objects. Full-universe: `rank_tbl`, `reversals`, `fwd_returns`,
`p_A/B/C`. Equity-only: `rank_tbl_eq`, `reversals_eq`, `fwd_returns_eq`,
`p_A_eq/B_eq/C_eq`. Canary: `canary_lead_tbl`, `canary_scores`,
`persistence_tbl`, `p_D`, `p_E`. Parameters: REBAL_FREQ = `r REBAL_FREQ` days,
LOOKBACK = `r LOOKBACK` days, PERSIST_N = `r PERSIST_N` periods,
TRANSIT_WIN = `r TRANSIT_WIN` periods, CANARY_WIN = `r CANARY_WIN` days.*
