---
title: "Market State Framework"
subtitle: "Documentation & Analysis — 3-Layer Regime Allocation Engine"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    fig_caption: true
params:
  min_persist: 5      # days new state must hold before confirmation
  t_fall:      0.10   # SPY/URTH fall threshold
  roll_win:    60     # rolling window for premium/appetite spreads (days)
  ma_win:      200    # moving-average window for sector breadth
  mom_win:     63     # sector rotation momentum window (days)
---

```{css, echo=FALSE}
/* ── Layout ─────────────────────────────────────────────────────────────── */
.state-badge {
  display: inline-block;
  padding: 6px 16px;
  border-radius: 4px;
  font-weight: bold;
  font-size: 1.05em;
  color: white;
  margin: 2px 4px;
}
.signal-box {
  background: #F0F4FA;
  border-left: 4px solid #1B2A4A;
  padding: 10px 16px;
  margin: 10px 0;
  font-size: 0.9em;
}
.doc-note {
  background: #fffbeb;
  border-left: 4px solid #d97706;
  padding: 8px 14px;
  margin: 10px 0;
  font-size: 0.88em;
  color: #444;
}
.finding-box {
  background: #f0fdf4;
  border-left: 4px solid #16a34a;
  padding: 10px 16px;
  margin: 12px 0;
  font-size: 0.9em;
}
.caption {
  color: #555;
  font-size: 0.82em;
  font-style: italic;
  border-left: 3px solid #1B3A6B;
  padding-left: 6px;
  margin-top: 4px;
}
.section-div { border-top: 2px solid #dee2e6; margin: 48px 0 24px 0; }
table { font-size: 0.88em; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 6
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(lubridate)
library(here)

select <- dplyr::select

if (!exists("params"))
  params <- list(min_persist = 5L, t_fall = 0.10,
                 roll_win = 60, ma_win = 200, mom_win = 63)

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
if (!exists("label_daily_regime"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

if (!exists("g_sector_tickers"))
  source(here("scripts/00_init_universe.R"))

source(here("scripts_market_state/ms_vector_state.R"))
source(here("scripts_market_state/ms_sector_pattern.R"))
source(here("scripts_market_state/ms_risk_allocation.R"))
source(here("scripts_market_state/ms_visuals.R"))

# ── Build histories ───────────────────────────────────────────────────────────
ms_hist <- build_market_state_history(
  xts_ret,
  roll_win    = params$roll_win,
  ma_win      = params$ma_win,
  t_fall      = params$t_fall,
  min_persist = params$min_persist
)

sp_hist <- build_sector_pattern_history(
  xts_ret,
  mom_win = params$mom_win
)

ra_hist <- build_risk_allocation_history(
  xts_ret,
  roll_win    = params$roll_win,
  ma_win      = params$ma_win,
  t_fall      = params$t_fall,
  mom_win     = params$mom_win,
  min_persist = params$min_persist
)

# ── Current snapshot ─────────────────────────────────────────────────────────
cur_ms <- ms_hist %>% filter(!is.na(market_state)) %>% tail(1)
cur_sp <- sp_hist %>% filter(!is.na(sector_pattern)) %>% tail(1)
cur_ra <- ra_hist %>% filter(!is.na(final_eq_weight)) %>% tail(1)

n_years <- as.numeric(diff(range(ms_hist$date))) / 365.25
```

---

# The Idea in Plain English

Think of this framework like **reading the weather before deciding what to wear**.

The basic question is simple: *how much of your portfolio should be in stocks vs bonds right now?*
We answer it in three steps.

---

**Layer 1 — "What's the weather globally?"**

We check 5 signals, like a weather station:

- Is the **world stock market** (URTH) falling or recovering?
- Is the **US market** (SPY) falling or recovering?
- Is the **US outperforming** the rest of the world over the last 60 days?
- Is **Emerging Markets** (riskier countries) keeping up with the US? If yes, investors are feeling brave globally.
- How many of the 10 US sectors are **above their 200-day average price**? If most are, the rally is broad and healthy. If only a few, it's fragile.

These 5 answers collapse into **one of 5 named states** — from best to worst:
**Expansion → US-Led → Late-Cycle → Deterioration → Contraction.**
Each state comes with a suggested equity weight (35% to 65%).

> **The persistence filter:** We don't react to a single bad day. The state has to hold for
> **`r params$min_persist` days in a row** before we believe it. Without this, the raw signal
> fires ~24 times per year — mostly noise. With it, we get ~7 real transitions per year
> (roughly quarterly).

---

**Layer 2 — "Which sectors are leading?"**

Inside the market, are **risky sectors** (tech, consumer, energy, industrials, materials)
outperforming **safe sectors** (utilities, staples, healthcare)?

- Risky sectors winning → investors are confident → nudge equity **up** 2–5%
- Safe sectors winning → investors are hiding → nudge equity **down** 2–5%

---

**Layer 3 — "Final answer"**

Layer 1 base weight + Layer 2 nudge = **final equity %**, clipped between 25% and 75%
so we never go to extremes.

---

**What this report shows:**

| Section | What you'll find |
|---|---|
| Current State Snapshot | Today's reading — state, signals, allocation vs passive 60/40 |
| Framework Architecture | How each signal is built and why |
| Historical Analysis | How the state has evolved since `r params$from_yr` |
| Persistence Analysis | How much noise the filter removes, and at what cost in lag |
| Transition Analysis | How often states change month-to-month, and by how much |
| Calibration Notes | Every threshold documented with its rationale and known limitations |

---

# Current State Snapshot {.tabset}

> As of **`r format(cur_ra$date, "%d %B %Y")`** &nbsp;|&nbsp;
> Persistence filter: **`r params$min_persist` days** &nbsp;|&nbsp;
> T_fall: **`r scales::percent(params$t_fall)`**

```{r snapshot-badges, results='asis'}
state_colours <- c(
  Expansion       = "#16a34a",
  `US-Led`        = "#2563eb",
  `Late-Cycle`    = "#d97706",
  Deterioration   = "#ea580c",
  Contraction     = "#dc2626"
)
pattern_colours <- c(
  `Strong Risk-On`  = "#15803d",
  `Risk-On`         = "#4ade80",
  `Neutral`         = "#64748b",
  `Risk-Off`        = "#fb923c",
  `Strong Risk-Off` = "#dc2626"
)

ms_col <- state_colours[as.character(cur_ms$market_state)]
sp_col <- pattern_colours[as.character(cur_sp$sector_pattern)]

cat(sprintf(
  "<div style='margin:18px 0 10px 0;'>
  <span style='font-size:0.9em;color:#555;'>L1 Market State</span><br>
  <span class='state-badge' style='background:%s;font-size:1.3em;'>%s</span>
  &nbsp;&nbsp;
  <span style='font-size:0.9em;color:#555;'>L2 Sector Pattern</span><br>
  <span class='state-badge' style='background:%s;font-size:1.3em;'>%s</span>
  &nbsp;&nbsp;&nbsp;&nbsp;
  <span style='font-size:0.9em;color:#555;'>Final EQ weight</span><br>
  <span style='font-size:2.2em;font-weight:bold;color:#1B2A4A;'>%.0f%%</span>
  &nbsp;
  <span style='font-size:1.1em;color:#888;'>/ %.0f%% FI</span>
  </div>",
  ms_col, as.character(cur_ms$market_state),
  sp_col, as.character(cur_sp$sector_pattern),
  cur_ra$final_eq_weight * 100,
  cur_ra$fi_weight * 100
))
```

## Signal readings

```{r snapshot-table}
snap <- tibble(
  Signal = c(
    "URTH regime", "SPY regime",
    "US premium (60d)", "EM appetite (60d)",
    "Sector breadth (200DMA)",
    "Cyclical momentum (63d)", "Defensive momentum (63d)",
    "Rotation score"
  ),
  Value = c(
    cur_ms$urth_regime, cur_ms$spy_regime,
    paste0(sprintf("%+.1f", cur_ms$us_premium_60d), "%  (",
           if (cur_ms$us_premium > 0) "US leading" else "US lagging", ")"),
    paste0(sprintf("%+.1f", cur_ms$em_appetite_60d), "%  (",
           if (cur_ms$em_appetite > 0) "EM leading \u2191" else "EM lagging \u2193", ")"),
    paste0(round(cur_ms$sector_breadth * 100), "% above 200DMA"),
    paste0(sprintf("%+.1f", cur_sp$cyclical_mom), "%"),
    paste0(sprintf("%+.1f", cur_sp$defensive_mom), "%"),
    paste0(sprintf("%+.2f", cur_sp$rotation_score), "%  (",
           as.character(cur_sp$sector_pattern), ")")
  ),
  Layer = c("L1","L1","L1","L1","L1","L2","L2","L2")
)

kable(snap, caption = paste0("Signal readings — ", format(cur_ra$date))) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  column_spec(3, color = "#888888")
```

## Allocation

```{r snapshot-alloc-table}
alloc <- tibble(
  ` ` = c("L1 base (market state)", "L2 adjustment (sector pattern)",
          "Final equity weight", "Final fixed income weight"),
  Weight = c(
    paste0(round(cur_ra$base_eq_weight * 100), "%"),
    paste0(sprintf("%+.0f", (cur_ra$adj_eq_weight - cur_ra$base_eq_weight) * 100), "%"),
    paste0("<strong>", round(cur_ra$final_eq_weight * 100), "%</strong>"),
    paste0("<strong>", round(cur_ra$fi_weight * 100), "%</strong>")
  ),
  vs_passive = c(
    paste0(sprintf("%+.0f", (cur_ra$base_eq_weight - 0.60) * 100), "pp vs 60/40"),
    "—",
    paste0(sprintf("%+.0f", (cur_ra$final_eq_weight - 0.60) * 100), "pp vs 60/40"),
    paste0(sprintf("%+.0f", (cur_ra$fi_weight - 0.40) * 100), "pp vs 60/40")
  )
)

kable(alloc, escape = FALSE,
      col.names = c("Component", "Weight", "vs Passive 60/40"),
      caption = "Allocation decomposition") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(3:4, bold = TRUE, background = "#F0F4FA") %>%
  column_spec(1, color = "#1B2A4A")
```

---

# Framework Architecture {.tabset}

<div class="doc-note">
This section documents the design decisions and signal definitions. It is intended as a
reference for future review and calibration.
</div>

## Overview

The framework produces a **daily equity/fixed-income allocation** from three independent layers:

```
URTH ──┐
SPY  ──┤  L1: Market State Vector   →  5 named states   →  base EQ weight (35–65%)
VWO  ──┤       (5 signals)
Sectors┘

Sectors ──  L2: Sector Rotation     →  5 pattern labels →  ±adjustment (−5% to +5%)
            (cyclical vs defensive momentum)

L1 + L2  →  L3: Risk Allocation    →  final EQ/FI weight  (clipped 25–75%)
```

**Design principles:**

1. **Hierarchy before breadth** — global (URTH) anchors the view; US (SPY) refines it; sectors confirm it.
2. **Confirmation before action** — `min_persist = `r params$min_persist`` days required before a state change is acted on (eliminates `r round((1 - 59/191)*100)`% of raw transitions).
3. **Additive, not multiplicative** — L2 adds ±pp on top of L1 base weight; the final weight is bounded [25%, 75%].
4. **Observable signals only** — all inputs are price-based, computable daily from public ETF data.

## L1 — Market State Vector

The five signals and their construction:

```{r l1-signal-table}
sig_tbl <- tibble(
  `#` = 1:5,
  Signal = c("URTH regime", "SPY regime", "US premium", "EM appetite", "Sector breadth"),
  Construction = c(
    "label_daily_regime(URTH, t_fall=10%) → Fall / Recovery / Consolidation",
    "label_daily_regime(SPY,  t_fall=10%) → Fall / Recovery / Consolidation",
    paste0("sign(rolling ", params$roll_win, "d cumret SPY − URTH)  → +1 (US leading) / −1 (US lagging)"),
    paste0("sign(rolling ", params$roll_win, "d cumret VWO − SPY)   → +1 (EM leading) / −1 (EM lagging)"),
    paste0("% of 10 GICS sector ETFs with price ≥ ", params$ma_win, "-day MA  → 0–1")
  ),
  `Why it matters` = c(
    "Global regime anchor — URTH falling means DM-wide risk-off, not just a US event",
    "US-specific regime — ~65% of URTH; divergence from URTH = regional decoupling",
    "Measures US exceptionalism premium; negative = broadening / non-US leadership",
    "EM outperformance = global risk appetite is open; EM lags in genuine risk-off",
    "Breadth is a participation measure: narrow rallies (few sectors) are fragile"
  )
)

kable(sig_tbl, caption = "Layer 1 signal definitions") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, width = "2em") %>%
  column_spec(2, bold = TRUE, color = "#1B2A4A", width = "12em") %>%
  column_spec(3, monospace = TRUE, width = "28em")
```

## L1 — State Classification

States are checked in **priority order** — most severe checked first:

```{r state-class-table}
state_tbl <- tibble(
  Priority = 1:5,
  State = c("Contraction", "Deterioration", "Late-Cycle", "US-Led", "Expansion"),
  Condition = c(
    "URTH = Fall AND SPY = Fall AND EM appetite < 0 AND breadth < 40%",
    "(URTH = Fall OR SPY = Fall) AND breadth < 55%",
    "breadth < 55%  OR  (breadth < 70% AND EM appetite < 0)",
    "US premium > 0 AND EM appetite ≤ 0",
    "(all else — broad participation, EM not lagging)"
  ),
  `EQ weight` = c("35%", "42%", "50%", "58%", "65%"),
  Interpretation = c(
    "Synchronised global sell-off with defensive rotation",
    "At least one major index in drawdown; breadth narrow",
    "Breadth fading; sector rotation toward defensives",
    "US driving — international drag reduces risk appetite",
    "Broad participation; EM affirms global risk appetite"
  )
)

state_colours_tbl <- c("#dc2626","#ea580c","#d97706","#2563eb","#16a34a")

kable(state_tbl, caption = "State classification logic (priority order, top to bottom)") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(2, bold = TRUE,
              color = "white",
              background = state_colours_tbl) %>%
  column_spec(4, bold = TRUE)
```

## L2 — Sector Rotation

```{r l2-table}
l2_tbl <- tibble(
  Group = c("Cyclicals", "Defensives", "Excluded"),
  Tickers = c(
    paste(intersect(SP_CYCLICALS,  colnames(xts_ret)), collapse = ", "),
    paste(intersect(SP_DEFENSIVES, colnames(xts_ret)), collapse = ", "),
    "XLF (ambiguous rate sensitivity), XLRE (rate sensitive)"
  ),
  `Momentum window` = c(
    paste0(params$mom_win, "d cumret"),
    paste0(params$mom_win, "d cumret"),
    "—"
  ),
  `Role` = c(
    "Growth/risk: outperform in expansion, underperform in stress",
    "Income/safety: outperform when markets derisking",
    "Omitted to keep signal clean"
  )
)

kable(l2_tbl, caption = "Layer 2 sector groupings") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

**Rotation score** = mean(cyclical 63d mom) − mean(defensive 63d mom)

Pattern labels and EQ adjustments:

```{r pattern-adj-table}
pat_tbl <- tibble(
  Pattern = c("Strong Risk-On", "Risk-On", "Neutral", "Risk-Off", "Strong Risk-Off"),
  Threshold = c("> +4%", "+1% to +4%", "−1% to +1%", "−4% to −1%", "< −4%"),
  `EQ adj` = c("+5pp", "+2pp", "0pp", "−2pp", "−5pp"),
  Colour = c("#15803d","#4ade80","#64748b","#fb923c","#dc2626")
)

kable(pat_tbl %>% select(-Colour),
      caption = "Layer 2 pattern labels and equity weight adjustments") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE,
              color = "white",
              background = pat_tbl$Colour) %>%
  column_spec(3, bold = TRUE)
```

---

# Historical Analysis {.tabset}

## State timeline

```{r timeline, fig.height=4.5}
plot_ms_timeline(ms_hist)
```

<p class="caption">White line = confirmed equity weight (y-axis right). Colour = confirmed market state after `r params$min_persist`-day persistence filter. Raw (unfiltered) state also computed but not shown.</p>

## Signal strips

```{r strips, fig.height=3.5}
plot_ms_signal_strip(ms_hist)
```

<p class="caption">Each row is one of the 5 L1 signals, mapped to a −1 → +1 scale. Red = risk-off / Fall. Green = risk-on / Expansion. Regime signals (URTH, SPY) encoded as Fall=−1, Recovery=+0.5, Consolidation=+1.</p>

## Allocation history

```{r alloc-chart, fig.height=5}
plot_ms_allocation(ra_hist)
```

<p class="caption">Stacked area = final EQ (blue) / FI (grey) weights after L1+L2 combination. Red/orange shading = Contraction / Deterioration periods. Dashed lines = passive 60/40 reference.</p>

---

# Persistence Analysis {.tabset}

<div class="doc-note">
The raw daily signal generates ~24 transitions/year — nearly 2 per month.
Most are 1–4 day noise runs. The persistence filter eliminates these before
they reach the allocation decision.
</div>

## Filter comparison

```{r persist-compare}
thresholds <- c(1, 3, 5, 10, 15, 21)

persist_tbl <- map_dfr(thresholds, function(p) {
  h <- build_market_state_history(
    xts_ret,
    roll_win = params$roll_win, ma_win = params$ma_win,
    t_fall = params$t_fall, min_persist = p
  ) %>% filter(!is.na(market_state))

  n_yr  <- as.numeric(diff(range(h$date))) / 365.25
  trans <- sum(h$market_state != lag(h$market_state), na.rm = TRUE)
  raw_t <- sum(h$market_state_raw != lag(h$market_state_raw), na.rm = TRUE)
  lag_p <- mean(as.character(h$market_state) != as.character(h$market_state_raw), na.rm = TRUE)

  tibble(
    `min_persist (days)` = p,
    `Transitions/yr`     = round(trans / n_yr, 1),
    `Noise filtered`     = paste0(round((1 - trans / raw_t) * 100), "%"),
    `Days lagged`        = paste0(round(lag_p * 100), "%"),
    `Approx cadence`     = case_when(
      trans / n_yr > 15 ~ "Weekly",
      trans / n_yr > 6  ~ "Monthly",
      trans / n_yr > 2  ~ "Quarterly",
      TRUE              ~ "Semi-annual"
    )
  )
})

kable(persist_tbl,
      caption = paste0("Effect of persistence filter on confirmed transitions  (full history: ",
                       round(n_years, 1), " years)")) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(which(thresholds == params$min_persist),
           bold = TRUE, background = "#F0F4FA") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  footnote(general = paste0("Highlighted row = current setting (min_persist = ", params$min_persist, ")"),
           general_title = "")
```

<div class="finding-box">
**Sweet spot: 5-day filter** eliminates ~69% of raw noise, leaves ~7 real transitions/year (quarterly cadence),
and lags confirmation by only ~18% of days — still fast enough to catch swift deteriorations (e.g. COVID-style drops
typically persist &gt;5 days before reversing).
</div>

## Run length distribution

```{r run-length, fig.height=4.5}
runs <- rle(as.character(ms_hist$market_state[!is.na(ms_hist$market_state)]))
run_df <- tibble(state = runs$values, days = runs$lengths) %>%
  mutate(state = factor(state, levels = MS_STATE_LEVELS))

ggplot(run_df, aes(x = days, fill = state)) +
  geom_histogram(bins = 40, colour = "white", linewidth = 0.15) +
  scale_fill_manual(values = MS_STATE_PAL, name = NULL) +
  scale_x_continuous(breaks = c(1, 5, 10, 21, 63, 126),
                     labels = c("1d","5d","10d","21d","63d","126d")) +
  geom_vline(xintercept = params$min_persist,
             linetype = "dashed", colour = "#1B2A4A", linewidth = 0.8) +
  annotate("text", x = params$min_persist + 1, y = Inf, vjust = 1.4, hjust = 0,
           label = paste0("persist=", params$min_persist), size = 3.2, colour = "#1B2A4A") +
  facet_wrap(~ state, ncol = 5, scales = "free_y") +
  labs(title = paste0("State run lengths — min_persist = ", params$min_persist),
       subtitle = "Each bar = one run of consecutive days in that confirmed state",
       x = "Run length (days)", y = "Count") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"))
```

## State distribution

```{r state-dist}
valid_ms <- ms_hist %>% filter(!is.na(market_state))

dist_tbl <- valid_ms %>%
  count(market_state) %>%
  mutate(
    pct     = n / sum(n),
    avg_dur = map_dbl(market_state, function(s) {
      r <- rle(as.character(valid_ms$market_state))
      mean(r$lengths[r$values == as.character(s)])
    })
  ) %>%
  transmute(
    State          = as.character(market_state),
    Periods        = map_int(market_state, function(s) {
      r <- rle(as.character(valid_ms$market_state))
      sum(r$values == as.character(s))
    }),
    `Trading days` = n,
    `% of history` = scales::percent(pct, accuracy = 0.1),
    `Avg run (days)` = round(avg_dur, 1)
  )

kable(dist_tbl, caption = paste0("State distribution — confirmed (min_persist = ",
                                  params$min_persist, ")")) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "white",
              background = MS_STATE_PAL[dist_tbl$State])
```

---

# Transition Analysis {.tabset}

## Monthly transition matrix

```{r monthly-trans, fig.height=5}
monthly <- ra_hist %>%
  filter(!is.na(market_state)) %>%
  mutate(ym = floor_date(date, "month")) %>%
  group_by(ym) %>%
  slice_tail(n = 1) %>%
  ungroup()

lvls <- MS_STATE_LEVELS

trans_counts <- monthly %>%
  mutate(
    from = factor(as.character(market_state), levels = lvls),
    to   = factor(as.character(lead(market_state)), levels = lvls)
  ) %>%
  filter(!is.na(to)) %>%
  count(from, to) %>%
  group_by(from) %>%
  mutate(prob = n / sum(n)) %>%
  ungroup()

ggplot(trans_counts, aes(x = to, y = fct_rev(from), fill = prob)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.0f%%\n(n=%d)", prob * 100, n)),
            size = 3, colour = "white", fontface = "bold") +
  scale_fill_gradient(low = "#e5e7eb", high = "#1B2A4A",
                      labels = percent_format(), name = "Probability") +
  scale_x_discrete(position = "top") +
  labs(
    title    = "Monthly transition matrix (end-of-month state)",
    subtitle = "Row = current month state | Column = next month state | Diagonal = stayed",
    x = "Next month →", y = "Current month ↓"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid   = element_blank(),
    axis.text    = element_text(face = "bold"),
    legend.position = "right"
  )
```

<p class="caption">Diagonal dominance confirms the signal has persistence — states rarely jump two levels in a single month. Most stress episodes (Deterioration → Contraction) follow an adjacent-state path.</p>

## Month-end change frequency

```{r mo-change}
mo_chg <- monthly %>%
  mutate(
    prev_state = lag(market_state),
    changed    = !is.na(prev_state) & (market_state != prev_state),
    delta_eq   = (final_eq_weight - lag(final_eq_weight)) * 100
  ) %>%
  filter(!is.na(prev_state))

chg_sum <- tibble(
  Metric = c(
    "Total months reviewed",
    "Months with state change",
    "Change frequency",
    "Avg |ΔEQ weight| on change",
    "Median |ΔEQ weight| on change",
    "Changes ≥ 5pp",
    "Changes ≥ 10pp"
  ),
  Value = c(
    nrow(mo_chg),
    sum(mo_chg$changed),
    scales::percent(mean(mo_chg$changed), accuracy = 0.1),
    paste0(round(mean(abs(mo_chg$delta_eq[mo_chg$changed]), na.rm=TRUE), 1), "pp"),
    paste0(round(median(abs(mo_chg$delta_eq[mo_chg$changed]), na.rm=TRUE), 1), "pp"),
    paste0(sum(abs(mo_chg$delta_eq[mo_chg$changed]) >= 5,  na.rm=TRUE),
           " (", round(mean(abs(mo_chg$delta_eq[mo_chg$changed]) >= 5,  na.rm=TRUE)*100), "%)"),
    paste0(sum(abs(mo_chg$delta_eq[mo_chg$changed]) >= 10, na.rm=TRUE),
           " (", round(mean(abs(mo_chg$delta_eq[mo_chg$changed]) >= 10, na.rm=TRUE)*100), "%)")
  )
)

kable(chg_sum, caption = "Month-end allocation change statistics") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, color = "#1B2A4A")
```

---

# Calibration Notes {.tabset}

<div class="doc-note">
This section records known design choices that may warrant revisiting as the data history grows.
</div>

## Thresholds

```{r calib-table}
calib <- tibble(
  Parameter = c(
    "t_fall", "roll_win", "ma_win", "mom_win",
    "Breadth: Contraction threshold",
    "Breadth: Deterioration threshold",
    "Breadth: Late-Cycle threshold",
    "Breadth: Late-Cycle (EM guard)",
    "Rotation: Strong threshold",
    "Rotation: Weak threshold",
    "EQ weight range",
    "L2 max adjustment",
    "min_persist"
  ),
  Value = c(
    paste0(params$t_fall * 100, "%"),
    paste0(params$roll_win, " days"),
    paste0(params$ma_win, " days"),
    paste0(params$mom_win, " days"),
    "< 40%", "< 55%", "< 55%", "< 70%",
    "± 4%", "± 1%",
    "25% – 75%",
    "± 5pp",
    paste0(params$min_persist, " days")
  ),
  `Set in` = c(
    "ms_vector_state.R", "ms_vector_state.R",
    "ms_vector_state.R", "ms_sector_pattern.R",
    ".classify_state()", ".classify_state()",
    ".classify_state()", ".classify_state()",
    "ms_sector_pattern.R", "ms_sector_pattern.R",
    "ms_risk_allocation.R", "ms_risk_allocation.R",
    "build_market_state_history()"
  ),
  `Rationale` = c(
    "Matches SPY DD regime engine — consistent across project",
    "~3 months: long enough for trend, short enough to adapt",
    "Standard institutional breadth convention",
    "~3 months momentum: same as standard CTA lookback",
    "Historically <40% breadth = full stress regime",
    "55% = majority of sectors breaking down",
    "55% = majority threshold; 70% = soft ceiling for healthy rally",
    "EM guard: <70% + EM lagging = late-cycle even if breadth borderline",
    "4% spread = economically meaningful cyclical leadership",
    "1% = noise floor; any score outside ±1% is directional",
    "Generous range — prevents mechanical extremes",
    "±5pp caps L2 influence at one state-step magnitude",
    "Sweet spot: eliminates 69% of noise, <18% lag cost"
  )
)

kable(calib, caption = "Framework calibration parameters") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 10) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  column_spec(2, monospace = TRUE)
```

## Known limitations

```{r limits-table}
lim <- tibble(
  `#` = 1:6,
  Limitation = c(
    "Breadth uses price level vs MA — not returns-based",
    "Sector momentum is simple mean (equal-weight across cyclicals/defensives)",
    "URTH/SPY regimes share the same t_fall threshold",
    "VWO as EM proxy is broad — includes China, which has idiosyncratic drivers",
    "No forward-looking inputs (no yield curve, credit spreads, earnings revisions)",
    "Confirmation lag is symmetric — same delay for both upgrades and downgrades"
  ),
  `Implication` = c(
    "MA crossover can lag price recovery by weeks in sharp V-rebounds",
    "XLK (largest cyclical) has outsized effective weight via market cap, not captured here",
    "In a US-only crash (spy falls, URTH holds) the Fall label fires correctly — URTH regime would stay Consolidation",
    "EM signal may fire late or early in China-specific events (e.g. regulatory crackdowns)",
    "Framework is backward-looking by construction — early-cycle regime detection may lag by 1–2 months",
    "Ideally Contraction/Deterioration entries should require fewer days than Expansion exits"
  ),
  Priority = c("Low","Low","Low","Medium","Medium","Medium")
)

kable(lim, caption = "Known limitations and implications") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, width = "2em") %>%
  column_spec(4, bold = TRUE,
              color = c("grey40","grey40","grey40","#ea580c","#ea580c","#ea580c"))
```

---

```{r footer}
cat(sprintf(
  "_Generated %s &nbsp;|&nbsp; min\\_persist = %d days &nbsp;|&nbsp; T\\_fall = %.0f%% &nbsp;|&nbsp; History: %.1f years_",
  format(Sys.Date(), "%Y-%m-%d"),
  as.integer(params$min_persist),
  params$t_fall * 100,
  n_years
))
```
