---
title: "Phase 2.2 — Overlay Backtest"
subtitle: "Type A / B / C applied to a ticker set vs 6040Classic"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
params:
  tickers:    !r c("QQQ", "SMH", "XLK", "GLD", "TLT", "DJP", "PDBC", "USMV")
  overlay_w:  0.10
  w_eq:       0.60
  w_fi:       0.40
  ma_win:     200
---

```{css, echo=FALSE}
.finding-box {
  background: #f0f9ff;
  border-left: 4px solid #2563eb;
  padding: 10px 16px;
  margin: 12px 0;
  font-size: 0.92em;
}
.type-a { background:#ede9fe; border-left:4px solid #7c3aed; padding:6px 14px; margin:6px 0; }
.type-b { background:#fef9c3; border-left:4px solid #ca8a04; padding:6px 14px; margin:6px 0; }
.type-c { background:#dcfce7; border-left:4px solid #16a34a; padding:6px 14px; margin:6px 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 14,
  fig.height = 6
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

if (!exists("project_tree"))       source(here("project_tree.R"))
if (!exists("etf_metadata"))       source(here(project_tree$scripts$init))
if (!exists("xts_ret"))            source(here(project_tree$scripts$wrangle))
if (!exists("build_regime_table")) source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
if (!exists("rt") || !is.data.frame(rt))
  rt <- build_regime_table(xts_ret[, "SPY"])

# Override params before sourcing
W_EQ           <- params$w_eq
W_FI           <- params$w_fi
OVERLAY_W      <- params$overlay_w
MA_WIN         <- as.integer(params$ma_win)
BASELINE_LABEL <- sprintf("%d/%d Classic", round(W_EQ * 100), round(W_FI * 100))

if (!exists(".build_baseline")) source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))
source(here("scripts_daa_saa_taa/09_overlay_backtest.R"))

tickers <- params$tickers

ob <- run_overlay_backtest(
  tickers        = tickers,
  xts_ret        = xts_ret,
  rt             = rt,
  w_eq           = W_EQ,
  w_fi           = W_FI,
  baseline_label = BASELINE_LABEL,
  overlay_w      = OVERLAY_W,
  ma_win         = MA_WIN,
  save_png       = FALSE
)

comp <- ob$comparison
```

---

## Framework

::: {.finding-box}
**Question answered:** For each ticker in the set, which trigger type — A, B, or C —
produces the best risk-adjusted outcome at a **`r round(OVERLAY_W * 100)`% overlay weight**
on the **`r BASELINE_LABEL`** baseline?
:::

### Baseline: `r BASELINE_LABEL`

```{r baseline-stats}
bl_stats <- purrr::map_dfr(comp$results, ~ .x$stats %>% filter(type == "Baseline")) %>%
  distinct(ann_ret, max_dd, ann_vol, ir)

tibble(
  Metric = c("Composition", "Annualised Return", "MaxDD", "Ann Vol", "IR"),
  Value  = c(
    sprintf("%.0f%% SPY + %.0f%% IEF", W_EQ * 100, W_FI * 100),
    sprintf("%+.1f%%", mean(bl_stats$ann_ret) * 100),
    sprintf("%.1f%%",  mean(bl_stats$max_dd)  * 100),
    sprintf("%.1f%%",  mean(bl_stats$ann_vol) * 100),
    sprintf("%.2f",    mean(bl_stats$ir))
  )
) %>%
  kable(align = "lr") %>%
  kable_styling(bootstrap_options = "condensed", full_width = FALSE, font_size = 12)
```

### Trigger type definitions

<div class="type-a">**Type A — Unconditional**
`(1 − w) × baseline + w × TICKER` — held always at fixed weight.
No signal, no switching. Suitable for tickers with persistent IR advantage.</div>

<div class="type-b">**Type B — SPY Regime-conditional**
Blended in Recovery + Consolidation; pure baseline in Fall.
Trigger: `build_regime_table()` → daily regime label. Lag risk: Fall fires after ~10% drop.</div>

<div class="type-c">**Type C — 200DMA signal-conditional**
Blended when TICKER is above its own `r MA_WIN`-day MA; pure baseline otherwise.
Trigger: cumulative price ≥ MA`r MA_WIN`. Whipsaw risk in choppy markets.</div>

---

## Scoreboard

Delta relative to `r BASELINE_LABEL` baseline. **Green = improvement.**

```{r scoreboard, fig.height=max(4, length(comp$results) * 0.6 + 3), echo=TRUE}
# plot_overlay_scoreboard(comp)
print(ob$p_scoreboard)
```

---

## Wealth comparison grid

One panel per ticker — `r round(OVERLAY_W * 100)`% overlay, all three trigger types vs baseline.

```{r grid, fig.height=ceiling(length(comp$results)/2)*3.8+2, echo=TRUE}
# plot_overlay_grid(comp, rt)
print(ob$p_grid)
```

---

## Per-ticker statistics table

```{r stats-table}
all_stats <- purrr::map_dfr(comp$results, ~ .x$stats) %>%
  select(ticker, type, ann_ret, max_dd, ann_vol, ir) %>%
  mutate(
    ann_ret = sprintf("%+.1f%%", ann_ret * 100),
    max_dd  = sprintf("%.1f%%",  max_dd  * 100),
    ann_vol = sprintf("%.1f%%",  ann_vol * 100),
    ir      = sprintf("%.2f",    ir)
  ) %>%
  arrange(ticker, factor(type, levels = c("Baseline", "A", "B", "C")))

all_stats %>%
  kable(
    col.names = c("Ticker", "Type", "Ann Ret", "MaxDD", "Ann Vol", "IR"),
    align = "llrrrr"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(which(all_stats$type == "Baseline"), background = "#f1f5f9", bold = TRUE) %>%
  row_spec(which(all_stats$type == "A"),        background = "#ede9fe") %>%
  row_spec(which(all_stats$type == "B"),        background = "#fef9c3") %>%
  row_spec(which(all_stats$type == "C"),        background = "#dcfce7")
```

---

## Session info

```{r session}
cat(sprintf("Report date   : %s\n", Sys.Date()))
cat(sprintf("Baseline      : %s (%.0f%% SPY + %.0f%% IEF)\n",
            BASELINE_LABEL, W_EQ * 100, W_FI * 100))
cat(sprintf("Overlay weight: %.0f%%\n", OVERLAY_W * 100))
cat(sprintf("MA window (C) : %d days\n", MA_WIN))
cat(sprintf("Tickers       : %s\n", paste(tickers, collapse = ", ")))
```
