---
title: "`r paste0('L/S Pair Factsheet — Long ', params$ticker, ' / Short ', params$bmk)`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: false
    theme: flatly
    highlight: tango
  pdf_document:
    latex_engine: xelatex
    toc: false
params:
  ticker:   "XLK"
  bmk:      "SPY"
  t_fall:   0.10
  t_cruise: 0.05
  rho_min:  0.75
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(knitr)
library(kableExtra)
library(here)

select <- dplyr::select   # PA masks dplyr::select

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

if (!exists("ticker_long_short"))
  source(here("scripts_saa_taa/taa_regime_engine.R"))

source(here("utility/factsheet_functions.R"))

# ── Regime table ──────────────────────────────────────────────────────────────
rt <- suppressWarnings(
  build_regime_table(xts_ret[, params$bmk], params$t_fall, params$t_cruise)
)

# ── Core analysis ─────────────────────────────────────────────────────────────
result <- ticker_long_short(
  ticker   = params$ticker,
  bmk      = params$bmk,
  xts_ret  = xts_ret,
  rt       = rt,
  t_fall   = params$t_fall,
  t_cruise = params$t_cruise,
  rho_min  = params$rho_min
)

s <- result$summary
r <- result$regime_tbl

# ── Colour helpers ────────────────────────────────────────────────────────────
REGIME_PAL <- c(Fall = "#D62728", Recovery = "#2CA02C", Consolidation = "#1F77B4")

conf_col <- function(x) dplyr::case_when(
  x == "High"   ~ "#1a7a3c",
  x == "Medium" ~ "#2171b5",
  x == "Low"    ~ "#fd8d3c",
  TRUE          ~ "#999999"
)

type_col <- function(x) dplyr::case_when(
  x == "ALPHA"  ~ "#1a7a3c",
  x == "HEDGE"  ~ "#2171b5",
  x == "STABLE" ~ "#6baed6",
  x == "LAG"    ~ "#fd8d3c",
  x == "LOSS"   ~ "#cb181d",
  TRUE          ~ "#999999"
)
```

---

## Section 1 — Pair Summary {.section}

```{r summary-header}
archetype_desc <- dplyr::case_when(
  s$archetype == "Cycle Amplifier"   ~ "Amplifies both SPY drawdowns and recoveries. Pure beta play.",
  s$archetype == "Recovery Sprinter" ~ "Underperforms in Fall, surges in Recovery, lags in Consolidation.",
  s$archetype == "Defensive Burden"  ~ "Cushions Fall but drags in Recovery — net opportunity cost.",
  s$archetype == "All-Regime Drag"   ~ "Underperforms in every regime. No TAA case.",
  s$archetype == "Corr-Fail"         ~ paste0("Correlation too low (ρ=", round(s$rho,3), " < ", params$rho_min, "). L/S pair structurally invalid."),
  TRUE                               ~ "No dominant cross-regime pattern detected."
)
```

```{r summary-table}
summary_df <- tibble(
  Metric = c(
    "Benchmark", "Archetype",
    "Correlation (full)", "Correlation (1y)",
    paste0(s$ticker, " Ann Return"), paste0(s$ticker, " Ann Vol"), paste0(s$ticker, " MaxDD"),
    "Spread Ann Return",  "Spread Ann Vol",  "Spread MaxDD",  "Information Ratio"
  ),
  Value = c(
    params$bmk, s$archetype,
    sprintf("%.3f", s$rho),   sprintf("%.3f", s$rho_1y),
    sprintf("%+.1f%%", s$ann_ret_ticker * 100),
    sprintf("%.1f%%",  s$ann_vol_ticker * 100),
    sprintf("%.1f%%",  s$max_dd_ticker  * 100),
    sprintf("%+.1f%%", s$ann_ret_spread * 100),
    sprintf("%.1f%%",  s$ann_vol_spread * 100),
    sprintf("%.1f%%",  s$max_dd_spread  * 100),
    sprintf("%+.2f",   s$ir)
  )
)

kable(summary_df, align = c("l","r"), col.names = c("", ""),
      caption = paste0(s$ticker, " vs ", s$bmk, " — Full history")) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(2, bold = TRUE, background = "#F0F4FA") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  column_spec(2, color = "#2C2C2C") %>%
  footnote(general = archetype_desc, general_title = "Archetype: ", footnote_as_chunk = TRUE)
```

---

## Section 2 — Regime Confidence Table {.section}

```{r regime-table}
r_display <- r %>%
  mutate(
    avg_alpha   = sprintf("%+.1f%%", avg_alpha   * 100),
    sd_alpha    = sprintf("%.1f%%",  sd_alpha    * 100),
    t_stat      = sprintf("%.2f",    t_stat),
    pct_correct = sprintf("%.0f%%",  pct_correct * 100),
    conf_score  = sprintf("%.2f",    conf_score)
  ) %>%
  select(Regime = regime, Type = modal_type, `Avg α` = avg_alpha,
         `SD α` = sd_alpha, `t-stat` = t_stat,
         `n episodes` = n_episodes, Consistency = pct_correct,
         Score = conf_score, Confidence = confidence)

conf_colours <- conf_col(r$confidence)
type_colours <- type_col(r$modal_type)

kable(r_display, align = c("l","c","r","r","r","r","r","r","c"),
      caption = "Per-regime alpha decomposition + confidence") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  column_spec(2, color = "white", background = type_colours) %>%
  column_spec(9, bold = TRUE, color = "white", background = conf_colours) %>%
  row_spec(which(r$regime == "Fall"),          background = "#FDF2F2") %>%
  row_spec(which(r$regime == "Recovery"),      background = "#F2FAF2") %>%
  row_spec(which(r$regime == "Consolidation"), background = "#F2F5FD")
```

> **Score formula:** `0.50 × min(|t|/2, 1) + 0.30 × consistency + 0.20 × min(n/6, 1)`
> **Tiers:** High ≥ 0.75 · Medium ≥ 0.50 · Low ≥ 0.30 · Weak < 0.30

---

## Section 3 — Performance Statistics {.section}

```{r pair-stats, results='hide', include=FALSE}
# Run fs_render once, capture return value — suppress raw cat/print output
fs_out <- fs_render(params$ticker, bmk = params$bmk, ret_xts = xts_ret, rt = rt)

# ── Rebuild internals needed for regime tables ─────────────────────────────
tk_r   <- as.numeric(xts_ret[, params$ticker])
bmk_r  <- as.numeric(xts_ret[, params$bmk])
common <- which(!is.na(tk_r) & !is.na(bmk_r))
tk_r   <- tk_r[common]
bmk_r  <- bmk_r[common]
dates  <- as.Date(index(xts_ret))[common]
spr_r  <- tk_r - bmk_r

# Rolling 250-day correlation
roll_cor <- as.numeric(zoo::rollapply(
  cbind(tk_r, bmk_r), width = 250,
  FUN = function(m) cor(m[,1], m[,2], use = "complete.obs"),
  by.column = FALSE, align = "right", fill = NA
))

# Rolling 250-day IR
mu_r   <- as.numeric(zoo::rollapply(spr_r, 250, mean,  align="right", fill=NA))
sd_r   <- as.numeric(zoo::rollapply(spr_r, 250, sd,    align="right", fill=NA))
roll_ir <- (mu_r / sd_r) * sqrt(252)

# Regime daily labels
regime_daily <- rt %>%
  rowwise() %>%
  mutate(date = list(seq(as.Date(xmin), as.Date(xmax), by = "day"))) %>%
  tidyr::unnest(date) %>%
  dplyr::select(date, regime) %>%
  ungroup() %>%
  mutate(regime = factor(regime, levels = c("Fall", "Recovery", "Consolidation")))

daily_df <- tibble(date = dates, roll_cor = roll_cor, roll_ir = roll_ir) %>%
  left_join(regime_daily, by = "date")

# Regime-conditional correlation
regime_cor_tbl <- daily_df %>%
  filter(!is.na(roll_cor), !is.na(regime)) %>%
  group_by(regime) %>%
  summarise(`Mean ρ` = round(mean(roll_cor), 3), .groups = "drop")

# IR by regime
regime_ir_tbl <- daily_df %>%
  filter(!is.na(roll_ir), !is.na(regime)) %>%
  group_by(regime) %>%
  summarise(
    `Mean IR`     = round(mean(roll_ir), 2),
    `% Positive`  = paste0(round(mean(roll_ir > 0) * 100, 1), "%"),
    .groups = "drop"
  )

full_ir  <- fs_out$full_ir
full_cor <- fs_out$full_cor
```

```{r stats-table-summary}
# ── 3a: Summary statistics ─────────────────────────────────────────────────
stats_df <- fs_out$stats %>%
  rename(Series = Series)

kable(stats_df,
      align   = c("l","r","r","r","r","r"),
      caption = paste0("Summary Statistics — ", params$ticker,
                       " vs ", params$bmk,
                       "  |  Common period: ",
                       format(min(dates)), " → ", format(max(dates)),
                       "  (", length(dates), " trading days)")) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(1, background = "#F0F4FA") %>%
  row_spec(3, bold = TRUE) %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

```{r stats-table-regime, fig.show='asis'}
# ── 3b: Regime-conditional correlation ────────────────────────────────────
kable(regime_cor_tbl,
      align   = c("l","r"),
      caption = sprintf("Regime-Conditional Rolling Correlation (250-day window)  |  Full-period ρ = %.3f",
                        full_cor)) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(which(regime_cor_tbl$regime == "Fall"),          background = "#FDF2F2") %>%
  row_spec(which(regime_cor_tbl$regime == "Recovery"),      background = "#F2FAF2") %>%
  row_spec(which(regime_cor_tbl$regime == "Consolidation"), background = "#F2F5FD") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

```{r dist-troika, fig.height=5}
# ── 3b.5: Three-view return distribution ──────────────────────────────────
print(fs_out$p_dist_troika)
```

```{r stats-table-ir}
# ── 3c: Rolling IR by regime ───────────────────────────────────────────────
kable(regime_ir_tbl,
      align   = c("l","r","r"),
      caption = sprintf("Rolling IR by SPY Regime (250-day window)  |  Full-period IR = %.2f",
                        full_ir)) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(which(regime_ir_tbl$regime == "Fall"),          background = "#FDF2F2") %>%
  row_spec(which(regime_ir_tbl$regime == "Recovery"),      background = "#F2FAF2") %>%
  row_spec(which(regime_ir_tbl$regime == "Consolidation"), background = "#F2F5FD") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

---

## Section 4 — 7-Panel Factsheet {.section}

*Cumulative wealth · L/S spread · Drawdown · Return distribution · Rolling ρ · Regime ρ · Rolling IR*

```{r factsheet, fig.height=14, results='hide'}
print(fs_out$combined)
```

```{r factsheet-troika, fig.height=5}
print(fs_out$p_dist_troika)
```

---

## Section 5 — Regime Overlay & Alpha Summary {.section}

*Cumulative return through regime episodes (left) · Average alpha per regime (right)*

```{r overlay-summary, fig.height=6}
plot(result, xts_ret = xts_ret, which = c("overlay", "summary"))
```

---

## Section 6 — Per-Episode Heatmap {.section}

*Each cell = one Fall/Recovery/Consolidation episode; coloured by outperformance type*

```{r heatmap, fig.height=5}
plot(result, xts_ret = xts_ret, which = "heatmap")
```

---

```{r footer}
cat(sprintf(
  "_Generated %s · %s vs %s · T_fall=%.0f%% · ρ_min=%.2f_",
  format(Sys.Date(), "%Y-%m-%d"),
  params$ticker, params$bmk,
  params$t_fall * 100, params$rho_min
))
```
