---
title: "Pair Analysis & Portfolio Construction"
subtitle: "Spread Volatility · Drawdown Theory · IR · 60/40 Decomposition"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    latex_engine: xelatex
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(MASS)
library(tidyverse)
library(xts)
library(scales)
library(patchwork)
library(here)

select <- dplyr::select

if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

spy_ret <- xts_ret[, "SPY"]
rt      <- build_regime_table(spy_ret, 0.10, 0.05)

source(here("utility/plot_xts_pair.R"))

# ── Shared helpers ────────────────────────────────────────────────────────────
T_yrs <- as.numeric(diff(range(index(xts_ret)))) / 365

magdon_mdd <- function(sigma, mu, T) {
  if (mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  (sigma / S) * (log(S * sqrt(T)) + 0.35)
}

pair_stats <- function(out) {
  s  <- out$stats
  sp <- s %>% filter(str_detect(Series, "−"))
  mu  <- parse_number(sp$`Ann Return`) / 100
  sig <- parse_number(sp$`Ann Vol`)    / 100
  mdd <- parse_number(sp$`Max DD`)     / 100
  list(
    mu       = mu,
    sig      = sig,
    mdd      = mdd,
    ir       = out$full_ir,
    rho      = out$full_cor,
    te_10pct = 0.10 * sig,
    mdd_10pct= 0.10 * abs(mdd),
    ret_10pct= 0.10 * mu,
    mdd_theory = magdon_mdd(sig, mu, T_yrs),
    stats    = s
  )
}

REGIME_COLORS <- c(Fall = "#D90429", Recovery = "#F77F00",
                   Consolidation = "#2D6A4F")
```

---

## 1. The Portfolio Decomposition Identity

Every active allocation decision relative to a benchmark decomposes as:

$$\boxed{\text{New PF} = \text{Benchmark} + \delta \times (r_{\text{new}} - r_{\text{funded from}})}$$

The **funded-from asset** determines the nature of the active bet:

| Funding source | Active bet type | Typical ρ | Spread vol | TE per 10% |
|---|---|---|---|---|
| Within equity (SPY) | Sector vs market | High (~0.85) | Low (~8–12%) | Small (~1%) |
| From bonds (AGG) | Equity vs fixed income | Negative (~−0.20) | High (~20–25%) | Large (~2.5%) |

We examine **five concrete decisions** against a 60% SPY / 40% AGG benchmark,
each at 10% notional.

```{r run-all-pairs, cache=TRUE, include=FALSE}
# ── Run all pairs ─────────────────────────────────────────────────────────────
pairs_list <- list(
  list(t1="XLK", t2="SPY", label="XLK vs SPY", type="Equity swap",
       decision="+10% XLK / −10% SPY"),
  list(t1="XLV", t2="SPY", label="XLV vs SPY", type="Equity swap",
       decision="+10% XLV / −10% SPY"),
  list(t1="XLF", t2="SPY", label="XLF vs SPY", type="Equity swap",
       decision="+10% XLF / −10% SPY"),
  list(t1="XLP", t2="AGG", label="XLP vs AGG", type="Cross-asset",
       decision="+10% XLP / −10% AGG"),
  list(t1="XLV", t2="AGG", label="XLV vs AGG", type="Cross-asset",
       decision="+10% XLV / −10% AGG")
)

results <- map(pairs_list, function(p) {
  out <- plot_xts_pair(
    xts_ret[, p$t1], xts_ret[, p$t2],
    label1 = p$t1, label2 = p$t2,
    rt     = rt
  )
  c(p, list(out = out, ps = pair_stats(out)))
})
```

---

## 2. Equity Sleeve Swaps — Sector vs SPY

These decisions keep total equity at 60% but replace part of the SPY
passive allocation with an active sector bet.

> **Decomposition:** 60% SPY + 40% AGG → 50% SPY + 10% **sector** + 40% AGG
> = 60/40 benchmark + 10% × (sector − SPY)

### 2.1 XLK +10% / SPY −10%  *(Technology)*

```{r xlk-spy, fig.height=35}
print(results[[1]]$out$combined)
```

### 2.2 XLV +10% / SPY −10%  *(Health Care)*

```{r xlv-spy, fig.height=35}
print(results[[2]]$out$combined)
```

### 2.3 XLF +10% / SPY −10%  *(Financials)*

```{r xlf-spy, fig.height=35}
print(results[[3]]$out$combined)
```

### 2.4 Equity Swap Comparison

```{r equity-comparison}
map_dfr(results[1:3], function(r) {
  ps <- r$ps
  tibble(
    Decision       = r$decision,
    `Sector`       = r$t1,
    `ρ (vs SPY)`   = round(ps$rho, 2),
    `Spread Vol`   = percent(ps$sig,      accuracy = 0.1),
    `Spread IR`    = round(ps$ir,         2),
    `E[Spread]`    = percent(ps$mu,       accuracy = 0.1),
    `Spread MaxDD` = percent(ps$mdd,      accuracy = 0.1),
    `TE added`     = percent(ps$te_10pct, accuracy = 0.01),
    `Active ret`   = percent(ps$ret_10pct,accuracy = 0.01),
    `Active MaxDD` = percent(ps$mdd_10pct,accuracy = 0.01)
  )
}) %>%
  knitr::kable(caption = "Equity sleeve swaps — 10% notional each")
```

**Reading the table:**

- **ρ** is the full-period rolling correlation between the sector and SPY.
  All three are high (> 0.80), confirming these are low-spread-vol bets.
- **TE added** is the annualised tracking error added to the 60/40 portfolio
  at 10% notional — all under 1.5%, a very small active risk budget.
- **IR** ranks the quality of each active bet. An IR > 0.50 is top-quartile.

---

## 3. Cross-Asset Additions — Equity funded from Bonds

These decisions increase total equity from 60% to 70%, funded by reducing
bonds from 40% to 30%.

> **Decomposition:** 60% SPY + 40% AGG → 60% SPY + 10% **sector** + 30% AGG
> = 60/40 benchmark + 10% × (sector − AGG)

The **active bet is now sector vs bonds** — a cross-asset spread where
equity and bonds are typically negatively correlated. This has a very
different risk profile from the equity swaps above.

### 3.1 XLP +10% / AGG −10%  *(Consumer Staples — defensive equity)*

```{r xlp-agg, fig.height=35}
print(results[[4]]$out$combined)
```

### 3.2 XLV +10% / AGG −10%  *(Health Care — vs bonds)*

```{r xlv-agg, fig.height=35}
print(results[[5]]$out$combined)
```

### 3.3 Cross-Asset Comparison

```{r cross-comparison}
map_dfr(results[4:5], function(r) {
  ps <- r$ps
  tibble(
    Decision       = r$decision,
    `Sector`       = r$t1,
    `ρ (vs AGG)`   = round(ps$rho, 2),
    `Spread Vol`   = percent(ps$sig,      accuracy = 0.1),
    `Spread IR`    = round(ps$ir,         2),
    `E[Spread]`    = percent(ps$mu,       accuracy = 0.1),
    `Spread MaxDD` = percent(ps$mdd,      accuracy = 0.1),
    `TE added`     = percent(ps$te_10pct, accuracy = 0.01),
    `Active ret`   = percent(ps$ret_10pct,accuracy = 0.01),
    `Active MaxDD` = percent(ps$mdd_10pct,accuracy = 0.01)
  )
}) %>%
  knitr::kable(caption = "Cross-asset additions — 10% notional each, funded from AGG")
```

---

## 4. Master Comparison — All Five Decisions

```{r master-table}
map_dfr(results, function(r) {
  ps <- r$ps
  tibble(
    Decision       = r$decision,
    Type           = r$type,
    `ρ`            = round(ps$rho,        2),
    `Spread Vol`   = percent(ps$sig,      accuracy = 0.1),
    `Spread IR`    = round(ps$ir,         2),
    `E[Spread]`    = percent(ps$mu,       accuracy = 0.1),
    `Spread MaxDD` = percent(ps$mdd,      accuracy = 0.1),
    `TE (10%)`     = percent(ps$te_10pct, accuracy = 0.01),
    `Active ret`   = percent(ps$ret_10pct,accuracy = 0.01)
  )
}) %>%
  arrange(Type, desc(`Spread IR`)) %>%
  knitr::kable(caption = "All five decisions ranked by type — 10% notional")
```

### 4.1 Spread Volatility Visualised

```{r spread-vol-bar, fig.height=5}
map_dfr(results, function(r) {
  ps <- r$ps
  tibble(
    decision     = r$decision,
    type         = r$type,
    spread_vol   = ps$sig,
    te           = ps$te_10pct,
    ir           = ps$ir
  )
}) %>%
  mutate(decision = fct_reorder(decision, spread_vol)) %>%
  ggplot(aes(x = decision, y = spread_vol, fill = type)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_text(aes(label = sprintf("IR=%.2f", ir)),
            hjust = -0.15, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("Equity swap"  = "#457B9D",
                                "Cross-asset"  = "#D90429")) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Spread Volatility by Decision (10% notional)",
       subtitle = "Blue = funded from SPY (equity swap)  |  Red = funded from AGG (cross-asset)\nIR labels show full-period Information Ratio",
       x = NULL, y = "Annualised Spread Vol", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "bottom",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

### 4.2 Tracking Error Added to 60/40

```{r te-bar, fig.height=5}
map_dfr(results, function(r) {
  ps <- r$ps
  tibble(
    decision = r$decision,
    type     = r$type,
    te       = ps$te_10pct,
    ret      = ps$ret_10pct,
    ir       = ps$ir
  )
}) %>%
  mutate(decision = fct_reorder(decision, te)) %>%
  pivot_longer(c(te, ret), names_to = "metric", values_to = "value") %>%
  mutate(metric = if_else(metric == "te", "TE added", "Active return")) %>%
  ggplot(aes(x = decision, y = value, fill = interaction(type, metric))) +
  geom_col(position = "dodge", width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = 0, colour = "grey30") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "Equity swap.Active return" = "#2D6A4F",
      "Equity swap.TE added"      = "#457B9D",
      "Cross-asset.Active return" = "#F4A261",
      "Cross-asset.TE added"      = "#D90429"
    ),
    labels = c("Equity swap — active return", "Equity swap — TE",
               "Cross-asset — active return", "Cross-asset — TE"),
    name = NULL
  ) +
  scale_y_continuous(labels = percent_format(accuracy = 0.01)) +
  labs(title    = "Active Return vs Tracking Error at 10% Notional",
       subtitle = "A good decision has active return > TE  |  i.e. IR > 1 at portfolio level",
       x = NULL, y = "Annualised %") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "bottom",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

### 4.3 Magdon-Ismail MaxDD — Theory vs Actual

```{r mdd-table}
map_dfr(results, function(r) {
  ps <- r$ps
  tibble(
    Decision        = r$decision,
    Type            = r$type,
    `E[Spread]`     = percent(ps$mu,          accuracy = 0.1),
    `Spread Vol`    = percent(ps$sig,          accuracy = 0.1),
    Sharpe          = round(ps$mu / ps$sig,   2),
    `MaxDD theory`  = percent(ps$mdd_theory,  accuracy = 0.1),
    `MaxDD actual`  = percent(abs(ps$mdd),    accuracy = 0.1),
    `Calmar actual` = round(ps$mu / abs(ps$mdd), 2)
  )
}) %>%
  knitr::kable(caption = "Magdon-Ismail E[MaxDD] vs realised — Sharpe² ≈ Calmar")
```

---

## 5. Spread Volatility Theory — Revisited

### 5.1 The Variance Formula

$$\sigma^2_{L/S} = \sigma^2_1 + \sigma^2_2 - 2\rho\,\sigma_1\sigma_2$$

The **funding asset's correlation with the new asset** is the single
most important driver of spread risk:

```{r rho-grid, fig.height=5}
# Illustrate σ_spread(ρ) for representative equity vs SPY and equity vs bonds
rho_seq <- seq(-1, 1, length.out = 300)

expand_grid(
  scenario = c("Equity sector vs SPY  (σ₁=20%, σ₂=17%)",
               "Equity sector vs AGG  (σ₁=20%, σ₂=5%)"),
  rho = rho_seq
) %>%
  mutate(
    s1 = if_else(str_detect(scenario, "SPY"), 0.20, 0.20),
    s2 = if_else(str_detect(scenario, "SPY"), 0.17, 0.05),
    sig_spread = sqrt(s1^2 + s2^2 - 2 * rho * s1 * s2)
  ) %>%
  ggplot(aes(x = rho, y = sig_spread, colour = scenario)) +
  geom_line(linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  annotate("text", x = -0.98, y = 0.37, hjust = 0, size = 3, colour = "grey40",
           label = "← ρ negative: risks compound") +
  annotate("text", x =  0.02, y = 0.37, hjust = 0, size = 3, colour = "grey40",
           label = "ρ positive: spread vol shrinks →") +
  scale_colour_manual(values = c("#457B9D", "#D90429"), name = NULL) +
  scale_x_continuous(breaks = seq(-1, 1, 0.25)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title    = "Spread Vol as a Function of Correlation",
       subtitle = "For L/S pairs: high ρ collapses spread vol → small TE\nFor cross-asset: negative ρ explodes spread vol → large TE",
       x = "Correlation ρ", y = "Annualised Spread Vol") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        plot.title       = element_text(face = "bold", size = 13),
        plot.subtitle    = element_text(colour = "grey50", size = 9))
```

### 5.2 Implied Correlation from Realised Data

```{r implied-rho}
map_dfr(results, function(r) {
  ps  <- r$ps
  s1  <- parse_number(r$out$stats %>%
           filter(Series == r$t1) %>% pull(`Ann Vol`)) / 100
  s2  <- parse_number(r$out$stats %>%
           filter(Series == r$t2) %>% pull(`Ann Vol`)) / 100
  sig_sp <- ps$sig
  rho_imp <- (s1^2 + s2^2 - sig_sp^2) / (2 * s1 * s2)
  tibble(
    Decision      = r$decision,
    `σ (asset 1)` = percent(s1,      accuracy = 0.1),
    `σ (asset 2)` = percent(s2,      accuracy = 0.1),
    `σ (spread)`  = percent(sig_sp,  accuracy = 0.1),
    `ρ implied`   = round(rho_imp,   2),
    `ρ rolling`   = round(ps$rho,    2)
  )
}) %>%
  knitr::kable(caption = "Implied correlation from σ_spread = √(σ₁²+σ₂²−2ρσ₁σ₂)")
```

---

## 6. MaxDD Theory

### 6.1 From Variance to Drawdown

Drawdown depends on the **path** of returns, not just variance.
For GBM with drift μ and vol σ over T years, Magdon-Ismail (2004) gives:

$$E[\text{MaxDD}] \approx \frac{\sigma}{S}\left[\ln(S\sqrt{T}) + 0.35\right]$$

where $S = \mu/\sigma$ is the Sharpe ratio of the spread.
This implies the **Sharpe ↔ Calmar identity:**

$$\text{Calmar} \approx S^2 \quad\Longleftrightarrow\quad
  \text{MaxDD} \approx \frac{\sigma^2}{\mu} = \frac{\sigma}{S}$$

Higher Sharpe → smaller MaxDD per unit of vol.

### 6.2 Sensitivity to drift

The log term makes MaxDD surprisingly insensitive to drift once μ > 0:

```{r drift-sensitivity, fig.height=5}
drift_grid <- seq(0.02, 0.30, length.out = 200)

tibble(
  mu    = drift_grid,
  GLD   = map_dbl(mu, ~ magdon_mdd(0.142, .x, T_yrs)),
  XLK   = map_dbl(mu, ~ magdon_mdd(0.229, .x, T_yrs))
) %>%
  pivot_longer(-mu, names_to = "asset", values_to = "mdd") %>%
  ggplot(aes(x = mu, y = mdd, colour = asset)) +
  geom_line(linewidth = 1.0) +
  geom_point(
    data = map_dfr(results[c(1,4)], function(r) {
      ps <- r$ps; tibble(asset = r$t1, mu = ps$mu, mdd = ps$mdd_theory)
    }),
    size = 3, shape = 21, fill = "white", stroke = 1.5
  ) +
  scale_colour_manual(values = c(GLD = "#2D6A4F", XLK = "#457B9D"),
                      name = NULL) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title    = "E[MaxDD] vs Ann Drift — Magdon-Ismail",
       subtitle = "Circles = realised drift from the data  |  MaxDD compresses slowly with drift (log scale)",
       x = "Annualised drift μ", y = "E[MaxDD]") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        plot.title       = element_text(face = "bold", size = 13),
        plot.subtitle    = element_text(colour = "grey50", size = 9))
```

---

## 7. Information Ratio

$$IR = \frac{E[r_1 - r_2]}{\sigma(r_1 - r_2)} \qquad
  P(\text{correct}) = 0.5 + \frac{IR}{2}$$

```{r ir-summary, fig.height=5}
ir_df <- map_dfr(results, function(r) {
  ps <- r$ps
  tibble(
    decision  = r$decision,
    type      = r$type,
    ir        = ps$ir,
    p_correct = 0.5 + ps$ir / 2,
    hurdle_ir_50 = 0.50 * ps$sig   # return needed for IR=0.5
  )
})

ir_df %>%
  mutate(decision = fct_reorder(decision, ir)) %>%
  ggplot(aes(x = decision, y = ir, fill = type)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = 0.50, linetype = "dashed",
             colour = "#2D6A4F", linewidth = 0.8) +
  geom_hline(yintercept = 0.30, linetype = "dotted",
             colour = "grey50", linewidth = 0.6) +
  annotate("text", x = 0.6, y = 0.52, hjust = 0, size = 3,
           colour = "#2D6A4F", label = "IR = 0.50 (top quartile)") +
  annotate("text", x = 0.6, y = 0.32, hjust = 0, size = 3,
           colour = "grey50",   label = "IR = 0.30 (median active)") +
  geom_text(aes(label = sprintf("P=%.0f%%", p_correct * 100)),
            hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("Equity swap" = "#457B9D",
                                "Cross-asset" = "#D90429")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  labs(title    = "Information Ratio by Decision",
       subtitle = "P labels = probability forecast must be correct to justify the bet\nDashed = top-quartile threshold (IR=0.50)",
       x = NULL, y = "IR", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "bottom",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

```{r ir-table}
ir_df %>%
  arrange(type, desc(ir)) %>%
  mutate(
    ir        = round(ir, 2),
    p_correct = percent(p_correct, accuracy = 0.1),
    hurdle_ir_50 = percent(hurdle_ir_50, accuracy = 0.1)
  ) %>%
  rename(Decision = decision, Type = type, IR = ir,
         `P(correct)` = p_correct,
         `E[spread] needed for IR=0.50` = hurdle_ir_50) %>%
  knitr::kable(caption = "IR summary — all five decisions")
```

---

## 8. Summary of Key Principles

```{r principles}
tibble(
  `#` = 1:8,
  Principle = c(
    "Every active allocation = benchmark + L/S overlay on (new asset − funded asset)",
    "Equity sleeve swaps (sector vs SPY): high ρ → low spread vol → small TE → low hurdle return",
    "Cross-asset additions (equity vs bonds): negative ρ → spread vol explodes → large TE → high hurdle return",
    "For L/S pairs: high ρ is desirable — it collapses spread vol to |σ₁−σ₂|",
    "For long-long portfolios: low ρ is desirable — it reduces combined variance",
    "MaxDD ≈ σ/S × (ln(S√T) + 0.35) — Sharpe² ≈ Calmar — higher Sharpe means smaller drawdown per unit of vol",
    "IR > 0.50 = top quartile; requires P(forecast correct) > 75% — a high bar for sustained calls",
    "Rolling IR by SPY regime reveals whether the active bet earns in stress — critical check for 60/40 tilts"
  )
) %>% knitr::kable(align = "cl")
```

---

*Benchmark: 60% SPY / 40% AGG. All decisions at 10% notional.
Data: `xts_ret_returns.rds` | SPY regimes: Fall ≥10% drawdown from peak.*
