---
title: "Rebalancing Drift — SPY/IEF 60/40"
subtitle: "A Pedagogical Guide to Volatility Harvesting, Drift, and Optimal Rebalancing Frequency"
author: "ETF Intelligence Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 14,
  fig.height = 20
)
library(here)
source(here("project_tree.R"))
source(here("00_libraries.R"))
source(here(project_tree$scripts$spy_dd_regime))   # defines build_regime_table()
source(here(project_tree$scripts$wrangle))
rt <- build_regime_table(xts_ret[, "SPY"])

# Source the full analysis script — builds df, panels, fig_* objects.
# Print guards (knitr.in.progress) suppress automatic output.
source(here("scripts_daa_saa_taa/plot_rebal_drift.R"), local = knitr::knit_global())
```

---

# 1  What Is Rebalancing Drift?

A **60/40 portfolio** holds 60% SPY (US equity) and 40% IEF (US 7–10Y Treasury).
Every portfolio management decision reduces to one question: *how often do you
reset that ratio?*

Three natural answers:

| Strategy | Rule | Weight resets? |
|---|---|---|
| **Buy & Hold (BaH)** | Invest once, never touch | Never — weights drift freely |
| **Annual Rebal** | Reset to 60/40 on 1 January each year | Once per year |
| **Daily Rebal** | Reset to 60/40 at the end of every trading day | Every day |

**Drift** is the daily return difference between a passive strategy and daily
rebalancing:

$$\text{Drift}_t = r_t^{\text{strategy}} - r_t^{\text{daily rebal}}$$

Cumulative drift tells you whether *not* rebalancing daily has been a net win
or loss over any horizon.

---

# 2  The Paradox: Why Does Daily Rebal Win Despite SPY's Outperformance?

## 2.1  The Seeming Contradiction

SPY compounded at roughly **10–11% per year** over the full history.
IEF compounded at roughly **3–4%**.
Buy & Hold lets the winning asset run — so BaH should accumulate more SPY
exposure and therefore more wealth, right?

**Wrong.** Daily rebalancing — which *sells SPY every day it rises* — beats
Buy & Hold in total wealth over the full history.

## 2.2  The Intuition: Volatility Harvesting

SPY outperforms IEF in **trend** (direction).
But SPY is also far more **volatile** (~15% annual vol vs ~6% for IEF).
Daily rebalancing exploits that **volatility**, not the trend.

> The rebalancing bonus is a *variance tax refund*: you sell the asset that
> has become more expensive relative to its long-run value and buy the cheaper
> one, collecting the spread between arithmetic and geometric returns.

## 2.3  The Mathematics

For any two-asset portfolio rebalanced continuously, the rebalancing bonus per
period is approximately:

$$\text{bonus} \approx \tfrac{1}{2} \times w_{\text{SPY}} \times w_{\text{IEF}}
\times \sigma^2(\text{SPY} - \text{IEF})$$

$$\approx \tfrac{1}{2} \times 0.60 \times 0.40 \times \text{var(daily spread)}$$

**Key properties:**

1. **Always positive** — as long as the two assets have any spread variance,
   the bonus is strictly positive regardless of which asset wins long-term.
2. **A free lunch from mean reversion** — relative prices mean-revert daily
   even while trending overall.
3. **Maximised at equal weights** — the product $w \times (1-w)$ peaks at 0.5.

The flip side is the **volatility drag** formula (Itô's lemma for log returns):

$$g \approx \mu - \tfrac{1}{2} \sigma^2$$

where $g$ is the geometric (compound) return, $\mu$ is the arithmetic return,
and $\sigma^2$ is variance.

Buy & Hold allows portfolio variance to grow as SPY weight drifts to ~80%.
Daily rebalancing keeps variance fixed at the target allocation level, so its
geometric compounding is more efficient *even though* its arithmetic return is
slightly lower.

## 2.4  Why BaH Loses Despite Riding SPY's Trend

| | Buy & Hold | Daily Rebal |
|---|---|---|
| Captures SPY trend | More | Less |
| Portfolio volatility | Grows to ~80% equity | Fixed at 60% |
| Volatility drag | Increasing over time | Stable |
| Fall entry weight | ~80% SPY | 60% SPY |
| **Net: geometric return** | **Lower** | **Higher** |

**The rebalancing bonus from harvesting SPY/IEF spread volatility > the
momentum cost of selling the winner.**

## 2.5  The Regime Dimension

It also matters *when* SPY outperforms.
BaH is maximally overweight SPY entering every **Fall regime** — exactly the
wrong time.
Daily rebalancing systematically trimmed equity on every up-day before the Fall,
so it always enters Falls at exactly 60%.

That asymmetry — *trim on the way up, buy on the way down* — compresses
drawdowns and lifts the geometric mean above BaH's.

---

# 3  What Drift Measures

## 3.1  Definition

$$\text{drift}_t = r_t^{\text{BaH}} - r_t^{\text{daily rebal}}$$

Each day:

- If drift > 0: carrying extra SPY exposure paid off (SPY beat IEF that day)
- If drift < 0: carrying extra SPY exposure hurt (IEF beat SPY that day)

Cumulative drift is the running total.

## 3.2  The Paradox Resolved: Why Cumulative Drift Is Negative When SPY Weight Reaches 80%

This is the key insight that trips most readers.

**Panel M6** shows BaH SPY weight drifting to ~80% during bull runs.
**Panel M5** shows cumulative drift staying negative.

These are not contradictory — they measure different things:

- **Panel M6** asks: *how much equity exposure has BaH accumulated?*
  → SPY drifted from 60% → 80% because equities outperformed. BaH was holding
  more equity — and was *richer* during the bull run.

- **Panel M5** asks: *was carrying that extra equity exposure profitable vs
  daily rebalancing?*
  → Drift = `(w_bah − 0.60) × (spy_r − ief_r)` each day.
  → During **Falls**, `w_bah − 0.60` is large and positive (~+20pp), but
  `spy_r − ief_r` is large and **negative** (SPY collapses, IEF rallies).
  → The product is a large negative spike.

The bull-run gains from drift (positive, gradual) are overwhelmed by the Fall
losses (negative, sudden, amplified by maximum overweight at the worst moment).

**Net result:** the rebalancing bonus (daily rebal advantage) exceeds the
momentum premium (BaH advantage) over the full history.

---

# 4  Overview: All Five Strategies

```{r fig-main, fig.height=26}
print(main_view)
```

**Reading the table (bottom strip):**

- `Final $1` — terminal wealth from $1 invested
- `Ann Ret` — compound annual growth rate
- `Ann Vol` — annualised daily volatility
- `Max DD` — peak-to-trough drawdown
- `Sharpe` — Ann Ret ÷ Ann Vol (no risk-free rate)
- `Calmar` — Ann Ret ÷ |Max DD| (return per unit of worst drawdown)

Green = best in metric | Yellow = middle | Red = worst.

---

# 5  Pedagogical Breakdown (Acts 1–4)

The appendix below traces the full causal chain across four acts.

```{r fig-appendix, fig.height=26}
print(appendix)
```

| Panel | Act | What it shows |
|---|---|---|
| A1 | Building blocks | SPY vs IEF individual returns |
| A2 | Building blocks | BaH SPY weight drifts to ~80% |
| A3 | Mechanism | Daily SPY − IEF spread — bad days cluster in Falls |
| A4 | Mechanism | Daily drift = (excess weight) × spread — Fall spikes |
| A5 | Consequence | Cumulative drift stays negative |
| A6 | Consequence | Daily rebal wins in cumulative wealth |
| A7 | When it happened | Annual drift cost: 2008 and 2022 the outliers |
| A8 | When it happened | Rolling 12m drift turns negative at Fall onset |

---

# 6  Pairwise Comparisons: Each Strategy vs Buy & Hold

Each figure below isolates one strategy against the Buy & Hold baseline.
The drift panels (right column) now measure **strategy − BaH** rather than
strategy − daily rebal.

**Positive drift** = strategy beats BaH on that day.
**Negative drift** = BaH beats strategy on that day.

## 6.1  Daily Rebal vs Buy & Hold

Daily rebalancing always enters Falls at exactly 60% equity — the minimum
possible overweight relative to BaH.
The SPY weight panel (D6) shows a **flat line** at 60%: the portfolio has no
drift by construction.

```{r fig-D, fig.height=26}
print(fig_D)
```

## 6.2  Quarterly Rebal vs Buy & Hold

Quarterly rebalancing resets four times per year — enough to prevent the most
extreme equity overweight while still capturing several months of momentum.
The SPY weight panel (Q6) shows a sawtooth pattern with amplitude roughly half
of BaH's drift.

```{r fig-Q, fig.height=26}
print(fig_Q)
```

## 6.3  Annual Rebal vs Buy & Hold

Annual rebalancing resets once per year on 1 January.
If a Fall occurs late in the year the portfolio may have drifted significantly
before the reset.
The SPY weight panel (AR6) shows the characteristic sawtooth: drift builds
through the year, snaps back at year-end.

```{r fig-AR, fig.height=26}
print(fig_AR)
```

## 6.4  Threshold Rebal ±10% vs Buy & Hold

Threshold rebalancing is **event-driven**: rebalance only when SPY weight
drifts outside [50%, 70%].
Unlike calendar strategies, it rebalances more during volatile periods (when
the bonus is largest) and less during calm trends (letting momentum run).
This is implicitly **regime-aware** without knowing the regime.

```{r fig-TR10, fig.height=26}
print(fig_TR10)
```

## 6.5  Threshold Rebal ±5% vs Buy & Hold

A tighter ±5% band [55%, 65%] triggers rebalancing more frequently.
It is closer to daily rebalancing in behaviour during volatile periods and
still captures some momentum during calm trends.
Compare TR5 and TR10 SPY weight panels to see how the band width controls
the degree of drift allowed.

```{r fig-TR5, fig.height=26}
print(fig_TR5)
```

---

# 7  Optimal Rebalancing Frequency

There is no single optimal frequency — it depends on three competing forces:

| Force | Direction | Explanation |
|---|---|---|
| Volatility harvesting | ↑ Higher frequency | More rebalancing = more bonus harvested |
| Momentum capture | ↓ Lower frequency | Selling the winner before it finishes rising has a cost |
| Transaction costs | ↓ Lower frequency | Bid-ask spread × number of trades |

**Theoretical sweet spot:**

$$\text{optimal frequency} \propto \frac{\sigma^2_{\text{spread}}}{\text{cost per trade}}$$

For low-cost ETFs (SPY, IEF), transaction costs are near zero today, shifting
the optimum toward higher frequency.
Empirically, **monthly to quarterly** rebalancing captures ~80–90% of the daily
bonus at ~1/20th of the trading activity.

**Threshold strategies are superior to calendar strategies** because they
rebalance *when the bonus is largest* (volatile periods = wide spread variance)
rather than on a fixed calendar regardless of market conditions.
They are effectively a **free regime signal**: the threshold fires exactly when
equity has been running hot (before a potential Fall) or has fallen hard
(Recovery buying opportunity).

---

# 8  Key Findings

1. **Daily rebalancing outperforms Buy & Hold in total wealth** over the full
   SPY/IEF history — not because it captures more trend, but because it
   harvests the spread variance via sell-high/buy-low every day.

2. **Drift = (excess SPY weight) × (SPY − IEF spread).** It is negative in
   aggregate because Fall-regime spikes (large overweight × large negative
   spread) outweigh bull-run gains (moderate overweight × moderate positive
   spread).

3. **The rebalancing bonus formula** $\approx \tfrac{1}{2} w(1-w)\sigma^2$
   is always positive, always a free lunch from variance, and always maximised
   at equal weights.

4. **BaH enters every Fall maximally overweight equity** (~80% vs 60% target)
   — the worst possible entry for a drawdown. Daily rebalancing always enters
   Falls at exactly the target.

5. **Threshold rebalancing dominates calendar rebalancing**: same volatility
   harvesting benefit, less unnecessary trading in calm markets, implicit
   regime timing.

6. **Optimal real-world frequency** for low-cost ETFs: monthly or quarterly
   calendar, or ±5% threshold — each captures most of the daily bonus.

---

# 9  `rebal_profile` — Choosing Rebalancing Frequency for Any Pair

The preceding sections studied five specific strategies for SPY/IEF 60/40.
`rebal_profile()` generalises that analysis to **any two-ticker portfolio**:
given a pair and target weights, it runs all three calendar-rebalancing
frequencies (Quarterly / Semi-Annual / Annual) plus the Buy & Hold baseline
and returns a full **12-metric characteristic vector** for each strategy.

**File:** `scripts_daa_saa_taa/rebal_profile.R`

**Three exported functions:**

| Function | Purpose |
|---|---|
| `rebal_profile(ticker1, ticker2, w1, ...)` | Runs the four strategies, returns 4 × 12 tibble |
| `print_rebal_profile(x)` | Formatted console table with ★ best-in-column marker |
| `best_rebal_freq(x, by)` | Returns the strategy name that wins on a named metric |

---

## 9.1  `rebal_profile()` — full signature

```r
source(here("scripts_daa_saa_taa/rebal_profile.R"))

rebal_profile(
  ticker1,                                       # character — must be in xts_ret
  ticker2,                                       # character — must be in xts_ret
  w1,                                            # numeric — target weight for ticker1
  w2      = 1 - w1,                              # numeric — complement (default)
  xts_ret = get("xts_ret", envir=parent.frame()) # xts of daily log returns
                                                 # auto-resolved from calling env
  rf      = 0                                    # annual risk-free rate (decimal)
)                                                # e.g. rf = 0.04 for 4% p.a.
```

**Returns:** a `tibble` with 4 rows (Quarterly / Semi-Annual / Annual / Buy & Hold)
and 12 metric columns, plus 6 attributes:

| Attribute | Content |
|---|---|
| `pair` | `"SPY/IEF"` |
| `weights` | `"60/40"` |
| `n_days` | number of overlapping trading days |
| `from` | first date in common history |
| `to` | last date in common history |
| `rf` | annual risk-free rate used |

### How it works internally

**Period boundary detection** uses integer keys — no string comparisons, no
`lubridate` dependency:

```r
yr  <- as.integer(format(dates, "%Y"))
mo  <- as.integer(format(dates, "%m"))
qtr <- yr * 4L + (mo - 1L) %/% 3L   # changes value at each quarter boundary
sem <- yr * 2L + (mo - 1L) %/% 6L   # changes value at each semi-annual boundary
```

A rebalance fires on day `i` when `qtr[i] != qtr[i-1]` (or `sem`, `yr` for the
other frequencies).  Buy & Hold never fires (`do_rebal = FALSE` always).

**Rebalancing mechanics** are identical to `plot_rebal_drift.R`:
each day the two weight buckets grow with their asset's gross return;
on a rebalancing day the combined total is redistributed to `w1` / `w2`.

---

## 9.2  The 12 metrics — formulas and interpretation

The metrics are grouped into four natural families:

### Return metrics

| Metric | Formula | What it tells you |
|---|---|---|
| `Final $1` | $\prod(1+r_t)$ — terminal value of \$1 | Absolute wealth creation |
| `Ann Ret` | $(V_T / V_0)^{252/T} - 1$ — CAGR | Compound annual growth rate; log-scale equivalent of wealth |

### Risk metrics

| Metric | Formula | What it tells you |
|---|---|---|
| `Ann Vol` | $\hat{\sigma}_{\text{daily}} \times \sqrt{252}$ | Total volatility — penalises both upside and downside swings |
| `Max DD` | $\min_t \left(\frac{V_t - \max_{s \le t} V_s}{\max_{s \le t} V_s}\right)$ | Worst peak-to-trough loss; the single number most relevant to investor psychology |

### Risk-adjusted return metrics

| Metric | Formula | Notes |
|---|---|---|
| `Sharpe` | $\frac{\text{Ann Ret} - r_f}{\text{Ann Vol}}$ | Standard; penalises all variance equally |
| `Sortino` | $\frac{\text{Ann Ret} - r_f}{\sigma_{\text{down}}}$ | Downside semi-deviation only: $\sigma_{\text{down}} = \sqrt{\text{mean}((r_t - r_{f,d})^2 \;\text{for}\; r_t < r_{f,d})} \times \sqrt{252}$. Better for non-normal return distributions — rebalancing strategies tend to be right-skewed |
| `Calmar` | $\frac{\text{Ann Ret}}{|\text{Max DD}|}$ | Return per unit of worst historical loss; most relevant to drawdown-sensitive mandates |
| `Omega` | $\frac{\sum \max(r_t - r_{f,d},\, 0)}{\sum \max(r_{f,d} - r_t,\, 0)}$ | Full distribution ratio (Keating & Shadwick). Values >1 mean gains outweigh losses relative to threshold. No distributional assumption |

### Tail risk metrics

| Metric | Formula | What it tells you |
|---|---|---|
| `VaR 5%` | $Q_{0.05}(r_t)$ — 5th percentile daily return | The daily loss threshold exceeded only 5% of days |
| `CVaR 5%` | $E[r_t \mid r_t \le \text{VaR 5\%}]$ — expected shortfall | Average loss on the worst 5% of days; more sensitive than VaR to extreme tail events |

### Trading-cost metrics

| Metric | Formula | What it tells you |
|---|---|---|
| `Rebal N` | Count of calendar rebalancing events | Direct measure of trading frequency and therefore transaction cost exposure |
| `Avg Slip` | $\text{mean}(|r_{i-1}|)$ for all rebalancing days $i$ | Mean absolute return on the day *before* each rebalance — the last day of unchecked drift. High values indicate the portfolio arrives at rebalancing day with a large position bias, implying higher market-impact risk on execution |

> **Note on `rf`:** Setting `rf > 0` shifts the Sharpe, Sortino, and Omega
> thresholds simultaneously. For comparison across pairs, use a consistent `rf`
> (e.g. current T-bill rate). The default `rf = 0` treats excess return over
> cash as the entire return.

---

## 9.3  `print_rebal_profile()` — formatted output

```r
print_rebal_profile(x, digits = 1)
```

- `x` — the tibble returned by `rebal_profile()`
- `digits` — decimal places for percentage columns (default 1)

**Output format:**

```
══════════════════════════════════════════════════════════════════════════════════
  Pair: SPY/IEF  |  Weights: 60/40  |  2003-04-07 → 2025-12-31  (5724 days)
══════════════════════════════════════════════════════════════════════════════════

   Strategy Final $1 Ann Ret Ann Vol  Max DD Sharpe Sortino Calmar  Omega VaR 5% CVaR 5% Rebal N Avg Slip Best
  Quarterly    $9.47   7.5%    9.3% -22.7%   0.81    1.17  0.331  1.617  -0.9%   -1.3%      88  0.551% ★★★★
Semi-Annual    $9.23   7.4%    9.4% -22.9%   0.78    1.13  0.322  1.598  -0.9%   -1.3%      44  0.551%  ★★
     Annual    $9.05   7.3%    9.5% -23.5%   0.76    1.10  0.311  1.576  -0.9%   -1.3%      22  0.552%
 Buy & Hold    $8.71   7.1%   10.1% -26.0%   0.70    1.02  0.274  1.531  -1.0%   -1.4%       0       0

  ★ = best among rebalanced strategies (excludes Buy & Hold)
══════════════════════════════════════════════════════════════════════════════════
```

The `Best` column concatenates one ★ per metric where that row wins among the
three rebalanced strategies.  Buy & Hold is always excluded from the ranking —
it is the *baseline*, not a rebalancing choice.

**Formatting rules applied:**
- `Ann Ret`, `Ann Vol`, `Max DD`, `VaR 5%`, `CVaR 5%` → shown as `%` (×100)
- `Avg Slip` → shown as `%` with 3 decimal places
- `Final $1` → shown as `$X.XX`
- Sharpe, Sortino, Calmar, Omega → 2 decimal places

---

## 9.4  `best_rebal_freq()` — single-criterion selector

```r
best_rebal_freq(x, by = "Sharpe")
```

Returns (invisibly) the strategy name that wins on `by`, and prints a one-line
summary.  Buy & Hold is always excluded.

**Higher-is-better metrics** (maximised): `Final $1`, `Ann Ret`, `Sharpe`,
`Sortino`, `Calmar`, `Omega`.

**Lower-is-better metrics** (minimised): `Ann Vol`, `Max DD`, `VaR 5%`,
`CVaR 5%`, `Rebal N`, `Avg Slip`.

```r
best_rebal_freq(p, by = "Sharpe")
#> Best rebalancing frequency by Sharpe: Quarterly  (Sharpe = 0.807)

best_rebal_freq(p, by = "Max DD")
#> Best rebalancing frequency by Max DD: Quarterly  (Max DD = -22.7%)

best_rebal_freq(p, by = "Rebal N")
#> Best rebalancing frequency by Rebal N: Annual  (Rebal N = 22)
```

---

## 9.5  Live examples

### SPY / IEF 60/40 — the canonical case

```{r rebal-profile-spy-ief}
source(here("scripts_daa_saa_taa/rebal_profile.R"))
p_6040 <- rebal_profile("SPY", "IEF", w1 = 0.60)
print_rebal_profile(p_6040)
best_rebal_freq(p_6040, by = "Sharpe")
best_rebal_freq(p_6040, by = "Max DD")
best_rebal_freq(p_6040, by = "Calmar")
```

**Quarterly** typically sweeps the risk-adjusted metrics because it rebalances
often enough to harvest the majority of the variance bonus while still allowing
~3 months of momentum to run.

### QQQ / AGG 65/35 — higher-volatility growth pair

```{r rebal-profile-qqq-agg}
p_qqq <- rebal_profile("QQQ", "AGG", w1 = 0.65)
print_rebal_profile(p_qqq)
best_rebal_freq(p_qqq, by = "Sharpe")
```

Higher QQQ volatility widens $\sigma^2(\text{spread})$ and amplifies the
rebalancing bonus.  The momentum effect of letting QQQ run is also stronger,
so the optimal frequency may shift toward Semi-Annual — the bonus from frequent
rebalancing competes more evenly against the cost of cutting winners early.

### With a non-zero risk-free rate

```r
p_rf <- rebal_profile("SPY", "IEF", w1 = 0.60, rf = 0.04)
print_rebal_profile(p_rf)
```

Setting `rf = 0.04` (4% p.a.) rescales Sharpe, Sortino, and Omega relative to
a cash hurdle.  Relative rankings between strategies are usually stable, but
absolute Sharpe values will be lower and Omega will be closer to 1.

---

## 9.6  Decision guide

| Primary objective | Look at | Prefer |
|---|---|---|
| Maximise long-run wealth | `Final $1`, `Ann Ret` | Higher frequency (Quarterly) |
| Maximise risk-adjusted return | `Sharpe`, `Sortino`, `Calmar` | ★ winner |
| Minimise worst drawdown | `Max DD` | ★ winner on Max DD |
| Minimise tail loss severity | `CVaR 5%` | ★ winner on CVaR 5% |
| Minimise trading activity | `Rebal N` | Annual (fewest resets) |
| Minimise market-impact exposure | `Avg Slip` | Lowest slip strategy |

> **Rule of thumb for low-cost ETFs:** Quarterly rebalancing captures ~85% of
> the daily-rebalancing bonus at ~1/20th of the trading activity and rarely
> loses to Annual on any risk-adjusted metric.  Semi-Annual is the pragmatic
> middle ground for higher-volatility pairs where momentum is a stronger force.

---

*Source: `scripts_daa_saa_taa/plot_rebal_drift.R` — SPY/IEF 60/40 universe.*
*Regime shading via `build_regime_table()` with T_FALL = 10%.*
*All returns are total-return log returns; no transaction costs assumed.*
