---
title: "Sovereign SAA Execution Report"
subtitle: "3-Book Portfolio · Spring-Load Signals · Sovereign 5 Roster"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
params:
  fs_ticker: "XLK"
  fs_bmk:    "SPY"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
    fig_width: 11
    fig_height: 6
---

```{css, echo=FALSE}
body       { font-family: "Inter", "Helvetica Neue", Arial, sans-serif;
             background: #f4f6fb; }
h1, h2, h3 { color: #1B2A4A; font-weight: 800; }
h2         { border-bottom: 2px solid #1B2A4A; padding-bottom: 6px; }
.section-label {
  display: inline-block; font-size: 10px; font-weight: 800;
  letter-spacing: .1em; text-transform: uppercase;
  padding: 3px 10px; border-radius: 12px; margin-bottom: 8px;
}
.lbl-navy  { background:#1B2A4A; color:#fff; }
.lbl-teal  { background:#1E6E7A; color:#fff; }
.lbl-gold  { background:#8B6510; color:#fff; }
.lbl-green { background:#2D6A4F; color:#fff; }
.lbl-red   { background:#D90429; color:#fff; }
.metric-row { display:flex; gap:18px; flex-wrap:wrap; margin-bottom:20px; }
.metric-box {
  flex:1; min-width:160px; background:#fff; border-radius:6px;
  padding:14px 18px; box-shadow:0 1px 5px rgba(0,0,0,.08);
  border-left: 5px solid #1B2A4A;
}
.metric-box.teal  { border-left-color:#1E6E7A; }
.metric-box.gold  { border-left-color:#8B6510; }
.metric-box.green { border-left-color:#2D6A4F; }
.metric-box.red   { border-left-color:#D90429; }
.metric-val  { font-size:28px; font-weight:800; color:#1B2A4A; line-height:1; }
.metric-lbl  { font-size:10px; font-weight:700; text-transform:uppercase;
               letter-spacing:.08em; color:#8a94a8; margin-top:5px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center")
library(here)
library(tidyverse)
library(scales)
library(DT)
library(knitr)
library(kableExtra)
library(ggrepel)
```

```{r run-pipeline, include=FALSE}
# ── Run the full SAA execution pipeline ──────────────────────────────────────
# Each script is idempotent: skips work if objects already in env.
source(here("scripts_saa_execution/saa_bridge_config.R"))
source(here("scripts_saa_execution/saa_quantdb_builder.R"))
source(here("scripts_saa_execution/saa_alpha_intelligence.R"))
source(here("scripts_saa_execution/saa_policy_weights.R"))
source(here("scripts_saa_execution/saa_drift_analysis.R"))
source(here("scripts_saa_execution/saa_master_allocator.R"))
source(here("scripts_saa_execution/saa_visual_dashboard.R"))
source(here("scripts_saa_execution/saa_perf_attribution.R"))
source(here("scripts_saa_execution/saa_quantdb_selector.R"))
source(here("utility/plot_sovereign_tapes.R"))
```

---

# Executive Summary

<div class="metric-row">

```{r exec-kpis, echo=FALSE, results='asis'}
# Book 3 metrics
n_positions <- nrow(execution_list)
long_wt     <- sum(execution_list$net_weight[execution_list$net_weight > 0])
short_wt    <- abs(sum(execution_list$net_weight[execution_list$net_weight < 0]))
n_sl        <- sum(selection_matrix$status == "Spring-Load", na.rm = TRUE)
n_ex        <- sum(selection_matrix$status == "Exhaustion",  na.rm = TRUE)
taa_active  <- nrow(policy_list$TAA[policy_list$TAA$book == "BOOK_2_TAA_LONG", ])

cat(sprintf('
<div class="metric-box">
  <div class="metric-val">%d</div>
  <div class="metric-lbl">Total Positions</div>
</div>
<div class="metric-box teal">
  <div class="metric-val">%s</div>
  <div class="metric-lbl">Long Exposure</div>
</div>
<div class="metric-box gold">
  <div class="metric-val">%d</div>
  <div class="metric-lbl">Spring-Loads</div>
</div>
<div class="metric-box green">
  <div class="metric-val">%d</div>
  <div class="metric-lbl">Active TAA Pairs</div>
</div>
<div class="metric-box red">
  <div class="metric-val">%d</div>
  <div class="metric-lbl">Exhaustions</div>
</div>',
n_positions,
scales::percent(long_wt, accuracy = 0.1),
n_sl, taa_active, n_ex))
```

</div>

> **Pipeline date:** `r Sys.Date()` · **Universe:** `r nrow(saa_metadata)` tickers ·
> **Reference AUM:** $1,000,000

---

# 1 · QuantDB — Ticker Intelligence

<span class="section-label lbl-navy">Stage 1</span>

The `etf_quantdb` is the foundation of the execution pipeline — an 8-layer
per-ticker database covering absolute performance, relative edge vs SPY,
and multi-horizon momentum confirmation.

## 1.1 Full QuantDB

```{r quantdb-table}
etf_quantdb %>%
  select(ticker, cluster, category,
         ann_ret, alpha_6y, corr_to_bmk, beta_spy,
         ann_vol, max_dd, ir_ratio,
         is_confirmed, mom_d) %>%
  mutate(across(c(ann_ret, alpha_6y, corr_to_bmk, beta_spy,
                  ann_vol, max_dd, ir_ratio, mom_d), ~ round(., 3))) %>%
  DT::datatable(
    filter    = "top",
    rownames  = FALSE,
    class     = "compact stripe",
    options   = list(pageLength = 15, scrollX = TRUE),
    caption   = "QuantDB — sorted by cumulative alpha (alpha_6y)"
  ) %>%
  DT::formatStyle("alpha_6y",
    background = DT::styleInterval(0, c("#fdecea", "#e8f5e9")),
    fontWeight = "bold") %>%
  DT::formatStyle("is_confirmed",
    color = DT::styleEqual(c("YES", "NO"), c("#2D6A4F", "#D90429")),
    fontWeight = "bold")
```

## 1.2 Alpha Landscape

```{r alpha-landscape, fig.height=7}
etf_quantdb %>%
  mutate(label_col = if_else(alpha_6y > 0, "#2D6A4F", "#D90429")) %>%
  ggplot(aes(x = corr_to_bmk, y = alpha_6y,
             colour = cluster, size = abs(ann_ret))) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.4) +
  geom_vline(xintercept = 0.5, colour = "grey80",
             linetype = "dashed", linewidth = 0.4) +
  geom_point(alpha = 0.75) +
  ggrepel::geom_text_repel(aes(label = ticker), size = 3,
                            max.overlaps = 20, show.legend = FALSE) +
  scale_colour_brewer(palette = "Dark2") +
  scale_size_continuous(range = c(2, 8), guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "Alpha Landscape: Cumulative Alpha vs Correlation to SPY",
    subtitle = "Size = |Ann. Return|  |  Dashed line = corr 0.50 (dampener boundary)",
    x = "Correlation to SPY", y = "Cumulative Alpha (full history)",
    colour = "Cluster"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank(),
        legend.position  = "right")
```

## 1.3 DNA Score Ranking

```{r dna-ranking, fig.height=6}
ranked <- rank_spider_dna()

ranked %>%
  head(20) %>%
  ggplot(aes(x = reorder(ticker, dna_score), y = dna_score,
             fill = dna_score)) +
  geom_col(width = 0.75) +
  geom_text(aes(label = round(dna_score, 2)),
            hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_gradient(low = "#d8e9ff", high = "#1B2A4A", guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Top 20 Tickers by Sovereign DNA Score",
       subtitle = "Weighted: Alpha ×0.4 · Stability ×0.3 · Momentum ×0.3",
       x = NULL, y = "DNA Score (0–1)") +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

---

# 2 · Alpha Signals — Spring-Load & Exhaustion

<span class="section-label lbl-teal">Stage 2</span>

Each ticker is classified daily along two dimensions:

| Signal | Condition | Interpretation |
|---|---|---|
| **Spring-Load** | range_pct < 30% AND above MA200 | Beaten down but structurally intact — coiled for recovery |
| **Exhaustion** | range_pct > 70% AND below MA200 | Extended but trend broken — potential mean reversion |
| **Neutral** | All other | No strong signal |

A **LONG quadrant** requires Spring-Load _and_ positive alpha velocity z-score (confirmed momentum building).

## 2.1 Signal Matrix

```{r signal-matrix}
selection_matrix %>%
  select(ticker, cluster, status, quadrant,
         alpha_velocity_z, range_percentile, dist_200, trend_regime) %>%
  mutate(across(c(alpha_velocity_z, range_percentile, dist_200),
                ~ round(., 3))) %>%
  DT::datatable(
    filter   = "top",
    rownames = FALSE,
    class    = "compact stripe",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "Selection Matrix — sorted by alpha velocity z-score"
  ) %>%
  DT::formatStyle("status",
    backgroundColor = DT::styleEqual(
      c("Spring-Load", "Exhaustion", "Neutral"),
      c("#e8f5e9",      "#fdecea",    "#f5f5f5")),
    fontWeight = "bold") %>%
  DT::formatStyle("quadrant",
    color = DT::styleEqual(
      c("SPRING_LOAD (LONG)", "EXHAUSTION (SHORT)", "NEUTRAL"),
      c("#2D6A4F",             "#D90429",             "#888888")),
    fontWeight = "bold")
```

## 2.2 Signal Map

```{r signal-map, fig.height=6}
selection_matrix %>%
  filter(!is.na(range_percentile), !is.na(dist_200)) %>%
  mutate(
    col = case_when(
      status == "Spring-Load" ~ "#2D6A4F",
      status == "Exhaustion"  ~ "#D90429",
      TRUE                    ~ "#aaaaaa"
    ),
    alpha_pt = if_else(status == "Neutral", 0.4, 0.9)
  ) %>%
  ggplot(aes(x = dist_200, y = range_percentile,
             colour = col, alpha = alpha_pt)) +
  annotate("rect", xmin = 0,    xmax = Inf,  ymin = -Inf, ymax = 0.30,
           fill = "#e8f5e9", alpha = 0.3) +
  annotate("rect", xmin = -Inf, xmax = 0,    ymin = 0.70,  ymax = Inf,
           fill = "#fdecea", alpha = 0.3) +
  geom_hline(yintercept = c(0.30, 0.70), linetype = "dashed",
             colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(size = 4) +
  ggrepel::geom_text_repel(
    aes(label = ticker), size = 3, max.overlaps = 25, show.legend = FALSE) +
  scale_colour_identity() +
  scale_alpha_identity() +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "Spring-Load / Exhaustion Signal Map",
    subtitle = "Green zone = Spring-Load candidates  |  Red zone = Exhaustion candidates",
    x = "Distance from MA200  (dist_200)",
    y = "52-Week Range Percentile"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

---

# 3 · Portfolio Construction — The 3 Books

<span class="section-label lbl-gold">Stages 3–5</span>

## 3.1 Book 1 — Sovereign SAA (60 / 40)

```{r book1}
policy_list$SAA %>%
  mutate(weight_pct = scales::percent(weight, accuracy = 0.1),
         dollar     = scales::dollar(weight * 1e6)) %>%
  kable(col.names = c("Ticker", "Weight", "Book", "Weight %", "$ Value"),
        align = "lrllr") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

## 3.2 Book 2 — TAA Long / Short Synthetics

```{r book2}
if (nrow(policy_list$TAA) > 0) {
  policy_list$TAA %>%
    mutate(weight_pct = scales::percent(weight, accuracy = 0.1),
           dollar     = scales::dollar(weight * 1e6)) %>%
    kable(col.names = c("Ticker", "Weight", "Book", "Weight %", "$ Value"),
          align = "lrllr") %>%
    kable_styling(bootstrap_options = c("striped", "condensed"),
                  full_width = FALSE)
} else {
  cat("> **Book 2 is flat** — no Spring-Loads qualified this cycle.")
}
```

## 3.3 Book 3 — Total Net Exposure

```{r book3-bar, fig.height=5}
print(p_book3_weights)
```

```{r book3-table}
execution_list %>%
  select(ticker, cluster, category, net_weight, weight_pct,
         target_value_fmt, action) %>%
  kable(col.names = c("Ticker", "Cluster", "Category",
                      "Net Wt", "Wt %", "$ Target", "Action"),
        align = "llllrrl") %>%
  kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  row_spec(which(execution_list$action == "SHORT/SELL"),
           color = "#D90429", bold = TRUE) %>%
  row_spec(which(execution_list$action == "BUY/HOLD"),
           color = "#2D6A4F")
```

## 3.4 Cluster Drift Audit

```{r drift-chart, fig.height=5}
crowding_summary %>%
  ggplot(aes(x = reorder(cluster, total_net_exposure),
             y = total_net_exposure, fill = total_net_exposure)) +
  geom_col(width = 0.75) +
  geom_text(aes(label = scales::percent(total_net_exposure, accuracy = 0.1)),
            hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_gradient2(low = "#D90429", mid = "#f5f5f5", high = "#1B2A4A",
                       midpoint = 0, guide = "none") +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Book 3: Net Exposure by Cluster",
       subtitle = "After applying 60/40 SAA + TAA LS synthetics",
       x = NULL, y = "Net Weight") +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

## 3.5 Treemap — SAA vs TAA (Book 3)

```{r treemap-setup, include=FALSE}
library(treemapify)
library(patchwork)
library(plotly)
source(here("scripts_saa_taa/saa_tree.R"))
source(here("key_plots/saa_treemap_templates.R"))

# Cluster → role mapping for execution universe
.cluster_to_role <- c(
  "Broad Market"      = "Anchor",
  "Sovereign/BMK"     = "Anchor",
  "Broad Bonds"       = "Core-Stabilizer",
  "Cash/Ultra-Short"  = "Core-Stabilizer",
  "Defensive Bulwark" = "Core-Stabilizer",
  "Factor Play"       = "Core-Growth",
  "Growth Laggards"   = "Core-Growth",
  "Cycle & Breadth"   = "Core-Growth",
  "Pipes & Power"     = "Core-Growth",
  "Multipolar Alpha"  = "Satellite"
)

# Build TAA (Book 3 long positions) compatible with treemap functions
.build_taa_tm_df <- function(drift_report, xts_ret) {
  ytd_tickers <- drift_report$ticker[drift_report$net_weight > 0]
  ytd_vals    <- calc_ytd(xts_ret, ytd_tickers)

  drift_report %>%
    filter(net_weight > 0) %>%
    mutate(
      weight      = net_weight,
      role        = coalesce(.cluster_to_role[cluster], "Core-Growth"),
      asset_class = case_when(
        category == "Fixed" ~ "Fixed Income",
        category == "Alt"   ~ "Alternatives",
        category == "Multi" ~ "Multi-Asset",
        TRUE                ~ "Equity"
      ),
      ytd       = ytd_vals[ticker],
      ytd_label = if_else(
        is.na(ytd), "n/a",
        paste0(if_else(ytd >= 0, "+", ""),
               scales::percent(ytd, accuracy = 0.1))
      ),
      leaf_fill  = case_when(
        is.na(ytd) ~ ROLE_PAL[role],
        ytd >= 0   ~ "#40916C",
        TRUE       ~ "#C1121F"
      ),
      cell_label = paste0(ticker, "\n",
                          scales::percent(weight, accuracy = 0.1),
                          "\n", ytd_label)
    )
}

saa_tm_df <- build_treemap_df(xts_ret) %>%
  mutate(
    leaf_fill = if ("ytd" %in% names(.)) {
      case_when(is.na(ytd) ~ unname(ROLE_PAL[role]),
                ytd >= 0   ~ "#40916C",
                TRUE       ~ "#C1121F")
    } else unname(ROLE_PAL[role])
  )
taa_tm_df <- .build_taa_tm_df(drift_report, xts_ret) %>%
  mutate(leaf_fill = unname(leaf_fill))
```

```{r treemap-side-by-side, fig.width=13, fig.height=5}
# ── SAA treemap (Book 1) ──────────────────────────────────────────────────────
p_tm_saa <- ggplot(saa_tm_df,
    aes(area = weight, fill = leaf_fill,
        label = cell_label, subgroup = asset_class)) +
  treemapify::geom_treemap(colour = "white", size = 2) +
  treemapify::geom_treemap_subgroup_border(colour = "white", size = 5) +
  treemapify::geom_treemap_subgroup_text(
    place = "topleft", colour = "white",
    fontface = "bold", size = 13, alpha = 0.7, grow = FALSE) +
  treemapify::geom_treemap_text(
    colour = "white", fontface = "bold",
    size = 11, place = "centre", grow = FALSE) +
  scale_fill_identity() +
  theme_void() +
  theme(plot.title    = element_text(face = "bold", size = 13,
                                     colour = "#1B2A4A",
                                     margin = margin(b = 4)),
        plot.subtitle = element_text(size = 9, colour = "grey50",
                                     margin = margin(b = 6))) +
  labs(title    = "SAA — Book 1 (60% SPY · 40% AGG baseline)",
       subtitle = paste0("Area = weight  |  Colour = YTD  |  ",
                         format(Sys.Date(), "%d %b %Y")))

# ── TAA treemap (Book 3 long positions) ───────────────────────────────────────
p_tm_taa <- ggplot(taa_tm_df,
    aes(area = weight, fill = leaf_fill,
        label = cell_label, subgroup = asset_class)) +
  treemapify::geom_treemap(colour = "white", size = 2) +
  treemapify::geom_treemap_subgroup_border(colour = "white", size = 5) +
  treemapify::geom_treemap_subgroup_text(
    place = "topleft", colour = "white",
    fontface = "bold", size = 13, alpha = 0.7, grow = FALSE) +
  treemapify::geom_treemap_text(
    colour = "white", fontface = "bold",
    size = 11, place = "centre", grow = FALSE) +
  scale_fill_identity() +
  theme_void() +
  theme(plot.title    = element_text(face = "bold", size = 13,
                                     colour = "#1B2A4A",
                                     margin = margin(b = 4)),
        plot.subtitle = element_text(size = 9, colour = "grey50",
                                     margin = margin(b = 6))) +
  labs(title    = "TAA — Book 3 Net (SAA + LS overlay, long positions)",
       subtitle = "Area = net weight  |  Colour = YTD  |  Excludes short legs")

p_tm_saa + p_tm_taa
```

## 3.6 Sunburst — SAA vs TAA (Book 3)

```{r sunburst-side-by-side, fig.width=13, fig.height=6}
# Helper: build plotly sunburst from any portfolio df
.make_sunburst <- function(df, title_text) {
  # 2-ring: asset_class (inner) → ticker (outer)
  ac_rows <- df %>%
    group_by(asset_class) %>%
    summarise(weight = sum(weight), .groups = "drop") %>%
    mutate(
      ids     = asset_class,
      labels  = asset_class,
      parents = "",
      values  = weight,
      text    = scales::percent(weight, accuracy = 0.1)
    )

  tk_rows <- df %>%
    mutate(
      ids     = paste0(asset_class, " - ", ticker),
      labels  = ticker,
      parents = asset_class,
      values  = weight,
      text    = if ("ytd_label" %in% names(.))
                  paste0(scales::percent(weight, accuracy = 0.1),
                         "  YTD ", ytd_label)
                else scales::percent(weight, accuracy = 0.1)
    ) %>%
    select(ids, labels, parents, values, text)

  all_rows <- bind_rows(ac_rows %>% select(ids, labels, parents, values, text),
                        tk_rows)

  plotly::plot_ly(
    ids     = all_rows$ids,
    labels  = all_rows$labels,
    parents = all_rows$parents,
    values  = all_rows$values,
    text    = all_rows$text,
    type    = "sunburst",
    branchvalues = "total",
    hovertemplate = "%{label}<br>%{text}<extra></extra>",
    textinfo = "label+percent entry",
    insidetextorientation = "radial"
  ) %>%
  plotly::layout(
    title = list(text = title_text, font = list(size = 13, color = "#1B2A4A")),
    margin = list(t = 40, b = 10, l = 10, r = 10)
  )
}

p_sb_saa <- .make_sunburst(saa_tm_df,  "SAA · Book 1")
p_sb_taa <- .make_sunburst(taa_tm_df,  "TAA · Book 3 Net")

plotly::subplot(p_sb_saa, p_sb_taa, nrows = 1) %>%
  plotly::layout(
    title = list(
      text = "Sunburst: SAA (Book 1)  vs  TAA Book 3 Net",
      font = list(size = 14, color = "#1B2A4A")
    )
  )
```

---

# 4 · Sovereign 5 Roster

<span class="section-label lbl-green">Selector</span>

The Sovereign 5 are selected from the QuantDB each cycle:
- **3 Enhancers** — highest `(alpha_6y × 0.7) + (mom_d × 0.3)` with confirmed momentum
- **2 Dampeners** — lowest tail risk `(−|kurtosis| − |skew|)` among tickers with `corr < 0.50`

## 4.1 Roster

```{r sovereign-roster}
sovereign_list %>%
  select(ticker, cluster, role, goal, alpha_6y, corr_to_bmk, is_confirmed) %>%
  mutate(alpha_6y = scales::percent(alpha_6y, accuracy = 0.1),
         corr_to_bmk = round(corr_to_bmk, 3)) %>%
  kable(col.names = c("Ticker", "Cluster", "Role", "Goal",
                      "Cum. Alpha", "Corr SPY", "Confirmed"),
        align = "lllllrr") %>%
  kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  row_spec(which(sovereign_list$role == "ENHANCER"),
           background = "#e8f5e9") %>%
  row_spec(which(sovereign_list$role == "DAMPENER"),
           background = "#e8f0fb")
```

## 4.2 Sovereign DNA — Vertical Spider

```{r sovereign-spider, fig.width=12, fig.height=14}
if (nrow(sovereign_list) > 0) {
  plot_etf_spider_vertical(c(sovereign_list$ticker, "SPY"))
}
```

---

# 5 · TAA Performance Attribution

<span class="section-label lbl-teal">Attribution</span>

```{r taa-stats}
if (nrow(taa_stats) > 0) {
  taa_stats %>%
    select(Metric, Value) %>%
    mutate(Value = round(Value, 4)) %>%
    kable(align = "lr") %>%
    kable_styling(bootstrap_options = c("striped", "condensed"),
                  full_width = FALSE)
} else {
  cat("> TAA Book 2 is flat — no attribution to show.")
}
```

```{r taa-chart, fig.height=5}
print(p_taa_cum_alpha)
```

---

# 6 · Regime Alpha Tape

<span class="section-label lbl-navy">Regime Context</span>

SPY row shows **absolute** returns; all other rows show **alpha** (relative return
vs SPY).  Vertical markers = SPY drawdown entry (dashed) and recovery exit
(dotted).  Quarterly x-axis for precision dating.

## 6.1 Defensive Bulwark

```{r tape-defensive, fig.height=6}
def_tickers <- defensive_tickers[defensive_tickers %in% colnames(xts_d_rel)]
if (length(def_tickers) > 0) {
  plot_sovereign_alpha_tape_v12(
    xts_abs        = xts_d_abs,
    xts_rel        = xts_d_rel,
    target_tickers = head(def_tickers, 5),
    dd_threshold   = 0.10
  )
}
```

## 6.2 Cycle & Breadth

```{r tape-cycle, fig.height=6}
cyc_tickers <- cycle_tickers[cycle_tickers %in% colnames(xts_d_rel)]
if (length(cyc_tickers) > 0) {
  plot_sovereign_alpha_tape_v12(
    xts_abs        = xts_d_abs,
    xts_rel        = xts_d_rel,
    target_tickers = head(cyc_tickers, 5),
    dd_threshold   = 0.10
  )
}
```

## 6.3 Sovereign 5 in Context

```{r tape-sovereign5, fig.height=6}
if (nrow(sovereign_list) > 0) {
  s5 <- sovereign_list$ticker[sovereign_list$ticker %in% colnames(xts_d_rel)]
  plot_sovereign_alpha_tape_v12(
    xts_abs        = xts_d_abs,
    xts_rel        = xts_d_rel,
    target_tickers = s5,
    dd_threshold   = 0.10
  )
}
```

---

# 7 · Execution List

<span class="section-label lbl-gold">Output</span>

```{r exec-list}
execution_list %>%
  select(ticker, cluster, category, net_weight,
         weight_pct, target_value_fmt, action) %>%
  DT::datatable(
    rownames = FALSE,
    class    = "compact stripe",
    options  = list(pageLength = 20, dom = "tp"),
    caption  = paste("Trade list generated:", Sys.Date())
  ) %>%
  DT::formatStyle("action",
    backgroundColor = DT::styleEqual(
      c("BUY/HOLD", "SHORT/SELL", "FLAT"),
      c("#e8f5e9",  "#fdecea",    "#f5f5f5")),
    fontWeight = "bold") %>%
  DT::formatStyle("net_weight",
    background = DT::styleColorBar(
      range(execution_list$net_weight), "#d0e8ff"),
    backgroundSize = "100% 70%",
    backgroundRepeat = "no-repeat",
    backgroundPosition = "center")
```

---

---

# 8 · IC Alignment — Institutional vs Sovereign Signals

<span class="section-label lbl-navy">IC Grid · 13-Mar-2026</span>

Compares the Investment Committee's TAA active weights (MAC USD · BAL profile)
against our own Spring-Load / Exhaustion signals and QuantDB momentum.
**Green = aligned · Red = conflicted · Grey = IC has no view on this ticker.**

```{r ic-load, include=FALSE}
source(here("input/ic_grid_signals.R"))
```

## 8.1 IC TAA Signal Map (USD · BAL)

```{r ic-bar, fig.height=6}
ic_bal_usd %>%
  filter(!is.na(etf_proxy)) %>%
  mutate(
    fill_col = case_when(
      direction == "BUY"  ~ "#2D6A4F",
      direction == "SELL" ~ "#D90429",
      TRUE                ~ "#aaaaaa"
    ),
    label = paste0(ifelse(delta > 0, "+", ""),
                   round(delta * 100, 1), "%")
  ) %>%
  ggplot(aes(x = reorder(etf_proxy, delta),
             y = delta, fill = fill_col)) +
  geom_col(width = 0.75) +
  geom_text(aes(label = label,
                hjust = ifelse(delta >= 0, -0.2, 1.2)),
            size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1),
                     expand = expansion(mult = c(0.15, 0.15))) +
  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.4) +
  labs(
    title    = "IC TAA Active Weights — USD / BAL",
    subtitle = "Green = IC overweight vs SAA  |  Red = IC underweight vs SAA",
    x = NULL, y = "TAA Δ (active weight vs SAA)",
    caption  = "Source: IC Active Weights Overview, 13-Mar-2026"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

## 8.2 Alignment Table — IC vs Sovereign Signals

```{r ic-alignment}
# Join IC BAL/USD signals with our selection matrix and quantdb
ic_joined <- ic_bal_usd %>%
  filter(!is.na(etf_proxy)) %>%
  select(etf_proxy, ic_name, asset_class, delta,
         direction, signal_tier, conviction) %>%
  left_join(
    selection_matrix %>%
      select(ticker, cluster, status, quadrant, alpha_velocity_z, dist_200),
    by = c("etf_proxy" = "ticker")
  ) %>%
  left_join(
    etf_quantdb %>% select(ticker, ann_ret, alpha_6y,
                            corr_to_bmk, is_confirmed, mom_d),
    by = c("etf_proxy" = "ticker")
  ) %>%
  mutate(
    # Alignment: does our signal agree with IC direction?
    our_direction = case_when(
      status == "Spring-Load" ~ "BUY",
      status == "Exhaustion"  ~ "SELL",
      mom_d > 0.5             ~ "BUY",
      mom_d < -0.5            ~ "SELL",
      TRUE                    ~ "NEUTRAL"
    ),
    alignment = case_when(
      our_direction == direction             ~ "ALIGNED",
      our_direction == "NEUTRAL"             ~ "NO SIGNAL",
      TRUE                                   ~ "CONFLICT"
    ),
    align_col = case_when(
      alignment == "ALIGNED"    ~ "#e8f5e9",
      alignment == "CONFLICT"   ~ "#fdecea",
      TRUE                      ~ "#f5f5f5"
    )
  ) %>%
  arrange(desc(abs(delta)))

ic_joined %>%
  select(etf_proxy, asset_class, ic_name,
         delta, direction, signal_tier,
         our_direction, alignment, alpha_6y, mom_d) %>%
  mutate(
    delta    = scales::percent(delta, accuracy = 0.1),
    alpha_6y = scales::percent(alpha_6y, accuracy = 0.1),
    mom_d    = round(mom_d, 2)
  ) %>%
  DT::datatable(
    rownames = FALSE,
    class    = "compact stripe",
    options  = list(pageLength = 20, dom = "tp"),
    caption  = "IC direction vs Sovereign signals (USD/BAL · 13-Mar-2026)"
  ) %>%
  DT::formatStyle("alignment",
    backgroundColor = DT::styleEqual(
      c("ALIGNED", "CONFLICT", "NO SIGNAL"),
      c("#e8f5e9",  "#fdecea",  "#f5f5f5")),
    fontWeight = "bold") %>%
  DT::formatStyle("direction",
    color = DT::styleEqual(
      c("BUY", "SELL", "HOLD"),
      c("#2D6A4F", "#D90429", "#888888")),
    fontWeight = "bold")
```

## 8.3 Alignment Summary

```{r ic-summary, fig.height=4}
align_summary <- ic_joined %>%
  count(alignment) %>%
  mutate(
    fill_col = case_when(
      alignment == "ALIGNED"   ~ "#2D6A4F",
      alignment == "CONFLICT"  ~ "#D90429",
      TRUE                     ~ "#aaaaaa"
    ),
    pct = scales::percent(n / sum(n), accuracy = 1)
  )

ggplot(align_summary, aes(x = reorder(alignment, n),
                           y = n, fill = fill_col)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(n, "  (", pct, ")")),
            hjust = -0.15, size = 4, fontface = "bold") +
  coord_flip() +
  scale_fill_identity() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  labs(title    = "Sovereign Signal Alignment with IC TAA",
       subtitle = "USD / BAL · 13-Mar-2026",
       x = NULL, y = "Number of ETF positions") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

## 8.4 IC Signal Strength by Profile (USD)

```{r ic-profile-heatmap, fig.height=5}
ic_signals %>%
  filter(ccy == "USD", !is.na(etf_proxy)) %>%
  mutate(profile = factor(profile, levels = c("INC", "BAL", "GRO"))) %>%
  ggplot(aes(x = profile, y = reorder(etf_proxy, abs(delta)),
             fill = delta)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = paste0(ifelse(delta > 0, "+", ""),
                               round(delta * 100, 1), "%")),
            size = 3, fontface = "bold",
            colour = ifelse(abs(ic_signals %>%
                                  filter(ccy=="USD", !is.na(etf_proxy)) %>%
                                  pull(delta)) > 0.04,
                            "white", "#1B2A4A")) +
  scale_fill_gradient2(low = "#D90429", mid = "#f5f5f5", high = "#2D6A4F",
                       midpoint = 0, name = "TAA Δ",
                       labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "IC TAA Active Weights — USD · All Profiles",
    subtitle = "INC = Income · BAL = Balanced · GRO = Growth",
    x = "Risk Profile", y = NULL,
    caption  = "Source: IC Active Weights Overview, 13-Mar-2026"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid       = element_blank(),
        axis.text.y      = element_text(face = "bold"))
```

## 8.5 Cumulative Return by Risk Profile — SAA vs TAA

Portfolio returns are computed from the **ETF-proxied subset** of each IC
profile's weights. Weights are rescaled to sum to 1 within the mapped universe
(untranslatable positions — AI Hedge Funds, S&P 500 hedge overlay — are
excluded). Both the SAA baseline and the TAA active allocation are shown,
benchmarked against SPY.

```{r ic-cumret, fig.height=7}
library(PerformanceAnalytics)

# ── Helper: build daily portfolio return series from IC weights ───────────────
.ic_portfolio_ret <- function(profile_id, weight_col = "taa") {

  w <- ic_signals %>%
    filter(ccy == "USD", profile == profile_id, !is.na(etf_proxy)) %>%
    select(etf_proxy, wt = all_of(weight_col)) %>%
    filter(wt != 0,
           etf_proxy %in% colnames(xts_d_abs))

  if (nrow(w) == 0) return(NULL)

  # Rescale to sum to 1 (proxy-universe normalisation)
  w <- w %>% mutate(wt = wt / sum(abs(wt)) * sign(sum(wt)))

  ret_subset <- xts_d_abs[, w$etf_proxy, drop = FALSE]

  # Weighted daily return: matrix multiply
  wt_vec <- setNames(w$wt, w$etf_proxy)
  port_ret <- xts(
    as.numeric(ret_subset %*% wt_vec[colnames(ret_subset)]),
    order.by = zoo::index(ret_subset)
  )
  port_ret
}

# Build all 6 series (3 profiles × SAA / TAA)
profiles <- c("INC", "BAL", "GRO")

cum_list <- purrr::map_df(profiles, function(p) {
  r_saa <- .ic_portfolio_ret(p, "saa")
  r_taa <- .ic_portfolio_ret(p, "taa")

  bind_rows(
    if (!is.null(r_saa)) data.frame(
      Date    = zoo::index(r_saa),
      CumRet  = as.numeric(cumprod(1 + zoo::coredata(r_saa)) - 1),
      Profile = p,
      Type    = "SAA"
    ),
    if (!is.null(r_taa)) data.frame(
      Date    = zoo::index(r_taa),
      CumRet  = as.numeric(cumprod(1 + zoo::coredata(r_taa)) - 1),
      Profile = p,
      Type    = "TAA"
    )
  )
})

# Add SPY benchmark
spy_ret  <- xts_d_abs[, "SPY"]
spy_cum  <- data.frame(
  Date    = zoo::index(spy_ret),
  CumRet  = as.numeric(cumprod(1 + zoo::coredata(spy_ret)) - 1),
  Profile = "SPY",
  Type    = "BMK"
)

plot_df <- bind_rows(cum_list, spy_cum) %>%
  mutate(
    Profile = factor(Profile, levels = c("INC", "BAL", "GRO", "SPY")),
    Type    = factor(Type,    levels = c("SAA", "TAA", "BMK")),
    series  = paste0(Profile, " ", Type),
    line_col = case_when(
      Type == "BMK"                   ~ "#333333",
      Type == "TAA" & Profile == "INC" ~ "#1E6E7A",
      Type == "SAA" & Profile == "INC" ~ "#7FC8D4",
      Type == "TAA" & Profile == "BAL" ~ "#1B2A4A",
      Type == "SAA" & Profile == "BAL" ~ "#7A90C0",
      Type == "TAA" & Profile == "GRO" ~ "#8B6510",
      Type == "SAA" & Profile == "GRO" ~ "#D4B870",
      TRUE                             ~ "#aaaaaa"
    ),
    line_type = if_else(Type == "SAA", "dashed", "solid"),
    line_size = if_else(Type == "BMK", 1.2, 0.9)
  )

ggplot(plot_df, aes(x = Date, y = CumRet,
                    colour    = line_col,
                    linetype  = line_type,
                    linewidth = line_size,
                    group     = series)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(alpha = 0.9) +
  scale_colour_identity() +
  scale_linetype_identity() +
  scale_linewidth_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  facet_wrap(~Profile, ncol = 1, scales = "free_y") +
  labs(
    title    = "IC Portfolio Cumulative Returns — USD · SAA vs TAA",
    subtitle = "Solid = TAA active  |  Dashed = SAA baseline  |  Black = SPY benchmark\nWeights rescaled to ETF-proxied universe (excludes AI HF & S&P hedge overlay)",
    x = NULL, y = "Cumulative Return",
    caption  = "IC weights: 13-Mar-2026  ·  Returns: full available history"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "#f0f3f8", colour = NA)
  )
```

```{r ic-cumret-combined, fig.height=5}
# ── Combined: all 3 TAA profiles + SPY on a single panel ─────────────────────
taa_only_df <- plot_df %>%
  filter(Type %in% c("TAA", "BMK"))

# Annualised return per series for end labels
.ann_ret_vec <- function(ret_xts) {
  rv <- as.numeric(zoo::coredata(ret_xts))
  mean(rv[!is.na(rv)]) * 252
}

ann_ret_map <- bind_rows(
  purrr::map_df(profiles, function(p) {
    r <- .ic_portfolio_ret(p, "taa")
    if (is.null(r)) return(NULL)
    tibble(series = paste0(p, " TAA"), ann_ret = .ann_ret_vec(r))
  }),
  tibble(series = "SPY BMK", ann_ret = .ann_ret_vec(spy_ret))
)

end_labels <- taa_only_df %>%
  group_by(series) %>%
  slice_max(Date, n = 1) %>%
  ungroup() %>%
  left_join(ann_ret_map, by = "series") %>%
  mutate(
    profile_short = sub(" .*", "", series),
    label = paste0(profile_short, "  ",
                   scales::percent(ann_ret, accuracy = 0.1))
  )

ggplot(taa_only_df, aes(x = Date, y = CumRet,
                        colour    = line_col,
                        linewidth = line_size,
                        group     = series)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(alpha = 0.9) +
  ggrepel::geom_text_repel(
    data          = end_labels,
    aes(label = label, colour = line_col),
    size          = 3.5,
    fontface      = "bold",
    direction     = "y",
    nudge_x       = 120,
    hjust         = 0,
    segment.size  = 0.3,
    segment.alpha = 0.5,
    show.legend   = FALSE
  ) +
  scale_colour_identity() +
  scale_linewidth_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               expand = expansion(mult = c(0.02, 0.12))) +
  labs(
    title    = "IC TAA All Profiles vs SPY — Cumulative Return",
    subtitle = "INC · BAL · GRO (TAA active weights)  |  Black = SPY benchmark\nLabels = annualised return over full history",
    x = NULL, y = "Cumulative Return",
    caption  = "IC weights: 13-Mar-2026  ·  Returns: full available history"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
    panel.grid.minor = element_blank()
  )
```

```{r ic-cumret-stats}
# Risk/return stats per series
# Base R max drawdown — avoids PerformanceAnalytics date-rowname requirement
.base_maxdd <- function(x) {
  cum      <- cumprod(1 + x) - 1
  high_wat <- cummax(cum)
  min(cum - high_wat)          # always <= 0
}

stats_df <- purrr::map_df(profiles, function(p) {
  purrr::map_df(c("saa", "taa"), function(wt_col) {
    r <- .ic_portfolio_ret(p, wt_col)
    if (is.null(r)) return(NULL)
    rv <- as.numeric(zoo::coredata(r))
    rv <- rv[!is.na(rv)]
    tibble(
      Profile = p,
      Type    = toupper(wt_col),
      Ann_Ret = mean(rv) * 252,
      Ann_Vol = sd(rv)   * sqrt(252),
      Sharpe  = (mean(rv) * 252) / (sd(rv) * sqrt(252)),
      Max_DD  = .base_maxdd(rv)
    )
  })
}) %>%
  mutate(across(c(Ann_Ret, Ann_Vol, Sharpe, Max_DD), ~ round(., 3)))

stats_df %>%
  mutate(
    Ann_Ret = scales::percent(Ann_Ret, accuracy = 0.1),
    Ann_Vol = scales::percent(Ann_Vol, accuracy = 0.1),
    Max_DD  = scales::percent(-Max_DD, accuracy = 0.1)
  ) %>%
  kable(col.names = c("Profile", "Type", "Ann. Return",
                      "Ann. Vol", "Sharpe", "Max DD"),
        align = "llrrrr",
        caption = "IC portfolio risk/return — ETF proxy universe · full history") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE) %>%
  row_spec(which(stats_df$Type == "TAA"), bold = TRUE)
```

## 8.6 BAL Sub-Portfolio Analysis — FI & Equity

<span class="section-label lbl-navy">BAL · USD · SAA weights</span>

Deep-dive on the two core building blocks of the BAL SAA: the **Fixed Income
sub-portfolio** (`bal_fi_pf`: LQD / IEF / HYG / EMB, normalised to 100%) and
the **Equity sub-portfolio** (`bal_eq_pf`: URTH / SPY / AAXJ, normalised to
100%). Each is benchmarked against its natural reference — AGG for FI, URTH for
Equity.

```{r bal-load, include=FALSE}
source(here("input/bal_saa_portfolios.R"))
```

```{r bal-weights, fig.width=11, fig.height=4}
# ── Weight bar charts side-by-side ────────────────────────────────────────────
fi_pal  <- c(LQD = "#1B2A4A", IEF = "#2E5073", HYG = "#C1121F", EMB = "#8B6510")
eq_pal  <- c(URTH = "#1E6E7A", SPY = "#2D6A4F", AAXJ = "#8B6510")

p_fi_bar <- ggplot(bal_fi_pf,
    aes(x = reorder(etf_proxy, -weight), y = weight, fill = etf_proxy)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(weight, accuracy = 0.1)),
            vjust = -0.4, fontface = "bold", size = 4) +
  scale_fill_manual(values = fi_pal) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(title = "bal_fi_pf — Fixed Income",
       subtitle = "Weights normalised to 100%",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank())

p_eq_bar <- ggplot(bal_eq_pf,
    aes(x = reorder(etf_proxy, -weight), y = weight, fill = etf_proxy)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(weight, accuracy = 0.1)),
            vjust = -0.4, fontface = "bold", size = 4) +
  scale_fill_manual(values = eq_pal) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(title = "bal_eq_pf — Equity",
       subtitle = "Weights normalised to 100%",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank())

p_fi_bar + p_eq_bar
```

```{r bal-cumret, fig.width=11, fig.height=5}
# ── Helper: build portfolio return series from a bal_*_pf tibble ──────────────
.pf_ret <- function(pf_df, ret_xts = xts_d_abs) {
  tickers_ok <- pf_df$etf_proxy[pf_df$etf_proxy %in% colnames(ret_xts)]
  w <- pf_df %>%
    filter(etf_proxy %in% tickers_ok) %>%
    mutate(weight = weight / sum(weight))        # re-normalise after any drops
  ret_subset <- ret_xts[, w$etf_proxy, drop = FALSE]
  wt_vec     <- setNames(w$weight, w$etf_proxy)
  xts(as.numeric(ret_subset %*% wt_vec[colnames(ret_subset)]),
      order.by = zoo::index(ret_subset))
}

# Build series
r_fi   <- .pf_ret(bal_fi_pf)
r_eq   <- .pf_ret(bal_eq_pf)
r_agg  <- xts_d_abs[, "AGG"]
r_urth <- xts_d_abs[, "URTH"]

.to_cum_df <- function(ret_xts, label) {
  data.frame(Date   = zoo::index(ret_xts),
             CumRet = as.numeric(cumprod(1 + zoo::coredata(ret_xts)) - 1),
             Series = label)
}

cum_bal <- bind_rows(
  .to_cum_df(r_fi,   "bal_fi_pf"),
  .to_cum_df(r_eq,   "bal_eq_pf"),
  .to_cum_df(r_agg,  "AGG (FI bmk)"),
  .to_cum_df(r_urth, "URTH (EQ bmk)")
) %>%
  mutate(
    colour    = case_when(
      Series == "bal_fi_pf"     ~ "#1B2A4A",
      Series == "AGG (FI bmk)"  ~ "#7A90C0",
      Series == "bal_eq_pf"     ~ "#1E6E7A",
      Series == "URTH (EQ bmk)" ~ "#7FC8D4"
    ),
    linetype  = if_else(grepl("bmk", Series), "dashed", "solid"),
    linewidth = if_else(grepl("bmk", Series), 0.7, 1.1)
  )

# End labels with Ann.Ret
.ann_ret <- function(r) mean(as.numeric(zoo::coredata(r)), na.rm = TRUE) * 252

end_lbl <- cum_bal %>%
  group_by(Series) %>% slice_max(Date, n = 1) %>% ungroup() %>%
  mutate(
    ann_ret = c(.ann_ret(r_fi), .ann_ret(r_eq),
                .ann_ret(r_agg), .ann_ret(r_urth))[
      match(Series, c("bal_fi_pf","bal_eq_pf","AGG (FI bmk)","URTH (EQ bmk)"))],
    label = paste0(Series, "  ", scales::percent(ann_ret, accuracy = 0.1))
  )

ggplot(cum_bal, aes(x = Date, y = CumRet,
                    colour = colour, linetype = linetype,
                    linewidth = linewidth, group = Series)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(alpha = 0.9) +
  ggrepel::geom_text_repel(
    data = end_lbl, aes(label = label, colour = colour),
    size = 3.3, fontface = "bold", direction = "y",
    nudge_x = 120, hjust = 0,
    segment.size = 0.3, segment.alpha = 0.5, show.legend = FALSE
  ) +
  scale_colour_identity() +
  scale_linetype_identity() +
  scale_linewidth_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               expand = expansion(mult = c(0.02, 0.14))) +
  labs(
    title    = "BAL Sub-Portfolios — Cumulative Return vs Benchmark",
    subtitle = "Solid = sub-portfolio  |  Dashed = benchmark  |  Labels = ann. return",
    x = NULL, y = "Cumulative Return",
    caption  = "bal_fi_pf: LQD/IEF/HYG/EMB  ·  bal_eq_pf: URTH/SPY/AAXJ  ·  Full history"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title       = element_text(face = "bold", colour = "#1B2A4A"),
        panel.grid.minor = element_blank())
```

```{r bal-stats}
# ── Risk/return stats for both sub-portfolios + benchmarks ────────────────────
.stats_row <- function(ret_xts, label) {
  rv <- as.numeric(zoo::coredata(ret_xts))
  rv <- rv[!is.na(rv)]
  tibble(Portfolio = label,
         Ann_Ret   = mean(rv) * 252,
         Ann_Vol   = sd(rv)   * sqrt(252),
         Sharpe    = (mean(rv) * 252) / (sd(rv) * sqrt(252)),
         Max_DD    = .base_maxdd(rv))
}

bal_stats <- bind_rows(
  .stats_row(r_fi,   "bal_fi_pf"),
  .stats_row(r_agg,  "AGG (FI bmk)"),
  .stats_row(r_eq,   "bal_eq_pf"),
  .stats_row(r_urth, "URTH (EQ bmk)")
) %>%
  mutate(across(c(Ann_Ret, Ann_Vol, Sharpe, Max_DD), ~ round(., 3)))

bal_stats %>%
  mutate(
    Ann_Ret = scales::percent(Ann_Ret, accuracy = 0.1),
    Ann_Vol = scales::percent(Ann_Vol, accuracy = 0.1),
    Max_DD  = scales::percent(-Max_DD, accuracy = 0.1)
  ) %>%
  kable(col.names = c("Portfolio", "Ann. Return", "Ann. Vol", "Sharpe", "Max DD"),
        align = "lrrrr",
        caption = "BAL sub-portfolio risk/return vs benchmarks · full available history") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
  row_spec(c(1, 3), bold = TRUE) %>%
  row_spec(c(2, 4), color = "grey")
```

---

# 9 · Pair Factsheet — `r params$fs_ticker` vs `r params$fs_bmk`

<span class="section-label lbl-navy">Factsheet</span>

Full 7-panel pair analysis: cumulative wealth, L/S spread, drawdown, return
distribution, 250-day rolling correlation, regime-conditional correlation, and
250-day rolling IR — all regime-shaded against the SPY Fall/Recovery/Consolidation
cycle.

Re-render for any ticker by passing `params`:
```r
rmarkdown::render(
  here("Rmd/saa_execution_report.Rmd"),
  params     = list(fs_ticker = "GLD", fs_bmk = "SPY"),
  output_file = "saa_execution_report_GLD.html",
  output_dir  = here("03_reports")
)
```

```{r factsheet-load, include=FALSE}
source(here("utility/factsheet_functions.R"))
```

```{r factsheet-plot, fig.width=13, fig.height=30, out.width="100%"}
fs_out <- fs_render(
  ticker  = params$fs_ticker,
  bmk     = params$fs_bmk,
  ret_xts = xts_d_abs,
  rt      = build_regime_table(xts_d_abs[, params$fs_bmk])
)
```

```{r factsheet-stats}
fs_out$stats %>%
  kable(caption = paste0(params$fs_ticker, " vs ", params$fs_bmk,
                         " — summary statistics · common date window"),
        align = "lrrrrr") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
  row_spec(1, bold = TRUE, color = "#1D3557") %>%
  row_spec(3, bold = TRUE, color = "#E63946")
```

---

*Report generated by `Rmd/saa_execution_report.Rmd` ·
`scripts_saa_execution/` pipeline · `r Sys.Date()`*
