---
title: "Safe Haven Regime Signal — Methods 1, 2 & 3"
subtitle: "Rolling Correlation · LASSO Logistic · Regime-Difference Eigenanalysis"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    latex_engine: xelatex
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 8
)

library(MASS)          # before tidyverse — prevents MASS::select masking dplyr::select
library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(scales)
library(patchwork)
library(here)
library(glmnet)
library(pROC)

select <- dplyr::select   # belt-and-suspenders namespace guard

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

# ── Shared parameters ─────────────────────────────────────────────────────────
MASTER    <- "SPY"
T_FALL    <- 0.10
T_CRUISE  <- 0.05
ROLL_WIN  <- 60
JUMP_MULT <- 2.0
HORIZON   <- 63
MAX_LAG   <- 120
CV_FOLDS  <- 5

# ── Sensitivity groups ────────────────────────────────────────────────────────
groups <- list(
  Credit     = "HYG",
  Equities   = c("XLF", "XLY", "XLK"),
  FX         = c("FXF", "FXY", "EMLC"),
  Safe_Haven = c("GLD", "IEF", "TLT", "UUP")
)
all_tickers <- unlist(groups)

group_colours <- c(
  Credit     = "#E63946",
  Equities   = "#457B9D",
  FX         = "#F4A261",
  Safe_Haven = "#2D6A4F"
)

REGIME_COLORS <- c(Fall = "#D90429", Recovery = "#F77F00", Consolidation = "#2D6A4F")

# ── Regime table ──────────────────────────────────────────────────────────────
spy_ret <- xts_ret[, MASTER]
rt      <- build_regime_table(spy_ret, T_FALL, T_CRUISE)

regime_daily <- rt %>%
  rowwise() %>%
  mutate(date = list(seq(as.Date(xmin), as.Date(xmax), by = "day"))) %>%
  unnest(date) %>%
  dplyr::select(date, regime) %>%
  ungroup()

shade_fall <- rt %>% filter(regime == "Fall")
shade_rec  <- rt %>% filter(regime == "Recovery")
```

---

## Overview

This report investigates whether a curated set of **sensitivity-group ETFs** carry
early regime signal about **SPY drawdown phases**.  The north star is a 3-state
SPY regime engine:

| State | Colour | Condition |
|---|---|---|
| **Fall** | <span style="color:#D90429">■</span> Red | Running drawdown from recent peak ≥ 10% |
| **Recovery** | <span style="color:#F77F00">■</span> Orange | After Fall trough, climbing back to peak |
| **Consolidation** | <span style="color:#2D6A4F">■</span> Green | Expansionary phase |

Three methods are applied in increasing sophistication:

| Method | Approach | Key Question |
|---|---|---|
| **1 — Rolling Correlation** | 60-day rolling corr vs SPY + derivatives | Does correlation break down before Fall? |
| **2 — LASSO Logistic** | Regularised logistic regression | Can feature signals predict P(Fall in 63 days)? |
| **3 — Eigenanalysis** | Regime-difference covariance matrix Δ | Which linear combinations change structure most at transitions? |

### Sensitivity Groups & Ticker Rationale

```{r ticker-rationale, echo=FALSE}
tribble(
  ~Group,       ~Ticker, ~`rc60 HIGH = ...`,                                ~`rc60 DROPS = ...`,
  "Credit",     "HYG",   "Credit spreads compressed — cycle healthy",       "Credit stress leads equity drawdowns",
  "Equities",   "XLF",   "Lending conditions normal, yield curve OK",        "Credit tightening, bank stress",
  "Equities",   "XLY",   "Consumer spending intact — everything is fine",    "Rotation out of cyclicals, consumer stress",
  "Equities",   "XLK",   "Growth expectations intact, rates benign",         "Rate pain or earnings-growth fears",
  "FX",         "FXF",   "Carry trades intact, no CHF safe-haven bid",       "Capital fleeing to Switzerland",
  "FX",         "FXY",   "Carry trades intact, risk appetite healthy",       "Carry unwind → leveraged positions unwinding",
  "FX",         "EMLC",  "Global risk appetite strong, USD stable",          "USD surge, capital flight from EM",
  "Safe Haven", "GLD",   "Inflation / reflation trade — not yet stress",     "Classic flight-to-safety activation",
  "Safe Haven", "IEF",   "Goldilocks — bonds and equities both bid",         "Rate cut bets, risk-off sovereign bond bid",
  "Safe Haven", "TLT",   "Duration risk not rewarded yet — risk-on",         "Textbook flight-to-safety trade",
  "Safe Haven", "UUP",   "Dollar-smile risk-on leg, global growth bid",      "Ambiguous — direction of decouple matters"
) %>%
  knitr::kable(align = "llll")
```

```{r data-summary, echo=TRUE}
cat(sprintf("SPY data  : %s → %s  (%d trading days)\n",
            as.character(start(spy_ret)),
            as.character(end(spy_ret)),
            nrow(spy_ret)))
cat(sprintf("Regime periods : %d  |  Fall periods: %d\n",
            nrow(rt),
            sum(rt$regime == "Fall")))
cat(sprintf("Current regime : %s  (since %s)\n",
            as.character(tail(rt$regime, 1)),
            as.character(as.Date(tail(rt$xmin, 1)))))
```

---

## Method 1 — Rolling Correlation Analysis

**Hypothesis:** sensitivity-group ETFs exhibit correlation *breakdown* with SPY
ahead of Fall regimes.  We track the 60-day rolling correlation, its first
derivative (rate of change), and *jump* events where the daily change exceeds
2× its own rolling standard deviation.

```{r m1-compute, include=FALSE}
compute_roll_corr <- function(ticker, win = ROLL_WIN) {
  tk_r  <- xts_ret[, ticker]
  dates <- index(spy_ret)

  roll_corr <- rollapply(
    merge(spy_ret, tk_r),
    width     = win,
    FUN       = function(m) cor(m[, 1], m[, 2], use = "complete.obs"),
    by.column = FALSE,
    align     = "right",
    fill      = NA
  )
  corr_vec <- as.numeric(roll_corr)
  d1       <- c(NA, diff(corr_vec))
  d2       <- c(NA, diff(d1))
  d1_sd    <- as.numeric(rollapply(d1, width = win, FUN = sd,
                                    na.rm = TRUE, align = "right", fill = NA))
  is_jump  <- abs(d1) > JUMP_MULT * d1_sd

  tibble(date = dates, ticker = ticker,
         corr = corr_vec, d1 = d1, d2 = d2, d1_sd = d1_sd, is_jump = is_jump)
}

roll_df <- map_dfr(all_tickers, compute_roll_corr) %>%
  left_join(
    rt %>% dplyr::select(regime, xmin, xmax) %>%
      rowwise() %>%
      mutate(date = list(seq(as.Date(xmin), as.Date(xmax), by = "day"))) %>%
      unnest(date) %>%
      dplyr::select(date, regime),
    by = "date"
  ) %>%
  mutate(
    group  = case_when(
      ticker %in% groups$Credit     ~ "Credit",
      ticker %in% groups$Equities   ~ "Equities",
      ticker %in% groups$FX         ~ "FX",
      ticker %in% groups$Safe_Haven ~ "Safe_Haven"
    ),
    group  = factor(group, levels = names(groups)),
    ticker = factor(ticker, levels = all_tickers),
    regime = factor(regime, levels = c("Fall", "Recovery", "Consolidation"))
  )
```

### 1.1 Rolling Correlation vs SPY

Circles mark correlation *jump* events — days where the daily change in
rolling correlation exceeds `r JUMP_MULT`× its own rolling standard deviation.
A clustering of jumps immediately before red shading would indicate the
breakdown is detectable in advance.

```{r m1-plot1, fig.height=14}
ggplot(roll_df %>% filter(!is.na(corr)),
       aes(x = date, y = corr, color = ticker)) +
  geom_rect(data = shade_fall,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#D90429", alpha = 0.08, inherit.aes = FALSE) +
  geom_hline(yintercept = 0,    color = "grey40", linewidth = 0.5) +
  geom_hline(yintercept = -0.5, color = "#D90429", linetype = "dashed",
             alpha = 0.4, linewidth = 0.4) +
  geom_hline(yintercept =  0.5, color = "#2D6A4F", linetype = "dashed",
             alpha = 0.4, linewidth = 0.4) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  geom_point(data = roll_df %>% filter(is_jump == TRUE, !is.na(corr)),
             aes(x = date, y = corr),
             shape = 21, fill = "white", size = 1.8, stroke = 0.8) +
  facet_wrap(~ group, ncol = 1, scales = "free_y") +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.5),
                     labels = number_format(accuracy = 0.1)) +
  scale_x_date(expand = expansion(mult = 0.01)) +
  labs(title    = sprintf("%d-Day Rolling Correlation vs %s — by Sensitivity Group",
                          ROLL_WIN, MASTER),
       subtitle = "Red shading = Fall regime  |  Circles = jumps (|Δcorr| > 2σ)  |  Dashed = ±0.5",
       x = NULL, y = "Rolling Correlation", color = "Ticker") +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold", size = 10),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 1.2 First Derivative of Rolling Correlation

The rate of change (Δcorr/day) shows how *fast* correlations are shifting.
Large spikes — particularly negative ones during pre-Fall windows — indicate
accelerating decoupling.

```{r m1-plot2, fig.height=14}
ggplot(roll_df %>% filter(!is.na(d1)),
       aes(x = date, y = d1, color = ticker)) +
  geom_rect(data = shade_fall,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#D90429", alpha = 0.08, inherit.aes = FALSE) +
  geom_hline(yintercept = 0, color = "grey40", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.7) +
  geom_point(data = roll_df %>% filter(is_jump == TRUE, !is.na(d1)),
             aes(x = date, y = d1),
             shape = 21, fill = "white", size = 2, stroke = 0.9) +
  facet_wrap(~ group, ncol = 1, scales = "free_y") +
  scale_x_date(expand = expansion(mult = 0.01)) +
  labs(title    = "First Derivative of Rolling Correlation vs SPY",
       subtitle = "Daily change in 60-day rolling corr  |  Circles = jumps (> 2σ move)",
       x = NULL, y = "Δ Correlation / day", color = "Ticker") +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold", size = 10),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 1.3 Regime-Conditional Correlation Distribution

Does each ticker's correlation with SPY *systematically differ* across regimes?
A strong safe-haven signal would show negative (or zero) correlation during Fall
and positive correlation during Consolidation.

```{r m1-plot3, fig.height=9}
ggplot(roll_df %>% filter(!is.na(corr), !is.na(regime)),
       aes(x = regime, y = corr, fill = regime)) +
  geom_hline(yintercept = 0, color = "grey40", linewidth = 0.5) +
  geom_boxplot(alpha = 0.8, outlier.size = 0.8, width = 0.6) +
  scale_fill_manual(values = REGIME_COLORS, guide = "none") +
  facet_wrap(~ group + ticker, ncol = 4, scales = "free_y") +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.5)) +
  labs(title    = "Correlation vs SPY — Distribution by Regime Phase",
       subtitle = sprintf("North Star: SPY Fall ≥%.0f%%  |  60-day rolling window", T_FALL * 100),
       x = NULL, y = "Rolling Correlation") +
  theme_minimal(base_size = 10) +
  theme(strip.text = element_text(face = "bold", size = 8),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 1.4 Jump Frequency by Regime

If correlation jumps are a genuine early warning, they should cluster
disproportionately in Fall regimes.

```{r m1-plot4, fig.height=5}
roll_df %>%
  filter(!is.na(regime), !is.na(is_jump)) %>%
  group_by(group, ticker, regime) %>%
  summarise(n_days   = n(),
            n_jumps  = sum(is_jump, na.rm = TRUE),
            jump_pct = n_jumps / n_days, .groups = "drop") %>%
  ggplot(aes(x = ticker, y = jump_pct, fill = regime)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = REGIME_COLORS) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  facet_wrap(~ group, scales = "free_x", nrow = 1) +
  labs(title    = "Correlation Jump Frequency by SPY Regime",
       subtitle = "Jump = |Δcorr| > 2σ  |  Higher Fall bar → early warning candidate",
       x = NULL, y = "Jump frequency (% of days)", fill = "SPY Regime") +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold", size = 10),
        axis.text.x = element_text(angle = 30, hjust = 1),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

---

## Method 2 — LASSO Logistic Regression

**Question:** can multi-dimensional signals from sensitivity-group ETFs
*predict* whether SPY will be in a Fall regime `r HORIZON` trading days ahead
(≈ 3 months)?

Seven feature families are engineered per ticker, then a LASSO-regularised
logistic regression selects the active subset via walk-forward cross-validation.

| Family | Features | Hypothesis |
|---|---|---|
| A. Rolling Corr | 60d/20d corr, Δcorr, jump count | Corr breaks down before Fall |
| B. Momentum Z | 20/60/120d momentum z-scored | Exhaustion signals regime end |
| C. Mom Divergence | ticker mom − SPY mom, z-scored | Safe havens decouple early |
| D. Vol Ratio | ticker vol / SPY vol, z-scored | Vol shift precedes Fall |
| E. Credit Spread | HYG − IEF momentum spread | Credit stress leads equity Fall |
| F. SH Composite | GLD + IEF + TLT convergence z-score | Multi-asset flight-to-safety |
| G. Regime Memory | current duration vs historical avg | Regimes mean-revert in length |

```{r m2-compute, include=FALSE, cache=TRUE}
roll_zscore <- function(x, win) {
  mu  <- as.numeric(rollapply(x, win, mean, na.rm = TRUE, align = "right", fill = NA))
  sig <- as.numeric(rollapply(x, win, sd,   na.rm = TRUE, align = "right", fill = NA))
  (as.numeric(x) - mu) / sig
}

roll_mom <- function(x, win)
  as.numeric(rollapply(x, win, FUN = function(r) prod(1 + r) - 1,
                        align = "right", fill = NA))

roll_corr_spy <- function(tk_r, win = ROLL_WIN)
  as.numeric(rollapply(
    merge(spy_ret, tk_r), width = win,
    FUN = function(m) cor(m[,1], m[,2], use = "complete.obs"),
    by.column = FALSE, align = "right", fill = NA))

dates <- as.Date(index(spy_ret))

feature_list <- list()
for (tk in all_tickers) {
  tk_r       <- as.numeric(xts_ret[, tk])
  rc60       <- roll_corr_spy(xts_ret[, tk], 60)
  rc20       <- roll_corr_spy(xts_ret[, tk], 20)
  d1_rc60    <- c(NA, diff(rc60))
  d1_sd      <- as.numeric(rollapply(d1_rc60, ROLL_WIN, sd, na.rm = TRUE,
                                      align = "right", fill = NA))
  jump_60    <- as.numeric(abs(d1_rc60) > JUMP_MULT * d1_sd)
  jump_sum20 <- as.numeric(rollapply(jump_60, 20, sum, na.rm = TRUE,
                                      align = "right", fill = NA))

  mom20  <- roll_mom(xts_ret[, tk], 20)
  mom60  <- roll_mom(xts_ret[, tk], 60)
  mom120 <- roll_mom(xts_ret[, tk], 120)
  mz20   <- roll_zscore(mom20,  60)
  mz60   <- roll_zscore(mom60,  120)
  mz120  <- roll_zscore(mom120, 252)

  spy_mom20 <- roll_mom(spy_ret, 20)
  spy_mom60 <- roll_mom(spy_ret, 60)
  div_z20   <- roll_zscore(mom20 - spy_mom20, 60)
  div_z60   <- roll_zscore(mom60 - spy_mom60, 120)

  vol_tk      <- as.numeric(rollapply(xts_ret[, tk], 20, sd, na.rm = TRUE,
                                       align = "right", fill = NA)) * sqrt(252)
  vol_spy     <- as.numeric(rollapply(spy_ret, 20, sd, na.rm = TRUE,
                                       align = "right", fill = NA)) * sqrt(252)
  vol_ratio_z <- roll_zscore(vol_tk / vol_spy, 60)

  feature_list[[tk]] <- tibble(
    date = dates, ticker = tk,
    rc60 = rc60, rc20 = rc20, d1_rc60 = d1_rc60, jump_sum20 = jump_sum20,
    mz20 = mz20, mz60 = mz60, mz120 = mz120,
    div_z20 = div_z20, div_z60 = div_z60,
    vol_ratio_z = vol_ratio_z)
}

features_long <- bind_rows(feature_list)

hyg_mom60       <- roll_mom(xts_ret[, "HYG"], 60)
ief_mom60       <- roll_mom(xts_ret[, "IEF"], 60)
credit_spread_z <- roll_zscore(hyg_mom60 - ief_mom60, 120)

sh_composite <- (
  roll_zscore(roll_mom(xts_ret[, "GLD"], 60), 120) +
  roll_zscore(roll_mom(xts_ret[, "IEF"], 60), 120) +
  roll_zscore(roll_mom(xts_ret[, "TLT"], 60), 120)
) / 3

regime_dur_avg <- rt %>%
  group_by(regime) %>%
  summarise(avg_days = mean(days), sd_days = sd(days), .groups = "drop")

regime_dur_daily <- regime_daily %>%
  left_join(
    rt %>% mutate(start = as.Date(xmin)) %>% dplyr::select(start, regime, days),
    by = "regime"
  ) %>%
  group_by(date) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  mutate(current_dur = as.numeric(date - as.Date(
    rt$xmin[findInterval(date, as.Date(rt$xmin))]))) %>%
  left_join(regime_dur_avg, by = "regime") %>%
  mutate(dur_z = (current_dur - avg_days) / sd_days) %>%
  dplyr::select(date, regime, dur_z)

features_wide <- features_long %>%
  pivot_wider(id_cols = date, names_from = ticker,
              values_from = c(rc60, rc20, d1_rc60, jump_sum20,
                              mz20, mz60, mz120, div_z20, div_z60, vol_ratio_z),
              names_glue = "{ticker}_{.value}") %>%
  mutate(credit_spread_z = credit_spread_z,
         sh_composite    = sh_composite) %>%
  left_join(regime_dur_daily %>% dplyr::select(date, dur_z), by = "date")

target_df <- regime_daily %>%
  mutate(future_date = date - HORIZON,
         in_fall     = as.integer(regime == "Fall")) %>%
  dplyr::select(future_date, in_fall) %>%
  rename(date = future_date)

model_df <- features_wide %>%
  inner_join(target_df, by = "date") %>%
  drop_na() %>%
  arrange(date)

X            <- model_df %>% dplyr::select(-date, -in_fall) %>% as.matrix()
y            <- model_df$in_fall
dates_model  <- model_df$date

n       <- nrow(X)
fold_sz <- floor(n / (CV_FOLDS + 1))

cv_results <- map_dfr(seq_len(CV_FOLDS), function(k) {
  train_end <- k * fold_sz
  test_idx  <- (train_end + 1):min(train_end + fold_sz, n)
  cv_fit    <- cv.glmnet(X[1:train_end, ], y[1:train_end],
                          family = "binomial", alpha = 1,
                          nfolds = 5, type.measure = "auc")
  pred      <- as.numeric(predict(cv_fit, X[test_idx, ],
                                   s = "lambda.min", type = "response"))
  tibble(date = dates_model[test_idx], prob = pred,
         actual = y[test_idx], fold = k)
})

cv_full  <- cv.glmnet(X, y, family = "binomial", alpha = 1,
                       nfolds = 10, type.measure = "auc")
coef_mat <- coef(cv_full, s = "lambda.min")

coef_df <- tibble(feature = rownames(coef_mat)[-1],
                   coef    = as.numeric(coef_mat)[-1]) %>%
  filter(coef != 0) %>%
  arrange(desc(abs(coef))) %>%
  mutate(direction = if_else(coef > 0, "Increases Fall P", "Decreases Fall P"),
         feature   = fct_reorder(feature, abs(coef)))
```

### 2.1 Predicted P(Fall) vs Actual Regimes

Each panel is one walk-forward CV fold.  Red shading marks *actual* Fall periods.
The line is the *predicted* probability.  A good model anticipates the red zones.

```{r m2-plot1, fig.height=16}
ggplot(cv_results, aes(x = date, y = prob)) +
  geom_rect(data = shade_fall,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#D90429", alpha = 0.12, inherit.aes = FALSE) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
  geom_line(color = "#1D3557", linewidth = 0.8) +
  geom_ribbon(aes(ymin = 0, ymax = prob), fill = "#1D3557", alpha = 0.15) +
  facet_wrap(~ fold, ncol = 1,
             labeller = labeller(fold = ~ paste("CV Fold", .x))) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_x_date(expand = expansion(mult = 0.01)) +
  labs(title    = sprintf("P(SPY Fall in %d days) — Walk-Forward CV", HORIZON),
       subtitle = "Red shading = actual SPY Fall  |  Line = predicted probability  |  Dashed = 0.5 threshold",
       x = NULL, y = "P(Fall)") +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold", size = 9),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 2.2 LASSO Feature Importance

Only features with **non-zero** LASSO coefficients are shown — the regulariser
has eliminated all others.  Bar length = coefficient magnitude.
Red = increases Fall probability; green = decreases it.

> **Reading the coefficient scale:** many features (e.g. rolling correlations)
> have very small standard deviations, so a large raw coefficient can translate
> to a modest *standardised* coefficient.  Multiply `coef × sd(feature)` to
> compare features on equal footing.

```{r m2-plot2, fig.height=7}
ggplot(coef_df, aes(x = feature, y = coef, fill = direction)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0, color = "grey30") +
  coord_flip() +
  scale_fill_manual(values = c("Increases Fall P" = "#D90429",
                                "Decreases Fall P" = "#2D6A4F")) +
  labs(title    = "LASSO Feature Importance — P(SPY Fall)",
       subtitle = sprintf("Horizon: %d days | Only non-zero coefficients shown", HORIZON),
       x = NULL, y = "Coefficient", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 2.3 ROC Curves — Walk-Forward CV Folds

AUC > 0.5 indicates genuine predictive power above a random classifier.
Each fold covers a different time window; consistent AUC across folds rules
out a lucky single period.

```{r m2-plot3, fig.height=6}
roc_df <- cv_results %>%
  group_by(fold) %>%
  group_map(~ {
    # Skip folds where only one class is present — roc() requires both 0 and 1
    if (length(unique(.x$actual)) < 2) return(NULL)
    r   <- roc(.x$actual, .x$prob, quiet = TRUE)
    auc <- as.numeric(auc(r))
    tibble(fpr = 1 - r$specificities, tpr = r$sensitivities,
           fold = .y$fold, auc = auc)
  }) %>%
  bind_rows()

auc_labels <- roc_df %>%
  distinct(fold, auc) %>%
  mutate(label = sprintf("Fold %d  AUC=%.2f", fold, auc))

ggplot(roc_df, aes(x = fpr, y = tpr, color = factor(fold))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_line(linewidth = 0.9) +
  scale_color_brewer(palette = "Set1", labels = auc_labels$label, name = NULL) +
  labs(title    = "ROC Curves — Walk-Forward CV Folds",
       subtitle = sprintf("Target: SPY Fall regime in %d trading days (~3 months)", HORIZON),
       x = "False Positive Rate", y = "True Positive Rate") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 2.4 Signal Family Importance

Which *family* of features carries the most collective weight?  Summing
`|coef|` within each family normalises for the number of features per group.

```{r m2-plot4, fig.height=5}
coef_df %>%
  mutate(family = case_when(
    str_detect(feature, "rc60|rc20|d1_rc|jump") ~ "A. Rolling Corr",
    str_detect(feature, "mz20|mz60|mz120")      ~ "B. Momentum Z",
    str_detect(feature, "div_z")                 ~ "C. Mom Divergence",
    str_detect(feature, "vol_ratio")             ~ "D. Vol Ratio",
    str_detect(feature, "credit_spread")         ~ "E. Credit Spread",
    str_detect(feature, "sh_composite")          ~ "F. SH Composite",
    str_detect(feature, "dur_z")                 ~ "G. Regime Memory",
    TRUE                                          ~ "Other"
  )) %>%
  group_by(family) %>%
  summarise(total_abs_coef = sum(abs(coef)), .groups = "drop") %>%
  arrange(desc(total_abs_coef)) %>%
  mutate(family = fct_reorder(family, total_abs_coef)) %>%
  ggplot(aes(x = family, y = total_abs_coef)) +
  geom_col(fill = "#1D3557", width = 0.7) +
  coord_flip() +
  labs(title    = "Feature Family Importance — Sum of |LASSO Coefs|",
       subtitle = "Which signal family matters most for predicting SPY Fall?",
       x = NULL, y = "Total |coefficient|") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

---

## Method 3 — Regime-Difference Covariance Eigenanalysis

**Mathematical framing:** three linear-algebra operators are applied to the
return matrix of all `r length(all_tickers)` sensitivity-group tickers.

$$\Delta = \Sigma_{\text{Fall}} - \Sigma_{\text{Normal}}$$

Eigendecomposing Δ reveals:

- **Positive eigenvalue** directions → variance *expands* in Fall (risk assets dispersing)
- **Negative eigenvalue** directions → variance *contracts* in Fall (safe-haven convergence)

The eigenvector with the most negative eigenvalue is, by construction, the
**data-optimal safe-haven composite** — the linear combination of tickers
that clusters most tightly during SPY Fall regimes.

```{r m3-compute, include=FALSE}
ret_mat   <- as.matrix(xts_ret[, all_tickers])
ret_dates <- as.Date(index(xts_ret))

fall_dates   <- regime_daily %>% filter(regime == "Fall")   %>% pull(date)
normal_dates <- regime_daily %>% filter(regime != "Fall")   %>% pull(date)
idx_fall     <- ret_dates %in% fall_dates
idx_normal   <- ret_dates %in% normal_dates

Sigma_fall   <- cov(ret_mat[idx_fall,   ], use = "complete.obs")
Sigma_normal <- cov(ret_mat[idx_normal, ], use = "complete.obs")
Delta        <- Sigma_fall - Sigma_normal
eig          <- eigen(Delta)

n_tickers  <- length(all_tickers)
ev_pos_idx <- which.max(eig$values)
ev_neg_idx <- which.min(eig$values)
ev_pos     <- eig$vectors[, ev_pos_idx]
ev_neg     <- eig$vectors[, ev_neg_idx]

sh_tickers    <- groups$Safe_Haven
sign_correct  <- sign(mean(ev_neg[all_tickers %in% sh_tickers]))
ev_neg_signed <- ev_neg * sign_correct

sh_composite_derived <- as.numeric(ret_mat %*% ev_neg_signed)
sh_composite_xts     <- xts(sh_composite_derived, order.by = ret_dates)
sh_roll_mu <- as.numeric(rollapply(sh_composite_xts, 60, mean, fill = NA, align = "right"))
sh_roll_sd <- as.numeric(rollapply(sh_composite_xts, 60, sd,   fill = NA, align = "right"))
sh_z       <- (sh_composite_derived - sh_roll_mu) / sh_roll_sd

y_fall  <- as.integer(ret_dates %in% fall_dates)
xcov_df <- map_dfr(seq(0, MAX_LAG, by = 5), function(h) {
  y_lagged <- c(y_fall[(h + 1):length(y_fall)], rep(NA, h))
  keep     <- !is.na(y_lagged)
  map_dfr(seq_along(all_tickers), function(i)
    tibble(lag = h, ticker = all_tickers[i],
           xcov = cov(ret_mat[keep, i], y_lagged[keep])))
})

Sigma_full <- cov(ret_mat, use = "complete.obs")
ridge_eps  <- 1e-6 * mean(diag(Sigma_full))
Precision  <- solve(Sigma_full + diag(ridge_eps, n_tickers))
rownames(Precision) <- colnames(Precision) <- all_tickers

cov_to_long <- function(mat, label) {
  as_tibble(mat, rownames = "row") %>%
    pivot_longer(-row, names_to = "col", values_to = "value") %>%
    mutate(panel = label,
           row   = factor(row, levels = all_tickers),
           col   = factor(col, levels = all_tickers))
}

weight_comparison <- tibble(
  ticker         = all_tickers,
  group          = rep(names(groups), lengths(groups)),
  derived_weight = round(ev_neg_signed, 4),
  hand_crafted   = if_else(all_tickers %in% c("GLD","IEF","TLT"), 1/3, 0)
) %>% arrange(desc(abs(derived_weight)))

# ── Verification signals ──────────────────────────────────────────────────────
roll_mom_vec <- function(tk, win)
  as.numeric(rollapply(xts_ret[, tk], win,
    FUN = function(r) prod(1 + r) - 1, align = "right", fill = NA))

roll_z_vec <- function(x, win) {
  mu  <- as.numeric(rollapply(x, win, mean, na.rm = TRUE, align = "right", fill = NA))
  sig <- as.numeric(rollapply(x, win, sd,   na.rm = TRUE, align = "right", fill = NA))
  (x - mu) / sig
}

xlk_z <- roll_z_vec(roll_mom_vec("XLK", 60), 120)
xlf_z <- roll_z_vec(roll_mom_vec("XLF", 60), 120)
gld_z <- roll_z_vec(roll_mom_vec("GLD", 60), 120)
ief_z <- roll_z_vec(roll_mom_vec("IEF", 60), 120)
tlt_z <- roll_z_vec(roll_mom_vec("TLT", 60), 120)

signal_df <- tibble(
  date  = ret_dates,
  sig_a = xlk_z - xlf_z,
  sig_b = (gld_z + ief_z + tlt_z) / 3,
  xlk_z = xlk_z,
  xlf_z = xlf_z
) %>%
  left_join(regime_daily, by = "date") %>%
  drop_na()
```

### 3.1 Eigenspectrum of Δ

How many directions change structure meaningfully at regime transitions?
The split between positive and negative eigenvalues reveals the balance
between dispersal and convergence forces.

```{r m3-eigenspectrum, fig.height=5}
tibble(index    = seq_along(eig$values),
       lambda   = eig$values,
       sign_lab = if_else(eig$values >= 0, "Dispersal (+)", "Safe-Haven (−)")) %>%
  ggplot(aes(x = index, y = lambda, fill = sign_lab)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0, colour = "grey30") +
  scale_fill_manual(values = c("Dispersal (+)" = "#D90429",
                                "Safe-Haven (−)" = "#2D6A4F")) +
  scale_x_continuous(breaks = seq_along(eig$values),
                     labels = sprintf("EV%d", seq_along(eig$values))) +
  labs(title    = "Eigenspectrum of Δ = Σ_Fall − Σ_Normal",
       subtitle = "Positive → directions more volatile in Fall  |  Negative → directions that converge in Fall",
       x = NULL, y = "Eigenvalue", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 3.2 Covariance Heatmaps: Σ_Fall | Σ_Normal | Δ

```{r m3-heatmap, fig.height=7}
long_df <- bind_rows(
  cov_to_long(Sigma_fall,   "Sigma[Fall]"),
  cov_to_long(Sigma_normal, "Sigma[Normal]"),
  cov_to_long(Delta,        "Delta == Sigma[Fall] - Sigma[Normal]")
) %>%
  left_join(
    bind_rows(
      cov_to_long(Sigma_fall,   "Sigma[Fall]"),
      cov_to_long(Sigma_normal, "Sigma[Normal]"),
      cov_to_long(Delta,        "Delta == Sigma[Fall] - Sigma[Normal]")
    ) %>% group_by(panel) %>%
      summarise(lim = max(abs(value)), .groups = "drop"),
    by = "panel")

ggplot(long_df, aes(x = col, y = fct_rev(row), fill = value)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  facet_wrap(~ panel, ncol = 3, labeller = label_parsed) +
  scale_fill_gradient2(low = "#2D6A4F", mid = "white", high = "#D90429",
                       midpoint = 0,
                       labels = scales::label_scientific(digits = 2)) +
  labs(title    = "Covariance Structure: Fall vs Normal vs Difference",
       subtitle = "Green = converges in Fall  |  Red = disperses in Fall",
       x = NULL, y = NULL, fill = NULL) +
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(face = "bold", size = 9),
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(2, "cm"),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 3.3 Top Eigenvectors of Δ

The loadings of the most positive and most negative eigenvectors reveal
*which asset combinations* drive the structural regime change.

```{r m3-eigenvectors, fig.height=9}
pos_idx <- order(eig$values, decreasing = TRUE)[1:3]
neg_idx <- order(eig$values, decreasing = FALSE)[1:3]

map_dfr(unique(c(pos_idx, neg_idx)), function(i) {
  lam <- eig$values[i]
  tibble(ticker   = all_tickers,
         loading  = eig$vectors[, i],
         lambda   = round(lam, 5),
         ev_label = sprintf("EV %d  λ=%.4f", i, lam),
         type     = if_else(lam > 0, "Dispersal (+λ)", "Safe-Haven Convergence (−λ)"),
         group    = rep(names(groups), lengths(groups)))
}) %>%
  mutate(ticker   = factor(ticker, levels = all_tickers),
         ev_label = fct_reorder(ev_label, lambda)) %>%
  ggplot(aes(x = ticker, y = loading, fill = group)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0, colour = "grey30") +
  facet_wrap(~ ev_label, ncol = 2) +
  scale_fill_manual(values = group_colours) +
  labs(title    = "Eigenvectors of Δ = Σ_Fall − Σ_Normal",
       subtitle = "Negative-eigenvalue vectors = safe-haven convergence directions  |  Positive = risk dispersal",
       x = NULL, y = "Loading", fill = "Group") +
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold", size = 9),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 3.4 Data-Derived Safe-Haven Composite vs SPY Regimes

The eigenvector with the most negative eigenvalue is projected onto daily
returns to produce a *data-optimal* composite.  Compare this to the
hand-crafted `(GLD + IEF + TLT) / 3`.

```{r m3-weights}
knitr::kable(weight_comparison,
             caption = "Derived eigenvector weights vs hand-crafted sh_composite",
             digits = 4, align = "llrr")
```

```{r m3-composite, fig.height=6}
tibble(date = ret_dates, z = sh_z) %>%
  drop_na() %>%
  ggplot(aes(x = date, y = z)) +
  geom_rect(data = shade_fall,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#D90429", alpha = 0.12, inherit.aes = FALSE) +
  geom_rect(data = shade_rec,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#F4A261", alpha = 0.10, inherit.aes = FALSE) +
  geom_hline(yintercept = c(-1, 0, 1), linetype = "dashed",
             colour = c("grey60","grey30","grey60"), linewidth = 0.4) +
  geom_ribbon(aes(ymin = 0, ymax = z), fill = "#2D6A4F", alpha = 0.15) +
  geom_line(colour = "#2D6A4F", linewidth = 0.7) +
  scale_x_date(expand = expansion(mult = 0.01)) +
  scale_y_continuous(breaks = seq(-4, 4, 1)) +
  labs(title    = "Data-Derived Safe-Haven Composite (Δ eigenvector projection)",
       subtitle = paste0("Weights: ",
                         paste(sprintf("%s=%.2f", all_tickers, ev_neg_signed), collapse=" | "),
                         "\nRed = Fall  |  Orange = Recovery  |  Series = 60d rolling z-score"),
       x = NULL, y = "Z-score") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 8))
```

### 3.5 Lagged Cross-Covariance: Cov(ticker_t, Fall_{t+h})

Which ticker return today has the strongest *lead* relationship with
SPY entering a Fall regime `h` days in the future?  No model needed —
pure linear algebra.

```{r m3-xcov, fig.height=7}
xcov_df %>%
  mutate(group  = rep(names(groups), lengths(groups))[match(ticker, all_tickers)],
         ticker = factor(ticker, levels = all_tickers)) %>%
  ggplot(aes(x = lag, y = xcov, colour = group, group = ticker)) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_vline(xintercept = HORIZON, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  annotate("text", x = HORIZON + 2, y = Inf, label = sprintf("h=%d", HORIZON),
           hjust = 0, vjust = 1.5, size = 3, colour = "grey40") +
  geom_line(linewidth = 0.7, alpha = 0.85) +
  facet_wrap(~ group, ncol = 2, scales = "free_y") +
  scale_colour_manual(values = group_colours, guide = "none") +
  scale_x_continuous(breaks = seq(0, MAX_LAG, 20)) +
  labs(title    = "Lagged Cross-Covariance: Cov(ticker_t, Fall_{t+h})",
       subtitle = "How strongly does each ticker return today predict SPY Fall h days ahead?",
       x = "Lead horizon h (trading days)", y = "Cross-covariance") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold", size = 9),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

### 3.6 Precision Matrix — Conditional Independence

The precision matrix Σ⁻¹ shows which tickers carry *unique* information
once all others are known.  High off-diagonal entries indicate strong
partial correlations — potential redundancy in the feature set.

```{r m3-precision, fig.height=6}
prec_long <- as_tibble(Precision, rownames = "row") %>%
  pivot_longer(-row, names_to = "col", values_to = "value") %>%
  mutate(row = factor(row, levels = all_tickers),
         col = factor(col, levels = all_tickers))

lim <- max(abs(prec_long$value[prec_long$row != prec_long$col]))

ggplot(prec_long, aes(x = col, y = fct_rev(row), fill = value)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  geom_text(data = prec_long %>% filter(row == col),
            aes(label = round(value, 0)), size = 2.5, colour = "grey20") +
  scale_fill_gradient2(low = "#2D6A4F", mid = "white", high = "#D90429",
                       midpoint = 0,
                       limits = c(-lim, lim), oob = scales::squish,
                       labels = scales::label_scientific(digits = 1)) +
  labs(title    = "Precision Matrix Σ⁻¹ (Conditional Independence)",
       subtitle = "Red off-diagonal = conditionally positively linked  |  Green = conditionally negatively linked",
       x = NULL, y = NULL, fill = NULL) +
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(2, "cm"),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

---

## Method 3 — Verification: XLK/XLF Spread vs GLD+IEF+TLT

The eigenvector analysis identified that the **dominant convergence direction**
is not the hand-crafted `GLD + IEF + TLT` composite but rather a
**Tech vs Financials spread (XLK − XLF)** combined with gold and short credit.

This section verifies that claim visually and statistically.

> **Signal A:** `XLK_z − XLF_z` (eigenvector-dominant)
>
> **Signal B:** `(GLD_z + IEF_z + TLT_z) / 3` (hand-crafted `sh_composite`)

```{r m3-verify-ts, fig.height=14}
shading <- list(
  geom_rect(data = shade_fall,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#D90429", alpha = 0.12, inherit.aes = FALSE),
  geom_rect(data = shade_rec,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#F4A261", alpha = 0.10, inherit.aes = FALSE),
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40", linewidth = 0.4)
)

base_theme <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(colour = "grey50", size = 8),
        axis.title.x = element_blank())

p1 <- ggplot(signal_df, aes(x = date, y = sig_a)) +
  shading +
  geom_ribbon(aes(ymin = 0, ymax = sig_a), fill = "#457B9D", alpha = 0.20) +
  geom_line(colour = "#457B9D", linewidth = 0.7) +
  scale_y_continuous(breaks = seq(-4, 4, 1)) +
  labs(title    = "Signal A: XLK_z − XLF_z  (Tech vs Financials spread)",
       subtitle = "Eigenvector-dominant term | Drops sharply into Fall") + base_theme

p2 <- ggplot(signal_df, aes(x = date, y = sig_b)) +
  shading +
  geom_ribbon(aes(ymin = 0, ymax = sig_b), fill = "#2D6A4F", alpha = 0.20) +
  geom_line(colour = "#2D6A4F", linewidth = 0.7) +
  scale_y_continuous(breaks = seq(-4, 4, 1)) +
  labs(title    = "Signal B: (GLD_z + IEF_z + TLT_z) / 3  (hand-crafted sh_composite)",
       subtitle = "Traditional safe-haven composite") + base_theme

p3 <- signal_df %>%
  dplyr::select(date, XLK = xlk_z, XLF = xlf_z) %>%
  pivot_longer(-date, names_to = "ticker", values_to = "z") %>%
  ggplot(aes(x = date, y = z, colour = ticker)) +
  shading +
  geom_line(linewidth = 0.65, alpha = 0.9) +
  scale_colour_manual(values = c(XLK = "#1D3557", XLF = "#E63946")) +
  scale_y_continuous(breaks = seq(-4, 4, 1)) +
  labs(title    = "XLK and XLF individually — the divergence during Fall",
       subtitle = "In Fall: XLF (red) collapses faster/deeper → spread widens",
       colour = NULL) +
  base_theme + theme(legend.position = "bottom")

(p1 / p2 / p3) +
  plot_annotation(
    title    = "Time-Series Comparison: Signal A vs Signal B vs Regime",
    subtitle = "Red shading = SPY Fall  |  Orange = Recovery",
    theme = theme(plot.title    = element_text(face = "bold", size = 14),
                  plot.subtitle = element_text(colour = "grey50", size = 9)))
```

### Regime-Conditional Distributions — Statistical Verdict

The violin + boxplot below is the definitive comparison.  A signal with a
clearly separated Fall distribution (lower mean, tighter band) is the stronger
regime indicator.

```{r m3-verify-dist, fig.height=6}
signal_df %>%
  filter(!is.na(regime)) %>%
  mutate(regime = factor(regime, levels = c("Fall", "Recovery", "Consolidation"))) %>%
  dplyr::select(regime, `A: XLK-XLF` = sig_a, `B: sh_composite` = sig_b) %>%
  pivot_longer(-regime, names_to = "signal", values_to = "value") %>%
  { . ->> dist_df_tmp } %>%
  ggplot(aes(x = regime, y = value, fill = regime)) +
  geom_violin(alpha = 0.35, colour = NA, trim = TRUE) +
  geom_boxplot(width = 0.22, outlier.size = 0.5, alpha = 0.8) +
  geom_text(
    data = dist_df_tmp %>%
      group_by(signal, regime) %>%
      summarise(m = mean(value, na.rm = TRUE), .groups = "drop"),
    aes(x = regime, y = m, label = sprintf("μ=%.2f", m)),
    vjust = -0.6, size = 3, fontface = "bold", inherit.aes = FALSE
  ) +
  facet_wrap(~ signal, ncol = 2) +
  scale_fill_manual(values = c(Fall = "#D90429", Recovery = "#F4A261",
                                Consolidation = "#457B9D"), guide = "none") +
  labs(title    = "Regime-Conditional Distributions: Signal A vs Signal B",
       subtitle = "The signal that separates Fall (red) most cleanly is the stronger regime indicator",
       x = NULL, y = "Z-score") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold", size = 10),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(color = "grey50", size = 9))
```

---

## Key Findings

```{r findings-table, echo=FALSE}
tibble(
  `#` = 1:6,
  Finding = c(
    "HYG and XLF show the largest correlation breakdown magnitude entering Fall — credit leads equity",
    "FXY (JPY) jump frequency is disproportionately elevated in Fall — carry unwind is an early signal",
    "LASSO selects XLY_rc60 (coef ≈ −15) as a dominant decreasing-Fall signal; large raw coef reflects compressed correlation scale",
    "The data-optimal safe-haven composite (Δ eigenvector) is dominated by XLK/XLF spread and GLD — not GLD+IEF+TLT",
    "TLT and IEF carry near-zero weight in the Δ eigenvector, consistent with bond/equity co-selloffs in 2022",
    "Lagged cross-covariance peaks at h < 63 for credit tickers (HYG), confirming credit leads the 63-day LASSO horizon"
  )
) %>%
  knitr::kable(align = "cl")
```

---

*North Star: SPY Fall defined as running drawdown from peak ≥ `r scales::percent(T_FALL)`.
All analysis uses daily returns from `xts_ret_returns.rds`.*
