---
title: "SAA Universe — Shiny App Executive Summary"
subtitle: "Core hierarchical universe at each detail level (Depth 0 → 3)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
    highlight: tango
    code_folding: hide
---

<style>
.depth-box {
  background-color: #f0f4f8;
  border-left: 4px solid #2c3e50;
  padding: 10px 16px;
  margin: 12px 0 20px 0;
  border-radius: 0 4px 4px 0;
  font-size: 13px;
}
.bucket-pill {
  display: inline-block;
  padding: 2px 8px;
  border-radius: 10px;
  font-size: 11px;
  font-weight: 700;
  margin-right: 4px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(plotly)
library(reactable)
library(scales)
library(here)
library(lubridate)

# ── Data pipeline (mirrors shiny_dashboard/global.R) ─────────────────────────
source(here("project_tree.R"))
source(here(project_tree$scripts$init))        # etf_metadata
source(here("shiny_dashboard/saa_config.R"))   # SAA_CONFIG

raw_data     <- readRDS(here(project_tree$products$raw_p_d))
tech_summary <- readRDS(here(project_tree$products$tech_summary))

today     <- Sys.Date()
ytd_start <- as.Date(paste0(year(today), "-01-01"))
mtd_start <- floor_date(today, "month")

perf_data <- raw_data %>%
  group_by(symbol) %>%
  arrange(date) %>%
  mutate(adjusted = as.numeric(adjusted)) %>%
  summarise(
    ret_ytd = (last(adjusted) / first(adjusted[date >= ytd_start])) - 1,
    ret_mtd = (last(adjusted) / first(adjusted[date >= mtd_start])) - 1,
    ret_1m  = (last(adjusted) / nth(adjusted, -22L)) - 1,
    ret_5d  = (last(adjusted) / nth(adjusted, -6L))  - 1,
    .groups = "drop"
  ) %>%
  left_join(etf_metadata %>% select(ticker, name, asset_class),
            by = c("symbol" = "ticker")) %>%
  left_join(tech_summary %>% select(ticker, trend_regime),
            by = c("symbol" = "ticker")) %>%
  left_join(SAA_CONFIG, by = c("symbol" = "ticker")) %>%
  filter(!is.na(ret_ytd), !is.na(saa_depth))

# ── Colour helpers ────────────────────────────────────────────────────────────
BUCKET_COLS <- c(Cash = "#3b82f6", FI = "#8b5cf6", EQ = "#10b981", Alt = "#f59e0b")

finviz_scale <- list(
  c(0.00, "#7f0000"), c(0.30, "#cc2222"), c(0.46, "#f4aaaa"),
  c(0.50, "#555555"),
  c(0.54, "#a3f0bc"), c(0.70, "#1a7a2e"), c(1.00, "#00AA44")
)

ret_col <- function(x) {
  case_when(x >= 0.05 ~ "#16a34a", x >= 0 ~ "#86efac",
            x >= -0.05 ~ "#fca5a5", TRUE ~ "#dc2626")
}

# ── Treemap builder ───────────────────────────────────────────────────────────
build_treemap <- function(df, depth_label) {
  dot <- function(x) case_when(x > .01 ~ "\U1F7E2", x < -.01 ~ "\U1F534", TRUE ~ "\u26AA")

  sectors <- df %>%
    group_by(grp = saa_bucket) %>%
    summarise(n = n(),
              color     = median(ret_ytd * 100, na.rm = TRUE),
              ret1m_med = median(ret_1m  * 100, na.rm = TRUE),
              .groups = "drop") %>%
    mutate(ids = grp, labels = grp, parents = "", values = n,
           tile_text = sprintf("<b>%s</b><br>YTD med: %+.1f%%<br>%s 1M: %+.1f%%",
                               grp, color, dot(ret1m_med / 100), ret1m_med),
           hover = tile_text)

  tickers <- df %>%
    mutate(ids = symbol, labels = symbol, parents = saa_bucket, values = 1,
           color = ret_ytd * 100,
           tile_text = sprintf("<b>%s</b><br>%+.1f%%", symbol, ret_ytd * 100),
           hover = sprintf("<b>%s</b><br>%s<br>YTD: %+.1f%%<br>MTD: %+.1f%%<br>1M: %+.1f%%",
                           symbol, coalesce(name, ""),
                           ret_ytd * 100, ret_mtd * 100, ret_1m * 100)) %>%
    select(ids, labels, parents, values, color, tile_text, hover)

  nodes <- bind_rows(
    sectors %>% select(ids, labels, parents, values, color, tile_text, hover),
    tickers
  )

  plot_ly() %>%
    add_trace(
      type          = "treemap",
      ids           = nodes$ids,
      labels        = nodes$labels,
      parents       = nodes$parents,
      values        = nodes$values,
      text          = nodes$tile_text,
      customdata    = nodes$hover,
      hovertemplate = "%{customdata}<extra></extra>",
      texttemplate  = "%{text}",
      textposition  = "top left",
      textfont      = list(color = "#ffffff", size = 12, family = "Inter"),
      marker        = list(
        colors     = nodes$color,
        colorscale = finviz_scale,
        cmid       = 0,
        showscale  = TRUE,
        colorbar   = list(title = list(text = "YTD %", font = list(size = 11)),
                          tickformat = "+.1f", ticksuffix = "%",
                          len = 0.6, thickness = 12),
        line       = list(width = 1.5, color = "#ffffff")
      ),
      sort   = FALSE,
      tiling = list(packing = "squarify"),
      pathbar = list(visible = TRUE, thickness = 20,
                     textfont = list(size = 11))
    ) %>%
    layout(
      margin = list(t = 10, b = 10, l = 10, r = 10),
      font   = list(family = "Inter")
    ) %>%
    config(displayModeBar = FALSE)
}

# ── Performance table builder ────────────────────────────────────────────────
build_table <- function(df) {
  df %>%
    arrange(saa_depth, saa_bucket, symbol) %>%
    select(symbol, name, saa_bucket, saa_depth,
           ret_ytd, ret_mtd, ret_1m, ret_5d, trend_regime) %>%
    reactable(
      compact      = TRUE,
      striped      = TRUE,
      highlight    = TRUE,
      defaultPageSize = 50,
      columns = list(
        symbol      = colDef(name = "Ticker",  width = 70,
                             style = list(fontWeight = "700")),
        name        = colDef(name = "Name",    minWidth = 160,
                             style = list(fontSize = "12px", color = "#555")),
        saa_bucket  = colDef(name = "Bucket",  width = 65,
                             cell = function(v) {
                               col <- BUCKET_COLS[[v]] %||% "#888"
                               htmltools::tags$span(
                                 style = paste0("background:", col,
                                                "; color:#fff; padding:1px 7px;",
                                                " border-radius:9px; font-size:11px;"),
                                 v)
                             }),
        saa_depth   = colDef(name = "Depth",  width = 55,
                             align = "center",
                             style = list(color = "#888", fontSize = "12px")),
        ret_ytd     = colDef(name = "YTD",    width = 70,  align = "right",
                             format = colFormat(percent = TRUE, digits = 1),
                             style  = function(v) list(
                               color      = ret_col(v),
                               fontWeight = "700", fontSize = "13px")),
        ret_mtd     = colDef(name = "MTD",    width = 70,  align = "right",
                             format = colFormat(percent = TRUE, digits = 1),
                             style  = function(v) list(color = ret_col(v))),
        ret_1m      = colDef(name = "1M",     width = 70,  align = "right",
                             format = colFormat(percent = TRUE, digits = 1),
                             style  = function(v) list(color = ret_col(v))),
        ret_5d      = colDef(name = "5D",     width = 60,  align = "right",
                             format = colFormat(percent = TRUE, digits = 1),
                             style  = function(v) list(color = ret_col(v))),
        trend_regime = colDef(name = "Trend", width = 75,
                              cell = function(v) {
                                if (is.na(v)) return("—")
                                col <- if (v == "Bullish") "#16a34a" else "#dc2626"
                                htmltools::tags$span(
                                  style = paste0("color:", col, "; font-weight:700;"),
                                  v)
                              })
      )
    )
}
```

---

## Depth 0 — Minimal Essential

<div class="depth-box">
**4 anchors — app launch default.**
One representative per bucket: the minimum coherent SAA snapshot.
<br><br>
<span class="bucket-pill" style="background:#3b82f6;color:#fff;">Cash</span> SGOV &nbsp;
<span class="bucket-pill" style="background:#8b5cf6;color:#fff;">FI</span> AGG &nbsp;
<span class="bucket-pill" style="background:#10b981;color:#fff;">EQ</span> URTH &nbsp;
<span class="bucket-pill" style="background:#f59e0b;color:#fff;">Alt</span> GLD
</div>

```{r treemap_d0, fig.height=4}
d0 <- perf_data %>% filter(saa_depth <= 0L)
htmltools::tagList(build_treemap(d0, "Depth 0"))
```

```{r table_d0}
build_table(d0)
```

---

## Depth 1 — First Expansion

<div class="depth-box">
**`r nrow(perf_data %>% filter(saa_depth <= 1L))` tickers.**
Full yield curve (Cash/FI), broad equity (US + DM + EM Europe).
</div>

```{r treemap_d1, fig.height=5}
d1 <- perf_data %>% filter(saa_depth <= 1L)
htmltools::tagList(build_treemap(d1, "Depth 1"))
```

```{r table_d1}
build_table(d1)
```

---

## Depth 2 — + US Sectors

<div class="depth-box">
**`r nrow(perf_data %>% filter(saa_depth <= 2L))` tickers.**
Full GICS-11 sector breakdown added inside EQ bucket.
</div>

```{r treemap_d2, fig.height=5}
d2 <- perf_data %>% filter(saa_depth <= 2L)
htmltools::tagList(build_treemap(d2, "Depth 2"))
```

```{r table_d2}
build_table(d2)
```

---

## Depth 3 — + Country Detail

<div class="depth-box">
**`r nrow(perf_data %>% filter(saa_depth <= 3L))` tickers.**
DM + EM country-level single-country ETFs. Full SAA tree.
</div>

```{r treemap_d3, fig.height=6}
d3 <- perf_data %>% filter(saa_depth <= 3L)
htmltools::tagList(build_treemap(d3, "Depth 3"))
```

```{r table_d3}
build_table(d3)
```

---

*Generated `r format(Sys.Date(), "%B %d, %Y")` · Data: Yahoo Finance via raw_data.rds*
