---
title: "SPY Tactical Replacement Study"
subtitle: "Cycle Amplifiers, Sector Archetypes & Factor ETFs — Full Equity Universe Screen"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 2
    latex_engine: xelatex
---

<style>
.insight-box {
  background-color: #f0f4f8;
  border-left: 4px solid #2c3e50;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
.finding-box {
  background-color: #eafaf1;
  border-left: 4px solid #27ae60;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
.warning-box {
  background-color: #fef9e7;
  border-left: 4px solid #f39c12;
  padding: 10px 16px;
  margin: 16px 0;
  border-radius: 0 4px 4px 0;
}
h2 { border-bottom: 2px solid #2c3e50; padding-bottom: 4px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(scales)
library(patchwork)
library(here)
library(knitr)
library(kableExtra)

# PerformanceAnalytics masks dplyr::select — restore
select <- dplyr::select

# ── Data ──────────────────────────────────────────────────────────────────────
xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources (spy_dd_regime_rel.R auto-loads spy_dd_regime.R + regime_multi_ticker.R)
source(here("scripts/00_init_universe.R"))
source(here("scripts_spy_dd_regime/spy_dd_regime_rel.R"))

# ── Constants ─────────────────────────────────────────────────────────────────
MASTER   <- "SPY"
T_FALL   <- 0.10
T_CRUISE <- 0.05
RHO_MIN  <- 0.75   # Corr-Fail threshold (mirrors KF2 L/S pair screen)

# ── Regime table ──────────────────────────────────────────────────────────────
rt <- build_regime_table(xts_ret[, MASTER], T_FALL, T_CRUISE)

# ── Equity universe: exclude SPY, EUR-hedged (dot suffix), IBIT (too short) ──
equity_tickers <- etf_metadata %>%
  filter(asset_class == "Equity",
         !grepl("\\.", ticker),
         ticker != MASTER,
         ticker != "IBIT") %>%
  pull(ticker) %>%
  intersect(colnames(xts_ret))
```

```{r screen-engine, cache=TRUE}
# ── Screening function: regime archetype per ticker ───────────────────────────
.screen_ticker <- function(tk, xts_ret, rt, master = "SPY") {
  tk_r  <- xts_ret[, tk]
  spy_r <- xts_ret[, master]

  # Align to common non-NA period
  both  <- which(!is.na(as.numeric(tk_r)) & !is.na(as.numeric(spy_r)))
  if (length(both) < 252) return(NULL)
  r_tk  <- as.numeric(tk_r)[both]
  r_spy <- as.numeric(spy_r)[both]

  spread <- r_tk - r_spy
  rho    <- cor(r_tk, r_spy)
  ir     <- if (sd(spread) > 0) (mean(spread) * 252) / (sd(spread) * sqrt(252)) else NA_real_
  max_dd <- -as.numeric(maxDrawdown(tk_r))   # stored negative for display

  # Per-episode outperf type + alpha
  ep_df <- rt %>%
    rowwise() %>%
    mutate(
      raw_ret = .period_ret(tk_r, xmin, xmax),
      spy_ret = period_return,
      alpha   = raw_ret - spy_ret,
      type    = .outperf_type(raw_ret, spy_ret, alpha)
    ) %>%
    ungroup()

  # Modal type + avg alpha per regime
  smry <- ep_df %>%
    group_by(regime) %>%
    summarise(
      modal_type = names(sort(table(type), decreasing = TRUE))[1],
      avg_alpha  = mean(alpha, na.rm = TRUE),
      .groups    = "drop"
    )

  get_val <- function(reg, col) {
    r <- smry %>% filter(regime == reg)
    if (nrow(r) == 0) return(NA)
    r[[col]]
  }

  tibble(
    ticker     = tk,
    rho        = rho,
    ir         = ir,
    max_dd     = max_dd,
    fall_type  = get_val("Fall",          "modal_type"),
    fall_alpha = get_val("Fall",          "avg_alpha"),
    rec_type   = get_val("Recovery",      "modal_type"),
    rec_alpha  = get_val("Recovery",      "avg_alpha"),
    con_type   = get_val("Consolidation", "modal_type"),
    con_alpha  = get_val("Consolidation", "avg_alpha")
  )
}

screen_raw <- map_dfr(equity_tickers, .screen_ticker,
                      xts_ret = xts_ret, rt = rt)

# ── Archetype classification ──────────────────────────────────────────────────
screen_tbl <- screen_raw %>%
  left_join(
    etf_metadata %>% select(ticker, name, asset_class, sub_block, pf_function),
    by = "ticker"
  ) %>%
  mutate(
    corr_fail = rho < RHO_MIN,
    archetype = case_when(
      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "ALPHA"                   &
        con_type  == "ALPHA"                   ~ "Cycle Amplifier",

      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "ALPHA"                   &
        con_type  == "LAG"                     ~ "Recovery Sprinter",

      !corr_fail &
        fall_type %in% c("STABLE", "HEDGE")    &
        rec_type  == "LAG"                     &
        con_type  %in% c("ALPHA", "LAG")       ~ "Defensive Burden",

      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "LAG"                     ~ "All-Regime Drag",

      corr_fail                                ~ "Corr-Fail",
      TRUE                                     ~ "Other"
    )
  ) %>%
  arrange(desc(ir))
```

---

## 1. Overview

<div class="insight-box">
**Question:** Across the full equity universe, which ETFs can serve as a tactical
SPY replacement or enhancement — and how do they behave across the three market regimes?

**Answer in one finding:** The cycle amplifier pattern (LOSS in Fall → ALPHA in Recovery →
ALPHA in Consolidation) is exclusive to **three tickers** in the 96-ticker Sovereign Universe:
SMH, XLK, and QQQ. No sector, factor, or regional equity ETF reproduces it.
</div>

This study screens every equity ticker in the Sovereign Universe against SPY across the
three-regime framework (Fall / Recovery / Consolidation). For each ticker we compute:

- **IR** — Information Ratio of the daily excess return spread vs SPY
- **ρ** — Full-history Pearson correlation vs SPY
- **MaxDD** — Standalone maximum drawdown
- **Regime archetype** — The modal outperformance type in each regime phase

The screen covers **`r length(equity_tickers)` equity tickers** from the Sovereign Universe,
spanning `r format(start(xts_ret), "%b %Y")` to `r format(end(xts_ret), "%b %Y")`.

---

## 2. Regime Framework

The SPY regime engine partitions history into three phases:

| Regime | Definition |
|--------|------------|
| **Fall** | SPY drawdown from recent peak ≥ `r scales::percent(T_FALL)` |
| **Recovery** | After a Fall — drawdown between 0 and `r scales::percent(T_FALL)`, returning to peak |
| **Consolidation** | All other periods — new-high territory or minor pullbacks < `r scales::percent(T_FALL)` |

For each ticker × regime episode, the outperformance type is classified as:

| Type | Condition | Meaning |
|------|-----------|---------|
| **HEDGE** | α > 0, abs > 0, SPY < 0 | Ticker rose while SPY fell |
| **STABLE** | α > 0, abs < 0, SPY < 0 | Both fell — ticker fell less |
| **ALPHA** | α > 0, SPY > 0 | Both rose — ticker rose more |
| **LAG** | α < 0, SPY > 0 | Underperformed rising SPY |
| **LOSS** | α < 0, SPY < 0 | Both fell — ticker fell worse |

The **modal type** across all episodes of a given regime is used to characterise each
ticker. The **average alpha** quantifies the magnitude.

```{r regime-episodes, fig.height=4}
rt %>%
  count(regime) %>%
  mutate(regime = factor(regime, levels = names(REGIME_PAL))) %>%
  ggplot(aes(x = regime, y = n, fill = regime)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.4, fontface = "bold", size = 4) +
  scale_fill_manual(values = REGIME_PAL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(face = "bold")) +
  labs(title = "Regime Episode Count",
       subtitle = sprintf("Fall ≥%.0f%%  |  %s–%s",
                          T_FALL * 100,
                          format(min(rt$xmin), "%b %Y"),
                          format(max(rt$xmax), "%b %Y")))
```

---

## 3. Full Equity Screen

```{r full-screen-table}
screen_tbl %>%
  mutate(
    ir      = round(ir, 2),
    rho     = round(rho, 3),
    max_dd  = scales::percent(max_dd, accuracy = 0.1),
    fall    = sprintf("%s  %s", fall_type,
                      scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec     = sprintf("%s  %s", rec_type,
                      scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con     = sprintf("%s  %s", con_type,
                      scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus")),
    corr_flag = if_else(corr_fail, "✗", "")
  ) %>%
  select(Ticker = ticker, Name = name, Archetype = archetype,
         IR = ir, ρ = rho, `Corr?` = corr_flag, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(
    caption = sprintf(
      "Full equity screen — %d tickers vs SPY  |  Fall ≥%.0f%%  |  ✗ = Corr-Fail (ρ < %.2f)",
      nrow(screen_tbl), T_FALL * 100, RHO_MIN
    ),
    align = c("l","l","l","r","r","c","r","l","l","l")
  ) %>%
  kable_styling(font_size = 9, full_width = TRUE) %>%
  row_spec(which(screen_tbl$archetype == "Cycle Amplifier"),
           bold = TRUE, background = "#eafaf1") %>%
  row_spec(which(screen_tbl$archetype == "Recovery Sprinter"),
           background = "#fef9e7") %>%
  row_spec(which(screen_tbl$corr_fail),
           color = "grey60")
```

<div class="finding-box">
**Screen summary:** `r sum(screen_tbl$archetype == "Cycle Amplifier")` Cycle Amplifiers  |
`r sum(screen_tbl$archetype == "Recovery Sprinter")` Recovery Sprinters  |
`r sum(screen_tbl$archetype == "Defensive Burden")` Defensive Burden  |
`r sum(screen_tbl$archetype == "All-Regime Drag")` All-Regime Drag  |
`r sum(screen_tbl$corr_fail)` Corr-Fail (ρ < `r RHO_MIN`)
</div>

---

## 4. Cycle Amplifiers — The Evergreen Pattern

<div class="finding-box">
**Pattern definition:** LOSS in Fall → ALPHA in Recovery → ALPHA in Consolidation.

These tickers **amplify** SPY's cycle in both directions. They fall harder than SPY in
downturns and recover faster and grind higher in every other phase.
**Never use them as downside protection.** They are cycle amplifiers, not hedges.
</div>

### 4.1 The Three-Gate Requirement

The Evergreen pattern requires **three conditions simultaneously**:

1. **β > 1 vs SPY** — amplifies moves in both directions (necessary for LOSS in Fall)
2. **ρ > 0.82 vs SPY** — tracks SPY regime phases closely enough to produce consistent ALPHA
3. **Structural positive alpha** — net positive drift above SPY over the full cycle

This combination exists only in **US tech / secular growth** within the 96-ticker universe.
High-β tickers without structural alpha (EM, Small Caps, Biotech) produce LOSS in Fall
but LAG in Recovery. Defensive tickers with positive structural alpha produce STABLE in
Fall but LAG in Recovery. Only US tech satisfies all three simultaneously.

### 4.2 Confirmed Evergreens

```{r evergreen-table}
screen_tbl %>%
  filter(archetype == "Cycle Amplifier") %>%
  mutate(
    ir     = round(ir, 2),
    rho    = round(rho, 3),
    max_dd = scales::percent(max_dd, accuracy = 0.1),
    fall   = sprintf("%s  (%s)", fall_type,
                     scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec    = sprintf("%s  (%s)", rec_type,
                     scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con    = sprintf("%s  (%s)", con_type,
                     scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus"))
  ) %>%
  select(Ticker = ticker, Name = name, IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "Confirmed Cycle Amplifiers") %>%
  kable_styling(full_width = FALSE) %>%
  row_spec(0, bold = TRUE)
```

```{r evergreen-heatmap, fig.height=6}
evergreen_tickers <- screen_tbl %>%
  filter(archetype == "Cycle Amplifier") %>%
  pull(ticker)

if (length(evergreen_tickers) > 0)
  print(plot_regime_rel_heatmap(xts_ret, evergreen_tickers, master = MASTER,
                                t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r evergreen-summary, fig.height=5}
if (length(evergreen_tickers) > 0)
  print(plot_regime_rel_summary(xts_ret, evergreen_tickers, master = MASTER,
                                t_fall = T_FALL, t_cruise = T_CRUISE))
```

### 4.3 Near-Miss: Conditional Candidates

```{r conditional-table}
# Tickers that are LOSS/ALPHA but con_type is LAG by a small margin
cond_tickers <- screen_tbl %>%
  filter(!corr_fail,
         fall_type == "LOSS",
         rec_type  == "ALPHA",
         con_type  == "LAG",
         con_alpha > -0.015)   # within 1.5pp of ALPHA

if (nrow(cond_tickers) > 0) {
  cond_tickers %>%
    mutate(
      ir     = round(ir, 2),
      rho    = round(rho, 3),
      max_dd = scales::percent(max_dd, accuracy = 0.1),
      fall   = sprintf("%s  (%s)", fall_type,
                       scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
      rec    = sprintf("%s  (%s)", rec_type,
                       scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
      con    = sprintf("%s  (%s)", con_type,
                       scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus"))
    ) %>%
    select(Ticker = ticker, Name = name, IR = ir, ρ = rho, MaxDD = max_dd,
           Fall = fall, Recovery = rec, Consolidation = con) %>%
    kable(caption = "Conditional candidates — LOSS/ALPHA with Consolidation within 1.5pp of ALPHA") %>%
    kable_styling(full_width = FALSE)
} else {
  cat("No conditional candidates within threshold.")
}
```

<div class="warning-box">
**AIQ (Global X AI & Big Data):** shows LOSS/ALPHA/ALPHA but its Consolidation alpha averages
−1.0% — below zero but within noise. It qualifies as a cycle amplifier *in spirit* but lacks
the structural Consolidation edge of SMH/XLK/QQQ. Appropriate only as a small satellite, not
a structural replacement.
</div>

---

## 5. US Sector Archetypes — 11 GICS Sectors

```{r sector-setup}
sector_tickers <- c("XLK","XLC","XLV","XLF","XLI","XLP","XLY","XLE","XLB","XLRE","XLU")
sector_tickers <- intersect(sector_tickers, colnames(xts_ret))

sector_screen <- screen_tbl %>%
  filter(ticker %in% sector_tickers) %>%
  arrange(desc(ir))
```

The 11 GICS sector ETFs divide into **five distinct behavioral archetypes** based on
their modal pattern across Fall / Recovery / Consolidation:

```{r sector-archetype-table}
sector_screen %>%
  mutate(
    ir       = round(ir, 2),
    rho      = round(rho, 3),
    max_dd   = scales::percent(max_dd, accuracy = 0.1),
    fall     = sprintf("%s (%s)", fall_type,
                       scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec      = sprintf("%s (%s)", rec_type,
                       scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con      = sprintf("%s (%s)", con_type,
                       scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus")),
    Archetype = case_when(
      ticker == "XLK"                             ~ "Cycle Amplifier",
      ticker %in% c("XLC","XLY")                  ~ "Recovery Sprinter",
      ticker %in% c("XLP","XLU","XLV","XLRE")     ~ "Defensive Burden",
      ticker %in% c("XLF","XLI")                  ~ "Late Cyclical",
      ticker == "XLB"                             ~ "All-Regime Drag",
      ticker == "XLE"                             ~ "Macro Satellite",
      TRUE                                        ~ archetype
    )
  ) %>%
  select(Ticker = ticker, Name = name, Archetype,
         IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "US Sector ETF regime archetypes — sorted by IR vs SPY") %>%
  kable_styling(font_size = 10, full_width = TRUE) %>%
  row_spec(which(sector_screen$ticker == "XLK"),
           bold = TRUE, background = "#eafaf1") %>%
  row_spec(which(sector_screen$ticker %in% c("XLC","XLY")),
           background = "#fef9e7") %>%
  row_spec(which(sector_screen$ticker == "XLB"),
           color = "#7B1010")
```

### Archetype Descriptions

**Cycle Amplifier — XLK only**
The only sector that delivers structural alpha above SPY in *both* Recovery and
Consolidation. Semiconductor/cloud/software earnings compound faster than the market
cycle. β > 1 makes it a LOSS in Fall, but the structural drift more than compensates.

**Recovery Sprinters — XLC, XLY**
These amplify Recovery (ALPHA) but fade in Consolidation (LAG). Consumer Discretionary
(Amazon, Tesla) and Communication Services (Meta, Alphabet) sprint hard off the
bottom driven by consumer spending re-acceleration. Once momentum normalises, the media
and telecom components drag. *Not structural alpha — it is mean-reversion beta.*

**Defensive Burden — XLP, XLU, XLV, XLRE**
Protect in Fall (STABLE), pay for it heavily in Recovery (LAG). The Fall cushion is
fully — and in the case of XLP and XLU, more than fully — clawed back in Recovery.

The **XLP/XLU protection premium** is a tactical illusion:

| | Fall α | Recovery α | Net (Fall + Recovery) |
|-|--------|------------|-----------------------|
| XLP Staples   | +9.6% | −15.6% | **−6.0%** |
| XLU Utilities | +8.6% | −17.7% | **−9.1%** |

XLV (Healthcare) has the same structure with a 6.7% Fall cushion and −14.8%
Recovery lag. Despite being marketed as "defensive", it is a negative-IR sector
across the full cycle (IR = −0.33).

**Late Cyclicals — XLF, XLI**
Mild or neutral Fall (STABLE with near-zero alpha), lag in Recovery, recover
modestly in Consolidation. XLF is the most interesting: −8.9% in Recovery
(credit cycle tightening) but +3.4% in Consolidation (yield curve steepening).
XLI is essentially SPY with a small permanent haircut — near-zero alpha in every phase.

**All-Regime Drag — XLB**
Materials fall harder than SPY in Fall, lag in Recovery, and lag in Consolidation.
No regime is favorable. Commodity input cost cycles dominate over earnings.

**Macro Satellite — XLE**
ρ = 0.61. Energy does not run on SPY drawdown cycles — it runs on oil price cycles.
The +5.9% Fall alpha reflects supply-shock episodes (2022) and is not repeatable as a
hedge. SPY regime classification is structurally unreliable for XLE.

```{r sector-heatmap, fig.height=9}
print(plot_regime_rel_heatmap(xts_ret, sector_tickers, master = MASTER,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r sector-summary, fig.height=6}
print(plot_regime_rel_summary(xts_ret, sector_tickers, master = MASTER,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

---

## 6. Factor & Smart-Beta ETFs

```{r factor-setup}
factor_tickers <- c("QUAL","MTUM","USMV","VTV","DGRW","COWZ","ACWV")
factor_tickers <- intersect(factor_tickers, colnames(xts_ret))

factor_screen <- screen_tbl %>%
  filter(ticker %in% factor_tickers) %>%
  arrange(desc(ir))
```

```{r factor-table}
factor_screen %>%
  mutate(
    ir     = round(ir, 2),
    rho    = round(rho, 3),
    max_dd = scales::percent(max_dd, accuracy = 0.1),
    fall   = sprintf("%s (%s)", fall_type,
                     scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec    = sprintf("%s (%s)", rec_type,
                     scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con    = sprintf("%s (%s)", con_type,
                     scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus")),
    Factor = case_when(
      ticker == "QUAL"  ~ "Quality",
      ticker == "MTUM"  ~ "Momentum",
      ticker == "USMV"  ~ "Min Volatility",
      ticker == "VTV"   ~ "Value",
      ticker == "DGRW"  ~ "Dividend Growth",
      ticker == "COWZ"  ~ "Cash Flow",
      ticker == "ACWV"  ~ "Min Vol (Global)",
      TRUE              ~ sub_block
    )
  ) %>%
  select(Ticker = ticker, Factor, IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "Factor & smart-beta ETFs — regime archetypes vs SPY") %>%
  kable_styling(font_size = 10, full_width = TRUE)
```

**No factor ETF clears the Evergreen screen.** All factor archetypes map onto patterns
already identified in the sector analysis:

- **QUAL** (ρ = 0.984): The highest correlation in the universe outside SPY itself. It is
  SPY minus alpha — a structural underperformer in every regime (IR = −0.20). The quality
  premium did not manifest vs SPY over this history.

- **MTUM** (Momentum): The most structurally interesting factor. LOSS in Fall (momentum
  crashes in sharp reversals), LAG in Recovery (beaten-down reversals steal from momentum
  names), then ALPHA in Consolidation (trending markets reward the factor). **Anti-cycle
  timing** — worst precisely when Recovery begins, best when volatility is absent.

- **VTV / COWZ** (Value / Cash Flow): Map onto the Late Cyclical archetype. STABLE or
  neutral in Fall, LAG in Recovery, modest Consolidation alpha. Value does not sprint off
  the bottom.

- **DGRW / USMV** (Dividend Growth / Min Vol): Map onto Defensive Burden. USMV in
  particular is the factor version of XLP — strong Fall protection (+5.5%), severe Recovery
  lag (−9.9%).

**Equal Weight S&P 500 (RSP) — not needed.** Equal-weight is structurally long
value/cyclicals and short mega-cap tech. It would show STABLE/LAG/LAG — already captured
by VTV + the sector spread in the universe. It introduces no new archetype.

```{r factor-heatmap, fig.height=8}
print(plot_regime_rel_heatmap(xts_ret, factor_tickers, master = MASTER,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r factor-summary, fig.height=6}
print(plot_regime_rel_summary(xts_ret, factor_tickers, master = MASTER,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

---

## 7. Portfolio Implications

<div class="finding-box">
**The master insight:** No equity ETF in the 96-ticker Sovereign Universe delivers
STABLE-in-Fall AND ALPHA-in-Recovery. The tradeoff is structural:

- Sectors that protect in Fall are defensives (low β) → they guarantee Recovery LAG.
- Sectors that amplify Recovery are growth (high β) → they guarantee Fall LOSS.

**XLK escapes the tradeoff** only because it has structural alpha above SPY *across
the full cycle* — not because it broke the β relationship.
</div>

### Key Rules from This Study

**1. Cycle Amplifiers are growth allocations only**
SMH, XLK, QQQ belong in the Growth sleeve of the SAA. Their LOSS in Fall is the *entry
price* for the Recovery/Consolidation ALPHA. Removing them tactically in Fall means missing
the sharpest part of the Recovery.

**2. The defensive trap (XLP / XLU)**
The Fall cushion of Staples (+9.6%) and Utilities (+8.6%) appears attractive.
The net of Fall + Recovery is −6.0% and −9.1% respectively. A full-cycle tactical rotation
into these sectors destroys more value in Recovery than it saves in Fall.

**3. XLV is not a hedge**
Healthcare is commonly labelled defensive. It protects in Fall (+6.7%) but gives back
−14.8% in Recovery. Over a full cycle its IR is −0.33. Use it as a satellite, never as
a defensive overlay.

**4. MTUM needs cycle awareness**
Momentum (IR = −0.09, better than most factors) delivers its ALPHA in Consolidation —
the longest and most persistent regime phase. But it crashes in Fall and lags in Recovery.
TAA rotation into MTUM should trigger *after* Recovery is confirmed, not during it.

**5. XLF is the best sector for late-cycle positioning**
The only sector with meaningful positive Consolidation alpha (+3.4%) while absorbing
modest Fall losses. Yield curve steepening in Consolidation benefits bank net interest
margins. Consider as a Consolidation-phase overweight in tactical allocation.

**6. SAA implications — the existing portfolio is correct**
The current SAA includes QQQ (5%) and XLK (5%) as Core-Growth satellites — both confirmed
Evergreens. The screen finds no additional sector or factor ETF that warrants inclusion as
a structural SPY replacement or enhancement. The universe is exhaustive.

```{r archetype-roster-table}
archetype_order <- c("Cycle Amplifier","Recovery Sprinter","Defensive Burden",
                     "Late Cyclical","All-Regime Drag","Corr-Fail","Other")

screen_tbl %>%
  mutate(archetype = factor(archetype, levels = archetype_order)) %>%
  arrange(archetype, desc(ir)) %>%
  group_by(archetype) %>%
  summarise(
    Tickers = paste(ticker, collapse = ", "),
    N       = n(),
    .groups = "drop"
  ) %>%
  rename(Archetype = archetype, `# Tickers` = N) %>%
  kable(caption = "Tickers by regime archetype — sorted by IR within each group") %>%
  kable_styling(font_size = 10, full_width = TRUE) %>%
  column_spec(2, width = "60%")
```

```{r archetype-summary-chart, fig.height=5}
archetype_order <- c("Cycle Amplifier","Recovery Sprinter","Defensive Burden",
                     "Late Cyclical","All-Regime Drag","Corr-Fail","Other")
archetype_pal <- c(
  "Cycle Amplifier"  = "#27ae60",
  "Recovery Sprinter"= "#f39c12",
  "Defensive Burden" = "#2980b9",
  "Late Cyclical"    = "#8e44ad",
  "All-Regime Drag"  = "#c0392b",
  "Corr-Fail"        = "#95a5a6",
  "Other"            = "#bdc3c7"
)

screen_tbl %>%
  mutate(archetype = factor(archetype, levels = archetype_order)) %>%
  count(archetype) %>%
  ggplot(aes(x = archetype, y = n, fill = archetype)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.4, fontface = "bold", size = 4.5) +
  scale_fill_manual(values = archetype_pal) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 12) +
  theme(
    axis.title         = element_blank(),
    axis.text.x        = element_text(face = "bold", size = 10),
    panel.grid.major.x = element_blank(),
    plot.title         = element_text(face = "bold", size = 13)
  ) +
  labs(
    title    = sprintf("Equity Universe Regime Archetypes — %d tickers vs SPY",
                       nrow(screen_tbl)),
    subtitle = sprintf("Fall ≥%.0f%%  |  Corr-Fail = ρ < %.2f  |  %s–%s",
                       T_FALL * 100, RHO_MIN,
                       format(start(xts_ret), "%b %Y"),
                       format(end(xts_ret), "%b %Y"))
  )
```

---

*Study date: `r Sys.Date()`  |  Universe: Sovereign Universe `r nrow(etf_metadata)` tickers
|  Screen: `r length(equity_tickers)` equity tickers vs SPY  |  Fall threshold: `r scales::percent(T_FALL)`*

---

# Part II — URTH as Benchmark

```{r urth-setup}
MASTER2  <- "URTH"
rt_urth  <- build_regime_table(xts_ret[, MASTER2], T_FALL, T_CRUISE)

equity_tickers2 <- etf_metadata %>%
  filter(asset_class == "Equity",
         !grepl("\\.", ticker),
         ticker != MASTER2,
         ticker != "IBIT") %>%
  pull(ticker) %>%
  intersect(colnames(xts_ret))
```

```{r urth-screen-engine, cache=TRUE}
screen_raw2 <- map_dfr(equity_tickers2, .screen_ticker,
                       xts_ret = xts_ret, rt = rt_urth, master = MASTER2)

screen_tbl2 <- screen_raw2 %>%
  left_join(
    etf_metadata %>% select(ticker, name, asset_class, sub_block, pf_function),
    by = "ticker"
  ) %>%
  mutate(
    corr_fail = rho < RHO_MIN,
    archetype = case_when(
      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "ALPHA"                   &
        con_type  == "ALPHA"                   ~ "Cycle Amplifier",
      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "ALPHA"                   &
        con_type  == "LAG"                     ~ "Recovery Sprinter",
      !corr_fail &
        fall_type %in% c("STABLE", "HEDGE")    &
        rec_type  == "LAG"                     &
        con_type  %in% c("ALPHA", "LAG")       ~ "Defensive Burden",
      !corr_fail &
        fall_type == "LOSS"                    &
        rec_type  == "LAG"                     ~ "All-Regime Drag",
      corr_fail                                ~ "Corr-Fail",
      TRUE                                     ~ "Other"
    )
  ) %>%
  arrange(desc(ir))
```

---

## II.1 Overview

<div class="insight-box">
**Benchmark:** URTH (iShares MSCI World) replaces SPY as the reference series.
All alpha, correlation, and regime classifications are computed **relative to URTH**.
The regime table is rebuilt from URTH's own drawdown history using the same
Fall ≥ `r scales::percent(T_FALL)` threshold.
</div>

URTH is a global developed-market benchmark (ρ ≈ 0.98 vs SPY over the full history).
Its regime episodes are structurally similar to SPY's but not identical — global market
stress sometimes diverges from US-only drawdowns.

```{r urth-regime-episodes, fig.height=4}
rt_urth %>%
  count(regime) %>%
  mutate(regime = factor(regime, levels = names(REGIME_PAL))) %>%
  ggplot(aes(x = regime, y = n, fill = regime)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.4, fontface = "bold", size = 4) +
  scale_fill_manual(values = REGIME_PAL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(face = "bold")) +
  labs(title = sprintf("URTH Regime Episode Count"),
       subtitle = sprintf("Fall ≥%.0f%%  |  %s–%s",
                          T_FALL * 100,
                          format(min(rt_urth$xmin), "%b %Y"),
                          format(max(rt_urth$xmax), "%b %Y")))
```

---

## II.2 Full Equity Screen vs URTH

```{r urth-screen-table}
screen_tbl2 %>%
  mutate(
    ir      = round(ir, 2),
    rho     = round(rho, 3),
    max_dd  = scales::percent(max_dd, accuracy = 0.1),
    fall    = sprintf("%s  %s", fall_type,
                      scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec     = sprintf("%s  %s", rec_type,
                      scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con     = sprintf("%s  %s", con_type,
                      scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus")),
    corr_flag = if_else(corr_fail, "✗", "")
  ) %>%
  select(Ticker = ticker, Name = name, Archetype = archetype,
         IR = ir, ρ = rho, `Corr?` = corr_flag, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(
    caption = sprintf(
      "Full equity screen — %d tickers vs URTH  |  Fall ≥%.0f%%  |  ✗ = Corr-Fail (ρ < %.2f)",
      nrow(screen_tbl2), T_FALL * 100, RHO_MIN
    ),
    align = c("l","l","l","r","r","c","r","l","l","l")
  ) %>%
  kable_styling(font_size = 9, full_width = TRUE) %>%
  row_spec(which(screen_tbl2$archetype == "Cycle Amplifier"),
           bold = TRUE, background = "#eafaf1") %>%
  row_spec(which(screen_tbl2$archetype == "Recovery Sprinter"),
           background = "#fef9e7") %>%
  row_spec(which(screen_tbl2$corr_fail),
           color = "grey60")
```

<div class="finding-box">
**Screen summary vs URTH:** `r sum(screen_tbl2$archetype == "Cycle Amplifier")` Cycle Amplifiers  |
`r sum(screen_tbl2$archetype == "Recovery Sprinter")` Recovery Sprinters  |
`r sum(screen_tbl2$archetype == "Defensive Burden")` Defensive Burden  |
`r sum(screen_tbl2$archetype == "All-Regime Drag")` All-Regime Drag  |
`r sum(screen_tbl2$corr_fail)` Corr-Fail (ρ < `r RHO_MIN`)
</div>

---

## II.3 Cycle Amplifiers vs URTH

```{r urth-evergreen-table}
evergreen_tickers2 <- screen_tbl2 %>%
  filter(archetype == "Cycle Amplifier") %>%
  pull(ticker)

screen_tbl2 %>%
  filter(archetype == "Cycle Amplifier") %>%
  mutate(
    ir     = round(ir, 2),
    rho    = round(rho, 3),
    max_dd = scales::percent(max_dd, accuracy = 0.1),
    fall   = sprintf("%s  (%s)", fall_type,
                     scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec    = sprintf("%s  (%s)", rec_type,
                     scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con    = sprintf("%s  (%s)", con_type,
                     scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus"))
  ) %>%
  select(Ticker = ticker, Name = name, IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "Confirmed Cycle Amplifiers vs URTH") %>%
  kable_styling(full_width = FALSE) %>%
  row_spec(0, bold = TRUE)
```

```{r urth-evergreen-heatmap, fig.height=6}
if (length(evergreen_tickers2) > 0)
  print(plot_regime_rel_heatmap(xts_ret, evergreen_tickers2, master = MASTER2,
                                t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r urth-evergreen-summary, fig.height=5}
if (length(evergreen_tickers2) > 0)
  print(plot_regime_rel_summary(xts_ret, evergreen_tickers2, master = MASTER2,
                                t_fall = T_FALL, t_cruise = T_CRUISE))
```

---

## II.4 Sector Archetypes vs URTH

```{r urth-sector-heatmap, fig.height=9}
sector_tickers2 <- intersect(sector_tickers, colnames(xts_ret))
print(plot_regime_rel_heatmap(xts_ret, sector_tickers2, master = MASTER2,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r urth-sector-summary, fig.height=6}
print(plot_regime_rel_summary(xts_ret, sector_tickers2, master = MASTER2,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r urth-sector-table}
screen_tbl2 %>%
  filter(ticker %in% sector_tickers) %>%
  arrange(desc(ir)) %>%
  mutate(
    ir     = round(ir, 2),
    rho    = round(rho, 3),
    max_dd = scales::percent(max_dd, accuracy = 0.1),
    fall   = sprintf("%s (%s)", fall_type,
                     scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec    = sprintf("%s (%s)", rec_type,
                     scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con    = sprintf("%s (%s)", con_type,
                     scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus"))
  ) %>%
  select(Ticker = ticker, Name = name, Archetype = archetype,
         IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "US Sector ETF regime archetypes vs URTH — sorted by IR") %>%
  kable_styling(font_size = 10, full_width = TRUE)
```

---

## II.5 Factor & Smart-Beta ETFs vs URTH

```{r urth-factor-heatmap, fig.height=8}
factor_tickers2 <- intersect(factor_tickers, colnames(xts_ret))
print(plot_regime_rel_heatmap(xts_ret, factor_tickers2, master = MASTER2,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r urth-factor-summary, fig.height=6}
print(plot_regime_rel_summary(xts_ret, factor_tickers2, master = MASTER2,
                              t_fall = T_FALL, t_cruise = T_CRUISE))
```

```{r urth-factor-table}
screen_tbl2 %>%
  filter(ticker %in% factor_tickers) %>%
  arrange(desc(ir)) %>%
  mutate(
    ir     = round(ir, 2),
    rho    = round(rho, 3),
    max_dd = scales::percent(max_dd, accuracy = 0.1),
    fall   = sprintf("%s (%s)", fall_type,
                     scales::percent(fall_alpha, accuracy = 0.1, style_positive = "plus")),
    rec    = sprintf("%s (%s)", rec_type,
                     scales::percent(rec_alpha,  accuracy = 0.1, style_positive = "plus")),
    con    = sprintf("%s (%s)", con_type,
                     scales::percent(con_alpha,  accuracy = 0.1, style_positive = "plus"))
  ) %>%
  select(Ticker = ticker, Name = name, Archetype = archetype,
         IR = ir, ρ = rho, MaxDD = max_dd,
         Fall = fall, Recovery = rec, Consolidation = con) %>%
  kable(caption = "Factor & smart-beta ETFs vs URTH — sorted by IR") %>%
  kable_styling(font_size = 10, full_width = TRUE)
```

---

## II.6 Summary — URTH Archetypes

```{r urth-archetype-roster-table}
archetype_order2 <- c("Cycle Amplifier","Recovery Sprinter","Defensive Burden",
                      "Late Cyclical","All-Regime Drag","Corr-Fail","Other")

screen_tbl2 %>%
  mutate(archetype = factor(archetype, levels = archetype_order2)) %>%
  arrange(archetype, desc(ir)) %>%
  group_by(archetype) %>%
  summarise(
    Tickers = paste(ticker, collapse = ", "),
    N       = n(),
    .groups = "drop"
  ) %>%
  rename(Archetype = archetype, `# Tickers` = N) %>%
  kable(caption = "Tickers by regime archetype vs URTH — sorted by IR within each group") %>%
  kable_styling(font_size = 10, full_width = TRUE) %>%
  column_spec(2, width = "60%")
```

```{r urth-archetype-summary-chart, fig.height=5}
archetype_pal2 <- c(
  "Cycle Amplifier"  = "#27ae60",
  "Recovery Sprinter"= "#f39c12",
  "Defensive Burden" = "#2980b9",
  "Late Cyclical"    = "#8e44ad",
  "All-Regime Drag"  = "#c0392b",
  "Corr-Fail"        = "#95a5a6",
  "Other"            = "#bdc3c7"
)

screen_tbl2 %>%
  mutate(archetype = factor(archetype, levels = archetype_order2)) %>%
  count(archetype) %>%
  ggplot(aes(x = archetype, y = n, fill = archetype)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.4, fontface = "bold", size = 4.5) +
  scale_fill_manual(values = archetype_pal2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 12) +
  theme(
    axis.title         = element_blank(),
    axis.text.x        = element_text(face = "bold", size = 10),
    panel.grid.major.x = element_blank(),
    plot.title         = element_text(face = "bold", size = 13)
  ) +
  labs(
    title    = sprintf("Equity Universe Regime Archetypes — %d tickers vs URTH",
                       nrow(screen_tbl2)),
    subtitle = sprintf("Fall ≥%.0f%%  |  Corr-Fail = ρ < %.2f  |  %s–%s",
                       T_FALL * 100, RHO_MIN,
                       format(start(xts_ret), "%b %Y"),
                       format(end(xts_ret), "%b %Y"))
  )
```

---

*URTH screen date: `r Sys.Date()`  |  Universe: Sovereign Universe `r nrow(etf_metadata)` tickers
|  Screen: `r length(equity_tickers2)` equity tickers vs URTH  |  Fall threshold: `r scales::percent(T_FALL)`*

---

## Appendix — Analytical Notes from Discovery Session

*The sections below preserve the reasoning and intermediate findings produced during the
original screening session. They document the analytical path: what was found, why the
boundaries were drawn where they were, and which near-misses were explicitly ruled out.*

---

### A1. Confirmed Evergreens — Raw Screen Output

**Three tickers pass all three gates across the full 61-ticker equity screen:**

| Ticker | Name | IR | ρ | Fall | Recovery | Consolidation |
|--------|------|----|---|------|----------|---------------|
| **SMH** | Semiconductors | 0.62 | 0.82 | LOSS −7.8% | ALPHA +25.7% | ALPHA +13.4% |
| **XLK** | Technology | 0.57 | 0.93 | LOSS −3.9% | ALPHA +16.8% | ALPHA +1.1% |
| **QQQ** | Nasdaq 100 | 0.47 | 0.94 | LOSS −2.5% | ALPHA +12.2% | ALPHA +0.4% |

**Conditional (borderline) — 1 ticker:**

| Ticker | Issue | Fall | Recovery | Consolidation |
|--------|-------|------|----------|---------------|
| **AIQ** | Consolidation alpha is −1.0% — pattern in spirit but shallow | LOSS −2.9% | ALPHA +13.8% | ALPHA −1.0% |

AIQ qualifies only if near-zero Consolidation alpha is accepted as sufficient. It is
not a structural Evergreen: the Consolidation edge is within noise.

---

### A2. Why Only 3 Pass — The Anatomy of the Pattern

The LOSS / ALPHA / ALPHA pattern requires three simultaneous conditions:

1. **β > 1 vs SPY** — amplifies moves in both directions (necessary for LOSS in Fall)
2. **ρ > 0.82 vs SPY** — tracks SPY regime phases closely enough to produce consistent ALPHA
3. **Structural positive alpha** — net positive drift above SPY over the full cycle

Each category of failures:

| Failure mode | Examples | What breaks |
|---|---|---|
| High β, low ρ | EM (EWY, FXI), Biotech (XBI, IBB), Cloud (WCLD) | β amplifies but ticker diverges from SPY phases → Recovery or Consolidation LAG |
| High β, negative structural alpha | IWM, EWY, FXI | Falls hard in Fall but LAGs in Recovery — no excess drift to capture |
| Low β, defensive | XLU, XLP, XLV, GLD | STABLE/HEDGE in Fall, then LAG — the mirror-image pattern |

The three Evergreens are **US secular growth stories with β > 1 AND positive structural
alpha above SPY**. SMH is the purest expression: the largest LOSS in Fall (−7.8% vs SPY)
paired with the most explosive Recovery (+25.7%) — semiconductor earnings leverage.

---

### A3. Notable Near-Misses and Why They Fail

| Ticker | Pattern | Where it breaks |
|--------|---------|-----------------|
| **CIBR** | LOSS / ALPHA / LAG | No Consolidation edge — cybersecurity lacks consistent alpha above the tech benchmark |
| **XLC** | LOSS / ALPHA / LAG | Communication Services drags in Consolidation — media and telecom dilute the growth component |
| **ITB** | LOSS / ALPHA / LAG | Homebuilders are rate-sensitive — stall in Consolidation as rates stay elevated |
| **IWM** | LOSS / LAG / LAG | Small caps recover slower than large-cap — no structural alpha, just beta without the drift |
| **IPO** | LOSS / ALPHA / LAG | IPO basket has explosive Recovery but reverts once momentum fades |

The Recovery Sprinter archetype (LOSS / ALPHA / LAG) is the most common near-miss.
These tickers have the beta engine but lack the structural alpha that sustains them
in Consolidation. **Growth beta ≠ growth alpha.**

---

### A4. Non-Equity Universe — Why It Was Not Screened

The screen was intentionally restricted to equities vs SPY.

Non-equity tickers (Fixed Income, Commodities, FX, Real Assets) are **structurally
defined by their divergence from SPY** — that is their purpose. Screening them for the
LOSS / ALPHA / ALPHA pattern would be a category error:

| Asset class | Expected pattern | Why |
|------------|-----------------|-----|
| FI (IEF, TLT, AGG) | HEDGE/STABLE in Fall → LAG in Recovery | Duration rally in stress; left behind when risk-on returns |
| GLD | HEDGE in Fall → LAG in Recovery | Canonical safe haven — the exact opposite archetype |
| HYG | STABLE in Fall → LAG in Recovery | Semi-equity credit; ρ ≈ 0.65, β < 1 — not an amplifier |
| VIXY | HEDGE in Fall → catastrophic LOSS in Recovery | Inverted cycle amplifier — never held long |

**Conclusion:** Zero Evergreens expected or found outside equities. The search is
complete within the scope of SPY-centric tactical allocation.

---

### A5. Sector Archetypes — Full Reasoning

#### Archetype 1 — Cycle Amplifier: XLK only

LOSS −3.9% / ALPHA +16.8% / ALPHA +1.1%  |  IR +0.57  |  ρ 0.93

The only sector with structural positive alpha above SPY in *both* Recovery and
Consolidation. Technology sector earnings (cloud, semiconductors, software) compound
faster than the S&P 500 average across the full market cycle.

#### Archetype 2 — Recovery Sprinters: XLC, XLY

| | Fall | Recovery | Consolidation | IR | ρ |
|--|------|----------|---------------|----|---|
| XLC | LOSS −0.0% | ALPHA +3.5% | LAG −2.5% | −0.13 | 0.85 |
| XLY | LOSS −2.7% | ALPHA +2.7% | LAG −1.7% | −0.19 | 0.90 |

Half-Evergreens. XLY (Amazon, Tesla) and XLC (Meta, Alphabet) sprint off the bottom
driven by consumer re-acceleration. They lose their edge in Consolidation as the media
and telecom components of XLC and the cyclical retail components of XLY normalise.
**Not structural alpha — mean-reversion beta.**

#### Archetype 3 — Defensive Burden: XLP, XLU, XLV, XLRE

**High-Yield Trap** (strongest Fall protection, harshest Recovery penalty):

| | Fall | Recovery | Consolidation | IR | ρ |
|--|------|----------|---------------|----|---|
| XLP Staples | STABLE +9.6% | LAG −15.6% | LAG −6.8% | −0.37 | 0.68 |
| XLU Utilities | STABLE +8.6% | LAG −17.7% | ALPHA −1.7% | −0.15 | 0.58 |

**The XLP/XLU protection premium is a tactical illusion:**

| | Fall α | Recovery α | Full-cycle net |
|--|--------|------------|----------------|
| XLP | +9.6% | −15.6% | **−6.0%** |
| XLU | +8.6% | −17.7% | **−9.1%** |

The protection is fully clawed back — and then some — on the very next regime transition.

**Defensive with Long Lag:**

| | Fall | Recovery | Consolidation | IR | ρ |
|--|------|----------|---------------|----|---|
| XLV Healthcare | STABLE +6.7% | LAG −14.8% | ALPHA −3.6% | −0.33 | 0.79 |
| XLRE Real Estate | STABLE +1.3% | LAG −11.3% | LAG −3.9% | −0.41 | 0.73 |

XLV is commonly labelled defensive. It is — in Fall. The −14.8% Recovery lag is as
punishing as XLP's. Over a full cycle its IR is −0.33. XLRE barely protects in Fall
(+1.3%) yet lags severely — the worst risk/reward in the defensive group.

#### Archetype 4 — Late Cyclicals: XLF, XLI

| | Fall | Recovery | Consolidation | IR | ρ |
|--|------|----------|---------------|----|---|
| XLI Industrials | STABLE −0.5% | LAG −0.5% | ALPHA −1.4% | −0.15 | 0.89 |
| XLF Financials | STABLE −1.5% | LAG −8.9% | ALPHA +3.4% | −0.28 | 0.85 |

XLI is near-zero in every regime — essentially SPY with a small permanent haircut and
no differentiated behavior. XLF is the most interesting sector outside XLK: it lags
hard in Recovery (credit cycle tightening, rising defaults) but delivers the best
Consolidation alpha of any non-tech sector (+3.4%, yield curve steepening).

#### Archetype 5 — All-Regime Drag: XLB

LOSS −0.2% / LAG −5.2% / LAG −4.3%  |  IR −0.43  |  ρ 0.84

No regime is favorable. Materials fall with SPY in Fall, then lag in both recovery
phases. Commodity input cost cycles dominate earnings relative to the S&P 500 mix.

#### Archetype 6 — Macro Satellite: XLE

LOSS +5.9% / LAG −13.0% / ALPHA +2.7%  |  IR −0.09  |  ρ 0.61

ρ = 0.61 — XLE does not run on SPY drawdown cycles. It runs on oil price cycles.
The +5.9% Fall alpha reflects supply-shock episodes (2022) where energy surged while
SPY dropped — that is not a repeatable hedge property. SPY regime classification is
structurally unreliable for XLE; it requires its own commodity regime model.

---

### A6. Factor ETFs — Mapping to Existing Archetypes

| Ticker | Factor | IR | ρ | Fall | Recovery | Consolidation | Maps to |
|--------|--------|----|---|------|----------|---------------|---------|
| MTUM | Momentum | −0.09 | 0.899 | LOSS −1.3% | LAG +0.3% | ALPHA −1.2% | Anti-cycle (unique) |
| QUAL | Quality | −0.20 | 0.984 | LOSS −0.6% | LAG +0.9% | LAG −1.1% | SPY minus alpha |
| COWZ | Cash Flow | −0.11 | 0.863 | LOSS +2.3% | LAG −6.5% | ALPHA +1.5% | Late Cyclical |
| VTV | Value | −0.32 | 0.916 | STABLE +3.1% | LAG −9.2% | ALPHA −0.5% | Late Cyclical |
| DGRW | Div Growth | −0.26 | 0.968 | STABLE +3.1% | LAG −3.5% | LAG −2.9% | Defensive Burden (mild) |
| USMV | Min Vol | −0.43 | 0.899 | STABLE +5.5% | LAG −9.9% | LAG −4.6% | Defensive Burden (strong) |

**QUAL (ρ = 0.984):** The highest correlation in the universe outside SPY itself.
SPY minus alpha — a structural underperformer in every regime. The quality premium did
not manifest vs SPY over this history.

**MTUM — the anti-cycle amplifier:** Crashes in Fall (momentum reversal), lags in
Recovery (beaten-down stocks sprint while momentum rotates slowly), delivers ALPHA in
Consolidation (trending markets). The opposite timing to XLK: worst when Recovery
begins, best when volatility is absent. TAA rotation into MTUM should trigger *after*
Recovery is confirmed, never during it.

**Equal Weight (RSP) — not needed:** Equal-weight S&P 500 is structurally long
value/cyclicals and short mega-cap tech. Pattern would be STABLE / LAG / LAG —
already captured by VTV plus the sector spread in the universe. No new archetype.
The equal-weight vs cap-weight debate since 2020 *is* the tech concentration question —
which the Evergreen screen already resolves: the cycle amplifier pattern is the
mega-cap tech premium.
