---
title: "Phase 2.1 — DD Stabilizer Screen"
subtitle: "Baseline: 6040Classic (60% SPY + 40% IEF)"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
params:
  w_eq:          0.60
  w_fi:          0.40
  screen_weight: 0.10
  roll_win:      63
  n_select:      6
  rho_max:       0.80
  full_corr_thr: -0.10
  part_corr_thr: 0.30
---

```{css, echo=FALSE}
.finding-box {
  background: #f0f9ff;
  border-left: 4px solid #2563eb;
  padding: 10px 16px;
  margin: 12px 0;
  font-size: 0.92em;
}
.type-full  { background:#dbeafe; border-left:4px solid #1d4ed8; padding:6px 14px; margin:6px 0; }
.type-part  { background:#fef9c3; border-left:4px solid #ca8a04; padding:6px 14px; margin:6px 0; }
.type-cond  { background:#fce7f3; border-left:4px solid #db2777; padding:6px 14px; margin:6px 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 6
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

# ── Data ──────────────────────────────────────────────────────────────────────
if (!exists("project_tree"))       source(here("project_tree.R"))
if (!exists("etf_metadata"))       source(here(project_tree$scripts$init))
if (!exists("xts_ret"))            source(here(project_tree$scripts$wrangle))
if (!exists("build_regime_table")) source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
if (!exists("rt") || !is.data.frame(rt))
  rt <- build_regime_table(xts_ret[, "SPY"])
if (!exists("plot_xts_pair"))      source(here("utility/plot_xts_pair.R"))

# Override script-level params from Rmd params before sourcing
W_EQ           <- params$w_eq
W_FI           <- params$w_fi
SCREEN_WEIGHT  <- params$screen_weight
ROLL_WIN       <- as.integer(params$roll_win)
N_SELECT       <- as.integer(params$n_select)
RHO_MAX        <- params$rho_max
FULL_CORR_THR  <- params$full_corr_thr
PART_CORR_THR  <- params$part_corr_thr
BASELINE_LABEL <- sprintf("%d/%d Classic",
                           round(params$w_eq * 100), round(params$w_fi * 100))

# Source the pipeline — knitr.in.progress guard prevents auto-run block from executing
source(here("scripts_daa_saa_taa/05_dd_stabilizer.R"))

# Run the three-stage pipeline
stab <- find_stabilizers(
  xts_ret        = xts_ret,
  rt             = rt,
  w_eq           = W_EQ,
  w_fi           = W_FI,
  baseline_label = BASELINE_LABEL,
  screen_weight  = SCREEN_WEIGHT,
  roll_win       = ROLL_WIN,
  n_select       = N_SELECT,
  rho_max        = RHO_MAX,
  full_corr_thr  = FULL_CORR_THR,
  part_corr_thr  = PART_CORR_THR
)

stabilizer_candidates  <- stab$candidates
stabilizer_scores      <- stab$scores
selected_stabilizers   <- stab$selected
baseline_xts           <- stab$baseline

# DD frontier sweep data
stab_frontier <- build_dd_frontier_data(
  stabilizer_candidates,
  xts_ret,
  baseline_xts
)

baseline_dd  <- .max_dd(baseline_xts)
baseline_ret <- .ann_ret(baseline_xts)
cur_regime   <- unique(stabilizer_scores$cur_regime)[1]
```

---

## Framework

::: {.finding-box}
**Phase 2.1 objective:** *Given 6040Classic as our SAA, find the minimum-cost overlay
that reduces MaxDD — with maximum Fall protection efficiency.*

**Key difference from Phase 2.2 (Enhancer):**
Phase 2.2 seeks return *above* the baseline. Phase 2.1 seeks drawdown *below* the baseline.
A stabilizer may sacrifice some return in exchange for a shallower crash profile.
:::

### Baseline: `r BASELINE_LABEL`

| Metric | Value |
|---|---|
| Composition | `r round(W_EQ*100)`% SPY + `r round(W_FI*100)`% IEF |
| Annualised Return | `r sprintf("%+.1f%%", baseline_ret * 100)` |
| MaxDD | `r sprintf("%.1f%%", baseline_dd * 100)` |
| Current SPY Regime | **`r cur_regime`** |

### Hedge type taxonomy

<div class="type-full">**FULL** — `fall_corr ≤ r FULL_CORR_THR`: negatively correlated with baseline in Fall.
True crisis hedge. Likely to *appreciate* when SPY falls (e.g. GLD, SHY, IEF-type assets).</div>

<div class="type-part">**PARTIAL** — `r FULL_CORR_THR < fall_corr ≤ r PART_CORR_THR`: uncorrelated to weakly positive in Fall.
Reduces DD through low-correlation portfolio math, but does not actively hedge (e.g. trend-following, commodities).</div>

<div class="type-cond">**CONDITIONAL** — `fall_corr > r PART_CORR_THR`: positive Fall correlation but still reduces MaxDD
at the portfolio level via dilution or timing effects. Use with caution — limited crisis protection.</div>

### Three-stage pipeline

| Stage | Function | Cadence | Purpose |
|---|---|---|---|
| 1 | `build_stabilizer_candidates()` | Quarterly | Screen all tickers: dd_reduction > 0 at `r round(SCREEN_WEIGHT*100)`%. Assign hedge type. |
| 2 | `score_stabilizer_context()` | Monthly | Score on 4 current signals. Composite → stabilizer_score 0–100. |
| 3 | `select_stabilizers()` | Monthly | Top-N with ρ ≤ `r RHO_MAX` pairwise guard. |

### Stage 2 signals (equal-weight rank sum)

| ID | Signal | Direction |
|---|---|---|
| s1 | Rolling 63d correlation vs baseline | Lower = better (inverted before ranking) |
| s2 | Historical Fall DD reduction at `r round(SCREEN_WEIGHT*100)`% | Higher = better |
| s3 | Own 200DMA trend signal | Higher = asset trending (optional — skipped if unavailable) |
| s4 | Own 63d annualised return | Higher = stabilizer currently "working" |

---

## Stage 1 — Candidates (`r nrow(stabilizer_candidates)` tickers)

```{r stage1-table}
type_colours <- c("FULL" = "#dbeafe", "PARTIAL" = "#fef9c3", "CONDITIONAL" = "#fce7f3")

stabilizer_candidates %>%
  mutate(
    dd_reduction    = sprintf("%+.1f pp", dd_reduction    * 100),
    fall_dd_red     = sprintf("%+.1f pp", fall_dd_reduction * 100),
    return_drag     = sprintf("%+.1f pp", return_drag     * 100),
    insurance_ratio = sprintf("%.2f",     insurance_ratio),
    fall_corr       = sprintf("%+.2f",    fall_corr),
    own_maxdd       = sprintf("%.1f%%",   own_maxdd       * 100),
    own_ret         = sprintf("%+.1f%%",  own_ret         * 100)
  ) %>%
  select(ticker, asset_class, hedge_type,
         dd_reduction, fall_dd_red, return_drag,
         insurance_ratio, fall_corr, own_maxdd, own_ret, n_days) %>%
  kable(
    col.names = c("Ticker", "Asset class", "Hedge type",
                  "DD Red", "Fall DD Red", "Return drag",
                  "Ins ratio", "Fall ρ", "Own MaxDD", "Own Ret", "N days"),
    align = "lllrrrrrrrl"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(
    which(stabilizer_candidates$hedge_type == "FULL"),
    background = "#dbeafe"
  ) %>%
  row_spec(
    which(stabilizer_candidates$hedge_type == "PARTIAL"),
    background = "#fef9c3"
  ) %>%
  row_spec(
    which(stabilizer_candidates$hedge_type == "CONDITIONAL"),
    background = "#fce7f3"
  )
```

---

## Stage 2 — Context scores (current regime: `r cur_regime`)

```{r stage2-table}
stabilizer_scores %>%
  arrange(desc(stabilizer_score)) %>%
  mutate(
    stabilizer_score = sprintf("%.1f",   stabilizer_score),
    rolling_corr_63d = sprintf("%+.2f",  rolling_corr_63d),
    fall_dd_red      = sprintf("%+.1f pp", fall_dd_red * 100),
    trend_f          = if_else(is.na(trend_on), "n/a", if_else(trend_on == 1, "ON", "OFF")),
    own_mom          = sprintf("%+.1f%%", own_momentum_63d * 100),
    dd_red_f         = sprintf("%+.1f pp", dd_reduction * 100),
    ins_f            = sprintf("%.2f",   insurance_ratio),
    fall_corr_f      = sprintf("%+.2f",  fall_corr)
  ) %>%
  select(ticker, asset_class, hedge_type, stabilizer_score,
         rolling_corr_63d, fall_dd_red, trend_f, own_mom,
         dd_red_f, ins_f, fall_corr_f) %>%
  kable(
    col.names = c("Ticker", "Asset class", "Hedge type", "Score",
                  "Roll ρ 63d", "Fall DD Red", "200DMA", "Own Mom 63d",
                  "DD Red", "Ins ratio", "Fall ρ"),
    align = "lllrrrrrrrr"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(
    which(stabilizer_scores %>% arrange(desc(stabilizer_score)) %>%
          pull(hedge_type) == "FULL"),
    background = "#dbeafe"
  ) %>%
  row_spec(
    which(stabilizer_scores %>% arrange(desc(stabilizer_score)) %>%
          pull(hedge_type) == "PARTIAL"),
    background = "#fef9c3"
  )
```

---

## Stage 3 — Selected stabilizers (`r nrow(selected_stabilizers)` tickers)

```{r stage3-table}
selected_stabilizers %>%
  mutate(
    stabilizer_score = sprintf("%.1f",   stabilizer_score),
    dd_red_f         = sprintf("%+.1f pp", dd_reduction    * 100),
    fall_dd_f        = sprintf("%+.1f pp", fall_dd_reduction * 100),
    drag_f           = sprintf("%+.1f pp", return_drag     * 100),
    ins_f            = sprintf("%.2f",   insurance_ratio),
    fall_corr_f      = sprintf("%+.2f",  fall_corr),
    roll_corr_f      = sprintf("%+.2f",  rolling_corr_63d),
    trend_f          = if_else(is.na(trend_on), "n/a", if_else(trend_on == 1, "ON", "OFF")),
    momentum_f       = sprintf("%+.1f%%", own_momentum_63d * 100)
  ) %>%
  select(rank, ticker, name, asset_class, hedge_type, stabilizer_score,
         dd_red_f, fall_dd_f, drag_f, ins_f, fall_corr_f, roll_corr_f,
         trend_f, momentum_f) %>%
  kable(
    col.names = c("#", "Ticker", "Name", "Asset class", "Hedge type", "Score",
                  "DD Red", "Fall DD Red", "Drag", "Ins ratio",
                  "Fall ρ", "Roll ρ 63d", "200DMA", "Own Mom"),
    align = "rlllllrrrrrrrr"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(
    which(selected_stabilizers$hedge_type == "FULL"),
    background = "#dbeafe"
  ) %>%
  row_spec(
    which(selected_stabilizers$hedge_type == "PARTIAL"),
    background = "#fef9c3"
  )
```

---

## Insurance Quadrant

```{r insurance-quad, fig.height=7, echo=TRUE}
# plot_stabilizer_quad(stabilizer_candidates)
plot_stabilizer_quad(stabilizer_candidates)
```

---

## DD Frontier — overlay weight 0% → 25%

```{r dd-frontier, fig.height=6, echo=TRUE}
# plot_dd_frontier(stab_frontier)
plot_dd_frontier(stab_frontier)
```

---

## Sleeve wealth charts

```{r sleeve-charts, fig.height=5, results='asis', echo=TRUE}
# plot_stabilizer_sleeves(selected_stabilizers, xts_ret, rt)
plot_stabilizer_sleeves(selected_stabilizers, xts_ret, rt)
```

---

## Ticker fingerprints

9-panel fingerprint for each selected stabilizer vs `r BASELINE_LABEL`.
Panels: 1. Wealth | 2. L/S Spread | 3. Rolling L/S | 4. DD | 5. 250d Return |
6. IR | 7. Rolling ρ | 8. Regime ρ | 9. Distribution

```{r fingerprints, fig.height=21, results='asis', echo=TRUE}
# plot_stabilizer_pairs(selected_stabilizers, xts_ret, rt, save_png = FALSE)
plot_stabilizer_pairs(
  selected_stabilizers = selected_stabilizers,
  xts_ret              = xts_ret,
  rt                   = rt,
  baseline_label       = BASELINE_LABEL,
  save_png             = FALSE
)
```

---

## Session info

```{r session}
cat(sprintf("Report date   : %s\n", Sys.Date()))
cat(sprintf("Baseline      : %s (%.0f%% SPY + %.0f%% IEF)\n",
            BASELINE_LABEL, W_EQ*100, W_FI*100))
cat(sprintf("Screen weight : %.0f%%\n", SCREEN_WEIGHT * 100))
cat(sprintf("Roll window   : %d days\n", ROLL_WIN))
cat(sprintf("N selected    : %d  (rho_max = %.2f)\n", N_SELECT, RHO_MAX))
cat(sprintf("Hedge types   : FULL ρ ≤ %.2f  |  PARTIAL ρ ≤ %.2f  |  CONDITIONAL otherwise\n",
            FULL_CORR_THR, PART_CORR_THR))
cat(sprintf("Current regime: %s\n", cur_regime))
cat(sprintf("Candidates    : %d  →  Scored: %d  →  Selected: %d\n",
            nrow(stabilizer_candidates), nrow(stabilizer_scores),
            nrow(selected_stabilizers)))
```
