---
title: "Stock Market Return Matrix"
subtitle: "S&P 500 Nominal Price Returns — Documentation & Analysis"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    fig_caption: true
params:
  from_yr:      1900   # first FROM year in the matrix
  to_yr:        null   # null = latest available; or e.g. 2024
  show_25yr:    true   # draw 25-year period diagonal
  show_text:    true   # show annualised return numbers in cells
  force_refresh: false # re-download Shiller data even if cached
---

```{css, echo=FALSE}
.doc-note {
  background: #fffbeb;
  border-left: 4px solid #d97706;
  padding: 8px 14px;
  margin: 10px 0;
  font-size: 0.88em;
  color: #444;
}
.finding-box {
  background: #f0fdf4;
  border-left: 4px solid #16a34a;
  padding: 10px 16px;
  margin: 12px 0;
  font-size: 0.9em;
}
.warn-box {
  background: #fef2f2;
  border-left: 4px solid #dc2626;
  padding: 8px 14px;
  margin: 10px 0;
  font-size: 0.88em;
}
.caption {
  color: #555;
  font-size: 0.82em;
  font-style: italic;
  border-left: 3px solid #1B3A6B;
  padding-left: 6px;
  margin-top: 4px;
}
.colour-swatch {
  display: inline-block;
  width: 18px; height: 18px;
  border-radius: 3px;
  vertical-align: middle;
  margin-right: 4px;
}
.section-div { border-top: 2px solid #dee2e6; margin: 48px 0 24px 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(readxl)
library(scales)
library(knitr)
library(kableExtra)
library(here)

select <- dplyr::select

if (!exists("params"))
  params <- list(from_yr = 1900, to_yr = NULL, show_25yr = TRUE,
                 show_text = TRUE, force_refresh = FALSE)

# ── Load chart function ───────────────────────────────────────────────────────
source(here("key_plots/chart_stock_matrix.R"))

# ── Build data ────────────────────────────────────────────────────────────────
annual <- .load_shiller_annual(
  cache_path    = here("input/shiller_annual.rds"),
  force_refresh = isTRUE(params$force_refresh)
)

to_yr   <- if (is.null(params$to_yr)) max(annual$year) else as.integer(params$to_yr)
from_yr <- as.integer(params$from_yr)

mat <- .build_return_matrix(annual, from_yr - 1, to_yr) %>%
  filter(from >= from_yr, to <= to_yr)

n_years  <- to_yr - from_yr + 1
n_cells  <- nrow(mat)
n_yrs_data <- as.numeric(diff(range(annual$year)))
```

---

# What This Chart Shows

The **Stock Market Return Matrix** answers a single question for every possible entry and exit year:

> *"If you invested in the S&P 500 at the start of year X and sold at the end of year Y, what was your annualised nominal return?"*

Each cell in the lower-right triangle shows the **annualised price-only return** (no dividends, no inflation adjustment) for that FROM→TO period, colour-coded by magnitude.

<div class="doc-note">
This is a reproduction of the <strong>Crestmont Research Stock Market Matrix</strong>
(www.CrestmontResearch.com), rebuilt from the same underlying data source (Robert Shiller / Yale).
Original copyright 2002–2025 Crestmont Research.
</div>

---

# Colour Legend & Reading the Matrix

```{r legend-table}
leg <- tibble(
  ` ` = c(
    "<span class='colour-swatch' style='background:#c0392b'></span>",
    "<span class='colour-swatch' style='background:#e8a0b4'></span>",
    "<span class='colour-swatch' style='background:#2155a0'></span>",
    "<span class='colour-swatch' style='background:#5aab5a'></span>",
    "<span class='colour-swatch' style='background:#1a6b1a'></span>"
  ),
  Band          = c("Red", "Pink", "Blue", "Light Green", "Dark Green"),
  `Return range` = c("< 0%", "0% to 3%", "3% to 7%", "7% to 10%", "> 10%"),
  Interpretation = c(
    "Loss — negative annualised return over the period",
    "Near-zero real return after inflation (~2–3%/yr CPI); wealth preserved, not grown",
    "Moderate nominal return; modest real gain after inflation",
    "Good nominal return; solid real gain",
    "Strong compounding — typically short bull market periods"
  )
)

kable(leg, escape = FALSE,
      col.names = c("", "Band", "Return range", "Interpretation"),
      caption = "Colour bands — annualised nominal price return") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, width = "2em") %>%
  column_spec(2, bold = TRUE) %>%
  column_spec(3, monospace = TRUE)
```

**How to read a cell:**
Pick a FROM year on the left axis, a TO year on the top axis, find the intersection — the number is the annualised % return for that holding period.

**White numbers** = P/E10 (CAPE) *decreased* over the period. The return was earned in a valuation headwind — price gains came from earnings growth alone, not multiple expansion. These are considered more "durable" returns by Crestmont.

**The 25-year diagonal line** (when shown) marks the boundary where holding periods equal 25 years. All cells to the right of the line are 25+ year periods.

---

# The Matrix {.tabset}

## Full matrix (`r from_yr`–`r to_yr`)

```{r full-matrix, fig.width=16, fig.height=15}
plot_stock_matrix(
  from_yr       = from_yr,
  to_yr         = to_yr,
  show_25yr     = isTRUE(params$show_25yr),
  show_text     = isTRUE(params$show_text),
  force_refresh = FALSE
)
```

<p class="caption">
Annualised nominal S&P 500 price returns (no dividends). Each cell = holding period from row year
to column year. Numbers = annualised return (%). White numbers = CAPE decreased (valuation headwind).
Black diagonal line = 25-year holding period boundary.
Save the plot as a large PNG for full legibility:
<code>ggsave(here("key_plots/stock_matrix.png"), width=22, height=20, dpi=200)</code>
</p>

## Post-war only (1950–`r to_yr`)

```{r postwar-matrix, fig.width=14, fig.height=13}
plot_stock_matrix(
  from_yr   = 1950,
  to_yr     = to_yr,
  show_25yr = TRUE,
  show_text = isTRUE(params$show_text)
)
```

<p class="caption">Post-war subset (1950 onwards). More readable at screen size. Same colour coding.</p>

## Recent decades (1990–`r to_yr`)

```{r recent-matrix, fig.width=10, fig.height=9}
plot_stock_matrix(
  from_yr   = 1990,
  to_yr     = to_yr,
  show_25yr = TRUE,
  show_text = TRUE
)
```

<p class="caption">1990 onwards — fully legible cell numbers at screen resolution.</p>

---

# Statistical Analysis {.tabset}

```{r stats-prep}
# Return distribution by holding period bucket
mat_analysis <- mat %>%
  mutate(
    hold_yrs = to - from,
    hold_bucket = cut(hold_yrs,
      breaks = c(0, 1, 5, 10, 20, 30, Inf),
      labels = c("1 yr", "2-5 yrs", "6-10 yrs", "11-20 yrs", "21-30 yrs", "30+ yrs"),
      right  = TRUE
    )
  )
```

## Return distribution by holding period

```{r dist-plot, fig.height=6}
ggplot(mat_analysis %>% filter(!is.na(hold_bucket)),
       aes(x = ann_ret_pct, fill = band)) +
  geom_histogram(bins = 40, colour = "white", linewidth = 0.15) +
  scale_fill_manual(values = SM_FILL, limits = SM_LABELS,
                    labels = c("< 0%","0-3%","3-7%","7-10%","> 10%"),
                    name = NULL) +
  geom_vline(xintercept = 0, colour = "#111", linewidth = 0.6, linetype = "dashed") +
  facet_wrap(~ hold_bucket, scales = "free_y", ncol = 3) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title    = "Distribution of annualised returns by holding period",
    subtitle = paste0("All FROM-TO pairs, ", from_yr, "-", to_yr,
                      " (n = ", scales::comma(n_cells), " periods)"),
    x = "Annualised return (%)", y = "Count"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor  = element_blank(),
    strip.text        = element_text(face = "bold"),
    legend.position   = "bottom"
  )
```

<p class="caption">As holding period lengthens the distribution narrows and shifts right — time diversification at work. Negative returns become rare beyond 10-year periods.</p>

## Probability of positive return

```{r prob-table}
prob_tbl <- mat_analysis %>%
  filter(!is.na(ann_ret_pct)) %>%
  group_by(`Holding period` = hold_bucket) %>%
  summarise(
    `N periods`        = n(),
    `% Positive`       = paste0(round(mean(ann_ret_pct > 0)  * 100, 1), "%"),
    `% Above 3%`       = paste0(round(mean(ann_ret_pct > 3)  * 100, 1), "%"),
    `% Above 7%`       = paste0(round(mean(ann_ret_pct > 7)  * 100, 1), "%"),
    `Median ann. ret`  = paste0(round(median(ann_ret_pct), 1), "%"),
    `Mean ann. ret`    = paste0(round(mean(ann_ret_pct),   1), "%"),
    `Worst`            = paste0(round(min(ann_ret_pct),    1), "%"),
    `Best`             = paste0(round(max(ann_ret_pct),    1), "%"),
    .groups = "drop"
  )

kable(prob_tbl,
      caption = paste0("Return statistics by holding period (", from_yr, "-", to_yr, ")")) %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

## Colour band frequency

```{r band-freq, fig.height=5}
band_freq <- mat_analysis %>%
  filter(!is.na(band)) %>%
  group_by(hold_bucket, band) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(hold_bucket) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  mutate(band = factor(band, levels = SM_LABELS))

ggplot(band_freq, aes(x = hold_bucket, y = pct, fill = band)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = SM_FILL, limits = SM_LABELS,
                    labels = c("< 0%","0-3%","3-7%","7-10%","> 10%"),
                    name = "Return band") +
  scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
  labs(
    title    = "Return band composition by holding period",
    subtitle = "How often does each return band appear as holding period lengthens?",
    x = "Holding period", y = "% of all periods"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position   = "right"
  )
```

<p class="caption">Red and Pink (low/negative returns) dominate short periods; Blue and Green dominate long periods. Beyond 20 years almost no Red remains in this history.</p>

---

# P/E Ratio & White Numbers {.tabset}

<div class="doc-note">
Crestmont's key insight: a market return has two components — <strong>earnings growth</strong>
and <strong>P/E expansion</strong>. White numbers show periods where all return came from
earnings alone (P/E headwind). These are considered more durable. Coloured numbers include
a P/E tailwind that may not repeat.
</div>

## P/E history

```{r pe-history, fig.height=4}
annual %>%
  filter(year >= from_yr, year <= to_yr, !is.na(cape_avg)) %>%
  ggplot(aes(x = year, y = cape_avg)) +
  geom_line(colour = "#2155a0", linewidth = 0.9) +
  geom_hline(yintercept = mean(annual$cape_avg, na.rm = TRUE),
             linetype = "dashed", colour = "#888", linewidth = 0.5) +
  annotate("text", x = from_yr + 2,
           y = mean(annual$cape_avg, na.rm = TRUE) + 1.5,
           label = paste0("Avg = ", round(mean(annual$cape_avg, na.rm=TRUE), 1)),
           size = 3, colour = "#888") +
  scale_x_continuous(breaks = seq(from_yr, to_yr, by = 10)) +
  labs(
    title    = "CAPE (P/E10) — annual average",
    subtitle = "Robert Shiller cyclically adjusted P/E ratio. Dashed = long-run average.",
    x = NULL, y = "CAPE"
  ) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank())
```

## White vs coloured number breakdown

```{r white-numbers}
pe_tbl <- mat %>%
  filter(!is.na(pe_fell), !is.na(band)) %>%
  group_by(`Holding period` = cut(to - from,
    breaks = c(0,1,5,10,20,30,Inf),
    labels = c("1 yr","2-5 yrs","6-10 yrs","11-20 yrs","21-30 yrs","30+ yrs"),
    right = TRUE
  )) %>%
  summarise(
    `N periods`        = n(),
    `P/E fell (white)` = paste0(round(mean(pe_fell) * 100, 1), "%"),
    `P/E rose`         = paste0(round(mean(!pe_fell) * 100, 1), "%"),
    `Avg ret (P/E fell)`  = paste0(round(mean(ann_ret_pct[pe_fell],  na.rm=TRUE), 1), "%"),
    `Avg ret (P/E rose)`  = paste0(round(mean(ann_ret_pct[!pe_fell], na.rm=TRUE), 1), "%"),
    .groups = "drop"
  )

kable(pe_tbl,
      caption = "P/E direction by holding period — white (fell) vs coloured (rose) numbers") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

---

# Data & Methodology {.tabset}

## Data source

```{r data-table}
data_tbl <- tibble(
  Item = c(
    "Price index", "P/E ratio", "Returns type",
    "Dividends", "Inflation", "Transaction costs",
    "Data source", "Annual price", "Update frequency"
  ),
  Detail = c(
    "S&P 500 Composite (and predecessor indices pre-1957)",
    "CAPE / P/E10 — cyclically adjusted using 10-year average earnings (Robert Shiller methodology)",
    "Nominal price-only (not total return)",
    "NOT included — price index only",
    "NOT adjusted — all figures are nominal",
    "NOT included",
    paste0("Robert Shiller / Yale University — ie_data.xls  |  ",
           "History: ", min(annual$year), " \u2013 ", max(annual$year),
           "  (", nrow(annual), " annual observations)"),
    "Annual average of monthly price observations (12 months per year)",
    "Annually, when Shiller updates ie_data.xls"
  )
)

kable(data_tbl, caption = "Data and methodology specification") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A", width = "18em")
```

## Return calculation

The annualised return for a FROM→TO period is:

$$\text{Ann. Return} = \left(\frac{\bar{P}_{TO}}{\bar{P}_{FROM}}\right)^{\frac{1}{TO - FROM}} - 1$$

where $\bar{P}_{year}$ is the **annual average price** (mean of all 12 monthly observations).

Key implications:
- A 1-year return uses consecutive annual averages, not January-to-January or December-to-December prices
- This smooths intra-year volatility and matches Shiller's published convention
- Dividends are excluded — total return would be roughly 4–5pp/yr higher over most of history

## Coverage summary

```{r coverage}
cov_tbl <- tibble(
  Metric = c(
    "FROM year range",
    "TO year range",
    "Total holding periods computed",
    "1-year periods",
    "5-year periods",
    "10-year periods",
    "20-year periods",
    "25-year periods",
    "Cells with P/E data (white/colour flag)"
  ),
  Value = c(
    paste0(from_yr, " \u2013 ", to_yr - 1),
    paste0(from_yr + 1, " \u2013 ", to_yr),
    scales::comma(n_cells),
    scales::comma(sum(mat$to - mat$from == 1)),
    scales::comma(sum(mat$to - mat$from == 5)),
    scales::comma(sum(mat$to - mat$from == 10)),
    scales::comma(sum(mat$to - mat$from == 20)),
    scales::comma(sum(mat$to - mat$from == 25)),
    paste0(scales::comma(sum(!is.na(mat$pe_fell))),
           " (", round(mean(!is.na(mat$pe_fell))*100), "%)")
  )
)

kable(cov_tbl, caption = "Matrix coverage") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "left", font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A")
```

---

# How to Render & Save

```{r render-code, echo=TRUE, eval=FALSE}
# ── From RStudio console ──────────────────────────────────────────────────────

# Quick view (loads functions + auto-runs quick-run block):
source(here::here("key_plots/chart_stock_matrix.R"))

# Full-resolution PNG (open in Finder/Preview for legibility):
p <- plot_stock_matrix(from_yr = 1900, show_25yr = TRUE)
ggsave(here::here("key_plots/stock_matrix.png"), p,
       width = 22, height = 20, dpi = 200, bg = "white")

# Post-war only (faster, more readable at screen size):
p <- plot_stock_matrix(from_yr = 1950)
ggsave(here::here("key_plots/stock_matrix_postwar.png"), p,
       width = 14, height = 13, dpi = 200, bg = "white")

# ── Render this report ────────────────────────────────────────────────────────
rmarkdown::render(
  here::here("Rmd/stock_matrix_report.Rmd"),
  output_file = here::here("03_reports/stock_matrix_report.html"),
  params = list(from_yr = 1900, show_25yr = TRUE, show_text = TRUE),
  envir  = new.env()
)

# ── Force data refresh ────────────────────────────────────────────────────────
rmarkdown::render(
  here::here("Rmd/stock_matrix_report.Rmd"),
  output_file = here::here("03_reports/stock_matrix_report.html"),
  params = list(force_refresh = TRUE),
  envir  = new.env()
)
```

---

```{r footer}
cat(sprintf(
  "_Generated %s &nbsp;|&nbsp; FROM: %d &nbsp;|&nbsp; TO: %d &nbsp;|&nbsp; %s holding periods &nbsp;|&nbsp; Data: Shiller Yale ie\\_data.xls_",
  format(Sys.Date(), "%Y-%m-%d"),
  from_yr, to_yr,
  scales::comma(n_cells)
))
```
