---
title: "TAA Momentum Screen"
subtitle: "Short-Term Relative Outperformance vs SPY · 6M / 12M / 2Y Horizons"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(scales)
library(knitr)
library(kableExtra)
library(here)

select <- dplyr::select

# ── Data ──────────────────────────────────────────────────────────────────────
xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
source(here("scripts/00_init_universe.R"))
source(here("scripts_saa_taa/taa_momentum_screen.R"))

# ── Build screen ──────────────────────────────────────────────────────────────
scr <- build_taa_screen(xts_ret, etf_metadata)

# ── Quadrant colour map (for tables) ─────────────────────────────────────────
QUAD_BG <- c(
  "TAA Only"   = "#FFF3E0",
  "SAA + TAA"  = "#E8F5E9",
  "SAA (cold)" = "#E3F2FD",
  "Avoid"      = "#FFEBEE"
)
QUAD_FG <- c(
  "TAA Only"   = "#E65100",
  "SAA + TAA"  = "#2E7D32",
  "SAA (cold)" = "#1565C0",
  "Avoid"      = "#C62828"
)
```

---

## Section 1 — Executive Summary

```{r exec-counts}
# ── Quadrant counts ─────────────────────────────────────────────────────────
counts_df <- scr %>%
  count(quadrant) %>%
  mutate(
    quadrant = factor(quadrant,
                      levels = c("SAA + TAA", "TAA Only",
                                 "SAA (cold)", "Avoid")),
    Description = case_when(
      quadrant == "SAA + TAA"  ~ "Structurally superior AND currently outperforming. Dual SAA/TAA case.",
      quadrant == "TAA Only"   ~ "Recently hot but no durable edge over SPY. Pure momentum/tactical play.",
      quadrant == "SAA (cold)" ~ "Full-period quality exceeds threshold but lagging short-term. Wait for entry.",
      quadrant == "Avoid"      ~ "Neither durable nor currently hot. No TAA or SAA case."
    )
  ) %>%
  arrange(quadrant) %>%
  rename(Quadrant = quadrant, `# Tickers` = n)

kable(counts_df, align = c("l", "r", "l"),
      caption = paste0("TAA Screen Summary — ", nrow(scr),
                       " tickers · Data as of ", format(Sys.Date()))) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  row_spec(which(counts_df$Quadrant == "SAA + TAA"),
           background = QUAD_BG["SAA + TAA"], color = QUAD_FG["SAA + TAA"],
           bold = TRUE) %>%
  row_spec(which(counts_df$Quadrant == "TAA Only"),
           background = QUAD_BG["TAA Only"]) %>%
  row_spec(which(counts_df$Quadrant == "SAA (cold)"),
           background = QUAD_BG["SAA (cold)"]) %>%
  row_spec(which(counts_df$Quadrant == "Avoid"),
           background = QUAD_BG["Avoid"]) %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2, bold = TRUE)
```

```{r exec-saa-taa-table}
# ── SAA + TAA candidates — detail ───────────────────────────────────────────
saa_taa <- scr %>%
  filter(quadrant == "SAA + TAA") %>%
  arrange(desc(ir_full)) %>%
  mutate(
    rel_6m      = percent(rel_6m,      accuracy = 1),
    rel_12m     = percent(rel_12m,     accuracy = 1),
    rel_2y      = percent(rel_2y,      accuracy = 1),
    ir_full     = sprintf("%+.2f", ir_full),
    ir_1y       = sprintf("%+.2f", ir_1y),
    consist_pct = percent(consist_pct, accuracy = 1)
  ) %>%
  select(Ticker = ticker, Name = name, `Asset Class` = asset_class,
         `6M α` = rel_6m, `12M α` = rel_12m, `2Y α` = rel_2y,
         `IR (full)` = ir_full, `IR (1Y)` = ir_1y,
         Consistency = consist_pct)

kable(saa_taa, align = c("l","l","l","r","r","r","r","r","r"),
      caption = "SAA + TAA Candidates — Structurally Superior AND Currently Outperforming") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  row_spec(seq_len(nrow(saa_taa)), background = QUAD_BG["SAA + TAA"])
```

```{r exec-saa-cold-table}
# ── SAA (cold) — worth watching ─────────────────────────────────────────────
saa_cold <- scr %>%
  filter(quadrant == "SAA (cold)") %>%
  arrange(desc(ir_full)) %>%
  mutate(
    rel_6m      = percent(rel_6m,      accuracy = 1),
    rel_12m     = percent(rel_12m,     accuracy = 1),
    rel_2y      = percent(rel_2y,      accuracy = 1),
    ir_full     = sprintf("%+.2f", ir_full),
    ir_1y       = sprintf("%+.2f", ir_1y),
    consist_pct = percent(consist_pct, accuracy = 1)
  ) %>%
  select(Ticker = ticker, Name = name, `Asset Class` = asset_class,
         `6M α` = rel_6m, `12M α` = rel_12m, `2Y α` = rel_2y,
         `IR (full)` = ir_full, `IR (1Y)` = ir_1y,
         Consistency = consist_pct)

kable(saa_cold, align = c("l","l","l","r","r","r","r","r","r"),
      caption = "SAA (Cold) — Structurally Superior but Lagging Short-Term (Potential Entry Points)") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                position = "center", font_size = 13) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  row_spec(seq_len(nrow(saa_cold)), background = QUAD_BG["SAA (cold)"])
```

> **Classification logic:** `TAA Only` = 6M relative return > 5% AND full-period IR < 0.10 ·
> `SAA + TAA` = both thresholds passed · `SAA (cold)` = quality IR ≥ 0.10 but 6M < 5% ·
> `Avoid` = neither criterion met.
> **Consistency** = % of rolling 126-day (6M) windows with positive relative return vs SPY.

---

## Section 2 — Quadrant Scatter

*Each dot is one ETF. Colour = asset class. Size = |trailing relative return|.*
*Dashed lines divide the four quadrants. Only TAA-relevant tickers are labelled.*

```{r scatter-6m, fig.height=7}
print(plot_taa_scatter(scr, "rel_6m"))
```

```{r scatter-12m, fig.height=7}
print(plot_taa_scatter(scr, "rel_12m"))
```

```{r scatter-2y, fig.height=7}
print(plot_taa_scatter(scr, "rel_2y"))
```

---

## Section 3 — Ranked Bar Charts

*Top 15 tickers per horizon. Colour = quadrant. Dashed line = 5% TAA threshold.*

```{r ranking, fig.height=6}
print(plot_taa_ranking(scr, n = 15L))
```

---

## Section 4 — Momentum Consistency

*Of the TAA Only + SAA+TAA tickers, how often have they beaten SPY on a rolling 6M basis?
High consistency = sustained momentum. Low consistency = a single lucky spike.*

```{r consistency, fig.height=7}
p_con <- plot_taa_consistency(scr, n = 25L)
if (!is.null(p_con)) print(p_con)
```

---

## Section 5 — TAA Only Candidates: Top 20

*Pure momentum plays — hot now but not structurally superior to SPY.*
*Use these for short-term tactical tilts only; do not size as SAA anchors.*

```{r taa-only-table}
taa_only <- scr %>%
  filter(quadrant == "TAA Only") %>%
  slice_max(rel_6m, n = 20) %>%
  mutate(
    rel_6m      = percent(rel_6m,      accuracy = 1),
    rel_12m     = percent(rel_12m,     accuracy = 1),
    rel_2y      = percent(rel_2y,      accuracy = 1),
    ir_full     = sprintf("%+.2f", ir_full),
    ir_1y       = sprintf("%+.2f", ir_1y),
    consist_pct = percent(consist_pct, accuracy = 1)
  ) %>%
  select(Ticker = ticker, Name = name, `Asset Class` = asset_class,
         `6M α` = rel_6m, `12M α` = rel_12m, `2Y α` = rel_2y,
         `IR (full)` = ir_full, `IR (1Y)` = ir_1y,
         Consistency = consist_pct)

kable(taa_only, align = c("l","l","l","r","r","r","r","r","r"),
      caption = "Top 20 TAA Only Candidates (ranked by trailing 6M α)") %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered"),
                font_size = 12) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  row_spec(seq_len(nrow(taa_only)), background = QUAD_BG["TAA Only"])
```

---

## Section 6 — Full Universe Table

*All `r nrow(scr)` tickers. Sort any column. Use browser search to filter.*

```{r full-table}
full_tbl <- scr %>%
  arrange(quadrant, desc(rel_6m)) %>%
  mutate(
    rel_6m      = percent(rel_6m,      accuracy = 1),
    rel_12m     = percent(rel_12m,     accuracy = 1),
    rel_2y      = percent(rel_2y,      accuracy = 1),
    ir_full     = sprintf("%+.2f", ir_full),
    ir_1y       = sprintf("%+.2f", if_else(is.na(ir_1y), 0, ir_1y)),
    consist_pct = percent(consist_pct, accuracy = 1)
  ) %>%
  select(Ticker = ticker, Name = name, `Asset Class` = asset_class,
         Quadrant = quadrant,
         `6M α` = rel_6m, `12M α` = rel_12m, `2Y α` = rel_2y,
         `IR (full)` = ir_full, `IR (1Y)` = ir_1y,
         Consistency = consist_pct)

# Row colouring by quadrant
quad_rows <- list(
  "SAA + TAA"  = which(full_tbl$Quadrant == "SAA + TAA"),
  "TAA Only"   = which(full_tbl$Quadrant == "TAA Only"),
  "SAA (cold)" = which(full_tbl$Quadrant == "SAA (cold)"),
  "Avoid"      = which(full_tbl$Quadrant == "Avoid")
)

tbl <- kable(full_tbl, align = c("l","l","l","l","r","r","r","r","r","r"),
             caption = paste0("Full TAA Universe — ", nrow(full_tbl),
                              " tickers · as of ", format(Sys.Date()))) %>%
  kable_styling(full_width = TRUE,
                bootstrap_options = c("hover", "condensed", "bordered", "striped"),
                font_size = 11) %>%
  row_spec(0, bold = TRUE, background = "#1B2A4A", color = "white") %>%
  column_spec(1, bold = TRUE, color = "#1B2A4A") %>%
  column_spec(4, bold = TRUE)

for (quad in names(quad_rows)) {
  rows <- quad_rows[[quad]]
  if (length(rows) > 0)
    tbl <- row_spec(tbl, rows,
                    background = QUAD_BG[quad],
                    color      = ifelse(quad %in% c("SAA + TAA", "TAA Only"),
                                        QUAD_FG[quad], "black"))
}

tbl
```

---

```{r footer}
cat(sprintf(
  "_Generated %s · Benchmark: SPY · IR threshold: %.2f · 6M α threshold: %.0f%% · Universe: %d tickers_",
  format(Sys.Date(), "%Y-%m-%d"),
  IR_SAA_THRESHOLD, REL_THRESHOLD * 100, nrow(scr)
))
```
