---
title: "Universe Pair Analysis — Category A & B Shortlists"
subtitle: "Top 10 Equity Swaps (vs SPY) · Top 10 Cross-Asset Additions (vs AGG)"
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(MASS)
library(tidyverse)
library(xts)
library(scales)
library(patchwork)
library(here)

select <- dplyr::select

if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

spy_ret <- xts_ret[, "SPY"]
rt      <- build_regime_table(spy_ret, 0.10, 0.05)

source(here("utility/plot_xts_pair.R"))

# ── Helpers shared with scoring ───────────────────────────────────────────────
T_yrs <- as.numeric(diff(range(index(xts_ret)))) / 365

max_dd_vec <- function(r) {
  w  <- cumprod(1 + r)
  dd <- (w - cummax(w)) / cummax(w)
  min(dd, na.rm = TRUE)
}

magdon_mdd <- function(sigma, mu, T) {
  if (is.na(mu) || mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  (sigma / S) * (log(S * sqrt(T)) + 0.35)
}

pair_stats <- function(out) {
  s  <- out$stats
  sp <- s %>% filter(str_detect(Series, "\u2212"))   # "−" minus sign
  mu  <- parse_number(sp$`Ann Return`) / 100
  sig <- parse_number(sp$`Ann Vol`)    / 100
  mdd <- parse_number(sp$`Max DD`)     / 100
  list(
    mu         = mu,
    sig        = sig,
    mdd        = mdd,
    ir         = out$full_ir,
    rho        = out$full_cor,
    te_10pct   = 0.10 * sig,
    mdd_10pct  = 0.10 * abs(mdd),
    ret_10pct  = 0.10 * mu,
    mdd_theory = magdon_mdd(sig, mu, T_yrs),
    stats      = s
  )
}

# ── Scoring components (reproduced from ticker_scoring.R) ─────────────────────
minmax <- function(x) {
  rng <- range(x, na.rm = TRUE)
  if (diff(rng) == 0) return(rep(0.5, length(x)))
  (x - rng[1]) / diff(rng)
}

spread_raw <- function(ticker, benchmark) {
  both <- merge(xts_ret[, ticker], xts_ret[, benchmark], join = "inner")
  both <- na.omit(both)
  r_t  <- as.numeric(both[, 1])
  r_b  <- as.numeric(both[, 2])
  r_s  <- r_t - r_b
  T_y  <- nrow(both) / 252
  mu_s <- mean(r_s) * 252;  sig_s <- sd(r_s) * sqrt(252)
  ir   <- if (sig_s > 0) mu_s / sig_s else NA_real_
  mdd  <- max_dd_vec(r_s)
  cal  <- if (abs(mdd) > 0) mu_s / abs(mdd) else NA_real_
  rho  <- cor(r_t, r_b, use = "complete.obs")
  tibble(ticker, benchmark, T_yrs = T_y,
         mu_spr = mu_s, sig_spr = sig_s, ir, mdd_spr = mdd,
         calmar = cal, rho,
         te_10pct = 0.10 * sig_s,
         mdd_theory = magdon_mdd(sig_s, mu_s, T_y))
}

# ── Category definitions (top-10 shortlists from scoring) ─────────────────────
top10_a_tickers <- c("SMH", "XLK", "QQQ", "AIQ", "DXJ",
                     "EWT", "EWC", "XLY", "ITA", "CIBR")

top10_b_tickers <- c("GLD", "SLV", "PDBC", "DJP", "URA",
                     "COPX", "DBA", "IGF", "MOO", "LIT")

# Score tables for summary sections
scores_a <- map_dfr(top10_a_tickers, ~spread_raw(.x, "SPY")) %>%
  mutate(ir_s = minmax(ir), cal_s = minmax(calmar), rho_s = minmax(rho),
         score = 0.50 * ir_s + 0.30 * cal_s + 0.20 * rho_s) %>%
  arrange(desc(score))

scores_b <- map_dfr(top10_b_tickers, ~spread_raw(.x, "AGG")) %>%
  mutate(ir_s = minmax(ir), cal_s = minmax(calmar), rho_s = minmax(-rho),
         score = 0.50 * ir_s + 0.30 * cal_s + 0.20 * rho_s) %>%
  arrange(desc(score))

REGIME_COLORS <- c(Fall = "#D90429", Recovery = "#F77F00",
                   Consolidation = "#2D6A4F")
```

---

## Scoring Framework

Every ticker is scored against its category benchmark (SPY for A, AGG for B)
using the spread between the ticker and the benchmark:

$$\text{Score} = 0.50 \times \widehat{IR} + 0.30 \times \widehat{\text{Calmar}} + 0.20 \times \widehat{\rho}_{\text{adj}}$$

where $\hat{\cdot}$ denotes min-max normalisation within the category, and:

- **IR** = $E[\text{spread}] / \sigma[\text{spread}]$ — quality of the active bet
- **Calmar** = $E[\text{spread}] / |\text{MaxDD}_{\text{spread}}|$ — drawdown efficiency
- **ρ_adj**: for A, reward **high** ρ (lower TE per notional);
  for B, reward **low / negative** ρ (stronger cross-asset diversification)

```{r score-tables}
bind_rows(
  scores_a %>% mutate(Category = "A — vs SPY"),
  scores_b %>% mutate(Category = "B — vs AGG")
) %>%
  transmute(
    Category,
    Ticker   = ticker,
    `E[spr]` = percent(mu_spr,  accuracy = 0.1),
    `σ(spr)` = percent(sig_spr, accuracy = 0.1),
    IR       = round(ir,      2),
    Calmar   = round(calmar,  2),
    `ρ`      = round(rho,     2),
    MaxDD    = percent(mdd_spr, accuracy = 0.1),
    `TE(10%)`= percent(te_10pct, accuracy = 0.01),
    Score    = round(score,   3)
  ) %>%
  knitr::kable(caption = "Scoring summary — top 10 per category")
```

---

## Category A — Equity Swaps (vs SPY) {.tabset}

> **Decision:** +10% ticker / −10% SPY — total equity unchanged, sector/style tilt added.
> High ρ with SPY → low spread vol → small TE per notional.

```{r run-cat-a, cache=TRUE, include=FALSE}
results_a <- map(top10_a_tickers, function(tk) {
  out <- plot_xts_pair(
    xts_ret[, tk], xts_ret[, "SPY"],
    label1 = tk, label2 = "SPY",
    rt     = rt
  )
  list(ticker = tk, out = out, ps = pair_stats(out))
})
names(results_a) <- top10_a_tickers
```

### A1 · SMH — Semiconductors

```{r a-smh, fig.height=35}
print(results_a[["SMH"]]$out$combined)
```

### A2 · XLK — Technology

```{r a-xlk, fig.height=35}
print(results_a[["XLK"]]$out$combined)
```

### A3 · QQQ — Nasdaq 100

```{r a-qqq, fig.height=35}
print(results_a[["QQQ"]]$out$combined)
```

### A4 · AIQ — AI & Big Data

```{r a-aiq, fig.height=35}
print(results_a[["AIQ"]]$out$combined)
```

### A5 · DXJ — Japan Hedged Equity

```{r a-dxj, fig.height=35}
print(results_a[["DXJ"]]$out$combined)
```

### A6 · EWT — Taiwan

```{r a-ewt, fig.height=35}
print(results_a[["EWT"]]$out$combined)
```

### A7 · EWC — Canada

```{r a-ewc, fig.height=35}
print(results_a[["EWC"]]$out$combined)
```

### A8 · XLY — Consumer Discretionary

```{r a-xly, fig.height=35}
print(results_a[["XLY"]]$out$combined)
```

### A9 · ITA — Aerospace & Defense

```{r a-ita, fig.height=35}
print(results_a[["ITA"]]$out$combined)
```

### A10 · CIBR — Cybersecurity

```{r a-cibr, fig.height=35}
print(results_a[["CIBR"]]$out$combined)
```

## {-}

### Category A — Summary Comparison

```{r cat-a-summary-table}
map_dfr(results_a, function(r) {
  ps <- r$ps
  sc <- scores_a %>% filter(ticker == r$ticker)
  tibble(
    Ticker         = r$ticker,
    `E[spread]`    = percent(ps$mu,        accuracy = 0.1),
    `σ(spread)`    = percent(ps$sig,       accuracy = 0.1),
    IR             = round(ps$ir,          2),
    Calmar         = round(ps$mu / abs(ps$mdd), 2),
    `ρ(SPY)`       = round(ps$rho,         2),
    `MaxDD spread` = percent(ps$mdd,       accuracy = 0.1),
    `TE(10%)`      = percent(ps$te_10pct,  accuracy = 0.02),
    `Active ret`   = percent(ps$ret_10pct, accuracy = 0.02),
    `MI MaxDD`     = percent(ps$mdd_theory,accuracy = 0.1),
    Score          = round(sc$score,       3)
  )
}) %>%
  arrange(desc(IR)) %>%
  knitr::kable(caption = "Category A — all ten tickers ranked by IR")
```

```{r cat-a-ir-bar, fig.height=5}
map_dfr(results_a, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, ir = ps$ir, rho = ps$rho,
         mu = ps$mu, sig = ps$sig)
}) %>%
  mutate(
    p_correct = 0.5 + ir / 2,
    colour    = if_else(ir > 0, "#2D6A4F", "#D90429"),
    ticker    = fct_reorder(ticker, ir)
  ) %>%
  ggplot(aes(x = ticker, y = ir, fill = colour)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = 0.50, linetype = "dashed",
             colour = "#2D6A4F", linewidth = 0.8) +
  geom_hline(yintercept = 0,    colour = "grey30") +
  annotate("text", x = 0.6, y = 0.52, hjust = 0, size = 3,
           colour = "#2D6A4F", label = "IR = 0.50 (top quartile)") +
  geom_text(aes(label = sprintf("P=%.0f%%", pmax(p_correct, 1 - p_correct) * 100),
                hjust = if_else(ir >= 0, -0.15, 1.15)),
            size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_identity() +
  scale_y_continuous(expand = expansion(mult = c(0.15, 0.2))) +
  labs(title    = "Category A — Information Ratio vs SPY",
       subtitle = "P label = probability of a correct daily directional call  |  Dashed = top-quartile IR",
       x = NULL, y = "IR") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

```{r cat-a-spread-vol, fig.height=5}
map_dfr(results_a, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, sig_spr = ps$sig, te = ps$te_10pct,
         ir = ps$ir, rho = ps$rho)
}) %>%
  mutate(ticker = fct_reorder(ticker, sig_spr)) %>%
  ggplot(aes(x = ticker, y = sig_spr)) +
  geom_col(aes(fill = rho), width = 0.65, alpha = 0.85) +
  geom_text(aes(label = sprintf("ρ=%.2f", rho)),
            hjust = -0.15, size = 3.2) +
  coord_flip() +
  scale_fill_gradient2(low = "#D90429", mid = "grey80", high = "#2D6A4F",
                       midpoint = 0.80,
                       name = "ρ(SPY)",
                       breaks = c(0.5, 0.7, 0.9)) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Category A — Spread Volatility and ρ(SPY)",
       subtitle = "Higher ρ (greener) = lower TE per 10% notional\nTE added = 10% × spread vol",
       x = NULL, y = "Annualised Spread Vol") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "right",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

```{r cat-a-mdd-theory, fig.height=5}
map_dfr(results_a, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, actual = abs(ps$mdd), theory = abs(ps$mdd_theory),
         ir = ps$ir)
}) %>%
  mutate(ticker = fct_reorder(ticker, actual)) %>%
  pivot_longer(c(actual, theory), names_to = "type", values_to = "mdd") %>%
  mutate(type = if_else(type == "actual", "Realised MaxDD", "Magdon-Ismail E[MaxDD]")) %>%
  ggplot(aes(x = ticker, y = mdd, fill = type)) +
  geom_col(position = "dodge", width = 0.65, alpha = 0.85) +
  coord_flip() +
  scale_fill_manual(values = c("Realised MaxDD" = "#D90429",
                                "Magdon-Ismail E[MaxDD]" = "#457B9D"),
                    name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title    = "Category A — Spread MaxDD: Realised vs Magdon-Ismail Theory",
       subtitle = "Realised > theory signals a path unfavourable to the spread",
       x = NULL, y = "MaxDD (absolute)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "bottom",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

---

## Category B — Cross-Asset Additions (vs AGG) {.tabset}

> **Decision:** +10% ticker / −10% AGG — equity exposure rises from 60% to 70%,
> funded by reducing bonds. Negative or low ρ with AGG → diversification benefit
> but large spread vol.

```{r run-cat-b, cache=TRUE, include=FALSE}
results_b <- map(top10_b_tickers, function(tk) {
  out <- plot_xts_pair(
    xts_ret[, tk], xts_ret[, "AGG"],
    label1 = tk, label2 = "AGG",
    rt     = rt
  )
  list(ticker = tk, out = out, ps = pair_stats(out))
})
names(results_b) <- top10_b_tickers
```

### B1 · GLD — Gold

```{r b-gld, fig.height=35}
print(results_b[["GLD"]]$out$combined)
```

### B2 · SLV — Silver

```{r b-slv, fig.height=35}
print(results_b[["SLV"]]$out$combined)
```

### B3 · PDBC — Broad Commodities (Optimum Yield)

```{r b-pdbc, fig.height=35}
print(results_b[["PDBC"]]$out$combined)
```

### B4 · DJP — Bloomberg Commodity Index

```{r b-djp, fig.height=35}
print(results_b[["DJP"]]$out$combined)
```

### B5 · URA — Uranium

```{r b-ura, fig.height=35}
print(results_b[["URA"]]$out$combined)
```

### B6 · COPX — Copper Miners

```{r b-copx, fig.height=35}
print(results_b[["COPX"]]$out$combined)
```

### B7 · DBA — Agriculture Futures

```{r b-dba, fig.height=35}
print(results_b[["DBA"]]$out$combined)
```

### B8 · IGF — Global Infrastructure

```{r b-igf, fig.height=35}
print(results_b[["IGF"]]$out$combined)
```

### B9 · MOO — Agribusiness

```{r b-moo, fig.height=35}
print(results_b[["MOO"]]$out$combined)
```

### B10 · LIT — Lithium & Battery Tech

```{r b-lit, fig.height=35}
print(results_b[["LIT"]]$out$combined)
```

## {-}

### Category B — Summary Comparison

```{r cat-b-summary-table}
map_dfr(results_b, function(r) {
  ps <- r$ps
  sc <- scores_b %>% filter(ticker == r$ticker)
  tibble(
    Ticker         = r$ticker,
    `E[spread]`    = percent(ps$mu,        accuracy = 0.1),
    `σ(spread)`    = percent(ps$sig,       accuracy = 0.1),
    IR             = round(ps$ir,          2),
    Calmar         = round(ps$mu / abs(ps$mdd), 2),
    `ρ(AGG)`       = round(ps$rho,         2),
    `MaxDD spread` = percent(ps$mdd,       accuracy = 0.1),
    `TE(10%)`      = percent(ps$te_10pct,  accuracy = 0.02),
    `Active ret`   = percent(ps$ret_10pct, accuracy = 0.02),
    `MI MaxDD`     = percent(ps$mdd_theory,accuracy = 0.1),
    Score          = round(sc$score,       3)
  )
}) %>%
  arrange(desc(IR)) %>%
  knitr::kable(caption = "Category B — all ten tickers ranked by IR")
```

```{r cat-b-ir-bar, fig.height=5}
map_dfr(results_b, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, ir = ps$ir, rho = ps$rho)
}) %>%
  mutate(
    p_correct = 0.5 + ir / 2,
    colour    = if_else(ir > 0, "#2D6A4F", "#D90429"),
    ticker    = fct_reorder(ticker, ir)
  ) %>%
  ggplot(aes(x = ticker, y = ir, fill = colour)) +
  geom_col(width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = 0.50, linetype = "dashed",
             colour = "#2D6A4F", linewidth = 0.8) +
  geom_hline(yintercept = 0,    colour = "grey30") +
  annotate("text", x = 0.6, y = 0.52, hjust = 0, size = 3,
           colour = "#2D6A4F", label = "IR = 0.50 (top quartile)") +
  geom_text(aes(label = sprintf("P=%.0f%%", pmax(p_correct, 1 - p_correct) * 100),
                hjust = if_else(ir >= 0, -0.15, 1.15)),
            size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_identity() +
  scale_y_continuous(expand = expansion(mult = c(0.15, 0.2))) +
  labs(title    = "Category B — Information Ratio vs AGG",
       subtitle = "P label = probability of a correct daily directional call  |  Dashed = top-quartile IR",
       x = NULL, y = "IR") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

```{r cat-b-spread-vol, fig.height=5}
map_dfr(results_b, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, sig_spr = ps$sig, rho = ps$rho, ir = ps$ir)
}) %>%
  mutate(ticker = fct_reorder(ticker, sig_spr)) %>%
  ggplot(aes(x = ticker, y = sig_spr)) +
  geom_col(aes(fill = rho), width = 0.65, alpha = 0.85) +
  geom_text(aes(label = sprintf("ρ=%.2f", rho)),
            hjust = -0.15, size = 3.2) +
  coord_flip() +
  scale_fill_gradient2(low = "#2D6A4F", mid = "grey80", high = "#D90429",
                       midpoint = 0,
                       name = "ρ(AGG)") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Category B — Spread Volatility and ρ(AGG)",
       subtitle = "Lower / more-negative ρ (greener) = stronger diversification from bonds\nTE added = 10% × spread vol",
       x = NULL, y = "Annualised Spread Vol") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "right",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

```{r cat-b-mdd-theory, fig.height=5}
map_dfr(results_b, function(r) {
  ps <- r$ps
  tibble(ticker = r$ticker, actual = abs(ps$mdd), theory = abs(ps$mdd_theory),
         ir = ps$ir)
}) %>%
  mutate(ticker = fct_reorder(ticker, actual)) %>%
  pivot_longer(c(actual, theory), names_to = "type", values_to = "mdd") %>%
  mutate(type = if_else(type == "actual", "Realised MaxDD", "Magdon-Ismail E[MaxDD]")) %>%
  ggplot(aes(x = ticker, y = mdd, fill = type)) +
  geom_col(position = "dodge", width = 0.65, alpha = 0.85) +
  coord_flip() +
  scale_fill_manual(values = c("Realised MaxDD" = "#D90429",
                                "Magdon-Ismail E[MaxDD]" = "#457B9D"),
                    name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title    = "Category B — Spread MaxDD: Realised vs Magdon-Ismail Theory",
       subtitle = "Realised > theory signals a path unfavourable to the spread",
       x = NULL, y = "MaxDD (absolute)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position    = "bottom",
        plot.title         = element_text(face = "bold", size = 13),
        plot.subtitle      = element_text(colour = "grey50", size = 9))
```

---

## Cross-Category Comparison

All 20 tickers on a common footing — spread IR vs spread vol,
coloured by category.

```{r cross-ir-vol, fig.height=6}
bind_rows(
  map_dfr(results_a, function(r) {
    ps <- r$ps
    tibble(ticker = r$ticker, category = "A — vs SPY",
           ir = ps$ir, sig = ps$sig, rho = ps$rho, mu = ps$mu)
  }),
  map_dfr(results_b, function(r) {
    ps <- r$ps
    tibble(ticker = r$ticker, category = "B — vs AGG",
           ir = ps$ir, sig = ps$sig, rho = ps$rho, mu = ps$mu)
  })
) %>%
  ggplot(aes(x = sig, y = ir, colour = category, label = ticker)) +
  geom_hline(yintercept = c(0, 0.50), linetype = c("solid","dashed"),
             colour = c("grey70","#2D6A4F"), linewidth = 0.7) +
  geom_point(aes(size = abs(mu)), alpha = 0.8) +
  ggrepel::geom_text_repel(size = 3.2, show.legend = FALSE) +
  scale_colour_manual(values = c("A — vs SPY" = "#457B9D",
                                  "B — vs AGG" = "#D90429"),
                      name = NULL) +
  scale_size_continuous(range = c(3, 9), name = "|E[spread]|") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(title    = "Spread IR vs Spread Vol — All 20 Shortlisted Tickers",
       subtitle = "Point size = |E[spread]|  |  Dashed = IR 0.50 top-quartile  |  Blue = equity swap, Red = cross-asset",
       x = "Annualised Spread Vol", y = "Information Ratio") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        plot.title       = element_text(face = "bold", size = 13),
        plot.subtitle    = element_text(colour = "grey50", size = 9))
```

```{r cross-master-table}
bind_rows(
  map_dfr(results_a, function(r) {
    ps <- r$ps; sc <- scores_a %>% filter(ticker == r$ticker)
    tibble(Category = "A — vs SPY", Ticker = r$ticker,
           `E[spr]` = percent(ps$mu,  accuracy = 0.1),
           `σ(spr)` = percent(ps$sig, accuracy = 0.1),
           IR       = round(ps$ir, 2),
           Calmar   = round(ps$mu / abs(ps$mdd), 2),
           ρ        = round(ps$rho, 2),
           MaxDD    = percent(ps$mdd, accuracy = 0.1),
           `TE(10%)`= percent(ps$te_10pct, accuracy = 0.02),
           Score    = round(sc$score, 3))
  }),
  map_dfr(results_b, function(r) {
    ps <- r$ps; sc <- scores_b %>% filter(ticker == r$ticker)
    tibble(Category = "B — vs AGG", Ticker = r$ticker,
           `E[spr]` = percent(ps$mu,  accuracy = 0.1),
           `σ(spr)` = percent(ps$sig, accuracy = 0.1),
           IR       = round(ps$ir, 2),
           Calmar   = round(ps$mu / abs(ps$mdd), 2),
           ρ        = round(ps$rho, 2),
           MaxDD    = percent(ps$mdd, accuracy = 0.1),
           `TE(10%)`= percent(ps$te_10pct, accuracy = 0.02),
           Score    = round(sc$score, 3))
  })
) %>%
  arrange(Category, desc(IR)) %>%
  knitr::kable(caption = "Master table — all 20 shortlisted tickers")
```

---

*Benchmark: SPY for Category A (equity swap), AGG for Category B (cross-asset addition).
All decisions at 10% notional. Data: `xts_ret_returns.rds` | SPY regimes: Fall ≥10% drawdown.*
