---
title: "Universe Pair Analysis — Cost–Benefit Vectors"
subtitle: "Category A (vs SPY) · Category B (vs AGG) · 10% Notional"
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 13,
  fig.height = 7
)

library(MASS)
library(tidyverse)
library(xts)
library(scales)
library(patchwork)
library(ggrepel)
library(here)

select <- dplyr::select

if (!exists("xts_ret"))
  xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

if (!exists("build_regime_table"))
  source(here("scripts_spy_dd_regime/spy_dd_regime.R"))

spy_ret <- xts_ret[, "SPY"]
rt      <- build_regime_table(spy_ret, 0.10, 0.05)

source(here("utility/plot_xts_pair.R"))

# ── Helpers ───────────────────────────────────────────────────────────────────
T_yrs <- as.numeric(diff(range(index(xts_ret)))) / 365

max_dd_vec <- function(r) {
  w  <- cumprod(1 + r)
  min((w - cummax(w)) / cummax(w), na.rm = TRUE)
}

magdon_mdd <- function(sigma, mu, T) {
  if (is.na(mu) || mu <= 0) return(sigma * sqrt(T))
  S <- mu / sigma
  (sigma / S) * (log(S * sqrt(T)) + 0.35)
}

pair_stats <- function(out) {
  s  <- out$stats
  sp <- s %>% filter(str_detect(Series, "\u2212"))
  mu  <- parse_number(sp$`Ann Return`) / 100
  sig <- parse_number(sp$`Ann Vol`)    / 100
  mdd <- parse_number(sp$`Max DD`)     / 100
  list(mu = mu, sig = sig, mdd = mdd,
       ir  = out$full_ir, rho = out$full_cor,
       te_10pct  = 0.10 * sig,
       mdd_10pct = 0.10 * abs(mdd),
       ret_10pct = 0.10 * mu,
       mdd_theory = magdon_mdd(sig, mu, T_yrs),
       stats = s)
}

# ── Shortlists ────────────────────────────────────────────────────────────────
top10_a <- c("SMH", "XLK", "QQQ", "AIQ", "DXJ",
             "EWT", "EWC", "XLY", "ITA", "CIBR")

top10_b <- c("GLD", "SLV", "PDBC", "DJP", "URA",
             "COPX", "DBA", "IGF", "MOO", "LIT")
```

```{r run-pairs, cache=TRUE, include=FALSE}
results_a <- set_names(
  map(top10_a, ~{
    out <- plot_xts_pair(xts_ret[, .x], xts_ret[, "SPY"],
                         label1 = .x, label2 = "SPY", rt = rt)
    list(ticker = .x, category = "A \u2014 vs SPY", out = out, ps = pair_stats(out))
  }),
  top10_a
)

results_b <- set_names(
  map(top10_b, ~{
    out <- plot_xts_pair(xts_ret[, .x], xts_ret[, "AGG"],
                         label1 = .x, label2 = "AGG", rt = rt)
    list(ticker = .x, category = "B \u2014 vs AGG", out = out, ps = pair_stats(out))
  }),
  top10_b
)
```

```{r vec-data, include=FALSE}
# Build vector tibble: one row per ticker
vec_df <- bind_rows(
  map_dfr(results_a, function(r) {
    ps <- r$ps
    tibble(ticker   = r$ticker,
           category = r$category,
           te       = ps$te_10pct,        # TE cost  at 10%
           ret      = ps$ret_10pct,       # active return at 10%
           mdd_cost = 0.10 * abs(ps$mdd), # MaxDD cost at 10%
           ir       = ps$ir,
           calmar   = ps$mu / abs(ps$mdd),
           rho      = ps$rho,
           sig      = ps$sig,
           mu       = ps$mu)
  }),
  map_dfr(results_b, function(r) {
    ps <- r$ps
    tibble(ticker   = r$ticker,
           category = r$category,
           te       = ps$te_10pct,
           ret      = ps$ret_10pct,
           mdd_cost = 0.10 * abs(ps$mdd),
           ir       = ps$ir,
           calmar   = ps$mu / abs(ps$mdd),
           rho      = ps$rho,
           sig      = ps$sig,
           mu       = ps$mu)
  })
) %>%
  mutate(
    angle_ir     = atan2(ret, te)       * 180 / pi,
    angle_calmar = atan2(ret, mdd_cost) * 180 / pi,
    magnitude_te = sqrt(te^2 + ret^2),
    magnitude_mdd= sqrt(mdd_cost^2 + ret^2),
    zone = case_when(
      ir >= 1.0  ~ "Exceptional  (IR \u2265 1.0)",
      ir >= 0.5  ~ "Top quartile  (IR \u2265 0.5)",
      ir >= 0.0  ~ "Positive but below hurdle",
      TRUE       ~ "Destroys value  (IR < 0)"
    )
  )

CAT_COLS <- c("A \u2014 vs SPY" = "#457B9D", "B \u2014 vs AGG" = "#D90429")

# Convenience filtered subsets
vec_a <- vec_df %>% filter(str_detect(category, "SPY"))
vec_b <- vec_df %>% filter(str_detect(category, "AGG"))

# Helper: position reference-line label at x = pct * max(x_var)
ref_label <- function(df, x_col, slope, pct = 0.75,
                      prefix = "IR") {
  x_at <- max(df[[x_col]], na.rm = TRUE) * pct
  tibble(x     = x_at,
         y     = slope * x_at,
         label = sprintf("%s = %.1f", prefix, slope))
}
```

```{r plot-fns, include=FALSE}
# ── Reusable plot functions — called with full or filtered vec_df ──────────────

plot_vec_te <- function(df, subtitle_extra = NULL) {
  x_max <- max(df$te)  * 1.22
  y_max <- max(max(df$ret) * 1.30, 0.001)
  y_min <- min(min(df$ret) * 1.45, -0.001)
  ref   <- bind_rows(ref_label(df, "te", 0.5, 0.78),
                     ref_label(df, "te", 1.0, 0.58))
  sub <- paste0("Arrow slope = IR  |  Steeper = more efficient  |  Below zero = destroys value",
                if (!is.null(subtitle_extra)) paste0("  |  ", subtitle_extra) else "")
  ggplot(df) +
    annotate("rect", xmin = 0, xmax = x_max, ymin = 0, ymax = y_max * 1.2,
             fill = "#2D6A4F", alpha = 0.04) +
    geom_abline(slope = 0.5, intercept = 0, linetype = "dashed",
                colour = "#2D6A4F", linewidth = 0.9) +
    geom_abline(slope = 1.0, intercept = 0, linetype = "dashed",
                colour = "#457B9D", linewidth = 0.9) +
    geom_hline(yintercept = 0, colour = "grey45", linewidth = 0.5) +
    geom_segment(aes(x = 0, y = 0, xend = te, yend = ret, colour = category),
                 arrow = arrow(length = unit(0.24, "cm"), type = "open"),
                 linewidth = 1.05, alpha = 0.88) +
    geom_point(aes(x = te, y = ret, colour = category), size = 3) +
    geom_text_repel(aes(x = te, y = ret, label = ticker),
                    size = 3.3, box.padding = 0.45,
                    min.segment.length = 0.3, segment.colour = "grey65") +
    geom_label(data = ref,
               aes(x = x, y = y, label = label,
                   colour = c("#2D6A4F", "#457B9D")),
               fill = "white", size = 3.1,
               label.padding = unit(0.18, "lines"), label.size = 0,
               show.legend = FALSE) +
    annotate("text", x = x_max * 0.01, y = y_max * 1.07,
             label = "active return > TE cost  (IR > 1.0)",
             hjust = 0, size = 3, colour = "#457B9D", fontface = "italic") +
    annotate("text", x = x_max * 0.01, y = y_max * 0.55,
             label = "top-quartile zone  (IR > 0.5)",
             hjust = 0, size = 3, colour = "#2D6A4F", fontface = "italic") +
    scale_colour_manual(values = CAT_COLS, name = NULL) +
    scale_x_continuous(labels = percent_format(accuracy = 0.01),
                       limits = c(0, x_max)) +
    scale_y_continuous(labels = percent_format(accuracy = 0.01),
                       limits = c(y_min, y_max * 1.18)) +
    labs(title    = "Panel 1 — TE Cost vs Active Return  (10% notional)",
         subtitle = sub,
         x = "Tracking Error added  =  10% \u00d7 \u03c3(spread)",
         y = "Active return  =  10% \u00d7 E[spread]") +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "bottom",
          plot.title    = element_text(face = "bold", size = 13),
          plot.subtitle = element_text(colour = "grey50", size = 9))
}

plot_vec_mdd <- function(df, subtitle_extra = NULL) {
  x_max <- max(df$mdd_cost) * 1.22
  y_max <- max(max(df$ret)  * 1.30, 0.001)
  y_min <- min(min(df$ret)  * 1.45, -0.001)
  ref   <- bind_rows(ref_label(df, "mdd_cost", 0.5, 0.78, "Calmar"),
                     ref_label(df, "mdd_cost", 1.0, 0.58, "Calmar"))
  sub <- paste0("Arrow slope = Calmar  |  X-axis = worst-case portfolio pain at 10% notional",
                if (!is.null(subtitle_extra)) paste0("  |  ", subtitle_extra) else "")
  ggplot(df) +
    annotate("rect", xmin = 0, xmax = x_max, ymin = 0, ymax = y_max * 1.2,
             fill = "#2D6A4F", alpha = 0.04) +
    geom_abline(slope = 0.5, intercept = 0, linetype = "dashed",
                colour = "#2D6A4F", linewidth = 0.9) +
    geom_abline(slope = 1.0, intercept = 0, linetype = "dashed",
                colour = "#457B9D", linewidth = 0.9) +
    geom_hline(yintercept = 0, colour = "grey45", linewidth = 0.5) +
    geom_segment(aes(x = 0, y = 0, xend = mdd_cost, yend = ret, colour = category),
                 arrow = arrow(length = unit(0.24, "cm"), type = "open"),
                 linewidth = 1.05, alpha = 0.88) +
    geom_point(aes(x = mdd_cost, y = ret, colour = category), size = 3) +
    geom_text_repel(aes(x = mdd_cost, y = ret, label = ticker),
                    size = 3.3, box.padding = 0.45,
                    min.segment.length = 0.3, segment.colour = "grey65") +
    geom_label(data = ref,
               aes(x = x, y = y, label = label,
                   colour = c("#2D6A4F", "#457B9D")),
               fill = "white", size = 3.1,
               label.padding = unit(0.18, "lines"), label.size = 0,
               show.legend = FALSE) +
    annotate("text", x = x_max * 0.01, y = y_max * 1.07,
             label = "active return > MaxDD cost  (Calmar > 1.0)",
             hjust = 0, size = 3, colour = "#457B9D", fontface = "italic") +
    annotate("text", x = x_max * 0.01, y = y_max * 0.55,
             label = "top-quartile zone  (Calmar > 0.5)",
             hjust = 0, size = 3, colour = "#2D6A4F", fontface = "italic") +
    scale_colour_manual(values = CAT_COLS, name = NULL) +
    scale_x_continuous(labels = percent_format(accuracy = 0.1),
                       limits = c(0, x_max)) +
    scale_y_continuous(labels = percent_format(accuracy = 0.01),
                       limits = c(y_min, y_max * 1.18)) +
    labs(title    = "Panel 2 — MaxDD Cost vs Active Return  (10% notional)",
         subtitle = sub,
         x = "MaxDD cost  =  10% \u00d7 |MaxDD(spread)|",
         y = "Active return  =  10% \u00d7 E[spread]") +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "bottom",
          plot.title    = element_text(face = "bold", size = 13),
          plot.subtitle = element_text(colour = "grey50", size = 9))
}

plot_vec_compass <- function(df, subtitle_extra = NULL) {
  angle_05 <- atan(0.5) * 180 / pi
  angle_10 <- atan(1.0) * 180 / pi
  r_max    <- max(df$magnitude_te) * 1.20
  sub <- paste0(
    "Horizontal = angle arctan(IR)  |  Vertical = vector magnitude  |  Green zone = IR > 0.5",
    if (!is.null(subtitle_extra)) paste0("  |  ", subtitle_extra) else ""
  )
  ggplot(df, aes(x = angle_ir, y = magnitude_te,
                 colour = category, label = ticker)) +
    annotate("rect", xmin = angle_05, xmax = 95,
             ymin = -Inf, ymax = Inf, fill = "#2D6A4F", alpha = 0.04) +
    geom_vline(xintercept = 0,        colour = "grey50",   linewidth = 0.5) +
    geom_vline(xintercept = angle_05, linetype = "dashed",
               colour = "#2D6A4F", linewidth = 0.9) +
    geom_vline(xintercept = angle_10, linetype = "dashed",
               colour = "#457B9D", linewidth = 0.9) +
    geom_hline(yintercept = 0, colour = "grey80") +
    geom_point(size = 3.5, alpha = 0.88) +
    geom_text_repel(size = 3.3, box.padding = 0.45,
                    segment.colour = "grey65", show.legend = FALSE) +
    annotate("text", x = angle_05 + 0.8, y = r_max * 0.97,
             hjust = 0, size = 3, colour = "#2D6A4F", label = "IR = 0.50") +
    annotate("text", x = angle_10 + 0.8, y = r_max * 0.97,
             hjust = 0, size = 3, colour = "#457B9D", label = "IR = 1.00") +
    annotate("text", x = -0.8, y = r_max * 0.97,
             hjust = 1, size = 3, colour = "grey50",
             label = "IR = 0\n(breakeven)") +
    scale_colour_manual(values = CAT_COLS, name = NULL) +
    scale_x_continuous(
      breaks = c(-45, 0, angle_05, angle_10, 45, 90),
      labels = c("-45\u00b0\n(IR=-1)", "0\u00b0\n(IR=0)",
                 sprintf("%.1f\u00b0\n(IR=0.5)", angle_05),
                 sprintf("%.0f\u00b0\n(IR=1.0)", angle_10),
                 "45\u00b0", "90\u00b0\n(IR=+inf)"),
      limits = c(-55, 95)
    ) +
    scale_y_continuous(labels = percent_format(accuracy = 0.01),
                       limits = c(0, r_max)) +
    labs(title    = "Panel 3 — Efficiency Compass  (TE-based vectors)",
         subtitle = sub,
         x = "Vector angle  (IR direction)",
         y = "Vector magnitude  =  sqrt(TE^2 + ActiveRet^2)") +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "bottom",
          plot.title    = element_text(face = "bold", size = 13),
          plot.subtitle = element_text(colour = "grey50", size = 9))
}
```

---

## The Vector Framework

Every active allocation decision at 10% notional creates a 2-D vector:

$$\vec{v} = \bigl(\underbrace{x}_{\text{cost at 10\%}},\;\underbrace{y}_{\text{active return at 10\%}}\bigr)$$

| Component | Panel 1 | Panel 2 |
|---|---|---|
| **x (cost)** | Tracking Error added = 10% × σ(spread) | MaxDD added = 10% × \|MaxDD(spread)\| |
| **y (benefit)** | Active return = 10% × E[spread] | Active return = 10% × E[spread] |
| **Slope = ratio** | **IR** = E[spread] / σ(spread) | **Calmar** = E[spread] / \|MaxDD\| |

The **steeper** the arrow, the better the risk-adjusted payoff.
The dashed reference lines at slope 0.5 and 1.0 are the key hurdles.

---

## Panel 1 — Tracking Error Cost vs Active Return {.tabset}

*Slope of arrow = Information Ratio*

### All

```{r vec-te-all, fig.height=8}
plot_vec_te(vec_df)
```

### Category A — vs SPY

```{r vec-te-a, fig.height=8}
plot_vec_te(vec_a, "Category A: equity swaps only")
```

### Category B — vs AGG

```{r vec-te-b, fig.height=8}
plot_vec_te(vec_b, "Category B: cross-asset additions only")
```

## {-}

---

## Panel 2 — MaxDD Cost vs Active Return {.tabset}

*Slope of arrow = Calmar ratio*

### All

```{r vec-mdd-all, fig.height=8}
plot_vec_mdd(vec_df)
```

### Category A — vs SPY

```{r vec-mdd-a, fig.height=8}
plot_vec_mdd(vec_a, "Category A: equity swaps only")
```

### Category B — vs AGG

```{r vec-mdd-b, fig.height=8}
plot_vec_mdd(vec_b, "Category B: cross-asset additions only")
```

## {-}

---

## Panel 3 — Efficiency Compass {.tabset}

Same vectors re-expressed in polar coordinates:
**angle** = IR quality (how steep the vector is),
**radius** = magnitude (how large the bet is).

### All

```{r vec-compass-all, fig.height=7}
plot_vec_compass(vec_df)
```

### Category A — vs SPY

```{r vec-compass-a, fig.height=7}
plot_vec_compass(vec_a, "Category A: equity swaps only")
```

### Category B — vs AGG

```{r vec-compass-b, fig.height=7}
plot_vec_compass(vec_b, "Category B: cross-asset additions only")
```

## {-}

---

## Summary Table

```{r vec-table}
vec_df %>%
  arrange(desc(ir)) %>%
  transmute(
    Category       = category,
    Ticker         = ticker,
    IR             = round(ir,     2),
    Calmar         = round(calmar, 2),
    `Angle (deg)`  = round(angle_ir, 1),
    `TE cost`      = percent(te,       accuracy = 0.01),
    `MDD cost`     = percent(mdd_cost, accuracy = 0.1),
    `Act. ret`     = percent(ret,      accuracy = 0.01),
    Zone           = zone
  ) %>%
  knitr::kable(caption = "All 20 tickers: IR angle, cost decomposition, and efficiency zone")
```

---

*Benchmark: SPY for Category A (equity swap), AGG for Category B (cross-asset).
All costs and returns scaled to 10% notional. Data: `xts_ret_returns.rds`.*
