---
title: "Long/Short Overlay Portfolio: Methodology & Results"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: show
date: "2026-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 11,
  fig.height = 7
)

library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(here)

# Load pre-computed relative returns (ETF daily return minus SPY daily return)
xts_rel <- readRDS(here("02_data_processed/xts_rel.rds"))
```

---

## 1. Overview

This document describes a **standalone Long/Short (L/S) Momentum Overlay** built on top of
a 59-ETF relative-return universe. The strategy is dollar-neutral: the long leg and short
leg each carry ±100% notional, so net market exposure is zero. All returns are expressed
as **relative returns vs SPY** (`xts_rel`), meaning:

- A long position earns when the ETF outperforms SPY
- A short position earns when the ETF underperforms SPY
- The portfolio P&L is **pure cross-sectional alpha** — market direction cancels out

The key result is that the strategy generates a **Sharpe ratio of ~1.50** on a
**restricted 15-ETF core universe** identified through PCA eigenvector analysis,
vs 1.62 on the full 59-ETF universe — confirming the signal is genuine and not
solely a product of breadth.

---

## 2. Data: Relative Returns (`xts_rel`)

`xts_rel` is produced by `scripts/01_etf_wrangle.R`. Each column is the arithmetic
daily return spread of one ETF vs SPY:

$$r^{rel}_{i,t} = r^{ETF}_{i,t} - r^{SPY}_t$$

The SPY column itself is identically zero. All other tickers represent a
**Long ETF / Short SPY** position. This is the raw material for the overlay.

```{r data-summary, echo=FALSE}
cat(sprintf("Universe : %d tickers\n", ncol(xts_rel)))
cat(sprintf("Period   : %s to %s\n",
            as.character(start(xts_rel)),
            as.character(end(xts_rel))))
cat(sprintf("Obs      : %d trading days\n", nrow(xts_rel)))
```

---

## 3. Portfolio Construction

### 3.1 Signal

At each monthly rebalance date, compute the **cumulative relative return** of every
ETF over the prior `signal_lookback` trading days (default: 63 days = 1 quarter):

$$\text{signal}_i = \prod_{t-L}^{t}(1 + r^{rel}_{i,t}) - 1$$

This is a cross-sectional momentum signal in relative-return space: ETFs that have
been beating SPY the most over the past quarter are expected to continue doing so.

### 3.2 Weight Assignment

Rank all ETFs by signal (descending). Assign:

- **Long**: top `n_long` ETFs → weight $w_i = +\frac{1}{n_{long}}$
- **Short**: bottom `n_short` ETFs → weight $w_i = -\frac{1}{n_{short}}$
- **Neutral**: all others → weight $0$

By construction: $\sum_i w_i = +1 - 1 = 0$ (dollar neutral).

### 3.3 Return Computation

Each day the portfolio return is the inner product of current weights and relative returns:

$$R^{port}_t = \sum_i w_i \cdot r^{rel}_{i,t}$$

> **Why not `Return.portfolio()`?**
> PerformanceAnalytics normalises by total portfolio value: $\frac{\sum w_i R_i}{\sum w_i}$.
> For a dollar-neutral book $\sum w_i = 0$, this produces $0/0$ — silently collapsed
> to zero for the entire history. Weights are instead broadcast to a daily matrix and
> returns computed directly via `rowSums(w_daily * xts_u)`.

### 3.4 Implementation

```{r build-ls-function}
build_ls_portfolio <- function(xts_rel,
                                n_long           = 5L,
                                n_short          = 5L,
                                signal_lookback  = 63L,
                                rebalance_freq   = "months",
                                min_hold_days    = 0L,
                                universe_tickers = NULL) {

  valid_freqs <- c("months", "quarters", "years")
  if (!(rebalance_freq %in% valid_freqs))
    stop("rebalance_freq must be one of: months, quarters, years")
  if (min_hold_days < 0L) stop("min_hold_days must be >= 0")
  if (n_long < 1 || n_short < 1) stop("n_long and n_short must be >= 1")

  # Universe
  if (!is.null(universe_tickers)) {
    universe_tickers <- intersect(universe_tickers, colnames(xts_rel))
  } else {
    universe_tickers <- setdiff(colnames(xts_rel), "SPY")
  }
  n_universe <- length(universe_tickers)
  xts_u      <- xts_rel[, universe_tickers]

  if (n_universe < n_long + n_short)
    stop(sprintf("Universe has %d tickers but n_long + n_short = %d.",
                 n_universe, n_long + n_short))

  # Rebalance endpoints
  ep <- endpoints(xts_u, on = rebalance_freq)
  ep <- ep[ep > 0]

  rebal_dates <- index(xts_u)[ep]
  w_mat <- matrix(NA_real_, nrow = length(ep), ncol = n_universe,
                  dimnames = list(as.character(rebal_dates), universe_tickers))

  signal_history <- list()
  last_rebal_idx <- NA_integer_

  for (k in seq_along(ep)) {
    ep_i <- ep[k]

    # Minimum holding period guard
    if (min_hold_days > 0L && !is.na(last_rebal_idx)) {
      if ((ep_i - last_rebal_idx) < min_hold_days) next
    }

    # Signal: cumulative return over lookback window
    lb_start <- max(1L, ep_i - signal_lookback + 1L)
    win      <- xts_u[lb_start:ep_i, ]
    cum_ret  <- vapply(universe_tickers, function(tk) {
      x <- as.numeric(win[, tk]); x <- x[!is.na(x)]
      if (length(x) == 0) return(NA_real_)
      prod(1 + x) - 1
    }, numeric(1))

    cum_valid <- cum_ret[!is.na(cum_ret)]
    if (length(cum_valid) < n_long + n_short) next

    ranked      <- sort(cum_valid, decreasing = TRUE)
    long_names  <- names(ranked)[seq_len(n_long)]
    short_names <- names(ranked)[seq(length(ranked) - n_short + 1L, length(ranked))]

    w_row <- setNames(rep(0, n_universe), universe_tickers)
    w_row[long_names]  <-  1 / n_long
    w_row[short_names] <- -1 / n_short
    w_mat[k, ]   <- w_row
    last_rebal_idx <- ep_i

    signal_history[[as.character(rebal_dates[k])]] <- data.frame(
      ticker = names(cum_valid), signal = as.numeric(cum_valid),
      rank   = rank(-cum_valid, ties.method = "first"),
      side   = ifelse(names(cum_valid) %in% long_names,  "Long",
               ifelse(names(cum_valid) %in% short_names, "Short", "Neutral")),
      stringsAsFactors = FALSE
    )
  }

  # Drop unassigned rows
  valid_rows  <- !is.na(w_mat[, 1])
  w_mat       <- w_mat[valid_rows, , drop = FALSE]
  rebal_dates <- rebal_dates[valid_rows]
  w_xts       <- xts(w_mat, order.by = rebal_dates)

  # Broadcast weights to daily and compute returns arithmetically
  all_dates <- index(xts_u)
  w_daily   <- matrix(NA_real_, nrow = length(all_dates), ncol = n_universe,
                      dimnames = list(NULL, universe_tickers))

  for (k in seq_len(nrow(w_mat))) {
    period_start <- rebal_dates[k]
    period_end   <- if (k < nrow(w_mat)) rebal_dates[k + 1L] else tail(all_dates, 1L)
    day_idx      <- which(all_dates >= period_start & all_dates <= period_end)
    if (length(day_idx) > 0)
      w_daily[day_idx, ] <- matrix(rep(w_mat[k, ], length(day_idx)),
                                   nrow = length(day_idx), byrow = TRUE)
  }

  w_daily_xts <- xts(w_daily, order.by = all_dates)
  first_idx   <- which(all_dates >= rebal_dates[1L])[1L]
  if (first_idx > 1L) w_daily_xts[seq_len(first_idx - 1L), ] <- NA

  port_ret_vec        <- rowSums(w_daily_xts * xts_u, na.rm = FALSE)
  port_ret_xts        <- xts(port_ret_vec, order.by = all_dates)
  port_ret_xts        <- port_ret_xts[!is.na(port_ret_xts)]
  colnames(port_ret_xts) <- "LS_Overlay"

  list(
    returns        = port_ret_xts,
    weights_bop    = w_daily_xts,
    weights_rebal  = w_xts,
    signal_history = signal_history
  )
}
```

---

## 4. Universe Selection: Eigenvector Analysis

### 4.1 Motivation

Running the overlay on all 59 ETFs inflates the Sharpe ratio through **breadth redundancy**:
many tickers load on the same underlying factor, so the portfolio is implicitly concentrating
in a handful of risk premia while appearing diversified. PCA on `xts_rel` reveals the true
factor structure.

### 4.2 PCA on the Relative-Return Correlation Matrix

```{r pca-analysis}
xts_u_full <- xts_rel[, setdiff(colnames(xts_rel), "SPY")]
xts_u_full <- xts_u_full[, apply(xts_u_full, 2, function(x) sd(x, na.rm = TRUE) > 0)]

pca <- prcomp(as.matrix(xts_u_full), scale. = TRUE, center = TRUE)

var_exp <- summary(pca)$importance[2, 1:8]
cum_var <- summary(pca)$importance[3, 1:8]
```

```{r pca-variance, echo=FALSE}
var_df <- data.frame(
  PC         = paste0("PC", 1:8),
  Prop_Var   = round(var_exp * 100, 1),
  Cumul_Var  = round(cum_var * 100, 1)
)

ggplot(var_df, aes(x = PC)) +
  geom_col(aes(y = Prop_Var), fill = "#2980b9", alpha = 0.8) +
  geom_line(aes(y = Cumul_Var, group = 1), color = "#e74c3c", linewidth = 1.2) +
  geom_point(aes(y = Cumul_Var), color = "#e74c3c", size = 3) +
  geom_text(aes(y = Prop_Var, label = paste0(Prop_Var, "%")),
            vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_y_continuous(
    name     = "Proportion of Variance (%)",
    sec.axis = sec_axis(~., name = "Cumulative Variance (%)")
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title    = "PCA on xts_rel: Variance Explained",
    subtitle = "Bars = individual PC | Red line = cumulative",
    x        = NULL
  )
```

### 4.3 Factor Interpretation

```{r pc-loadings, echo=FALSE}
pc_labels <- c(
  PC1 = "PC1 (35%): Bonds vs Equity — all FI ETFs load negatively (underperform SPY in bull markets)",
  PC2 = "PC2 (10%): International vs Domestic — developed & EM equity ETFs",
  PC3 = "PC3 (8%):  Growth vs Cyclical — QQQ/XLK short, XLF/XLI/XLE long"
)

for (pc in 1:3) {
  top10 <- sort(abs(pca$rotation[, pc]), decreasing = TRUE)[1:10]
  signed <- pca$rotation[names(top10), pc]
  df <- data.frame(
    ETF     = names(signed),
    Loading = round(signed, 3),
    direction = ifelse(signed > 0, "Positive", "Negative")
  ) %>% arrange(desc(abs(Loading)))

  p <- ggplot(df, aes(x = reorder(ETF, abs(Loading)), y = Loading, fill = direction)) +
    geom_col(width = 0.7) +
    coord_flip() +
    scale_fill_manual(values = c("Positive" = "#27ae60", "Negative" = "#e74c3c")) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none", panel.grid.minor = element_blank()) +
    labs(title = pc_labels[pc], x = NULL, y = "Loading")
  print(p)
}
```

### 4.4 Eigenvector Centrality

The leading eigenvector of the correlation matrix (which is identical to PC1 loadings)
measures how "connected" each ETF is to the rest of the network. High centrality = the
ETF moves with many others simultaneously.

```{r eigenvector-centrality, echo=FALSE}
cor_mat    <- cor(as.matrix(xts_u_full), use = "pairwise.complete.obs")
centrality <- abs(eigen(cor_mat)$vectors[, 1])
names(centrality) <- colnames(xts_u_full)
centrality <- sort(centrality, decreasing = TRUE)

data.frame(
  Rank       = 1:20,
  ETF        = names(centrality)[1:20],
  Centrality = round(centrality[1:20], 4)
) %>%
  ggplot(aes(x = reorder(ETF, Centrality), y = Centrality)) +
  geom_col(fill = "#2c3e50", alpha = 0.85) +
  geom_text(aes(label = round(Centrality, 3)), hjust = -0.1, size = 3.2) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title    = "Eigenvector Centrality: Top 20 ETFs",
    subtitle = "Most connected nodes in the xts_rel correlation network",
    x        = NULL, y        = "Centrality Score"
  )
```

### 4.5 Restricted Core Universe (15 ETFs)

After removing redundant pairs (AGG ≈ BND, SHY ≈ SGOV, IEFA ≈ ACWX ≈ URTH, dropping
composite balanced funds AOR/AOK), the **15-ETF core universe** covers all three
principal factors:

| Factor | ETFs |
|---|---|
| PC1 — Rates / Bonds | `AGG`, `IEF`, `TLT`, `HYG`, `TIP` |
| PC2 — International | `IEFA`, `FEZ`, `VWO` |
| PC3 — Equity Rotation | `QQQ`, `XLK`, `XLF`, `XLI`, `XLE` |
| Tail / Regime | `GLD`, `VIXY` |

---

## 5. Results

### 5.1 Build Both Portfolios

```{r build-portfolios}
# Full 59-ticker universe
ls_full <- build_ls_portfolio(
  xts_rel         = xts_rel,
  n_long          = 5L,
  n_short         = 5L,
  signal_lookback = 63L,
  rebalance_freq  = "months"
)

# Restricted 15-ETF core universe
core_universe <- c(
  "AGG", "IEF", "TLT", "HYG", "TIP",
  "IEFA", "FEZ", "VWO",
  "QQQ", "XLK", "XLF", "XLI", "XLE",
  "GLD", "VIXY"
)

ls_core <- build_ls_portfolio(
  xts_rel          = xts_rel,
  n_long           = 5L,
  n_short          = 5L,
  signal_lookback  = 63L,
  rebalance_freq   = "months",
  universe_tickers = core_universe
)

# Align and merge
common_idx <- intersect(index(ls_full$returns), index(ls_core$returns))
comp <- merge(ls_full$returns[common_idx], ls_core$returns[common_idx])
colnames(comp) <- c("LS_Full_59", "LS_Core_15")
```

### 5.2 Weight History — Core Universe

```{r weight-heatmap, echo=FALSE}
w_rebal <- ls_core$weights_rebal

w_df <- as.data.frame(w_rebal) %>%
  rownames_to_column("date") %>%
  mutate(date = as.Date(date)) %>%
  pivot_longer(-date, names_to = "ticker", values_to = "weight") %>%
  filter(weight != 0) %>%
  mutate(
    side   = if_else(weight > 0, "Long", "Short"),
    ticker = factor(ticker, levels = core_universe)
  )

ggplot(w_df, aes(x = date, y = ticker, fill = weight)) +
  geom_tile(color = "white", linewidth = 0.25) +
  scale_fill_gradient2(
    low = "#c0392b", mid = "white", high = "#27ae60",
    midpoint = 0, name = "Weight",
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x      = element_text(angle = 45, hjust = 1),
    panel.grid       = element_blank(),
    legend.position  = "right"
  ) +
  labs(
    title    = "Core 15-ETF L/S Overlay: Monthly Weight History",
    subtitle = "Green = Long (+20%) | Red = Short (-20%) | White = Neutral",
    x        = NULL, y = NULL
  )
```

### 5.3 Annualized Performance

```{r perf-table, echo=FALSE}
tbl <- table.AnnualizedReturns(comp, scale = 252, geometric = FALSE)
knitr::kable(round(tbl, 4), caption = "Annualized Performance: Full vs Core Universe")
```

### 5.4 Performance Summary Chart

```{r perf-chart, echo=FALSE, fig.height=9}
charts.PerformanceSummary(
  comp,
  main       = "L/S Overlay: Full Universe (59) vs Core Universe (15)",
  colorset   = c("#bdc3c7", "#2c3e50"),
  lwd        = c(2, 3),
  legend.loc = "topleft",
  geometric  = FALSE
)
```

### 5.5 Drawdown Analysis

```{r drawdown-table, echo=FALSE}
dd_full <- table.Drawdowns(ls_full$returns, top = 5)
dd_core <- table.Drawdowns(ls_core$returns, top = 5)

cat("Full Universe (59 ETFs) — Top 5 Drawdowns\n")
print(dd_full)
cat("\nCore Universe (15 ETFs) — Top 5 Drawdowns\n")
print(dd_core)
```

---

## 6. Key Findings

```{r findings-table, echo=FALSE}
data.frame(
  Metric              = c("Ann. Return", "Ann. Volatility", "Sharpe Ratio",
                          "Universe size", "Rebalances", "Signal"),
  `Full (59 ETFs)`    = c("34.4%", "21.3%", "1.62", "59", "99 (monthly)",
                          "63-day cumulative rel. return"),
  `Core (15 ETFs)`    = c("26.4%", "17.7%", "1.50", "15", "99 (monthly)",
                          "63-day cumulative rel. return"),
  check.names = FALSE
) %>% knitr::kable(caption = "Summary Comparison")
```

**Interpretation:**

1. **Sharpe degrades only 7%** (1.62 → 1.50) when universe shrinks from 59 to 15 ETFs.
   The 15 ETFs genuinely capture the signal — the full universe gains from breadth, not
   from independent information.

2. **Volatility drops from 21.3% to 17.7%** because the core universe eliminates
   redundant correlated positions, producing a cleaner factor exposure.

3. **Same drawdown regime** across both universes — the three major drawdown events
   (Sep 2018–Jan 2020, Sep 2024–Oct 2025, Jul–Sep 2024) occur at identical dates,
   confirming both versions are exposed to the same underlying risk factor.

4. **PC3 (Growth vs Cyclical) drives equity alpha.** The most important rotation
   within equities is QQQ/XLK vs XLF/XLI/XLE — these four tickers alone define
   the equity L/S signal.

---

## 7. Next Steps

- **Phase 2**: Add L/S overlay on top of the 50/50 core (SPY + AGG) to combine
  structural beta with pure alpha
- **Robustness**: Check for signal lag bias, sub-period stability (pre/post COVID),
  and sensitivity to `signal_lookback` (21d, 63d, 126d)
- **Transaction costs**: Estimate turnover-adjusted returns given monthly rebalancing
  on 15 ETFs

---

*Generated: `r Sys.Date()` | Data: 2018-01-01 to present | Signal: 63-day momentum on xts_rel*
