---
title: "SPY Drawdown Regime & State Machine — Methodology and Results"
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 8
)

library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(scales)
library(patchwork)
library(here)

# ── Data ──────────────────────────────────────────────────────────────────────
xts_ret <- readRDS(here("02_data_processed/xts_ret_returns.rds"))

# ── Sources ───────────────────────────────────────────────────────────────────
source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
source(here("scripts_spy_dd_regime/regime_multi_ticker.R"))
source(here("scripts_spy_dd_regime/spy_dd_regime_rel.R"))
source(here("scripts_spy_dd_regime/spy_statemachine_diagram_rel.R"))

# ── Universe ──────────────────────────────────────────────────────────────────
core_universe <- c("IEF", "HYG", "IEFA", "XLK", "XLF", "XLI", "GLD")

spy_ret <- xts_ret[, "SPY"]
rt      <- build_regime_table(spy_ret)
```

---

## 1. Overview

This report documents the **SPY Drawdown Regime Engine** — a 3-state classification
system that labels every calendar day as one of:

| State | Colour | Condition |
|---|---|---|
| **Fall** | <span style="color:#D90429">■</span> Red | Running drawdown from recent peak ≥ 10% |
| **Recovery** | <span style="color:#F77F00">■</span> Orange | After Fall trough — climbing back toward peak |
| **Consolidation** | <span style="color:#2D6A4F">■</span> Green | Everything else — expansionary phase |

The engine is fully data-driven (no manual labels), uses only `xts_ret[,"SPY"]`,
and is parameterised by a single threshold `t_fall` (default 10%).

Results are visualised through four scripts:

| Script | Purpose |
|---|---|
| `spy_dd_regime.R` | Single-ticker regime overlay + stats |
| `regime_multi_ticker.R` | Multi-ticker stacked bar overlay |
| `spy_dd_regime_rel.R` | Mixed mode — SPY absolute, others α vs SPY |
| `spy_statemachine_diagram_rel.R` | State machine flow diagram + performance panels |

```{r data-summary, echo=TRUE}
cat(sprintf("SPY data  : %s → %s  (%d trading days)\n",
            as.character(start(spy_ret)),
            as.character(end(spy_ret)),
            nrow(spy_ret)))
cat(sprintf("Regime periods identified : %d\n", nrow(rt)))
cat(sprintf("Current regime            : %s  (since %s)\n",
            as.character(tail(rt$regime, 1)),
            format(tail(rt$xmin, 1), "%d %b %Y")))
```

---

## 2. Regime Definitions

### 2.1 Three-State Model

The engine identifies macro drawdown events using `PerformanceAnalytics::table.Drawdowns()`
and classifies every period as follows:

$$\text{Fall}_i : \quad DD_t \leq -t_{fall}$$

$$\text{Recovery}_i : \quad \text{From trough of Fall}_i \text{ until drawdown recovered}$$

$$\text{Consolidation}_i : \quad \text{All remaining expansion periods}$$

where the running drawdown at time $t$ is:

$$DD_t = \frac{P_t}{\max_{s \leq t} P_s} - 1$$

### 2.2 Empirical Period Summary

```{r regime-table, echo=FALSE}
rt %>%
  as_tibble() %>%
  mutate(
    regime        = as.character(regime),
    xmin          = format(xmin, "%d %b %Y"),
    xmax          = format(xmax, "%d %b %Y"),
    period_return = scales::percent(period_return, accuracy = 0.1),
    ann_return    = scales::percent(ann_return,    accuracy = 0.1)
  ) %>%
  select(regime, xmin, xmax, days, period_return, ann_return) %>%
  knitr::kable(
    col.names = c("Regime", "From", "To", "Days", "Period Return", "Ann. Return"),
    align = c("l","l","l","r","r","r")
  )
```

---

## 3. Time-Series Regime Chart

`plot_regime_overlay()` — cumulative SPY return with regime bar row and `▶` transition
arrows coloured by the incoming state.

```{r x1, fig.height=7}
plot_regime_overlay(spy_ret, t_fall = 0.10, asset_name = "SPY")
```

---

## 4. Regime Statistics

`plot_regime_stats()` — average period return, annualised return, duration, and
fraction of time spent in each state.

```{r x2, fig.height=5}
plot_regime_stats(spy_ret, asset_name = "SPY")
```

---

## 5. Monthly Calendar Heatmap

`plot_regime_calendar()` — each month coloured by dominant regime. Useful for
identifying seasonal patterns in drawdown cycles.

```{r x3, fig.height=6}
plot_regime_calendar(spy_ret, asset_name = "SPY")
```

---

## 6. State Machine Flow Diagram

### 6.1 Pure State Machine

`plot_sm_diagram()` — directed graph of the 3-state cycle. Each node shows:
*N periods · avg duration · avg return · **speed** (annualised return) · % of time*.
Each edge shows the transition rule and empirical probability (`p = XX%`).

```{r x4, fig.height=12, fig.width=16}
plot_sm_diagram(xts_ret, master = "SPY", t_fall = 0.10)
```

### 6.2 State Machine + Relative Performance

`plot_sm_diagram_perf()` — same triangle layout with a **performance panel** below
each node. One horizontal bar per ticker:

- **SPY** (navy): absolute avg return for that regime
- **Others** (green/red): avg alpha vs SPY — width proportional to magnitude

```{r x5, fig.height=14, fig.width=16}
plot_sm_diagram_perf(xts_ret, tickers = core_universe,
                     master = "SPY", t_fall = 0.10)
```

---

## 7. Ticker × Regime Alpha Heatmap

`plot_sm_alpha_heatmap()` — compact matrix: rows = tickers, columns = regime types.
SPY row shows absolute avg return; all other rows show avg alpha vs SPY.
Navy border distinguishes SPY from relative rows.

```{r x6, fig.height=5, fig.width=9}
plot_sm_alpha_heatmap(xts_ret, tickers = core_universe,
                      master = "SPY", t_fall = 0.10)
```

---

## 8. Regime Transition Probability Matrix

`plot_sm_transition_matrix()` — empirical 4×4 (3×3 here) transition heatmap.
Each cell: observed probability + count. Row = From, Column = To.

```{r x7, fig.height=5, fig.width=7}
plot_sm_transition_matrix(xts_ret, master = "SPY", t_fall = 0.10)
```

---

## 9. Multi-Ticker Regime Overlay

`plot_regime_multi_overlay()` — SPY cumulative return at top; stacked bar rows below,
one per ticker, showing absolute return per regime period.

```{r x8, fig.height=9}
plot_regime_multi_overlay(
  xts_ret  = xts_ret,
  tickers  = core_universe,
  master   = "SPY",
  t_fall   = 0.10,
  relative = FALSE
)
```

---

## 10. Relative Mode — Mixed Overlay

`plot_regime_rel_overlay()` — SPY row shows **absolute** return per period;
all other rows show **alpha vs SPY** (+ prefix = outperformed).

```{r x9, fig.height=9}
plot_regime_rel_overlay(
  xts_ret = xts_ret,
  tickers = core_universe,
  master  = "SPY",
  t_fall  = 0.10
)
```

---

## 11. Relative Mode — Per-Regime Heatmap

`plot_regime_rel_heatmap()` — each column is one historical regime period (not a
regime type). SPY row bordered in navy; diverging red/green scale centred at zero.

```{r x10, fig.height=5}
plot_regime_rel_heatmap(
  xts_ret = xts_ret,
  tickers = core_universe,
  master  = "SPY",
  t_fall  = 0.10
)
```

---

## 12. Relative Mode — Summary by Regime Type

`plot_regime_rel_summary()` — average return/alpha per ticker faceted by regime type.
Sorted within each facet; `+` prefix indicates outperformance of SPY.

```{r x11, fig.height=6}
plot_regime_rel_summary(
  xts_ret = xts_ret,
  tickers = core_universe,
  master  = "SPY",
  t_fall  = 0.10
)
```

---

## 13. Key Findings

```{r x12, echo=FALSE}
stats_tbl <- rt %>%
  as_tibble() %>%
  mutate(regime = as.character(regime)) %>%
  group_by(regime) %>%
  summarise(
    N           = n(),
    avg_days    = round(mean(days)),
    avg_ret     = mean(period_return),
    avg_ann_ret = mean(ann_return, na.rm = TRUE),
    pct_time    = sum(days) / sum(rt$days),
    .groups     = "drop"
  )

cat("── Regime Summary ──────────────────────────────────────────────────────\n")
for (i in seq_len(nrow(stats_tbl))) {
  r <- stats_tbl[i, ]
  cat(sprintf(
    "  %-16s  N=%-2d  avg %3d days  avg return %+5.1f%%  speed %+5.1f%%/yr  %4.0f%% of time\n",
    r$regime, r$N, r$avg_days,
    r$avg_ret * 100, r$avg_ann_ret * 100, r$pct_time * 100
  ))
}
cat("────────────────────────────────────────────────────────────────────────\n")

current <- tail(rt, 1)
cat(sprintf("\nCurrent regime : %s\n", as.character(current$regime)))
cat(sprintf("Started        : %s\n", format(current$xmin, "%d %b %Y")))
cat(sprintf("Duration so far: %d days\n", as.numeric(Sys.Date() - current$xmin)))
```

**Structural observations:**

- **Fall** is the shortest and most intense state — large negative returns over a
  relatively small number of days, with high annualised velocity.
- **Recovery** is the mirror — the market recoups losses quickly once the trough is
  confirmed; speed in Recovery typically exceeds Fall speed in absolute terms.
- **Consolidation** dominates calendar time (~60%+). Ticker alpha is most dispersed
  here — sector rotation signals are strongest in this state.
- **Transition probabilities** are highly deterministic: Fall → Recovery is nearly
  always 100%; Recovery → Consolidation is the dominant path.

---

## 14. Next Steps

| Priority | Item |
|---|---|
| 1 | Wire regime labels into SAA/TAA signal layer as a regime overlay |
| 2 | Extend relative analysis to full 59-ETF universe |
| 3 | Backtest sector-rotation strategy conditioned on regime state |
| 4 | Add Consolidation sub-classification (early vs late cycle) |
| 5 | Embed real-time regime badge in Shiny dashboard |

---

*Generated with `spy_statemachine_diagram_rel.R` · `spy_dd_regime.R` ·
`regime_multi_ticker.R` · `spy_dd_regime_rel.R`*
