---
title: "Phase 2 — Overlay Backtest Report"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
params:
  role:         "Enhancer"   # "Enhancer" | "Stabilizer"
  tickers:      !r NULL      # NULL → auto-loaded from find_enhancers() / find_stabilizers()
  overlay_w:    0.10
  w_eq:         0.60
  w_fi:         0.40
  ma_win:       200
---

```{css, echo=FALSE}
.exec-box {
  background: #f0f9ff;
  border-left: 5px solid #1d4ed8;
  padding: 14px 18px;
  margin: 16px 0;
  font-size: 0.94em;
  border-radius: 3px;
}
.rec-box {
  background: #f0fdf4;
  border-left: 5px solid #15803d;
  padding: 14px 18px;
  margin: 16px 0;
  font-size: 0.94em;
  border-radius: 3px;
}
.warn-box {
  background: #fefce8;
  border-left: 5px solid #ca8a04;
  padding: 10px 16px;
  margin: 10px 0;
  font-size: 0.92em;
}
.type-a { background:#ede9fe; border-left:4px solid #7c3aed; padding:5px 12px; margin:4px 0; }
.type-b { background:#fef9c3; border-left:4px solid #ca8a04; padding:5px 12px; margin:4px 0; }
.type-c { background:#dcfce7; border-left:4px solid #16a34a; padding:5px 12px; margin:4px 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 14,
  fig.height = 6
)

library(tidyverse)
library(xts)
library(zoo)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

# ── Environment ───────────────────────────────────────────────────────────────
if (!exists("project_tree"))       source(here("project_tree.R"))
if (!exists("etf_metadata"))       source(here(project_tree$scripts$init))
if (!exists("xts_ret"))            source(here(project_tree$scripts$wrangle))
if (!exists("build_regime_table")) source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
if (!exists("rt") || !is.data.frame(rt))
  rt <- build_regime_table(xts_ret[, "SPY"])

W_EQ           <- params$w_eq
W_FI           <- params$w_fi
OVERLAY_W      <- params$overlay_w
MA_WIN         <- as.integer(params$ma_win)
BASELINE_LABEL <- sprintf("%d/%d Classic", round(W_EQ * 100), round(W_FI * 100))
ROLE           <- params$role   # "Enhancer" | "Stabilizer"

if (!exists(".build_baseline")) source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))
source(here("scripts_daa_saa_taa/09_overlay_backtest.R"))

# ── Resolve tickers ───────────────────────────────────────────────────────────
# Priority: (1) explicit params$tickers, (2) run the relevant screen
tickers <- params$tickers

if (is.null(tickers)) {
  if (ROLE == "Enhancer") {
    if (!exists("find_enhancers")) source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))
    enh     <- find_enhancers(xts_ret, rt)
    tickers <- enh$selected$ticker
  } else {
    if (!exists("find_stabilizers")) source(here("scripts_daa_saa_taa/05_dd_stabilizer.R"))
    stab    <- find_stabilizers(xts_ret, rt)
    tickers <- stab$selected$ticker
  }
}

# ── Run backtest ──────────────────────────────────────────────────────────────
ob <- run_overlay_backtest(
  tickers        = tickers,
  xts_ret        = xts_ret,
  rt             = rt,
  role           = ROLE,
  w_eq           = W_EQ,
  w_fi           = W_FI,
  baseline_label = BASELINE_LABEL,
  overlay_w      = OVERLAY_W,
  ma_win         = MA_WIN,
  save_png       = FALSE
)

comp    <- ob$comparison
optimal <- ob$optimal
rec     <- optimal$recommendations

# Colour / accent by role
accent_col  <- if (ROLE == "Enhancer") "#1d4ed8" else "#15803d"
accent_fill <- if (ROLE == "Enhancer") "#eff6ff"  else "#f0fdf4"
phase_tag   <- if (ROLE == "Enhancer") "Phase 2.2 — Enhancer Overlay" else "Phase 2.1 — Stabilizer Overlay"
crit_label  <- if (optimal$criterion == "dd") "lowest MaxDD" else "highest IR"

cur_regime <- {
  last_date <- as.Date(max(index(xts_ret)))
  rd <- rt %>%
    rowwise() %>%
    mutate(date = list(seq(as.Date(xmin), as.Date(xmax), by = "day"))) %>%
    unnest(date) %>%
    ungroup() %>%
    filter(date <= last_date) %>%
    slice_tail(n = 1) %>%
    pull(regime)
  if (length(rd) == 0) "Consolidation" else rd
}

bl_stats <- purrr::map_dfr(comp$results, ~ .x$stats %>% filter(type == "Baseline")) %>%
  summarise(ann_ret = mean(ann_ret), max_dd = mean(max_dd), ir = mean(ir))
```

---

## `r phase_tag` — `r ROLE` Trigger Type Analysis

::: {.exec-box}
**Baseline:** `r BASELINE_LABEL` (`r round(W_EQ*100)`% SPY + `r round(W_FI*100)`% IEF)
**Overlay weight:** `r round(OVERLAY_W*100)`%  per ticker
**Tickers evaluated:** `r paste(tickers, collapse = "  |  ")`
**Current SPY regime:** **`r cur_regime`**
**Selection criterion:** `r crit_label`
:::

---

## Executive Summary & Recommendations

```{r exec-summary-table}
rec_display <- rec %>%
  mutate(
    ann_ret_f  = sprintf("%+.1f%%", ann_ret  * 100),
    max_dd_f   = sprintf("%.1f%%",  max_dd   * 100),
    ir_f       = sprintf("%.2f",    ir),
    ret_delta_f= sprintf("%+.1f pp",ret_delta * 100),
    dd_delta_f = sprintf("%+.1f pp",dd_delta  * 100),
    type_label = case_when(
      type == "A" ~ "A — Unconditional",
      type == "B" ~ "B — Regime-conditional",
      type == "C" ~ "C — 200DMA signal"
    )
  ) %>%
  select(Ticker = ticker, `Recommended type` = type_label,
         `Ann Ret` = ann_ret_f, MaxDD = max_dd_f, IR = ir_f,
         `ΔRet` = ret_delta_f, `ΔDD` = dd_delta_f,
         Reason = reason)

rec_display %>%
  kable(align = "llrrrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(which(rec$type == "A"), background = "#ede9fe") %>%
  row_spec(which(rec$type == "B"), background = "#fef9c3") %>%
  row_spec(which(rec$type == "C"), background = "#dcfce7") %>%
  column_spec(2, bold = TRUE) %>%
  column_spec(8, italic = TRUE, color = "grey40")
```

```{r optimal-stats}
if (!is.null(optimal$optimal_stats)) {
  optimal$optimal_stats %>%
    mutate(
      ann_ret = sprintf("%+.1f%%", ann_ret * 100),
      max_dd  = sprintf("%.1f%%",  max_dd  * 100),
      ann_vol = sprintf("%.1f%%",  ann_vol * 100),
      ir      = sprintf("%.2f",    ir)
    ) %>%
    kable(
      col.names = c("Portfolio", "Ann Ret", "MaxDD", "Ann Vol", "IR"),
      align = "lrrrr",
      caption = sprintf("Optimal combination vs %s (equal-weight, %.0f%% each ticker)",
                        BASELINE_LABEL, OVERLAY_W * 100)
    ) %>%
    kable_styling(bootstrap_options = c("condensed", "hover"), full_width = FALSE,
                  font_size = 12) %>%
    row_spec(1, bold = TRUE, background = "#f1f5f9") %>%
    row_spec(2, bold = TRUE, background = accent_fill)
}
```

::: {.rec-box}
**Implementation spec — `r ROLE` overlay at `r round(OVERLAY_W*100)`% per ticker:**

```{r impl-spec, results='asis'}
for (i in seq_len(nrow(rec))) {
  r <- rec[i, ]
  tag <- switch(r$type,
    "A" = "**Type A — hold always**",
    "B" = "**Type B — hold in Recovery + Consolidation; exit in Fall**",
    "C" = "**Type C — hold when above 200DMA; exit when below**"
  )
  cat(sprintf("- **%s** → %s  *(IR: %.2f | MaxDD: %.1f%% | ΔDD: %+.1f pp)*\n\n",
              r$ticker, tag, r$ir, r$max_dd * 100, r$dd_delta * 100))
}
```
:::

### Trigger type legend

<div class="type-a">**A — Unconditional:** blend held always at fixed weight. No switching cost. Suitable when the ticker adds value in all regimes.</div>
<div class="type-b">**B — Regime-conditional:** blend held in Recovery + Consolidation; revert to baseline in Fall. Avoids drawdown amplification. Lag risk: Fall fires after ~10% already absorbed.</div>
<div class="type-c">**C — 200DMA signal:** blend held when ticker is above its `r MA_WIN`-day MA. Smoother entry/exit than hard regime switch. Whipsaw risk in choppy markets.</div>

---

## Scoreboard

Delta vs `r BASELINE_LABEL`. **Green = improvement.**

```{r scoreboard, fig.height=max(4, length(comp$results) * 0.65 + 3), echo=TRUE}
# ob$p_scoreboard
print(ob$p_scoreboard)
```

---

## Wealth comparison grid

`r round(OVERLAY_W*100)`% overlay, all three trigger types vs `r BASELINE_LABEL`.

```{r grid, fig.height=length(comp$results)*3.8+2, echo=TRUE}
# ob$p_grid
print(ob$p_grid)
```

---

## Full statistics table

```{r full-stats}
all_stats <- purrr::map_dfr(comp$results, ~ .x$stats) %>%
  select(ticker, type, ann_ret, max_dd, ann_vol, ir) %>%
  mutate(
    ann_ret = sprintf("%+.1f%%", ann_ret * 100),
    max_dd  = sprintf("%.1f%%",  max_dd  * 100),
    ann_vol = sprintf("%.1f%%",  ann_vol * 100),
    ir      = sprintf("%.2f",    ir)
  ) %>%
  arrange(ticker, factor(type, levels = c("Baseline", "A", "B", "C")))

all_stats %>%
  kable(col.names = c("Ticker", "Type", "Ann Ret", "MaxDD", "Ann Vol", "IR"),
        align = "llrrrr") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  row_spec(which(all_stats$type == "Baseline"), background = "#f1f5f9", bold = TRUE) %>%
  row_spec(which(all_stats$type == "A"),        background = "#ede9fe") %>%
  row_spec(which(all_stats$type == "B"),        background = "#fef9c3") %>%
  row_spec(which(all_stats$type == "C"),        background = "#dcfce7")
```

---

## Session info

```{r session}
cat(sprintf("Report date      : %s\n", Sys.Date()))
cat(sprintf("Role             : %s\n", ROLE))
cat(sprintf("Baseline         : %s (%.0f%% SPY + %.0f%% IEF)\n",
            BASELINE_LABEL, W_EQ * 100, W_FI * 100))
cat(sprintf("Overlay weight   : %.0f%% per ticker\n", OVERLAY_W * 100))
cat(sprintf("MA window (C)    : %d days\n", MA_WIN))
cat(sprintf("Current regime   : %s\n", cur_regime))
cat(sprintf("Selection crit.  : %s\n", crit_label))
cat(sprintf("Tickers          : %s\n", paste(tickers, collapse = ", ")))
```
