---
title: "SAA Overlay Workflow"
subtitle: "Build → Classify → Fingerprint → Backtest — End-to-End Guide"
author: "ETF Intelligence Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: xelatex
params:
  ticker1:            "SPY"    # first SAA component
  ticker2:            "IEF"    # second SAA component; set to null for single-ticker SAA
  w1:                 0.60     # target weight for ticker1 (w2 = 1 - w1 automatically)
  enh_pct:            0.10     # enhancer overlay fraction α
  stab_pct:           0.10     # stabilizer overlay fraction α
  top_n:              5        # number of candidates to rank
  print_fingerprints: false    # true = run screen_ticker() per candidate (slow)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 14,
  fig.height = 8
)
library(here)
source(here("project_tree.R"))
source(here("00_libraries.R"))
source(here(project_tree$scripts$spy_dd_regime))
source(here(project_tree$scripts$wrangle))
rt <- build_regime_table(xts_ret[, "SPY"])

# Load the workflow — knitr.in.progress guard suppresses auto-print
source(here("scripts_daa_saa_taa/saa_overlay_workflow.R"), local = knitr::knit_global())

# Resolve ticker2: YAML null arrives as the string "null" — convert to R NULL
.t2 <- params$ticker2
p_ticker2 <- if (is.null(.t2) || identical(.t2, "null") || identical(trimws(.t2), "")) {
  NULL
} else {
  .t2
}
rm(.t2)

# Human-readable SAA label for headings
saa_label <- if (is.null(p_ticker2)) {
  sprintf("100%% %s", params$ticker1)
} else {
  sprintf("%.0f%%/%.0f%% %s/%s",
          params$w1 * 100, (1 - params$w1) * 100,
          params$ticker1, p_ticker2)
}
```

---

# 1  Purpose

This workflow answers a single practical question:

> Given a binary SAA (two tickers + target weights), which assets from the
> full universe best **enhance** its return or **stabilise** its drawdown —
> and how much does adding each one improve the portfolio?

It runs five steps automatically:

| Step | Function | Output |
|---|---|---|
| 1 | `build_saa_ret()` | Daily-rebalanced SAA return series → injected into `xts_ret` |
| 2 | `rank_enhancers()` / `rank_stabilizers()` | Top 5 candidates ranked by criterion |
| 3 | `overlay_summary_heatmap()` | Two-panel ranking table |
| 4 | `fingerprint_candidates()` | `screen_ticker()` fingerprint per candidate |
| 5 | `backtest_ls_overlay()` | Wealth + L/S spread + stats for each set |

All steps are orchestrated by a single call to `run_saa_overlay_workflow()`.

---

# 2  Design Choices (v1)

| Dimension | v1 Choice | Rationale |
|---|---|---|
| SAA type | Binary (two tickers) | Simplest starting point |
| SAA rebalancing | Daily — $r_{\text{SAA}} = w_1 r_1 + w_2 r_2$ | Constant-weight; no loop needed |
| SAA injection | Added to `xts_ret` as `"saa{W1}{W2}"` | Allows `classify_universe(baseline = "saa6040")` with no code changes |
| Enhancer ranking | `full_ir` — information ratio over full period | Single interpretable number; already computed by `classify_universe()` |
| Stabilizer ranking | `dd_reduction` — MaxDD reduction in pp | Most direct measure of the overlay's defensive value |
| Universe | Full Sovereign Universe (default); pass `universe` vector to restrict | Consistent with existing `classify_universe()` interface |
| Overlay formula | **Option B**: $r_{\text{ov}} = r_{\text{SAA}} + \alpha (r_{\text{candidate}} - r_{\text{SAA}})$ | Keeps pure L/S framing; maps to existing `build_ls_portfolio()` pattern |
| Backtest | Long overlay-SAA / Short original-SAA | Spread PnL = $\alpha (r_{\text{candidate}} - r_{\text{SAA}})$ |
| Fingerprint | `screen_ticker()` per candidate + condensed summary | Reuses existing 7-panel fingerprint infrastructure |

---

# 3  Overlay Formula

$$r_t^{\text{overlay}} = r_t^{\text{SAA}} + \alpha \times \left(r_t^{\text{candidate}} - r_t^{\text{SAA}}\right)$$

Expanding:

$$= (1 - \alpha)\, r_t^{\text{SAA}} + \alpha\, r_t^{\text{candidate}}$$

So the overlay portfolio is simply **$(1-\alpha)$ of SAA plus $\alpha$ of the
candidate** — but the L/S framing makes clear what you are *adding*: a spread
trade on top of unchanged SAA exposure.

**L/S spread daily PnL:**

$$\text{spread}_t = \alpha \times \left(r_t^{\text{candidate}} - r_t^{\text{SAA}}\right)$$

Cumulative spread > 1 means the candidate is beating the SAA over that period.

---

# 4  Function Reference

## 4.1  `build_saa_ret()`

```r
build_saa_ret(
  ticker1,          # character — first component (e.g. "SPY")
  ticker2,          # character — second component (e.g. "IEF")
  w1,               # numeric  — target weight for ticker1 (e.g. 0.60)
  w2  = 1 - w1,     # numeric  — complement (default)
  name = NULL       # character — column name; default "saa{W1}{W2}" e.g. "saa6040"
)
```

**What it does:**

1. Computes $r_{\text{SAA}, t} = w_1 r_{1,t} + w_2 r_{2,t}$ over the common date range.
2. Assigns the result to the global `xts_ret` as a new column.
3. Prints a one-line confirmation with the column name, date range, and day count.

If the column already exists, it is silently skipped (safe to re-run).

## 4.2  `rank_enhancers()` / `rank_stabilizers()`

```r
rank_enhancers(
  baseline,           # SAA column name in xts_ret (e.g. "saa6040")
  universe  = NULL,   # character vector; NULL = all cols except baseline
  top_n     = 5L,
  overlay_w = 0.10,   # overlay weight passed to screen_ticker() inside classify_universe()
  verbose   = FALSE
)
```

Both functions call `classify_universe(baseline = saa_name, ...)` which runs
`screen_ticker()` for every ticker in the universe and returns the full
classification table. Then:

- **`rank_enhancers()`** filters to `classification %in% c("Enhancer","Dual")`,
  sorts descending by `full_ir`, returns top N.
- **`rank_stabilizers()`** filters to `classification %in% c("Stabilizer","Dual")`,
  sorts descending by `dd_reduction`, returns top N.

**Returned columns:** `ticker`, `classification`, `trigger_type`, `hedge_type`,
`full_ir`, `dd_reduction`, `ins_ratio`, `fall_corr`, `n_days`.

**Key metrics:**

| Metric | Definition | Units |
|---|---|---|
| `full_ir` | Information ratio: excess Ann Ret / tracking error vs SAA | ratio |
| `dd_reduction` | Reduction in SAA MaxDD when overlay_w of candidate is added | pp (percentage points) |
| `ins_ratio` | dd_reduction / return_drag — insurance per unit of cost | ratio |
| `fall_corr` | Correlation with SAA *during Fall regimes only* | −1 to +1 |

## 4.3  `fingerprint_candidates()`

```r
fingerprint_candidates(
  tickers,            # character vector — tickers to screen
  baseline,           # SAA column name
  overlay_w = 0.10,
  print_plot = TRUE,  # print 7-panel fingerprint to Plots pane
  ...                 # additional args passed to screen_ticker()
)
```

Loops `screen_ticker(tk, baseline = saa_name)` for each ticker. Returns a
named list; each element is a `screen_ticker()` result:

```r
res$enh_screens[["GLD"]]$plot        # 7-panel fingerprint plot
res$enh_screens[["GLD"]]$enhancer    # enhancer verdict tibble
res$enh_screens[["GLD"]]$stabilizer  # stabilizer verdict tibble
```

Set `print_fingerprints = FALSE` in `run_saa_overlay_workflow()` to skip this
step during exploratory runs (it is the slowest step).

## 4.4  `overlay_summary_heatmap()`

```r
overlay_summary_heatmap(enh_tbl, stab_tbl, baseline = "")
```

Two-panel patchwork figure:

- **Left — Enhancers:** Rank | Ticker | Class | Trigger | Full IR (traffic-light: green = #1, red = #N)
- **Right — Stabilizers:** Rank | Ticker | Class | Hedge | DD Red (pp) | Ins Ratio | Fall ρ

## 4.5  `backtest_ls_overlay()`

```r
backtest_ls_overlay(
  saa_name,                           # SAA column in xts_ret
  tickers,                            # candidate tickers
  pct,                                # α — overlay fraction
  type = c("enhancer","stabilizer")   # controls colour palette and title
)
```

Returns a list with five elements:

| Element | Content |
|---|---|
| `$fig` | Patchwork figure: P1 wealth + P2 spread + P3 stats table |
| `$scatter` | Two-panel scatter: absolute (Ann Vol vs Ann Ret, iso-Sharpe) + relative (TE vs excess return, iso-IR) |
| `$stats` | Tibble: Ann Ret / Ann Vol / Max DD / Sharpe / Calmar for SAA + each overlay |
| `$ret` | xts: SAA + all overlay return series on common dates |
| `$spread` | data.frame: daily cumulative spread for each candidate |

**Figure panels:**

- **P1 (wealth, tall):** SAA in purple + each overlay in its colour; end-of-line labels show ticker + Ann Ret; regime shading.
- **P2 (spread, medium):** Cumulative $\alpha(r_{\text{candidate}} - r_{\text{SAA}})$ for each candidate; dashed zero line; regime shading.
- **P3 (stats, short):** Traffic-light table — green = best metric, red = worst; strategy names in their respective colours.

## 4.6  `run_saa_overlay_workflow()` — master orchestrator

```r
run_saa_overlay_workflow(
  ticker1,
  ticker2,
  w1,
  w2                 = 1 - w1,
  saa_name           = NULL,    # auto: "saa{W1}{W2}"
  universe           = NULL,    # NULL = full Sovereign Universe
  top_n              = 5L,
  enh_pct            = 0.10,    # α for enhancer overlays
  stab_pct           = 0.10,    # α for stabilizer overlays
  overlay_w          = 0.10,    # classification overlay weight
  print_fingerprints = TRUE,    # FALSE = skip screen_ticker() (faster)
  verbose            = FALSE
)
```

Runs all five steps in sequence and returns a named list:

```r
res$saa_name       # "saa6040"
res$enhancers      # tibble — top 5 enhancers with metrics
res$stabilizers    # tibble — top 5 stabilizers with metrics
res$summary_fig    # patchwork heatmap (Step 3)
res$enh_screens    # named list of screen_ticker() results (Step 4)
res$stab_screens   # named list of screen_ticker() results (Step 4)
res$enh_backtest   # list(fig, stats, ret, spread) — enhancer backtest
res$stab_backtest  # list(fig, stats, ret, spread) — stabilizer backtest
```

---

# 5  Live Example — `r saa_label`

## 5.1  Run the full workflow

Parameters active in this render (set via `params` in YAML or `rmarkdown::render()`):

| Param | Value | Meaning |
|---|---|---|
| `ticker1` | `r params$ticker1` | First SAA component |
| `ticker2` | `r if (is.null(p_ticker2)) "*(none — single-ticker SAA)*" else p_ticker2` | Second SAA component |
| `w1` | `r sprintf("%.0f%%", params$w1 * 100)` | Weight of ticker1 |
| `enh_pct` | `r sprintf("%.0f%%", params$enh_pct * 100)` | Enhancer overlay α |
| `stab_pct` | `r sprintf("%.0f%%", params$stab_pct * 100)` | Stabilizer overlay α |
| `top_n` | `r params$top_n` | Candidates per type |
| `print_fingerprints` | `r params$print_fingerprints` | Run screen_ticker()? |

```{r workflow-run, fig.height=6}
res <- run_saa_overlay_workflow(
  ticker1            = params$ticker1,
  ticker2            = p_ticker2,          # NULL for single-ticker SAA
  w1                 = params$w1,
  enh_pct            = params$enh_pct,
  stab_pct           = params$stab_pct,
  top_n              = params$top_n,
  print_fingerprints = params$print_fingerprints
)
```

## 5.2  Ranking summary

```{r summary-heatmap, fig.height=4}
print(res$summary_fig)
```

**Reading the table:**

- **Enhancers** are ranked by `full_ir` — the annualised excess return per unit
  of tracking error vs the SAA. Positive IR means the ticker beat the SAA on a
  risk-adjusted basis over the full history.
- **Stabilizers** are ranked by `dd_reduction` (in pp) — how many percentage
  points of MaxDD the SAA loses when 10% of the portfolio is replaced by the
  candidate. `Ins Ratio` = DD reduction / return drag; higher = cheaper insurance.
- **Hedge type:**
  - `FULL` — Fall correlation ≤ −0.10: true crisis hedge, rallies when SAA falls.
  - `PARTIAL` — Fall correlation between −0.10 and +0.30: uncorrelated in Falls.
  - `CONDITIONAL` — Fall correlation > +0.30: improves DD only through portfolio maths.

## 5.3  Enhancer raw rankings

```{r enh-table}
res$enhancers[, c("ticker","classification","trigger_type","full_ir","fall_corr","n_days")]
```

## 5.4  Stabilizer raw rankings

```{r stab-table}
res$stabilizers[, c("ticker","classification","hedge_type","dd_reduction","ins_ratio","fall_corr")]
```

---

# 6  Enhancer Backtest

```{r enh-backtest, fig.height=13}
print(res$enh_backtest$fig)
```

**Reading the figure:**

- **P1 (wealth):** Purple = SAA baseline; coloured lines = each enhancer overlay
  at $\alpha = 10\%$. End-of-line annotation shows ticker + annualised return.
  Outperformance vs SAA means the overlay-SAA portfolio beat the base portfolio.
- **P2 (spread):** Cumulative PnL of the pure L/S leg:
  $\alpha \times (r_{\text{enhancer}} - r_{\text{SAA}})$. Above 1 = enhancer
  beating SAA since inception; below 1 = SAA winning.
- **P3 (stats):** Traffic-light colouring per metric column (green = best across
  all rows, red = worst). Strategy names in their respective colours.

**Stats table:**

```{r enh-stats}
res$enh_backtest$stats %>%
  dplyr::mutate(
    `Ann Ret` = sprintf("%.1f%%", `Ann Ret` * 100),
    `Ann Vol` = sprintf("%.1f%%", `Ann Vol` * 100),
    `Max DD`  = sprintf("%.1f%%", `Max DD`  * 100)
  )
```

---

# 7  Stabilizer Backtest

```{r stab-backtest, fig.height=13}
print(res$stab_backtest$fig)
```

**Reading the figure:** Same structure as the enhancer figure. For stabilizers,
the key panels are:

- **P1:** Does the overlay reduce drawdown in Fall regimes (red-shaded bands)?
- **P2:** Is the spread PnL roughly flat outside Falls (low drag in good times)?

The ideal stabilizer profile: spread hovers near zero in Consolidation/Recovery,
then spikes sharply positive during Falls — insurance that costs little in calm
markets.

**Stats table:**

```{r stab-stats}
res$stab_backtest$stats %>%
  dplyr::mutate(
    `Ann Ret` = sprintf("%.1f%%", `Ann Ret` * 100),
    `Ann Vol` = sprintf("%.1f%%", `Ann Vol` * 100),
    `Max DD`  = sprintf("%.1f%%", `Max DD`  * 100)
  )
```

---

# 8  Fingerprints

When `print_fingerprints = TRUE`, each candidate generates a 7-panel
`screen_ticker()` fingerprint:

| Panel | Content |
|---|---|
| P1 | Wealth: ticker vs SAA |
| P2 | L/S spread: ticker minus SAA (cumulative) |
| P3 | Drawdown comparison |
| P4 | Return distribution |
| P5 | Rolling 63-day correlation with SAA |
| P6 | Regime-conditional correlation (Fall / Recovery / Consolidation) |
| P7 | Rolling 63-day information ratio |

```r
# Access individual fingerprints after running with print_fingerprints = TRUE
res2 <- run_saa_overlay_workflow("SPY", "IEF", w1 = 0.60, print_fingerprints = TRUE)
res2$enh_screens[["GLD"]]$plot       # 7-panel fingerprint for GLD
res2$enh_screens[["GLD"]]$enhancer   # enhancer verdict + scores
res2$stab_screens[["TLT"]]$stabilizer  # stabilizer verdict + dd metrics
```

---

# 9  Advanced Usage

## 9.1  Custom universe

Restrict classification to a thematic subset:

```r
equity_hedges <- c("GLD","SLV","BIL","SHY","TLT","BNDX","VXX","DBMF")

res_eq <- run_saa_overlay_workflow(
  ticker1  = "SPY", ticker2 = "IEF", w1 = 0.60,
  universe = equity_hedges,
  enh_pct  = 0.10, stab_pct = 0.10,
  print_fingerprints = FALSE
)
```

## 9.2  Different SAA pairs

The workflow is pair-agnostic. Any two tickers in `xts_ret` work:

```r
# Growth / bond
res_qqq <- run_saa_overlay_workflow("QQQ", "AGG", w1 = 0.65)

# International equity / gold
res_intl <- run_saa_overlay_workflow("URTH", "GLD", w1 = 0.70)
```

Each call creates a new named column in `xts_ret` (e.g. `"saa6535"`, `"saa7030"`),
so multiple SAA variants can coexist.

## 9.3  Varying α

To compare overlay sizes, run `backtest_ls_overlay()` directly:

```r
for (alpha in c(0.05, 0.10, 0.15, 0.20)) {
  bt <- backtest_ls_overlay(
    saa_name = "saa6040",
    tickers  = res$enhancers$ticker,
    pct      = alpha,
    type     = "enhancer"
  )
  print(bt$fig)
}
```

## 9.4  Accessing the classification table

`rank_enhancers()` and `rank_stabilizers()` return only the top N rows.
To see the full table:

```r
source(here("scripts_daa_saa_taa/screen_ticker.R"))
full_cls <- classify_universe(
  baseline  = "saa6040",
  xts_ret   = xts_ret,
  rt        = rt,
  overlay_w = 0.10,
  verbose   = TRUE
)
# All tickers with their classification, full_ir, dd_reduction, ins_ratio, etc.
```

---

# 10  Risk-Return Scatter

Each `backtest_ls_overlay()` call returns a two-panel scatter stored in
`$scatter`. The orchestrator prints both automatically; in Rmd access them
via `res$enh_backtest$scatter` and `res$stab_backtest$scatter`.

## 10.1  Enhancer scatter

```{r scatter-enh, fig.width=13, fig.height=6}
res$enh_backtest$scatter
```

## 10.2  Stabilizer scatter

```{r scatter-stab, fig.width=13, fig.height=6}
res$stab_backtest$scatter
```

## 10.3  How to read the two panels

**Left panel — Absolute (Ann Vol vs Ann Ret)**

Iso-Sharpe reference lines (SR = 0.5 / 1.0 / 1.5) fan from the origin.
The SAA baseline appears in purple; overlays are coloured by type.

| Position relative to SAA | Interpretation |
|---|---|
| Up and left | Higher return, lower vol — strictly dominant |
| Up and right | Higher return, higher vol — Sharpe gain if above SAA's iso-line |
| Down and left | Lower vol, lower return — acceptable when DD reduction is the goal |
| Down and right | Strictly dominated — avoid |

**Right panel — Relative / L/S (Tracking Error vs Excess Return)**

Axes are the spread portfolio (overlay − SAA): tracking error on x,
excess return on y. Iso-IR lines at IR = −0.5 / 0.5 / 1.0 / 1.5.
Each label shows the ticker and its Information Ratio
(`IR = excess_ret / tracking_error`).

Key properties:

- **IR is scale-invariant** — doubling α doubles both excess return and
  tracking error, leaving IR unchanged. Candidates rank the same at any
  overlay size.
- **Stabilizers appear below the zero line by design** — they drag
  return to buy DD insurance. Their value shows up in the absolute
  panel (lower vol, shallower drawdowns).
- **IR > 0.5** is a practical minimum for an enhancer worth including;
  **IR > 1.0** is a strong signal.

## 10.4  Scatter interpretation workflow

1. Absolute scatter — confirm the overlay is not strictly dominated
   (right and down of SAA).
2. Relative scatter — check IR. IR > 0.5 is the minimum bar for
   enhancers.
3. Cross-check with the spread chart (`$fig` panel 2). A high IR driven
   by one or two Fall episodes is less reliable than a flat, steady
   spread.
4. Stabilizers: expect negative excess return. Their case is purely
   in the absolute panel (lower vol/DD) and the spread panel (positive
   spikes during Falls).

---

# 11  Interpretation Guide

## What makes a good enhancer?

An enhancer consistently beats the SAA on a risk-adjusted basis across all
three regimes. The strongest enhancers show:

- `full_ir > 0.3` — meaningful outperformance per unit of tracking risk
- `trigger_type = "ALPHA"` — both the candidate and SAA rise, but the candidate
  rises more (Consolidation/Recovery leadership)
- Fall excess ≥ 0 — doesn't underperform significantly during drawdowns

> High `full_ir` with negative Fall excess is an *enhancer in good times only*
> — valuable for return but adds risk concentration at the worst moment.

## What makes a good stabilizer?

A stabilizer reduces the SAA's MaxDD at low cost to return. The strongest
stabilizers show:

- `dd_reduction > 2pp` — meaningful drawdown protection at 10% overlay
- `ins_ratio > 5` — each pp of return drag buys ≥5pp of DD reduction
- `fall_corr < 0` — genuinely negatively correlated with SAA during Falls (`FULL` hedge)
- Spread PnL: flat in calm periods, positive spikes during Falls

> `CONDITIONAL` hedge types improve DD only through portfolio variance maths
> (low correlation but positive Fall correlation) — cheaper but less reliable
> during crisis episodes than `FULL` hedges.

## Decision table

| Objective | Primary metric | Secondary metric |
|---|---|---|
| Add return without increasing risk | `full_ir` | Sharpe vs SAA baseline |
| Reduce worst drawdown | `dd_reduction` | `ins_ratio` |
| Reduce tail loss | `CVaR 5%` improvement | `fall_corr` |
| Cheapest insurance | `ins_ratio` | `return_drag` |
| Regime-aware protection | `hedge_type = FULL` | `fall_corr` < −0.10 |

---

*Source: `scripts_daa_saa_taa/saa_overlay_workflow.R`*
*Classification via `classify_universe()` in `screen_ticker.R`*
*Fingerprints via `screen_ticker()` in `screen_ticker.R`*
*Regime shading via `build_regime_table()` with T_FALL = 10%*
