---
title: "KF4 — Phase 2 Overlay Framework"
subtitle: "Stabilizer + Enhancer overlays on 6040Classic: design decisions, taxonomy, and methodology"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
---

<style>
h2 { border-bottom: 3px solid #1d3461; padding-bottom: 6px; margin-top: 40px; }
h3 { color: #1d3461; border-left: 4px solid #3b82f6; padding-left: 10px; margin-top: 24px; }
h4 { color: #374151; margin-top: 16px; }

.finding-box {
  background: #fffbeb; border-left: 5px solid #f59e0b;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.conclusion-box {
  background: #f0fdf4; border-left: 5px solid #16a34a;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.warn-box {
  background: #fef2f2; border-left: 5px solid #dc2626;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.phase-box {
  background: #eff6ff; border-left: 5px solid #1d4ed8;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.stab-box {
  background: #f0fdf4; border-left: 5px solid #15803d;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.enh-box {
  background: #eff6ff; border-left: 5px solid #1d4ed8;
  padding: 12px 18px; margin: 14px 0;
  border-radius: 0 4px 4px 0;
}
.fn-box {
  background: #f8f9fa; border-left: 4px solid #6b7280;
  padding: 8px 14px; margin: 8px 0 14px 0;
  font-family: monospace; font-size: 13px;
  border-radius: 0 4px 4px 0;
}
.type-a { background:#ede9fe; border-left:4px solid #7c3aed; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
.type-b { background:#fef9c3; border-left:4px solid #ca8a04; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
.type-c { background:#dcfce7; border-left:4px solid #16a34a; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
.hedge-full  { background:#dbeafe; border-left:4px solid #1d4ed8; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
.hedge-part  { background:#fef9c3; border-left:4px solid #ca8a04; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
.hedge-cond  { background:#fce7f3; border-left:4px solid #db2777; padding:6px 14px; margin:5px 0; border-radius:0 3px 3px 0; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 13,
  fig.height = 5
)
library(tidyverse)
library(xts)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(here)

if (!exists("project_tree"))       source(here("project_tree.R"))
if (!exists("etf_metadata"))       source(here(project_tree$scripts$init))
if (!exists("xts_ret"))            source(here(project_tree$scripts$wrangle))
if (!exists("build_regime_table")) source(here("scripts_spy_dd_regime/spy_dd_regime.R"))
if (!exists("rt") || !is.data.frame(rt))
  rt <- build_regime_table(xts_ret[, "SPY"])
if (!exists(".build_baseline"))    source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))

baseline_xts <- .build_baseline(xts_ret)
.ann_ret <- function(r) { r <- as.numeric(r[!is.na(r)]); prod(1+r)^(252/length(r))-1 }
.max_dd  <- function(r) { w <- cumprod(1+as.numeric(r[!is.na(r)])); min((w-cummax(w))/cummax(w)) }
.ann_vol <- function(r) sd(as.numeric(r), na.rm=TRUE) * sqrt(252)
.ir      <- function(r) { v <- .ann_vol(r); if(v==0) NA_real_ else .ann_ret(r)/v }

bl_ret <- .ann_ret(baseline_xts)
bl_dd  <- .max_dd(baseline_xts)
bl_vol <- .ann_vol(baseline_xts)
bl_ir  <- .ir(baseline_xts)
```

---

## 1. Strategic Context

### The core problem

The project starts from a confirmed empirical finding (KF6):

<div class="warn-box">
**KF6 — Regime-rotation TAA does NOT beat passive 60/40**
Walk-forward DAA produced ~6.7% vs passive ~7.6% over full history.
Root causes: (1) international diversification drag during US-dominated decade,
(2) regime detection lag — Fall fires only after 10% drawdown already absorbed,
(3) tilts too small to overcome base drag.

**Implication:** Do not attempt to improve return via regime rotation.
The right objective is **drawdown reduction at minimum return sacrifice.**
</div>

This finding locks the project objective:

> *"Given we accept SPY/IEF 60/40 return, what is the minimum-cost overlay that reduces MaxDD from ~`r sprintf("%.0f%%", abs(bl_dd)*100)` to ~15%?"*

### Phase architecture

The framework is split into three phases with clean separation of concerns:

```{r phase-table}
tibble(
  Phase   = c("1.0", "2.1", "2.2"),
  Name    = c("SAA Baseline", "Stabilizer Overlay", "Enhancer Overlay"),
  File    = c("01_saa_baseline.R", "05_dd_stabilizer.R", "08_enhancer_screen.R"),
  Objective = c(
    "Define the fixed benchmark — 6040Classic",
    "Reduce MaxDD with minimum return sacrifice",
    "Add alpha above baseline without harming MaxDD"
  ),
  Trigger = c("None — structural", "Fall correlation + hedge type", "Regime + 200DMA + IR signal")
) %>%
  kable(align = "llllll") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), font_size = 12) %>%
  row_spec(1, background = "#f1f5f9", bold = TRUE) %>%
  row_spec(2, background = "#f0fdf4") %>%
  row_spec(3, background = "#eff6ff")
```

---

## 2. Phase 1.0 — Baseline: 6040Classic

<div class="phase-box">
**6040Classic = 60% SPY + 40% IEF, buy-and-hold, no rebalancing signal.**

This is the benchmark all Phase 2 overlays must beat or at minimum not harm.
Built via `.build_baseline(xts_ret, W_EQ=0.60, W_FI=0.40)`.
</div>

```{r baseline-stats}
tibble(
  Metric = c("Composition", "Annualised Return", "MaxDD", "Ann Vol", "IR"),
  Value  = c(
    "60% SPY + 40% IEF",
    sprintf("%+.1f%%", bl_ret * 100),
    sprintf("%.1f%%",  bl_dd  * 100),
    sprintf("%.1f%%",  bl_vol * 100),
    sprintf("%.2f",    bl_ir)
  )
) %>%
  kable(align = "lr") %>%
  kable_styling(bootstrap_options = "condensed", full_width = FALSE, font_size = 12)
```

### Why SPY/IEF specifically

- **KF5 (international drag):** International diversification consistently underperformed during the US-dominated post-2010 decade. The optimizer would converge on SPY/IEF anyway.
- **KF1 (anti-diversification):** Low-ρ tickers do NOT reduce MaxDD — they often worsen it. The right hedge is a ticker with its own shallow DD, not simply low correlation.
- **Grounding by evidence:** Rather than leaving the baseline to judgement, KF5 + KF1 together justify SPY/IEF explicitly.

### Design decision: no rebalancing

The baseline is strict buy-and-hold. This is intentional:

- Rebalancing introduces a timing decision that complicates attribution
- All Phase 2 overlays are layered on top — the baseline drift is part of the "cost of doing nothing"
- Keeps the benchmark clean and reproducible

---

## 3. Phase 2.1 — Stabilizer Overlay

<div class="stab-box">
**Objective:** Find instruments that, when held at a small overlay weight w,
reduce the MaxDD of 6040Classic with minimum annualised return sacrifice.

**Formula:** `(1 − w) × 6040Classic + w × TICKER`

**Primary screen metric:** `dd_reduction = MaxDD(baseline) − MaxDD(blended at w=10%)`
**Efficiency metric:** `insurance_ratio = dd_reduction / max(return_drag, 0.001)`
</div>

### Hedge type taxonomy

The Fall correlation with the baseline is the primary classifier:

<div class="hedge-full">**FULL** — `fall_corr ≤ −0.10`
Negatively correlated with 6040Classic during Fall episodes. These instruments *appreciate*
(or fall much less) precisely when SPY is in freefall. True crisis hedges.
Examples: GLD, long-duration government bonds (TLT), managed futures.
</div>

<div class="hedge-part">**PARTIAL** — `−0.10 < fall_corr ≤ 0.30`
Uncorrelated to weakly positive in Fall. They reduce MaxDD through portfolio diversification
math rather than active crisis protection. They do not amplify drawdowns but do not offset them.
Examples: trend-following, broad commodities, low-vol equity.
</div>

<div class="hedge-cond">**CONDITIONAL** — `fall_corr > 0.30`
Positive Fall correlation — they move with SPY in a crisis. They may still reduce MaxDD at
the portfolio level through dilution (lower own-MaxDD), but offer no genuine crisis hedge.
Use with caution; return drag likely not compensated by protection in extreme events.
</div>

### Three-stage pipeline

```{r stab-pipeline}
tibble(
  Stage = c("Stage 1", "Stage 2", "Stage 3"),
  Function = c("build_stabilizer_candidates()", "score_stabilizer_context()", "select_stabilizers()"),
  Cadence  = c("Quarterly", "Monthly", "Monthly"),
  Purpose  = c(
    "Screen all tickers: dd_reduction > 0 at w=10%. Assign hedge type (FULL/PARTIAL/CONDITIONAL).",
    "Score each candidate on 4 current signals → stabilizer_score 0–100.",
    "Greedy top-N selection with pairwise ρ ≤ 0.80 guard."
  )
) %>%
  kable(align = "lllll") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), font_size = 12)
```

### Stage 2 signals — stabilizer context score

| ID | Signal | Direction | Rationale |
|---|---|---|---|
| s1 | Rolling 63d correlation vs baseline | **Lower = better** (inverted) | Is the ticker decorrelating from baseline RIGHT NOW? |
| s2 | Historical Fall DD reduction at 10% | **Higher = better** | Does it consistently protect in crises? |
| s3 | Own 200DMA trend signal | **Higher = better** | Is the ticker trending — likely to continue its protective role? |
| s4 | Own 63d annualised return | **Higher = better** | Is the stabilizer currently "working" (appreciating as hedge)? |

<div class="finding-box">
**Design note:** s1 is inverted before ranking — a stabilizer with *lower* rolling correlation right now is more valuable, not less. This is the opposite of an enhancer where higher rolling IR is better.
</div>

### Key analytical insight — stabilizer

<div class="conclusion-box">
**The KF1 anti-diversification finding directly informs stabilizer design.**
MaxDD of the blended portfolio is NOT minimised by minimising correlation.
It is minimised by finding tickers with (a) low own-MaxDD AND (b) negative Fall correlation.
A ticker with fall_corr = −0.6 and own_maxdd = −15% beats fall_corr = 0.0 and own_maxdd = −45%.
`insurance_ratio` captures this: it rewards DD reduction per unit of return drag, not raw correlation.
</div>

---

## 4. Phase 2.2 — Enhancer Overlay

<div class="enh-box">
**Objective:** Find instruments that, at overlay weight w, generate alpha *above*
the 6040Classic baseline — improving the risk-adjusted return profile.

**Formula:** `(1 − w) × 6040Classic + w × TICKER`

**Primary screen metric:** `full_ir = IR of spread (ticker − baseline)` over full history
**Admission logic:** OR-gate — pass if `full_ir > 0.20` OR `fall_excess > 0` OR `recov_excess > 0`
</div>

### Trigger type taxonomy

All spreads, IRs, and excess returns are computed versus **6040Classic** — not raw SPY.

<div class="type-a">**Type A — Unconditional**
`fall_excess ≥ −2pp vs baseline` → hold TICKER at fixed weight always.
No signal, no switching. Suitable for tickers with persistent IR advantage in all regimes.
Risk: no fall protection — the ticker is blended even during SPY crashes.
Examples: QQQ (historically), USMV.
</div>

<div class="type-b">**Type B — SPY Regime-conditional**
`fall_excess < −2pp vs baseline` → hold only in Recovery + Consolidation; revert to baseline in Fall.
Trigger: `build_regime_table()` → daily regime label via SPY drawdown.
Risk: **regime detection lag** — Fall fires only after ~10% drop already absorbed.
Examples: SMH, XLK (ALPHA in good regimes, LOSS in Fall).
</div>

<div class="type-c">**Type C — 200DMA signal-conditional**
Type B ticker with `trend_signals` data available → use 200DMA filter instead of hard regime switch.
Trigger: ticker price above its own 200DMA → blend ON; below → revert to baseline.
Risk: whipsaw cost in choppy markets (quantified in `06_trend_quality.R`).
Advantage over B: smoother entry/exit, fires earlier than regime detection lag.
</div>

**Threshold:** `FALL_TOLERANCE = −0.02` — the A/B boundary. Tickers losing more than 2pp vs baseline in Fall are classified B.

### Three-stage pipeline

```{r enh-pipeline}
tibble(
  Stage = c("Stage 1", "Stage 2", "Stage 3"),
  Function = c("build_enhancer_candidates()", "score_current_context()", "select_enhancers()"),
  Cadence  = c("Quarterly", "Monthly", "Monthly"),
  Purpose  = c(
    "OR-screen: admit if full_ir > floor OR fall/recov excess > 0. Assign trigger_type A/B.",
    "Score on 4 current signals → context_score 0–100. Refine B → C when trend_signals available.",
    "Greedy top-N selection with pairwise ρ guard (default ρ_max = 1.0, uncapped)."
  )
) %>%
  kable(align = "lllll") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), font_size = 12)
```

### Stage 2 signals — Context Score v1

| ID | Signal | Direction | Rationale |
|---|---|---|---|
| s1 | Rolling IR 63d of spread vs baseline | Higher = better | Is alpha present in the last quarter? |
| s2 | Historical excess return in current SPY regime | Higher = better | Does the ticker historically outperform in today's regime? |
| s3 | Own 200DMA trend signal (1/0) | Higher = better | Is the ticker trending — alpha likely to continue? |
| s4 | Spread momentum (20d mean > 63d mean) | Higher = better | Is alpha accelerating or fading? |

<div class="finding-box">
**This is Context Score Version 1.** The label "v1" is intentional — future versions may use different signals, windows, or non-linear combinations. Always label results with the version to keep comparisons unambiguous.
</div>

### Key analytical insight — enhancer

<div class="conclusion-box">
**Unconditional vs Conditionality — the central strategic theme.**

The 9-panel ticker fingerprint (especially Panel 3: Rolling L/S and Panel 6: Rolling IR) reveals
whether alpha is **stable across time** (Type A candidate) or **concentrated in specific regimes**
(Type B/C candidate).

- **Type A:** Rolling IR stays positive across all regimes. Example: QQQ in a rising-rate Consolidation.
- **Type B:** Rolling IR collapses or inverts during Fall. Example: SMH in 2022 — deep LOSS phase.
- **Type C:** Rolling IR tracks the 200DMA signal closely. The trend filter captures the same regime
  shift as Type B but fires earlier and exits cleaner.

A ticker that looks Type A in full-history stats may reveal Type B behaviour in the rolling panels.
**Always examine the rolling panels before assigning trigger type.**
</div>

---

## 5. Overlay Backtest Methodology

Once the screen has identified *which* tickers to use, the overlay backtest (`09_overlay_backtest.R`)
answers *how* each ticker should be held.

### Three return series per ticker

For each ticker at overlay weight w:

| Type | Return formula | Signal |
|---|---|---|
| A | `(1−w) × baseline + w × ticker` always | None |
| B | Blend in Recovery+Consolidation; `baseline` in Fall | `build_regime_table()` |
| C | Blend when ticker above 200DMA; `baseline` otherwise | `price ≥ MA(200)` |

### Recommendation criterion

| Role | Criterion | Rationale |
|---|---|---|
| Enhancer | Highest IR | Maximise risk-adjusted alpha above baseline |
| Stabilizer | Lowest MaxDD | Maximise drawdown protection |

### Optimal combination

`build_optimal_overlay()` picks the best type per ticker and assembles an equal-weight
combined portfolio, allowing direct comparison of the "smart" implementation vs pure baseline.

### Known limitations

<div class="warn-box">
1. **Type B lag risk:** The Fall regime fires after ~10% SPY drawdown is already absorbed.
   The switch-off is backward-looking. Type C partially mitigates this by using price momentum.

2. **Type C whipsaw:** In choppy markets (e.g. 2015–2016 range-bound SPY), the 200DMA
   trigger generates false exits and re-entries. Quantified in `06_trend_quality.R`.

3. **Look-ahead in screening:** The 3-stage screen uses full-history stats to identify candidates.
   Walk-forward validation is not yet implemented. Treat results as ex-post evidence,
   not as a live trading signal.

4. **Equal-weight combination:** The optimal portfolio uses equal weight across selected tickers.
   An optimised weight vector (e.g. minimum-variance or IR-weighted) is not yet implemented.
</div>

---

## 6. Summary of Analytical Conclusions

```{r conclusions}
tibble(
  `#`  = 1:7,
  Conclusion = c(
    "Regime-rotation TAA does not beat passive 60/40 (KF6). DD reduction — not return enhancement — is the right Phase 2 objective.",
    "The 6040Classic baseline (60% SPY + 40% IEF) is justified empirically by KF1 (anti-diversification) and KF5 (international drag). No further optimisation needed at Phase 1.",
    "Stabilizer hedge type is determined by Fall correlation: FULL (ρ ≤ −0.10) provides true crisis protection; PARTIAL and CONDITIONAL rely on portfolio math only.",
    "Enhancer trigger type is determined by Fall excess return vs baseline: Type A tickers are structurally robust; Type B/C tickers require switching to avoid Fall losses.",
    "The 200DMA signal (Type C) is generally superior to the hard regime switch (Type B): it fires earlier, exits cleaner, and avoids regime-detection lag at the cost of whipsaw in choppy markets.",
    "Context Score v1 (4-signal rank-sum) captures the CURRENT opportunity in each enhancer candidate. It must be re-run monthly; the quarterly candidate list is stable but the score is not.",
    "Rolling panels in the ticker fingerprint (Rolling L/S, Rolling IR) are the primary diagnostic for Type A vs B/C classification — full-history stats alone can mask regime-concentrated alpha."
  )
) %>%
  kable(align = "rl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%
  column_spec(1, bold = TRUE, width = "2em") %>%
  column_spec(2, width = "90%")
```

---

## 7. File Map

```{r file-map}
tibble(
  File = c(
    "scripts_daa_saa_taa/01_saa_baseline.R",
    "scripts_daa_saa_taa/05_dd_stabilizer.R",
    "scripts_daa_saa_taa/08_enhancer_screen.R",
    "scripts_daa_saa_taa/09_overlay_backtest.R",
    "scripts_daa_saa_taa/overlay_backtest_report.Rmd",
    "Rmd/stabilizer_screen_report.Rmd",
    "Rmd/enhancer_screen_report.Rmd",
    "utility/plot_xts_pair.R"
  ),
  Purpose = c(
    "Phase 1.0 — 6040Classic baseline definition",
    "Phase 2.1 — Stabilizer pipeline: candidates → scores → selection",
    "Phase 2.2 — Enhancer pipeline: candidates → scores → selection",
    "Type A/B/C backtest for any ticker set; exec summary + recommendations",
    "Self-contained HTML report: enhancer or stabilizer role, exec summary",
    "Parameterised stabilizer screen report",
    "Parameterised enhancer screen report",
    "9-panel ticker fingerprint (plot_xts_pair); role header box; extended flag"
  ),
  `Key function` = c(
    ".build_baseline()",
    "find_stabilizers()",
    "find_enhancers()",
    "run_overlay_backtest()",
    "params: role / tickers / overlay_w",
    "params: screen_weight / roll_win / n_select",
    "params: ir_floor / fall_tolerance / n_select",
    "plot_xts_pair(..., role=, extended=)"
  )
) %>%
  kable(align = "lll") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), font_size = 11) %>%
  column_spec(1, monospace = TRUE, color = "#1d3461") %>%
  column_spec(3, monospace = TRUE, color = "#15803d")
```

---

## 8. How to Run

### Interactive (RStudio console)

```r
source(here("00_libraries.R"))
source(here("scripts/00_init_universe.R"))
source(here("scripts/01_etf_wrangle.R"))
rt <- build_regime_table(xts_ret[, "SPY"])

# Phase 2.1 — Stabilizer
source(here("scripts_daa_saa_taa/05_dd_stabilizer.R"))
# → auto-runs find_stabilizers() + saves fingerprints to ticker_fingerprints/

# Phase 2.2 — Enhancer
source(here("scripts_daa_saa_taa/08_enhancer_screen.R"))
# → auto-runs find_enhancers() + saves fingerprints to ticker_fingerprints/

# Overlay backtest — Enhancer
source(here("scripts_daa_saa_taa/09_overlay_backtest.R"))
ob <- run_overlay_backtest(enh$selected$ticker, xts_ret, rt, role = "Enhancer")
```

### Rendered reports

```r
# Enhancer overlay report
rmarkdown::render(
  here::here("scripts_daa_saa_taa/overlay_backtest_report.Rmd"),
  output_file = here::here("scripts_daa_saa_taa/enhancer_overlay_report.html"),
  params = list(role = "Enhancer", overlay_w = 0.10),
  envir  = new.env()
)

# Stabilizer overlay report
rmarkdown::render(
  here::here("scripts_daa_saa_taa/overlay_backtest_report.Rmd"),
  output_file = here::here("scripts_daa_saa_taa/stabilizer_overlay_report.html"),
  params = list(role = "Stabilizer", overlay_w = 0.10),
  envir  = new.env()
)
```

---

*Document generated: `r Sys.Date()` — ETF Intelligence Project, Phase 2 session.*
